### 2016-04-04 CixPAG/CixPAG.Rmd
Initial scripts and results for ClimaTrans and CixPAG

*CixPAG results: 
-----------------------------------
title: "Extreme precip and heat waves: empirical-statistical downscaling and analysis"
author: "Rasmus Benestad"
date: "April 1, 2016"
output: pdf_document
---

CiXPAG will investigate the complex interactions between climate extremes, air pollution and  
agricultural ecosystems. Climate extremes (e.g., droughts, floods, heatwaves) and air pollution events often co-occur causing substantial losses in agricultural productivity. We do not yet fully understand how these stresses interact and what the impacts of the combined climate - air pollution effects may be for agricultural ecosystems in some of the most vulnerable parts of the world. This lack of knowledge is particularly challenging considering the threats that climate change and food security pose to society.  
The novel research proposed in CiXPAG will collect new experimental data and develop new  
modelling techniques to integrate knowledge on changes in climate extremes and air pollution to assess effects on agricultural productivity. Integration of farmers’ knowledge will enable the results to be translated into agricultural adaptation options within the particular socio-economic and political context.  
CiXPAG addresses several goals of the call - climate change, pollution and ecosystems - and its
objectives are on the forefront of international research. CiXPAG will build up expertise in a relatively new, but important research field for Norwegian climate research. Generated knowledge and methods can readily be applied to regions beyond the regional focus of this project. CiXPAG is an international collaboration of interdisciplinary experts, and is a gender balanced project with a female project leader.

**Initial analysis**

Analyse three northern cluseters of GHDN rain gauges: maps of mean, historical trends, mean seasonal variations and inter-annual variability in total precipitation, wet-day mean, and wet-day frequency. Monsoon duration. PCA-analysis. Explore teleconnections to MSLP and T2M/super-saturation-pressure.

```{r}  
## CixPAG:
library(esd)
## Retrieve daily precipitation from the GHCN data set:
load('gdcn.india.rda')

## Inspect the data
map(gdcn.india,cex=-2,FUN='wetfreq',showaxis=TRUE,
colbar=list(pal='precip',breaks=seq(0,0.5,by=0.025)))
```

Select various sub-regions for more detailed analysis

```{r} 
y1 <- subset(gdcn.india,is=list(lon=c(69,75),lat=c(20,25)))
y2 <- subset(gdcn.india,is=list(lon=c(75,85),lat=c(15,20)))
y3 <- subset(gdcn.india,is=list(lon=c(73,78),lat=c(26,37)))
y4 <- subset(gdcn.india,is=list(lon=c(83,89),lat=c(21,28)))
y5 <- subset(gdcn.india,is=list(lon=c(89,92),lat=c(23,28)))
```

Take one of the sub-regions and extract the monsoon season over the period 1901-1970 (these data do not cover the recent decades). Exclude stations with large gaps of missing data.

```{r} 
y <- subset(y4,it=c(1901,1970))
y <- subset(y,it=month.abb[6:9])
nv <- apply(coredata(y),2,'nv')
print(summary(nv))
y <- subset(y,is=(nv > quantile(nv,0.3)))
```

The group of station data series represents the predictand. Estimate the wet-day mean and frequency of each season and allow estimates for seasons with some missing data. At least 90 days are required with valid data to give a seasonal estimate. For the wet-day mean:

```{r}
mu <- annual(y,'wetmean',nmin=90)
mu <- pcafill(mu)
mu.pca <- PCA(mu)
plot(mu.pca)
```
Repeat the aggregation for the wet-day frequency:

```{r}
fw <- annual(y,'wetfreq',nmin=90)
fw <- pcafill(fw)
fw.pca <- PCA(fw)
plot(fw.pca)
```
**Get the description of large-scale

```{r}
## Get predictors
#t2m <- retrieve('data/ncep/air.mon.mean.nc',lon=c(30,120),lat=c(-10,30))
t2m <- retrieve('data/reanalysis20c/air.tropo.mon.mean.nc',lon=c(30,120),lat=c(-10,30))
slp <- retrieve('data/reanalysis20c/prmsl.mon.mean.nc',param='prmsl',lon=c(30,120),lat=c(-10,30))
## Estimate the vapour saturation pressure (es) over sea and take the mean over the season:
es <- aggregate(subset(C.C.eq(mask(t2m,land=TRUE)),it=month.abb[5:9]),year,'mean')
## Take the mean surface temperature
tas <- aggregate(subset(t2m,it=month.abb[5:9]),year,'mean')
eof.tas <- EOF(es)
eof.es <- EOF(tas)
eof.slp <- EOF(aggregate(subset(slp,it=month.abb[5:9]),year,'mean'))
index(eof.es) <- year(eof.es)
index(eof.tas) <- year(eof.tas)
index(eof.slp) <- year(eof.slp)
```

Use canonical correlation ananlysis (CCA) to assess the tele-connection between the wet-day mean and the large-scale vapour saturration pressure.

```{r, fig.width=6, fig.height=6}
mu.cca <- CCA(mu.pca,eof.es)
plot(mu.cca)
```

Repeat with the wet-day frequency and SLP, for which a stronger connection is expected:

```{r, fig.width=6, fig.height=6}
fw.cca <- CCA(fw.pca,eof.slp)
plot(fw.cca)
```

The predictand and predictor are ready for being used in downscaling. Do the analysis for the wet-day frequency:

```{r}
## Not many overlapping years: 1948-1970
ds.fw<- DS(fw.pca,eof.slp)
plot(ds.fw)
```

```{r}
## Not many overlapping years: 1948-1970
ds.mu <- DS(mu.pca,eof.es,eofs=1:7)
#ds2 <- DS(mu.pca,eof2,eofs=1:3)
#ds.fw<- DS(fw.pca,eof.slp)
plot(ds.mu)
#plot(ds2)
#plot(ds.fw)
```

Gridded daily maximum temperature data (IMD?): explore dependency of number of hot days to mean seasonal temperature and the shape of PDF. Near-normal? Possible to infer heat wave statistics using DSensemble. Aggregated spatial temperrature over India from the CMIP5 runs.

### 2016-04-04 CixPAG/paper58.Rmd
Initial scripts and results for ClimaTrans and CixPAG

*Benestad et al - Upper-limit estimation of future precipitation return-values* 
-----------------------------------
title: "Analysis and calculations in manuscript"
author: "Rasmus Benestad"
date: "April 4, 2016"
output: pdf_document
---

First pre-amble that checks whether the esd-package is installed and installs it if needed.

```{r}
rm(list=ls())
#xlim <- c(-10,40); ylim <- c(35,70)
xlim <- c(0,35); ylim <- c(45,70)
readecad <- FALSE
figshare=TRUE
nmin=50

## Check if you need to get the esd-package:
install.esd <- ("esd" %in% rownames(installed.packages()) == FALSE)

if (install.esd) {
  print('Need to install the esd package')
  ## Need online access.
  ## Use the devtools-package for simple facilitation of installing.
  if ("devtools" %in% rownames(installed.packages()) == FALSE)
    install.packages('devtools')
  library(devtools)
  ## Install esd directly from github
  install_github('metno/esd')
  print('The latest version of esd has been installed')
}

## Start the esd-library:
library(esd)
```

Below are definition of a number of functions used in the data processing and analysis. The first are simple functions used in ploting.

**Simple

```{r}
## Standardisation
s <- function(x) {
    z <- coredata(x)
    y <- (z - mean(z,na.rm=TRUE))/sd(z,na.rm=TRUE)
    attributes(y) <- NULL
    coredata(x) <- y
    return(x)
}
```

**Functions which estimate various statistics**

```{r}
## Return statistics for mean,min,max,wettest month,driest month
muclim <- function(x) {
  y <- coredata(x)
  iX <- mean((1:12)[is.element(y,max(y))])
  iN <- mean((1:12)[is.element(y,min(y))])
  stats <- c(mean(y,na.rm=TRUE),min(y,na.rm=TRUE),max(y,na.rm=TRUE),iX,iN)
  names(stats) <- c('mean','min','max','wettest month','driest month')
  stats
}

## Produce a set of percentiles and their counterparts for the exponential distribution
qqexp <- function(x,x0=1) {
  x[x < x0] <- NA
  if (sum(is.finite(x))>0) {
    mu <- mean(x,na.rm=TRUE)
    qx <- quantile(x,probs=seq(0,1,length=101),na.rm=TRUE)
    qy <- -log(1-seq(0,1,length=101))*mu
  } else {
    qx <- rep(NA,101); qy <- rep(NA,101)
  }
  return(cbind(qx,qy))
}

## Skill associated with predicting the wet-day mean mu
muskill <- function(x) {
  ## Estimate the skill of the calibration:
  r2 <- round(summary(lm(y ~ x,data=x))$r.squared,3)
  ## Negative slopes are not credible:
  if (summary(lm(y ~ x,data=x))$coefficients[2] < 0) r2 <- 0
  r2
}

## A function for extracting the regression coefficients and their
## error terms.
beta <- function(x,verbose=FALSE) {
  wc.model <- lm(y ~ x, data=x)
  if (verbose) print(summary(wc.model))
  beta <- summary(wc.model)$coefficients[c(2,4)]
  return(beta)
}
```

**Data input and procesing**

## Read the data from the CMIP5 GCMs
```{r}
readGCMs <- function(path='CMIP5.monthly/rcp45/',pattern='tas',
                     lon=c(-100,30),lat=c(0,40)) {
    ncfiles <- list.files(path=path,pattern=pattern,full.names=TRUE)
    n <- length(ncfiles)
    print(paste(n,'netCDF files'))
    X <- matrix(rep(NA,n*201),n,201)
    for (i in 1:n) {      print(paste(i,ncfiles[i]))
      ## Spatial average:                    
      gcm <- annual(spatial.avg.field(C.C.eq(retrieve(ncfiles[i],
                                                      lon=lon,lat=lat))))
      i1 <- is.element(1900:2100,year(gcm))
      i2 <- is.element(year(gcm),1900:2100)
      X[i,i1] <- coredata(gcm)[i2]
    }
    ## Extract the 5 & 95 percentile and the ensemble mean:
    print('Extract the 5 & 95 percentile and the ensemble mean')
    x <- apply(X,2,function(x) c(quantile(x,probs=c(0.05,0.95),na.rm=TRUE),mean(x,na.rm=TRUE)))
    print(dim(x))
    names(x) <- c('q05','q95','mean')
    x <- zoo(t(x),order.by=1900:2100)
    w2000 <- X[,is.element(1900:2100,2000)]
    w2050 <- X[,is.element(1900:2100,2020)]
    w2100 <- X[,is.element(1900:2100,2100)]
    plot(x,plot.type='single')
    attr(x,'path') <- path
    attr(x,'2000') <- w2000
    attr(x,'2050') <- w2050
    attr(x,'2100') <- w2100
    attr(x,'N') <- n
    attr(x,'region') <- paste(lon,lat,collapse=' ')
    return(x)
}
```

**model calibration and regression analysis**

```{r}
## Calibrate a model for the wet-day mean mu using temperature as input
mucal <- function(x,pre=NULL,lon=c(-100,30),lat=c(0,40),
                  plot=FALSE,verbose=FALSE) {

## If no pre, use the crude NCEP-reanalysis provided in esd 
  if (is.null(pre)) {
    if (verbose) print('default predictor')
      t2m <- t2m.NCEP(lon=lon,lat=lat)
      pre <- spatial.avg.field(C.C.eq(t2m))
      if (plot) plot(EOF(t2m))
  } else
      if (is.character(pre))
          pre <- spatial.avg.field(C.C.eq(retrieve(ncfile=pre,lon=lon,lat=lat))) else
  if (inherits(pre,'field')) {
      if (is.T(pre)) pre <- spatial.avg.field(C.C.eq(pre)) else
      pre <- spatial.avg.field(pre)
  } else if (inherits(pre,'station')) pre <- pre
  z <- aggregate(pre,by=month,FUN='mean')

  cal <- data.frame(y=coredata(x),x=coredata(z))
  attr(cal,'standard.error') <- attr(x,'standard.error')
  stats <- cor.test(cal$y,cal$x)
  wc.model <- lm(y ~ x, data=cal)
  if (plot) {
      dev.new()
      par(bty='n',cex.sub=0.7,col.sub='grey40')
      ylim <- range(cal$y,na.rm=TRUE); xlim=range(cal$x,na.rm=TRUE)
      dy <- diff(ylim)/25
      plot(cal$x,cal$y,pch=19,cex=1.5,col='grey',
           ylab=expression(paste(mu,' (mm/day)')),
         xlab=expression(paste(e[s],' (Pa)')),
         ylim=ylim,xlim=xlim,
         main='"Worst-case" fit based on seasonal variations',
         sub=paste(loc(x),' (',round(lon(x),2),'E/',round(lat(x),2),'N; ',
             alt(x),'m.a.s.l.)',sep=''))
      segments(x0=cal$x,y0=cal$y,x1=cal$x,y1=cal$y+2*attr(x,'standard.error'),
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x,y1=cal$y-2*attr(x,'standard.error'),
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x+2*attr(z,'standard.error'),y1=cal$y,
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x-2*attr(z,'standard.error'),y1=cal$y,
               col='grey')
      points(cal$x,cal$y,pch=19,cex=1.5,col='grey')
      grid()
      abline(wc.model)
      text(xlim[1],ylim[2],paste('Correlation=',round(stats$estimate,2),
                                 '(','p-value=',
                                 100*round(stats$p.value,4),'%)'),
           pos=4,cex=0.7,col='grey')
      text(xlim[1],ylim[2]-dy,paste('Regression: y=',
                                    round(wc.model$coeff[1],4), '+',
                                    round(wc.model$coeff[2],4), 'x (R2=',
                                    round(summary(wc.model)$r.squared,2),')'),
           pos=4,cex=0.7,col='grey')
      par(new=TRUE,fig=c(0.5,0.97,0.1,0.5),yaxt='n',xaxt='n',xpd=TRUE,
          cex.axis=0.7,col.axis='grey')
      plot((cal$x - mean(cal$x))/sd(cal$x),type='l',lwd=2,
           ylab='',xlab='',col=rgb(0.6,0.3,0))
      lines((cal$y - mean(cal$y))/sd(cal$y),type='l',lwd=2,col=rgb(0,0.3,0.6))
      par(xaxt = "s")
      axis(1,at=1:12,labels=month.abb,col='grey')
  }
  invisible(cal)
}
```

Functions for projection and prediction for the future/past

```{r}
## Projection based on the calibration with the annual cycle:
muproject <- function(x,gcm,verbose=FALSE,prct=TRUE) {
  #print('projection')
  wcmodel <- lm(y ~ x, data=x)
  if (verbose) {
    print(summary(wcmodel))
    print(dim(gcm))
  }
  pq05 <- data.frame(x=coredata(gcm[,1]))
  pq95 <- data.frame(x=coredata(gcm[,2]))
  pmea <- data.frame(x=coredata(gcm[,3]))
  y <- cbind(predict(wcmodel,newdata=pq05),
             predict(wcmodel,newdata=pq95),
             predict(wcmodel,newdata=pmea))
  if (verbose) print(dim(y))
  if (prct) {
#    ii <- is.element(year(gcm),1961:1990)
    ii <- is.element(year(gcm),2000:2010)
    bline <- mean(y[ii,3])
    #bline <- colMeans(y[ii,])
    #browser()
    #print(length(bline))
    #y <- t(100*t(y)/bline)
    y <- 100*y/bline
  }
  y <- zoo(y,order.by=index(gcm))
  names(y) <- names(gcm)
  return(y)          
}

## Predict values of the wet-day mean mu taking a given predictor
mupredict <- function(x,pre,verbose=FALSE,prct=TRUE) {
  if (verbose) print('projection')
  wcmodel <- lm(y ~ x, data=x)
  if (verbose) {
    print(summary(wcmodel))
  }
  eval <- data.frame(x=coredata(pre))
  y <- predict(wcmodel,newdata=eval)
  if (prct) {
    ii <- is.element(year(pre),1961:1990)
    bline <- mean(y[ii])
    if (verbose) print(length(bline))
    y <- 100*y/bline
  }
  y <- zoo(y,order.by=index(pre))
  names(y) <- 'mu'
  return(y)          
}
```

**Generation of graphics/maps**

```{r}
## Generate a map for the PC weights for the different modes of the
## wet-day mean mu annual cycle.
mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
  #dev.new()
  r2 <- as.numeric(r2)
  colpc <- rev(colscal("t2m",n=100))
  cz <- round(100*abs(pca$v[,ipca])/quantile(abs(pca$v[,ipca]),0.95))
  cz[cz > 100] <- 100; cz[cz < 1] <- 1
  #print(c(length(cz),sum(is.finite(cz))))
  col <- colpc[cz]
  pch <- rep(19,length(r2))
  mo <- c(r2) > 0.4
  hi <- c(r2) > 0.6
  pch[!hi] <- 1
  pch[!mo] <- 4
  cex <- 1.5*c(r2) + 0.2
  #print(c(sum(is.finite(pch)),length(col),sum(is.finite(cex))))

  map(mu, xlim=xlim,ylim=ylim,bg='grey80',new=FALSE)
  points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
  #text(lon(mu),lat(mu),r2,cex=0.7); print(table(col))
  par(xpd=TRUE)
  text(10,73,paste('Annual cycle in PC',ipca,
      'with variance of',round(100*pca$d[ipca]^2/sum(pca$d^2)),'%'),pos=4)

  colbar(pretty(pca$v[,ipca],n=100),colpc,fig = c(0.05, 0.1, 0.05, 0.2))

  par(xaxt = "n", yaxt = "s", fig = c(0.05,0.25,0.80,0.95),
      mar = c(0, 1, 0, 0), new = TRUE, las = 1, cex.axis = 0.5,bty='n')
  plot(pca$u[,ipca],type='l',lwd=3,col="red")
  par(xaxt = "s")
  axis(1,at=1:12,labels=month.abb)
}

## Plot shaded regions
shade <- function(x,col=rgb(0.5,0.5,0.5,0.3),border=NULL) {
  t <- index(x)
  if (is.null(border)) border <- col
  y <- coredata(x)
  polygon(c(t,rev(t)),c(y[,1],rev(y[,2])),col=col,border=border)
  lines(t,y[,3],lwd=5,col=col)
}

## Estimate correlation between different sets of wet-day mean mu
## The function is designed to be used in apply for best efficiency
cormu <- function(x) {
  n <- length(x); nh <- n/2
  x1 <- x[1:nh]; x2 <- x[(nh+1):n]
  ok <- is.finite(x1) & is.finite(x2)
  return(cor(x1[ok],x2[ok]))
}
```

The definition of functions is followed by code that carry out the analysis based on these.  First get the data needed:

```{r}
##-----------------------------------------------------------------------------------------

## Need to obtain some of the data files - fetch from Figshare:

if (!file.exists("mu.worstcasemu.rda") & figshare) {
  download.file("http://files.figshare.com/2193033/mu.worstcasemu.rda",destfile="mu.worstcasemu.rda") 
}

if (!file.exists("pre.worstcasemu.rda") & figshare) {
  download.file("http://files.figshare.com/2193038/pre.worstcasemu.rda",destfile="pre.worstcasemu.rda") 
}

if (!file.exists("cmip5.rda") & figshare) {
  download.file("http://files.figshare.com/2193041/cmip5.rda",destfile="cmip5.rda") 
}

## Preparations that only needs to be done once. 
if (readecad) {
## This section generates a processed data file from scratch using ECA&D data:
  pr <- station(src=c('metnod','ecad'),param='precip',nmin=nmin,it=c(1961,2010),lon=xlim,lat=ylim)
  pr <- subset(pr,it=c(1961,2014))
  nt <- apply(pr,2,FUN='nv')
  pr <- subset(pr,is=(nt >=19000))
  cpr <- coredata(pr)
  cpr[cpr > 250] <- NA
  coredata(pr) <- cpr
  save(file='pr.worstcasemu.rda',pr)
  file.remove('mu.worstcasemu.rda')
}

if (!file.exists('mu.worstcasemu.rda')) {

## pr.worstcasemu.rda is a huge file with daily precipdata based on ECA&D - generated above
  load('pr.worstcasemu.rda')
```

Then the predictand data is processed: estimate annual mean aggregates and the mean seasonal cycle. Remove stations with large gaps of missing values.

```{r}
  ## Strip away stations with a lot of missing data

## Time series of the annual wet-day freq & mean - for evaluation
## Randomly sub-sample due to excessive volume:
  FW <- annual(pr,FUN='wetfreq',nmin=350)
  MU <- annual(pr,FUN='wetmean',nmin=350)
  fw <- aggregate(pr,month,FUN='wetfreq')
  mu <- aggregate(pr,month,FUN='wetmean')

  n <- apply(pr,2,FUN='nv')
  y1 <- subset(subset(pr,is=(n==max(n))),is=1)
  save(file='mu.worstcasemu.rda',mu,MU,fw,FW,y1,n)
} else load('mu.worstcasemu.rda')

## Keep stations with no missing data and stations
## without suspect outlier trends
nok <- n
ok <- apply(mu,2,function(x) sum(!is.finite(x))==0) &
      apply(fw,2,function(x) sum(!is.finite(x))==0) &
      abs(apply(MU,2,FUN='trend.coef')) <= 1 &
      abs(apply(FW,2,FUN='trend.coef')) <= 0.02
mu <- subset(mu,is=ok) # REB 2015-03-01
MU <- subset(MU,is=ok) # REB 2015-03-01
fw <- subset(fw,is=ok)
FW <- subset(FW,is=ok)

## Remove stations with little data
nval <- apply(MU,2,nv)
mu <- subset(mu,is=(nval > 50))
MU <- subset(MU,is=(nval > 50))
fw <- subset(fw,is=(nval > 50))
FW <- subset(FW,is=(nval > 50))
nval <- apply(FW,2,nv)
mu <- subset(mu,is=(nval > 50))
MU <- subset(MU,is=(nval > 50))
fw <- subset(fw,is=(nval > 50))
FW <- subset(FW,is=(nval > 50))
```

Get and process the predictor data.

```{r}
print("predictor")
if (!file.exists('pre.worstcasemu.rda')) {
  t2m <- retrieve('data/ncep/air.mon.mean.nc',lon=c(-100,30),lat=c(0,40))
  pre <- spatial.avg.field(C.C.eq(t2m))
  attr(pre,'region') <- '100W,30E/0N,40N'
  save(file='pre.worstcasemu.rda',pre)
} else load('pre.worstcasemu.rda')
```

Now the data is ready for the analysis. Calibrate the regression models and extract the regression coeffieients. Use 'apply' to speed up the process for multiple stations.

```{r}
## Extract the monthly aggregates for all stations
print('apply mucal')
V <- apply(mu,2,FUN='mucal',pre=pre)
Beta <- lapply(V,FUN='beta')

## Collect the R-squared statistics from lm(y ~ x) for each site
print('muskill')
r2 <- lapply(V,muskill)
```

**Plot figure 1**

```{r,echo=TRUE}
## Extract the mean, min, max, wettest month, and driest months in terms of mu
## for all stations:
print('muclim')
X <- apply(mu,2,FUN='muclim')
print(table(X[4,]))
wmns <- as.numeric(rownames(table(X[4,])))
nc <- max(wmns) - min(wmns) + 1
cols <- colscal(n=nc)
col1 <- cols[c(X[4,])]
cex <- 1.5*n/max(n)

## The relationship between mu and e_s for one station
## to show the calibration procedure
is <- (1:length(n))[X[4,]==8][1]
print(paste('plot mucal - is=',is))
mucal(subset(mu,is=is),pre=pre,verbose=TRUE,plot=TRUE)

figlab('Figure 1')
#dev.copy2pdf(file='figure1.pdf')
```

**Figure 2a+b**

```{r}
## Remove locations with missing values for PCA 
print('Mu - matrix for PCA')
Mu <- as.matrix(coredata(mu))

## Anomalies wrt the mean value at each location.
Mu <- apply(Mu,2,function(x) (x - mean(x)))
pca <- svd(Mu)

## Plot maps with PCs:
print('Maps with PCs')

mupcamap(mu,pca,1,xlim,ylim,r2)
figlab('Figure 2a')
#dev.copy2pdf(file='figure2a.pdf')
mupcamap(mu,pca,2,xlim,ylim,r2)
figlab('Figure 2b')
#dev.copy2pdf(file='figure2b.pdf')
```

**Figure 3**

```{r,echo=TRUE}
print('Variance accounted for by the modes:')
print(round(100*pca$d**2/sum(pca$d**2),1))

## The sign of PCs is arbitrary...
print(cor.test(pca$v[,1],as.numeric(r2)))
print(cor.test(pca$v[,2],as.numeric(r2)))
print(cor.test(pca$v[,3],as.numeric(r2)))

## Estimate trend statistics for both fw and mu:
## Use subset - in the PCA there is something strange in the 1990s.
fw.trend <- 100*apply(subset(FW,it=c(1950,1990)),2,'trend.coef')/
                      apply(FW,2,'mean',na.rm=TRUE)
mu.trend <- 100*apply(subset(MU,it=c(1950,1990)),2,'trend.coef')/
                      apply(MU,2,'mean',na.rm=TRUE)

#dev.new(); diagnose(MU)

## Select only the sites which have an R2 greater than 0.6:
is <- (1:length(r2))[as.numeric(r2)> 0.6]
mux <- subset(mu,is=is)
MUx <- subset(MU,is=is)
FWx <- subset(MU,is=is)
Xx <- X[,is]
Vx <- V[is]

## Read the GCM ensembles andestimate the 5 and 95 percentiles as well as
## the ensemble mean.

## Get the annual mean temperature from GCM ensembles:
if (!file.exists('cmip5.rda')) {
  print('get CMIP5 RCP4.5')
  rcp4.5 <- readGCMs(path='CMIP5.monthly/rcp45/',pattern='tas_Amon_ens_rcp')
  print('get CMIP5 RCP8.5')
  rcp8.5 <- readGCMs(path='CMIP5.monthly/rcp85/',pattern='tas_Amon_ens_rcp')
  print('get CMIP5 RCP2.6')
  rcp2.6 <- readGCMs(path='CMIP5.monthly/rcp26/',pattern='tas_Amon_ens_rcp')
  save(file='cmip5.rda',rcp4.5,rcp8.5,rcp2.6)
} else load('cmip5.rda')

## derive time series for each location:
Z.rcp4.5 <- lapply(Vx,'muproject',rcp4.5)
Z.rcp8.5 <- lapply(Vx,'muproject',rcp8.5)
Z.rcp2.6 <- lapply(Vx,'muproject',rcp2.6)
t <- index(rcp4.5)

## Map showing RCP4.5 ensemble mean and the upper 95% change in the
## outer part of the symbol. Also an insert with box-plot diagram
## showin the other RCPs.

## Estimates for mu in 2010:
mu2010.rcp4.5 <- lapply(Z.rcp4.5,'window',start=2010,end=2010)
mu2010.rcp8.5 <- lapply(Z.rcp8.5,'window',start=2010,end=2010)
mu2010.rcp2.6 <- lapply(Z.rcp2.6,'window',start=2010,end=2010)
x2010 <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[3]]))
x2010u <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[2]]))
y2010 <- as.numeric(lapply(mu2010.rcp8.5,function(x) x[[3]]))
y2010u <- as.numeric(lapply(mu2010.rcp8.5,function(x) x[[2]]))
z2010 <- as.numeric(lapply(mu2010.rcp2.6,function(x) x[[3]]))
z2010u <- as.numeric(lapply(mu2010.rcp2.6,function(x) x[[2]]))

## Repeat for 2100:
mu2100.rcp4.5 <- lapply(Z.rcp4.5,'window',start=2100,end=2100)
mu2100.rcp8.5 <- lapply(Z.rcp8.5,'window',start=2100,end=2100)
mu2100.rcp2.6 <- lapply(Z.rcp2.6,'window',start=2100,end=2100)
x2100 <- as.numeric(lapply(mu2100.rcp4.5,function(x) x[[3]]))
x2100u <- as.numeric(lapply(mu2100.rcp4.5,function(x) x[[2]]))
y2100 <- as.numeric(lapply(mu2100.rcp8.5,function(x) x[[3]]))
y2100u <- as.numeric(lapply(mu2100.rcp8.5,function(x) x[[2]]))
z2100 <- as.numeric(lapply(mu2100.rcp2.6,function(x) x[[3]]))
z2100u <- as.numeric(lapply(mu2100.rcp2.6,function(x) x[[2]]))

## Data frams with changes in percentages:
mu.2100 <- data.frame(mean.RCP4.5=x2100 - x2010,q95.RCP4.5=x2100u - x2010u,
                      mean.RCP2.6=z2100 - z2010,q95.RCP4.5=z2100u - z2010u,
                      mean.RCP8.5=y2100 - y2010,q95.RCP8.5=y2100u - y2010u)

## Fig 3
## Plot a map of projected values:
#dev.new()
cols <- colscal(n=100,col='precip')
cx2100 <- round(x2100 - x2010)
cx2100u <- round(y2100u - y2010u)
cx2100[cx2100 < 1] <- 1; cx2100[cx2100 > 100] <- 100
cx2100u[cx2100u < 1] <- 1; cx2100u[cx2100u > 100] <- 100
colx <- cols[cx2100]
coly <- cols[cx2100u]
Cex <- 1.25

print("Plot a map of projected values:")
map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',gridlines=FALSE,
    colbar=list(col=cols,n=12,type="p",h=0.6,v=1))
points(lon(mux),lat(mux),pch=19,col=colx,cex=Cex)
points(lon(mux),lat(mux),pch=21,col=coly,cex=Cex,lwd=2)
par(xpd=TRUE)
text(20,73,'Wet-day mean: 2100')
legend(20,32,c(expression(bar(x)),expression(q[95])),
       pch=c(21,19),bty='n',col='grey',text.col='grey',horiz=TRUE)

colbar(pretty(c(x2100u- x2010u, x2100- x2010),n=15),cols,
       fig = c(0.05, 0.1, 0.05, 0.2))

par(new=TRUE,fig = c(0.05, 0.4, 0.75, 0.975),
    cex.axis=0.75,mar=c(1,1,0.1,0.1),xaxt='n')
boxplot(mu.2100,col=c(rep(rgb(0.5,0.5,0.5,0.3),2),
                      rep(rgb(0.5,1,0.5,0.3),2),
                      rep(rgb(1,0.5,0.5,0.3),2)))
par(xaxt='s')
axis(1,c(1,3,5),labels=c('RCP4.5','RCP2.6','RCP8.5'))
grid()

figlab('Figure 3')
#dev.copy2pdf(file='figure3.pdf')
```

**Supporting material/analysis (SM)**

```{r,echo=TRUE}
## Evaluation:
## Use the calibration strategy to predict the annual mu based on the
## predictor (t2m -> e_s)
print('evaluation: correlation')
#mu.eval <- lapply(Vx,'mupredict',annual(pre))
mu.eval <- lapply(Vx,'mupredict',annual(pre),prct=FALSE)
m <- length(mu.eval); n <- length(mu.eval[[1]])
MUz <- matrix(unlist(mu.eval),n,m)
MUz <- zoo(MUz,order.by=index(mu.eval[[1]]))
MUz <- subset(MUz,it=MUx)
MUx <- subset(MUx,it=MUz)
FWx <- subset(FWx,it=MUz)
r.eval <- apply(rbind(coredata(MUz),coredata(MUx)),2,'cormu')
print(summary(r.eval))

## ----------- Supporting Material ------------------------

#dev.new()
par(xaxt='n',yaxt='n',bty='n')
plot(c(0,1),c(0,1),type='n',xlab='',ylab='')
text(0.5,0.5,'Supporting figures',cex=2,font=2)
dev.copy2pdf(file='figureSM0.pdf')

## Plot the statistics of R2:
#dev.new()
hist(100*as.numeric(r2),breaks=seq(0,100,by=5),lwd=2,col=rgb(0,0.3,0.5),
     xlab=expression(paste(R^2,' (%)')),freq=TRUE,
     main="Summary of regression scores")
grid()
figlab('Figure SM1')
#dev.copy2pdf(file='figureSM1.pdf')

## Example of estimates for 2050:
print('Example plot - evolution')
#dev.new()
par(bty='n')
i <- 1
N <- length(Z.rcp4.5)
plot(Z.rcp4.5[[i]],plot.type='single',lty=c(2,2,1),lwd=c(1,1,2),
     ylab='%',ylim=c(80,150),
     main=paste('Wet-day mean at',loc(subset(mux,is=i))))
shade(Z.rcp4.5[[i]],col=rgb(0.5,0.5,0.5,0.3))
shade(Z.rcp8.5[[i]],col=rgb(1,0.5,0.5,0.3))
shade(Z.rcp2.6[[i]],col=rgb(0.5,1,0.5,0.3))
grid()
figlab('Figure SM2')
#dev.copy2pdf(file='figureSM2.pdf')

## Trend in projected wet-day mean
trendbeta <- unlist(lapply(Z.rcp4.5,function(x) trend.coef(x[,3])))
## Strange results:
print((1:N)[trendbeta < 0])

## test: Does the model predict observed trends?
print('evaluation: trends')

## Make sure to compare series with data for same times
mask <- !is.finite(coredata(MUx))
class(MUz) <- class(MUx)
muz <- coredata(MUz)
muz[mask] <- NA; dim(muz) <- dim(MUz)
coredata(MUz) <- muz

## Only look at stations with more than 50 years with data
ok <- (apply(coredata(MUx),2,nv) > 50)
MUz <- subset(MUz,is=ok)
MUx <- subset(MUx,is=ok)
FWx <- subset(FWx,is=ok)

#dev.new()
trend.mux <- apply(MUx,2,'trend.coef')
trend.pre <- apply(MUz,2,'trend.coef')
trenderr.mux <- apply(MUx,2,'trend.err')
trenderr.pre <- apply(MUz,2,'trend.err')

## Need to get a picture whether the predictions gives a plausible
## upper limit.
trend.sense <- data.frame(x=c(-trend.mux,trend.mux),
                          y=c(-trend.pre,trend.pre))

xylim <- max(abs(c(trend.mux,trend.pre)))*c(-1,1)

par(bty='n',col.sub='grey',cex.sub=0.8)
plot(trend.mux,trend.pre,pch=19,col=rgb(0.6,0.2,0,0.3),cex=1.5,
     main=expression(paste('Trends in ',mu,': observed and predicted upper limit')),
     xlab='Observed trend (mm/decade)',ylab='predicted trend (mm/decade)',
     xlim=xylim,ylim=xylim,
     sub=paste('Mean correlation for local year-to-year variations over t=[',
       start(MUx),',',end(MUx),
       '] is ',round(mean(r.eval),2),' (',round(quantile(r.eval,0.05),2),', ',
       round(quantile(r.eval,0.95),2),
       ')',sep='') )
grid()

polygon(c(xylim[1],xylim[2],xylim[1],xylim[1]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
        col=rgb(0.2,0.6,1,0.1),border=NA)
polygon(c(xylim[1],xylim[2],xylim[2],xylim[2]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
        col=rgb(1,0.2,0.2,0.1),border=NA)
points(trend.mux,trend.pre,pch=1,col=rgb(0,0,0,0.1),cex=1.5)

## Plot error bars 
apply(rbind(trend.mux,trend.pre,trenderr.mux,trenderr.pre),2,
      FUN=function(x) {lines(x[1]+c(-2,2)*x[3],x[2]+c(0,0),col=rgb(0.6,0.2,0,0.1))
                       lines(x[1]+c(0,0),x[2]+c(-2,2)*x[4],col=rgb(0.6,0.2,0,0.1))
                       lines(x[1]+c(-1,1)*0.01,x[2]+c(2,2)*x[4],col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(-1,1)*0.01,x[2]+c(-2,-2)*x[4],col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(-2,-2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(2,2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
                     })

figlab('Figure SM3')
#dev.copy2pdf(file='figureSM3.pdf')

## Map showing trends in mu
#dev.new()
map(MU,FUN='trend',colbar=list(breaks=seq(-1,1,length=21),rev=TRUE),cex=cex)
figlab('Figure SM4',ypos=0.999)
figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
#dev.copy2pdf(file='figureSM4.pdf')

## 99-percentile wet-day mean and typical wet-day frequency
## The Tellus paper on specification approximately annual maximum.
## Prob(X>x) for annual maximum for 24hr precip
## Pr(X > x) = 1/365.25 = fw*(1-p): p = 1 - 1/(365.25*fw)
#dev.new()
par(bty='n')
W <- apply(fw,2,FUN='muclim')
hist(100*(1-1/(W[1,]*365.25)),col='steelblue',lwd=2,breaks=seq(90,100,by=0.1),
     main='Wet-day percentile for annual maximum 24-precipitation',
     xlab='p (%)')
grid()

figlab('Figure SM5')
dev.copy2pdf(file='figureSM5.pdf')

## Statistics of trend in wet-day frequency
print('Wet-day frequency statistics')
#dev.new()
hist(fw.trend,breaks=seq(-50,50,by=1),col='grey',lwd=2,
     main='Trend in wet-day frequency',
     xlab=expression(paste(f[w],' (%/decade)')))
grid()
figlab('Figure SM6')
#dev.copy2pdf(file='figureSM6.pdf')

## Map showing trends in fw
#dev.new()
map(FW,FUN='trend',colbar=list(breaks=seq(-0.05,0.05,length=21),rev=TRUE),cex=cex)
figlab('Figure SM7',ypos=0.999)
figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
#dev.copy2pdf(file='figureSM7.pdf')

## Figure SM8.
## test: See if the quantiles are consistent when the mean mu varies.
#dev.new()
qtest <- aggregate(y1,year,FUN='qqexp')
qx <- c(coredata(qtest[,1:101]))
qy <- c(coredata(qtest[,102:202]))

par(bty='n')
plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
     main='Test: exponential distribution & changing mean',
     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
lines(c(0,40),c(0,40),col='grey')
figlab('Figure SM8')
#dev.copy2pdf(file='figureSM8.pdf')

pr.mean <- aggregate(y1,by=month,FUN='mean')
pr.mu <- aggregate(y1,by=month,FUN='wetmean')
pr.fw <- aggregate(y1,by=month,FUN='wetfreq')
y1.l <- spell(y1,threshold=1)
pr.wet <- aggregate(subset(y1.l,is=1),by=month,FUN='mean')
pr.dry <- aggregate(subset(y1.l,is=2),by=month,FUN='mean')

#dev.new()
par(bty='n',xaxt='n')
plot(merge(pr.mean,pr.mu,10*pr.fw,pr.wet,pr.dry),plot.type='single',
     col=c('steelblue','darkblue','grey','darkgreen','red'),
     lwd=c(3,3,1,1,1),ylab="",xlab="Calendar month",main=loc(y1))
grid()
par(yaxt='s',xaxt='s')
axis(1,at=1:12,labels=month.abb,cex.lab=0.7, col='grey')
axis(4,at=10*pretty(pr.fw),pretty(pr.fw),col='grey')

legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
               expression(bar(n[c*w*d])),expression(bar(n[c*d*d]))),bty='n',
       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1))
figlab('Figure SM9')
#dev.copy2pdf(file='figureSM9.pdf')

## Map of the future return values

print('Map of return values for 2100 with RCP4.5')
rv <- MUx
## x2100u gives the percentage of mu,
coredata(rv) <- t((x2100u/100)*apply(coredata(MUx),2,'mean')*log(365.25*t(coredata(FWx))))
cexr2 <- 1.5*c(as.numeric(r2)) + 0.2
dev.new()
map(rv,FUN='mean',cex=cexr2,colbar=list(breaks=seq(30,100,by=1)))
figlab('Figure SM10',ypos=0.999)
figlab('Return values for 2100 assuming RCP4.5',xpos=0.3,ypos=0.999)
#dev.copy2pdf(file='figureSM10.pdf')

## Print the numbers:
print('--- Changes in 20% returnvalue in terms of % from 2010:')
print(lapply(mu.2100,summary))
print(paste('Summary for',length(mu.2100[[1]]),'locations'))

## Compare the regression coefficients derived from individual
## seasonal cycle with that derived from mean climatology.

data(mu.eq.f.tx)

col <- rgb(0.1,0.1,0.7,0.25)
mutx <- summary(mu.eq.f.tx)$coefficients[c(2,4)]
b1 <- as.numeric(lapply(Beta,function(x) x[1]))
e1 <- as.numeric(lapply(Beta,function(x) x[2]))

#dev.new()
par(bty='n')
plot(b1,pch=19,col=col,cex=cexr2,
     main=expression(paste('Scaling coefficient for ',mu,' and ',e[s])),
     xlab='',ylab=expression(beta))
grid()
for (i in 1:length(b1)) {
  lines(rep(i,2),b1[i]+e1[i]*c(-2,2),col=col)
  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(-2,-2),col=col)
  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(2,2),col=col)
}
polygon(c(1,rep(length(b1),2),rep(1,2)),
        mutx[1]+mutx[2]*c(-2,-2,2,2,-2),
        border=rgb(0.5,0.5,0.5,0.4),col=rgb(0.5,0.5,0.5,0.3))
lines(c(1,length(b1)),rep(mutx[1],2),lwd=3,col=rgb(0.5,0.5,0.5,0.3))
figlab('Figure SM12')
#dev.copy2pdf(file='figureSM12.pdf')

## Show the predictor area:
X <- retrieve('data/ncep/air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
figlab('Figure SM13',xpos=0.8,ypos=0.999)
##dev.copy2pdf(file='figureSM13.pdf')
```

### 2016-04-04 ClimaTrans/climatrans.Rmd
Initial scripts and results for ClimaTrans and CixPAG

*Climatrans results: Work package 3: Outline of the development trends and climate change impact up to 2050* 
-----------------------------------
title: "Extreme precip and heat waves: empirical-statistical downscaling and analysis"
author: "Rasmus Benestad"
date: "March 30, 2016"
output: pdf_document
    fig_width:  8
    fig_height: 8
---

**Background**
The main objectives of the project are: 1) Assess climate change and environmental impacts in urban areas in India related to the transport sector. 2) Develop mitigation and adaption strategies related to the transport sector in urban areas in India. The objectives are set responding to the following objectives in the KLIMAFORK call: Improve knowledge about the impacts of climate change on the natural environment and society; and Enhance knowledge about how society can and should adapt to the challenges of climate change.

The CLIMATRANS project proposal responds to the research themes-Strategies for reduced human made climate change" and "Strategies for climate change adaption" of the KLMAFORSK call. Assessing risk, uncertainty and irreversibility as inherent to the climate problem, this project will provide knowledge that makes it possible to improve institutional capacity in climate decision-making strategies. The project is interdisciplinary within the realm of social sciences (economics, political science, sociology), and natural sciences (meteorology and civil engineering). It is also a collaboration between Norwegian (Institute of Transport Economics and Norway Meteorological Institute) and four Indian research environments, thereby responding to the INDNOR program.

India’s three largest cities are selected as case cities: New Delhi, Mumbai, and Bangalore.

**WP3: Outline of the development trends and climate change impact up to 2050**
WP leader: Dr. Fagerli, MET. Main contributors: TERI and TOI. Duration: Months 11-18.
In WP3 we will attempt to outline the trends (of aspects named in WP2), and project the climate change impact in the case cities in 2050.
Evaluation of the climate change impact by 2050 (base scenario) will assess the likely impact in the case cities given a base scenario (-business as usual‖) with no new policy interventions implemented.
Task 1 through 6 will address the same respective issues as in WP2, but with focus on the situation in India in 2050. Partners responsible for the respective tasks in WP3 are the same as the corresponding tasks in WP2. In addition, the following impact assessments will be done:
Task 7: Climate change impact on the transport infrastructure. (TERI, MET, TOI)
Task 8: Social impact of climate change on various population groups, health and wellbeing effects, etc. In terms of health effects, the downscaled future climate scenario and the derived transport scenarios for the cities, combined with appropriate emission scenarios for India/Asia, will be used as basis for deriving health effects of ozone and PM (including climate change). The downscaled future climate scenario and the present day emission scenarios will be used to derive increases in health effects only due to climate change. (MET, TERI, TOI)
7
Task 9: Economic impact of climate change in the case cities, and for the population (e.g., household income, wealth distribution, poverty levels, etc.)

**Analysis** 
This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. To execute this type:

```{r,eval=FALSE}
library(rmarkdown); render('climatrans.Rmd',pdf_document())
```

Load the esd and ncdf packages.

```{r}
library(esd)
library(ncdf)
```

---------------------------------

##*Precipitation*

#### Function used to estimate proportional change (for precipitation):

```{r,eval=FALSE}
propchange <- function(X,it0=c(1979,2013)) {
  z <- coredata(X)
  if (is.null(dim(z)))
      z <- 100*(z/mean(coredata(subset(X,it=it0)),na.rm=TRUE)) else
      z <- 100*t(t(z)/apply(coredata(subset(X,it=it0)),2,'mean',na.rm=TRUE))
  attributes(z) <- NULL
  z -> coredata(X)  
  X
}
```

Function that converts the surface temperature into saturation vapour pressure:

```{r}
tas2es <- function(x,mask=FALSE,land=TRUE,season=6:9,FUN='min') {
  if (mask) x <- mask(x,land=land)
  es <- subset(C.C.eq(x),it=month.abb[season])
  es <- aggregate(es,year,FUN=FUN)
  invisible(es)
}
```

#### Pre-processing: adapting the netCDF data to esd: set up station objects: 
```{r,eval=FALSE}
## Precipitation
## time origin: minutes since 1901-01-01 00:00
ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/delhi.nc')
delhi <- get.var.ncdf(ncid,varid='p')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
delhi <- zoo(x=delhi,order.by=as.Date(time/(24*60),origin='1901-01-01'))
delhi <- as.station(delhi,loc='Delhi',param='precip',unit='mm/day',lon=77.2,lat=28.6,
                    cntr='India',longname='precipitation',
                    info='extracted rainfall data from IMD gridded daily data',
                    ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bombay.nc')
bombay <- get.var.ncdf(ncid,varid='p')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
bombay <- zoo(x=bombay,order.by=as.Date(time/(24*60),origin='1901-01-01'))
bombay <- as.station(bombay,loc='Bombay',param='precip',unit='mm/day',lon=72.9,lat=19.0,
                cntr='India',longname='precipitation',
                info='extracted rainfall data from IMD gridded daily data',
                ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bangalore.nc')
bangalore <- get.var.ncdf(ncid,varid='p')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)

bangalore <- zoo(x=bangalore,order.by=as.Date(time/(24*60),origin='1901-01-01'))
bangalore <- as.station(bangalore,loc='Bangalore',param='precip',unit='mm/day',
                    lon=77.6,lat=13.0,cntr='India',longname='precipitation',
                    info='extracted rainfall data from IMD gridded daily data',
                    ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

## Combine the individual stations into a group of stations
climatrans.pr <- combine(delhi,bombay,bangalore)
## Save the results for future use
save(file='~/Dropbox/Public/ClimaTrans/climatrans.pr.rda',climatrans.pr)

## Maximum temperature
## time origin: minutes since 1969-01-01 00:00
ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/delhi_tmax.nc')
delhi <- get.var.ncdf(ncid,varid='temp')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
delhi <- zoo(x=delhi,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.delhi <- as.station(delhi,loc='Delhi',param='tmax',unit='degC',lon=77.2,lat=28.6,
                   cntr='India',longname='daily maximum temperature',
                   info='extracted from IMD gridded daily data',
                   ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bombay_tmax.nc')
bombay <- get.var.ncdf(ncid,varid='temp')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
bombay <- zoo(x=bombay,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.bombay <- as.station(bombay,loc='Bombay',param='tmax',unit='degC',lon=72.9,lat=19.0,
                    cntr='India',longname='daily maximum temperature',
                    info='extracted from IMD gridded daily data',
                    ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bangalore_tmax.nc')
bangalore <- get.var.ncdf(ncid,varid='temp')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
bangalore <- zoo(x=bangalore,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.bangalore <- as.station(bangalore,loc='Bangalore',param='tmax',unit='degC',
                    lon=77.6,lat=13.0,cntr='India',longname='daily maximum temperature',
                    info='extracted from IMD gridded daily data',
                    ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

## Combine the individual stations into a group of stations
climatrans.tx <- combine(tx.delhi,tx.bombay,tx.bangalore)
## Save the results for future use
save(file='~/Dropbox/Public/ClimaTrans/climatrans.tx.rda',climatrans.tx)
```

## Analysis

The analysis and downscaling can be carried out once the data has been converted to esd-station object. First load the precipitation data:

### **Daily station data**
```{r,echo=TRUE}
## Load precipitation data extracted from the gridded IMD data set:
load('~/Dropbox/Public/ClimaTrans/climatrans.pr.rda')
## Display the structure of the data as
str(climatrans.pr)
```

Examine the mean seasonal cycle to see if it responds to systematic forcings - the simples and most obvious being the seasonal variation in the solar inclination and the monsoon season:

```{r, fig.width=6, fig.height=6}
## Wet-day frequency
plot(aggregate(climatrans.pr,month,'wetfreq'),new=FALSE)
```

```{r, fig.width=6, fig.height=6}
## Wet-day mean
plot(aggregate(climatrans.pr,month,'wetmean'),new=FALSE)
```

```{r, fig.width=6, fig.height=6}
## Aggregate to mean seasonal cycle and annual mean values
plot(aggregate(climatrans.pr,month,'sum'),new=FALSE)
```

Estimate annual aggregated statistics for the months May-October (the wet season):

```{r}
fw <- aggregate(subset(climatrans.pr,it=month.abb[5:10]),year,'wetfreq')
mu <- aggregate(subset(climatrans.pr,it=month.abb[5:10]),year,'wetmean')
q95 <- aggregate(subset(climatrans.pr,it=month.abb[5:10]),year,'quantile',probs=0.95)

## check the relationship between wet-day 95-percentile and wet-day mean
plot(zoo(mu[,2]),zoo(q95[,2]),pch=19,xlim=c(0,40),ylim=c(0,120),
     main='Wet-day mean v.s. 95th wet percentile',
     xlab=expression(paste(mu,' (mm/day)')), ylab=expression(paste(q[95],' (mm/day)')))
points(zoo(mu[,1]),zoo(q95[,1]),pch=19,col=rgb(1,0,0,0.2))
points(zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
grid()
legend(0,120,loc(climatrans.pr),pch=19,col=c(rgb(1,0,0,0.2),'black',rgb(0,0,1,0.2)),bty='n')
```

Now prepare for the downscaling: Retrieve and process the predictors. For the wet-day mean we try temperature (or sea surface temperature, SST). There are different possibilities. We can estimate the vapour saturation pressure based on the Clausius-Clapeyron equation (`C.C.eq()`) or we can use the temperature directly (the former translates more directly into moisture mass). The aggregation of the temperature over the wet-season may involve the average, minimum or the maximum values. Trial anderror to see what works and what doesn't:

```{r}
## Aggregate sub-season values to annual mean values
if (!exists("es")) {
  #sst <- retrieve('data/ncep/sst.mon.mean.nc',lon=c(50,100),lat=c(0,40))
  #es <- aggregate(C.C.eq(subset(sst,it=month.abb[6:9])),year,'max')
  t2m <- retrieve('data/ncep/air.mon.mean.nc',lon=c(50,100),lat=c(5,30))
  es <- annual(t2m,'tas2es')
}
```

The mean seal-level pressure (SLP) is usually a promising predictor for the wet-day frequency.

```{r}
if (!exists("slp")) {
  slp <- retrieve('data/ncep/slp.mon.mean.nc',lon=c(50,100),lat=c(0,40))
  slp <- aggregate(subset(slp,it=month.abb[6:9]),year,'mean')
}

```

### **Examine potential large-scale tele-connections: correlation fields**

```{r}
## Correlation between the wet-day mean and the temperature-based predictor
y <- subset(mu,is=2)
corfield(y,es,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)))
```

```{r}
## Correlation between the wet-day frequency and the SLP
z <- subset(fw,is=2)
corfield(z,slp,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)))
```

Principal component analysis (PCA) can improve the quality of empirical-statistical downscaling (ESD) by reorganising the data so that the large-scale variability is enhanced (<http://dx.doi.org/10.3402/tellusa.v67.28326>).

### **Estimate PCA for the predicands**
```{r}
pca.mu <- PCA(mu)
pca.fw <-PCA(fw)
```

### **Estimate EOFs for the predictands**
```{r}
eof.es <- EOF(es)
eof.slp <- EOF(slp)
index(eof.es) <- year(eof.es)
index(eof.slp) <- year(eof.slp)
```

### **Plot the EOFs for the temperature-based predictor**
```{r, fig.width=6, fig.height=6}
plot(eof.es,new=FALSE)
```

### **Plot the EOFs for the SLP**
```{r, fig.width=6, fig.height=6}
plot(eof.slp,new=FALSE)
```

## **Do the downscaling**

### 
```{r,echo=FALSE, fig.width=6, fig.height=6}
## The function DS uses multiple regression by default
ds.mu <- DS(pca.mu,eof.es,eofs=1:20)
ds.fw <- DS(pca.fw,eof.slp,eofs=1:20)
plot(ds.mu,colbar1=list(breaks=seq(0,1,by=0.1),rev=2),
     colbar2=list(breaks=seq(-500,500,by=25),rev=TRUE))
plot(ds.fw)
```

#### **Compare the histogram of the preciptation data with an exponential distribution**
```{r}
z <- climatrans.pr[,3]; z <- z[z > 1]
hist(z,breaks=seq(0,500,by=2),xlim=c(0,100),freq=FALSE)
lines(seq(0,250,by=1),dexp(seq(0,250,by=1),rate=1/mean(z)),lwd=3,col='red')
```

#### **Do the downscaling based on the CMIP5 RCP4.5 ensemble**

Here include the argument \code{xfuns='tas2es'} to use the option \code{annual(tas2es(gcm))} rather than the option \code{annual(gcm,FUN='tas2es')} which will fail. Also

```{r}
if (!file.exists('dse.mu.climatrans.rda')) {
  dse <- DSensemble(pca.mu,predictor=es,FUNX='tas2es',xfuns='tas2es',nmin=1)
  save(file='dse.mu.climatrans.rda',dse)
} else load('dse.mu.climatrans.rda')
dse.stations<-as.station(dse)
```

#### *Maximum temperature*

Check the long-term change: historic trend in the mean

```{r}
  load('~/Dropbox/Public/ClimaTrans/climatrans.tx.rda')
  plot(as.4seasons(anomaly(climatrans.tx)))
```

Downscale the seasonal seasonal mean maximum temperature

```{r}
  predictor <- retrieve('data/ncep/air.mon.mean.nc',
                        param='air',lon=c(60,90),lat=c(10,35))

  dev.new()
  trh <- c(40,30,32)
  for (i in 1:3) {
    y <- subset(climatrans.tx,is=i)
    if (!file.exists(paste('dse.tx.climatrans.',loc(y),'.rda',sep=''))) {
      dse.tx <- DSensemble.t2m(y,biascorrect=TRUE,type='ncdf4',
                           predictor=predictor,nmin=60,verbose=FALSE)
      save(file=paste('dse.tx.climatrans.',loc(y),'.rda',sep=''),dse.tx)
    } else load(paste('dse.tx.climatrans.',loc(y),'.rda',sep=''))

    ## The function hotsummerdays estimates the number of hot days based on
    ## the assumption that the temperature is close to being normally distributed
    hw <- hotsummerdays(x=y,dse=dse.tx,threshold=trh[i],plot=FALSE)
    plot(hw)
  }
```

Check the correlation between the seasonal percentiles and the seasonal mean temperature, assuming the temperature is normally distributed.

```{r}
  txq95 <- diagnose(subset(climatrans.tx,is=1))
```

Check whether the maximum temperature is close to being normally distributed

```{r}
## New Delhi:	
  new.delhi.tx <- subset(climatrans.tx,is=1)
## Bangalore:
  bangalore.tx <- subset(climatrans.tx,is=3)
## Bombay:     
  bombay.tx <- subset(climatrans.tx,is=2)

  y <- new.delhi.tx
  par(mfcol=c(2,2))
  qqnorm(coredata(subset(y,it='djf')),col='grey',main='DJF')
  qqline(coredata(subset(y,it='djf')))
  qqnorm(coredata(subset(y,it='mam')),col='grey',main='MAM')
  qqline(coredata(subset(y,it='mam')))
  qqnorm(coredata(subset(y,it='jja')),col='grey',main='JJA')
  qqline(coredata(subset(y,it='jja')))
  qqnorm(coredata(subset(y,it='son')),col='grey',main='SON')
  qqline(coredata(subset(y,it='son')))
```

Check the long-term change: historic trend in the seasonal spread (standard deviation)

```{r}
  load('~/Dropbox/Public/ClimaTrans/climatrans.tx.rda')
  plot(as.4seasons(anomaly(climatrans.tx),FUN='sd'))
```

### 2016-04-06 KSS/dr-pr4KSS.Rmd
KSS dr-pr4KSS.Rmd

*Downscale precipitation for the Norwegian climate services (KSS)*
-----------------------------------
title: Downscling wet-day frequency and wet-day mean precipitation
author: "Rasmus Benestad"
date: "April 5, 2016"
output: pdf_document
    fig_width:  8
    fig_height: 8
---

This analysis is based on publications on how to downscale wet-day mean precipitation iteratively from several predictors (representing different moisture sources) and mean sea-level pressure to capture the wet-day frequency <http://dx.doi.org/10.3402/tellusa.v67.25954> and making use of a group of stations together with principal component analysis (PCA) to maximise the use of available information <http://dx.doi.org/10.3402/tellusa.v67.28326>.

**Analysis** 
The downscaling of the wet-day mean precipitation used the surface temperature as predictor.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. To execute this type:

```{r,eval=FALSE}
library(rmarkdown); render('climatrans.Rmd',pdf_document())
```

Load the esd and ncdf packages.

```{r}
library(esd)
```

---------------------------------

Function that converts the surface temperature into saturation vapour pressure:

##*Predictands*

```{r}
fname.predictand <- 'mufw.nordic.rda'
if (!file.exists(fname.predictand)) {
  ## Read daily precipitation from the ECA&D data set
  ## Long stations within the Nordic countries++
  ss <- select.station(param='precip',src='ecad',
                       lon=c(5,35),lat=c(55,72),nmin=60)
  precip <- station(ss)
  ## Remove stations with large holes of missing data and limit the data
  ## to a common interval with good data coverage

  diagnose(precip)
  precip <- subset(precip,it=c(1961,2013))
  nv <- apply(coredata(precip),2,nv)
  precip <- subset(precip,is=(nv > 50*365.25))
  mu.c <- annual(subset(precip,it=month.abb[c(1:3,10:12)]),'wetmean',nmin=180)
  mu.w <- annual(subset(precip,it=month.abb[c(4:9)]),'wetmean',nmin=180)
  fw.c <- annual(subset(precip,it=month.abb[c(1:3,10:12)]),'wetfreq',nmin=180)
  fw.w <- annual(subset(precip,it=month.abb[c(4:9)]),'wetfreq',nmin=180)
  save(file=fname.predictand,mu.c,mu.w,fw.c,fw.w)
} else load(fname.predictand)
```

**Process the predictands. Interploate for missing data, as PCA cannot be applied to data matrices with missing data:

```{r}
## Fill in small gaps of missing data using PCA-based regression
mu.c <- pcafill(mu.c)
mu.w <- pcafill(mu.w)
fw.c <- pcafill(fw.c)
fw.w <- pcafill(fw.w)
```
PCAs can be computed once the data matrices are complete. The PCAs will give an impression about the main features in the precipitation statistics from the region. The amount of precipitation varies systematically from place to place, with higher amounts over the west coast of Norway. Hence the PCA will emphasise the variations over western Norway. But the PCA can also be applied to relative changes (in percentage), where all locations weigh the same (more emphasis is now placed on the region with denser observations).

```{r}
## Estimate the PCAs - first for data in terms of mm/day
pca.mu.c <- PCA(mu.c)
pca.mu.w <- PCA(mu.w)
pca.fw.c <- PCA(fw.c)
pca.fw.w <- PCA(fw.w)
## Also estimate PCA for precipitation variations in terms of %
pca.mu.C <- PCA(propchange(mu.c))
pca.mu.W <- PCA(propchange(mu.w))
pca.fw.C <- PCA(propchange(fw.c))
pca.fw.W <- PCA(propchange(fw.w))
```

```{r}
## The main mode of variability for the absolute precipitation intencity
plot(pca.mu.c)
```

```{r}
## The main mode of variability for the relative precipitation intencity
plot(pca.mu.C)
```

```{r}
## Trend analysis
cols <- rainbow(4)
plot.zoo(pca.mu.c[,1:4],plot.type='single',col=cols)
for (i in 1:4) lines(trend(pca.mu.c[,i]),col=cols[i])
```

```{r}
## Trend analysis
cols <- rainbow(4)
plot.zoo(pca.fw.c[,1:4],plot.type='single',col=cols)
for (i in 1:4) lines(trend(pca.fw.c[,i]),col=cols[i])
```

##**Predictors**

Now prepare for the downscaling: Retrieve and process the predictors. For the wet-day mean we try temperature (or sea surface temperature, SST). There are different possibilities. We can estimate the vapour saturation pressure based on the Clausius-Clapeyron equation (`C.C.eq()`) or we can use the temperature directly (the former translates more directly into moisture mass). The aggregation of the temperature over the wet-season may involve the average, minimum or the maximum values. Trial anderror to see what works and what doesn't:

```{r}
## Aggregate sub-season values to annual mean values
if (!exists("es.c")) {
  slp <- retrieve('slp.mon.mean.nc',lon=c(-30,50),lat=c(30,75))
  slp.w <- annual(subset(slp,it=month.abb[4:9]),'mean',nmin=6)
  slp.c <- annual(subset(slp,it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  t2m <- retrieve('air.mon.mean.nc',lon=c(-30,50),lat=c(30,75))
  t2m.w <- annual(subset(t2m,it=month.abb[4:9]),'mean',nmin=6)
  t2m.c <- annual(subset(t2m,it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  es.w <- annual(subset(C.C.eq(t2m),it=month.abb[4:9]),'mean',nmin=6)
  es.w <- mask(es.w,land=TRUE)
  es.c <- annual(subset(C.C.eq(t2m),it=month.abb[c(1:3,10:12)]),'mean',nmin=6)
  es.c <- mask(es.w,land=TRUE)
}
## Represent the predictors in terms of EOFs
eof.t2m.w <- EOF(t2m.w)
eof.t2m.c <- EOF(t2m.c)
eof.slp.w <- EOF(slp.w)
eof.slp.c <- EOF(slp.c)
eof.es.w <- EOF(es.w)
eof.es.c <- EOF(es.c)
```

Explorative downscaling analysis: need to fine some connections to the large-scale temperatures and then find the residuals from this. The residuals are then used in a secondary downscaling analysis against another set of predictors.

```{r}
## Downscaling analysis
  ds.mu.c <- DS(pca.mu.c,eof.es.c)
  plot(ds.mu.c)
```

We have now identified one connection between the large-scale conitions - vapour saturation pressure over the North Atlantic (a main moisture source) and the wet-day mean. There may be other fators affecting thestatistics too, and we can search for remaining structure in the data in terms of spatio-temporal variability. If the reminder looks onisy (e.g. 'flat' eigenvalues), then the remainder may be assumed to be from local effects and can be represented by a noise model.

```{r}
## Examine the PCA of residuals to see if there are any structures left
  mu.c.2 <- as.residual(as.station(ds.mu.c))
  mu.c.2 <- attrcp(mu.c,mu.c.2); class(mu.c.2) <- class(mu.c)
  pca.mu.c.2 <- PCA(mu.c.2)
  plot(pca.mu.c.2)
```

```{r}
## Downscaling analysis
  dse.mu.c <- DSensemble(pca.mu.c,predictor=es.c,
                         it=month.abb[c(1:3,10:12)],
                         FUNX='C.C.eq',plot=TRUE,verbose=FALSE)

```

### 2016-04-07 ClimaTrans/climatrans.Rmd
Results for Climatrans: precipitation and hot summer days

@@
 **Analysis** 
+The downscaling of the wet-day mean precipitation used the surface temperature as predictor.
+
 This is an R Markdown document.
@@
 
-##*Precipitation*
+##*Functions*
 
-#### Function used to estimate proportional change (for precipitation):
+Define a function that converts the surface temperature into saturation vapour pressure:
 
-```{r,eval=FALSE}
-propchange <- function(X,it0=c(1979,2013)) {
-  z <- coredata(X)
-  if (is.null(dim(z)))
-      z <- 100*(z/mean(coredata(subset(X,it=it0)),na.rm=TRUE)) else
-      z <- 100*t(t(z)/apply(coredata(subset(X,it=it0)),2,'mean',na.rm=TRUE))
-  attributes(z) <- NULL
-  z -> coredata(X)  
-  X
+```{r}
+tas2es <- function(x,mask=TRUE,land=TRUE,season=5:10,FUN='min') {
+  if (mask) x <- mask(x,land=land)
+  es <- subset(C.C.eq(x),it=month.abb[season])
+  es <- aggregate(es,year,FUN=FUN)
+  index(es) <- year(es)
+  invisible(es)
 }
@@
 
-Function that converts the surface temperature into saturation vapour pressure:
+Function to estimate probabilty for daily amount of exceeding a threshold x0 given that the distribution follows an exponential distribution and the records of the wet-day mean precipitation and frequency.
 
 ```{r}
-tas2es <- function(x,mask=FALSE,land=TRUE,season=6:9,FUN='min') {
-  if (mask) x <- mask(x,land=land)
-  es <- subset(C.C.eq(x),it=month.abb[season])
-  es <- aggregate(es,year,FUN=FUN)
-  invisible(es)
+PrexpPr <- function(mu,fw,x0=10) {
+  Pr <- zoo(coredata(fw)*exp(-x0/coredata(mu)),order.by=year(fw))
+
+  Pr.mu <- zoo(mean(coredata(fw))*exp(-x0/coredata(mu)),order.by=year(fw))
+  Pr.fw <- zoo(coredata(fw)*exp(-x0/mean(coredata(mu))),order.by=year(fw))
+  attr(Pr,'prob.f(mu)') <- Pr.mu
+  attr(Pr,'prob.f(fw)') <- Pr.fw
+  Pr <- attrcp(mu,Pr)
+  attr(Pr,'variable') <- 'Pr'
+  attr(Pr,'unit') <- 'probability'
+  attr(Pr,'longname') <- paste('Probability of exceeding',x0)
+  class(Pr) <- class(mu)
+  invisible(Pr)
 }
 ```
+##*Precipitation*
 
@@
 ## Save the results for future use
-save(file='~/Dropbox/Public/ClimaTrans/climatrans.pr.rda',climatrans.pr)
+save(file='climatrans.pr.rda',climatrans.pr)
 
@@
 ## Save the results for future use
-save(file='~/Dropbox/Public/ClimaTrans/climatrans.tx.rda',climatrans.tx)
+save(file='climatrans.tx.rda',climatrans.tx)
 ```
@@
 ## Load precipitation data extracted from the gridded IMD data set:
-load('~/Dropbox/Public/ClimaTrans/climatrans.pr.rda')
+load('climatrans.pr.rda')
 ## Display the structure of the data as
@@
 ```
+##**Sensitivity tests**
+
+How do the precipitation respond to varying forcings/conditions? Such iformation may provide useful clues as to how the precipitation is expected to change in the future.
 
@@
 ```{r, fig.width=6, fig.height=6}
-## Aggregate to mean seasonal cycle and annual mean values
+## Monthly precipitation totals
 plot(aggregate(climatrans.pr,month,'sum'),new=FALSE)
@@
 
-Estimate annual aggregated statistics for the months May-October (the wet season):
+Estimate annual aggregated statistics for the months May-October (the wet/Monsoon season):
 
 ```{r}
-fw <- aggregate(subset(climatrans.pr,it=month.abb[5:10]),year,'wetfreq')
-mu <- aggregate(subset(climatrans.pr,it=month.abb[5:10]),year,'wetmean')
-q95 <- aggregate(subset(climatrans.pr,it=month.abb[5:10]),year,'quantile',probs=0.95)
+season<-5:10
+fw <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetfreq')
+mu <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetmean')
+q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'quantile',probs=0.95)
+```
 
+Compare the data to the exponential distribution - if the data is exponentially distributed, then the wet-day mean precipitation $\mu$ and the 95-percentile $q_{95}$ are propportional by $q_p = -ln(1-p)\mu$.
+
+```{r}
 ## check the relationship between wet-day 95-percentile and wet-day mean
-plot(zoo(mu[,2]),zoo(q95[,2]),pch=19,xlim=c(0,40),ylim=c(0,120),
+plot(-log(0.05)*zoo(mu[,2]),zoo(q95[,2]),pch=19,xlim=c(0,120),ylim=c(0,120),
      main='Wet-day mean v.s. 95th wet percentile',
-     xlab=expression(paste(mu,' (mm/day)')), ylab=expression(paste(q[95],' (mm/day)')))
-points(zoo(mu[,1]),zoo(q95[,1]),pch=19,col=rgb(1,0,0,0.2))
-points(zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
+     xlab=expression(paste(-ln(1-0.95)*mu,' (mm/day)')), ylab=expression(paste(q[95],' (mm/day)')))
+points(-log(0.05)*zoo(mu[,1]),zoo(q95[,1]),pch=19,col=rgb(1,0,0,0.2))
+points(-log(0.05)*zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
 grid()
@@
 ```
+
+Compare the histogram of the preciptation data with an exponential distribution
+```{r}
+z <- climatrans.pr[,3]; z <- z[z > 1]
+hist(z,breaks=seq(0,500,by=2),xlim=c(0,100),freq=FALSE)
+lines(seq(0,250,by=1),dexp(seq(0,250,by=1),rate=1/mean(z)),lwd=3,col='red')
+```
+
+##**Historical long-term trends**
+
+Estimate the long-term trends in the precipitation statistics at the three megacities: first the wet-day frequency $f_w$ - does the number of rainy days change over time?
+
+```{r, fig.width=6, fig.height=6}
+## Wet-day frequency
+plot(fw,new=FALSE)
+for(i in 1:3) lines(trend(subset(fw,is=i)))
+```
+
+Does the wet-day mean precipitation $\mu$ change over time?
+
+```{r, fig.width=6, fig.height=6}
+## Wet-day mean
+plot(mu,new=FALSE)
+for(i in 1:3) lines(trend(subset(mu,is=i)))
+```
+
+If the 24-hr preciptation amounts is exponential (when it rains), then the probability of exceeding a threshold is $Pr(X>x_0) = f_w e^{-x_0/\mu}$. Since we know both
+
+```{r, fig.width=6, fig.height=6}
+## Wet-day mean
+Pr <- PrexpPr(mu,fw,x0=50) 
+plot(Pr,new=FALSE,errorbar=FALSE)
+for(i in 1:3) lines(trend(subset(Pr,is=i)))
+```
+
+Also, it's interesting to see how the annual total precipitation amount changes over time:
+
+```{r, fig.width=6, fig.height=6}
+## Aggregate to mean seasonal cycle and annual mean values
+plot(annual(climatrans.pr,'sum'),new=FALSE)
+for(i in 1:3) lines(trend(subset(annual(climatrans.pr,'wetfreq'),is=i)))
+```
+
+##**Predictors**
 
@@
 if (!exists("es")) {
-  #sst <- retrieve('data/ncep/sst.mon.mean.nc',lon=c(50,100),lat=c(0,40))
-  #es <- aggregate(C.C.eq(subset(sst,it=month.abb[6:9])),year,'max')
-  t2m <- retrieve('data/ncep/air.mon.mean.nc',lon=c(50,100),lat=c(5,30))
-  es <- annual(t2m,'tas2es')
+  t2m <- retrieve('data/ncep/air.mon.mean.nc',lon=c(60,100),lat=c(0,23))
+  es <- aggregate(tas2es(t2m,season=season),year,'mean')
 }
@@
 if (!exists("slp")) {
-  slp <- retrieve('data/ncep/slp.mon.mean.nc',lon=c(50,100),lat=c(0,40))
-  slp <- aggregate(subset(slp,it=month.abb[6:9]),year,'mean')
+  slp <- retrieve('data/ncep/slp.mon.mean.nc',lon=c(60,100),lat=c(0,23))
+  slp <- aggregate(subset(slp,it=month.abb[season]),year,'mean')
 }
@@
 plot(ds.mu,colbar1=list(breaks=seq(0,1,by=0.1),rev=2),
-     colbar2=list(breaks=seq(-500,500,by=25),rev=TRUE))
-plot(ds.fw)
+     colbar2=list(breaks=seq(-500,500,by=25),rev=TRUE),new=FALSE)
+plot(ds.fw,new=FALSE)
 ```
 
-#### **Compare the histogram of the preciptation data with an exponential distribution**
+## **Examine the residuals from the downscaling**
+
+First check the residuals for the wet-day mean precipitation:
+
 ```{r}
-z <- climatrans.pr[,3]; z <- z[z > 1]
-hist(z,breaks=seq(0,500,by=2),xlim=c(0,100),freq=FALSE)
-lines(seq(0,250,by=1),dexp(seq(0,250,by=1),rate=1/mean(z)),lwd=3,col='red')
+## Examine the PCA of residuals for the cold monthsto see if there are any structures left
+  mu.2 <- as.residual(as.station(ds.mu))
+  mu.2 <- attrcp(mu,mu.2); class(mu.2) <- class(mu)
+  pca.mu.2 <- PCA(mu.2)
+  plot(pca.mu.2)
+```
+
+The residuals for the wet-day frequency:
+
+```{r}
+## Examine the PCA of residuals for the cold monthsto see if there are any structures left
+  fw.2 <- as.residual(as.station(ds.fw))
+  fw.2 <- attrcp(fw,fw.2); class(fw.2) <- class(fw)
+  pca.fw.2 <- PCA(fw.2)
+  plot(pca.fw.2)
 ```
@@
 
-Here include the argument \code{xfuns='tas2es'} to use the option \code{annual(tas2es(gcm))} rather than the option \code{annual(gcm,FUN='tas2es')} which will fail. Also
+Here include the argument xfuns='tas2es' to use the option annual(tas2es(gcm)) rather than the option annual(gcm,FUN='tas2es') which will fail. Also
 
@@
 if (!file.exists('dse.mu.climatrans.rda')) {
-  dse <- DSensemble(pca.mu,predictor=es,FUNX='tas2es',xfuns='tas2es',nmin=1)
-  save(file='dse.mu.climatrans.rda',dse)
+  index(pca.mu) <- as.Date(paste(year(pca.mu),'01-01',sep='-'))
+  index(pca.fw) <- as.Date(paste(year(pca.fw),'01-01',sep='-'))
+  dse.mu <- DSensemble(pca.mu,predictor=es,FUNX='tas2es',xfuns='tas2es',plot=TRUE)
+  dse.fw <- DSensemble(pca.fw,predictor=slp,pattern="psl_Amon_ens_",plot=TRUE)
+  save(file='dse.mu.climatrans.rda',dse.mu,dse.fw)
 } else load('dse.mu.climatrans.rda')
-dse.stations<-as.station(dse)
+mu.stations<-as.station(dse.mu)
+fw.stations<-as.station(dse.fw)
+
+## Plot the downscaled results for the three megacities:
+for (i in 1:3) {
+  plot(mu.stations[[i]],new=FALSE)
+  plot(fw.stations[[i]],new=FALSE)
+}
 ```
@@
 
-Check the long-term change: historic trend in the mean
+Check the long-term change: historic trend in the mean - include all seasons, but to get a clearer ipression, plot the anomalies (i.e. exclude the mean seasonal cycle).
 
 ```{r}
-  load('~/Dropbox/Public/ClimaTrans/climatrans.tx.rda')
+  load('climatrans.tx.rda')
   plot(as.4seasons(anomaly(climatrans.tx)))
@@
 
-Downscale the seasonal seasonal mean maximum temperature
+Downscale the seasonal seasonal mean maximum temperature and then use the mean temperature to estimate the number of hot days assuming that the daily distribution is approximately Gaussian for a given season.
 
 ```{r}
-  predictor <- retrieve('data/ncep/air.mon.mean.nc',
-                        param='air',lon=c(60,90),lat=c(10,35))
+  predictor <- retrieve('air.mon.mean.nc',param='air',lon=c(60,90),lat=c(10,35))
 
-  dev.new()
   trh <- c(40,30,32)
@@
     hw <- hotsummerdays(x=y,dse=dse.tx,threshold=trh[i],plot=FALSE)
-    plot(hw)
+    plot(hw,new=FALSE)
   }
@@
 
-Check the long-term change: historic trend in the seasonal spread (standard deviation)
+Check the long-term change: historic trend in the seasonal spread (standard deviation) - the above estimates assumed that the spread is constant.
+The number of hot days will be even higher if the spread increases over time.
 
 ```{r}
-  load('~/Dropbox/Public/ClimaTrans/climatrans.tx.rda')
-  plot(as.4seasons(anomaly(climatrans.tx),FUN='sd'))
+  plot(as.4seasons(anomaly(climatrans.tx),FUN='sd'),new=FALSE)
 ```

### 2016-04-07 EU-Circle/eu-circle_case-study-3.Rmd
Results for Climatrans: precipitation and hot summer days

**EU-Circle case study 3
-----------------------------------
title: Coastal Flooding (surface water, highway, sewer and watercourse flooding) across Torbay, UK
author: "Rasmus Benestad"
date: "April 6, 2016"
output: pdf_document
    fig_width:  8
    fig_height: 8
---

Torbay is located in South Devon (UK) and covers an area of approximately 62 km2 . The area has suffered from flooding over many years from a number of different sources, including surface water run-off, highway flooding, sewer flooding, main river and ordinary watercourse flooding during intense rainfall events. In addition the coastal areas of Torbay  suffer coastal flooding due to overtopping of the sea defences during high tides that coincide with easterly winds. It should be noted that the surface water, highway, sewer, main river and watercourse flooding is exacerbated in the low lying areas around the coast of Torquay, Paignton and Brixham during high tidal cycles when the capacity of the surface water outfalls discharging to coastal waters is impeded. In addition to the property flooding, during all of these flood events numerous roads have flooded to some extent, with some of the roads having to be closed to traffic until the flood water has subsided.

The main coast road linking Torquay to Paignton and Brixham has to be closed on a regular basis due to overtopping of the sea wall during high tides that coincide with easterly wind conditions. These closures result in long traffic diversions and delays. In addition, as a result of surface water flooding from watercourses and main rivers during intense rainfall events the main inland route linking Torquay, Paignton and Brixham has also been closed for periods resulting in no major roads being available between the towns until flooding has subsided. During the last major flooding event, which occurred on 24 th October 1999, over 200 properties across all three towns were affected by flooding. Approximately 50% of these properties were commercial properties including shops, restaurants, hotels, bars and a cinema. In addition to the property flooding both the major roads linking Torquay to Paignton and Brixham were closed for a significant period making travel within the Bay extremely difficult and affecting emergency response.

Obviously climate change can affect local flood risk in several ways. Impacts will depend on local conditions and vulnerability. Wetter winters and more of this rain falling in wet spells will increase river and watercourse flooding. More intense rainfall causes more surface run-off, increasing localised flooding and erosion. In turn this will increase pressure on drains, sewers and water quality. Rising sea levels, as a result of climate change, will increase local flood risk both in coastal regions from increased risk of overtopping of the sea wall and inland from main rivers and watercourses due to the interaction with drains, sewers and smaller watercourses. As sea level is predicted to rise by over 1m in Torbay over the next 100 years the frequency and impact of overtopping of the sea defences will increase resulting in more infrastructure and properties being affected by flooding.

**Analysis** 
The downscaling of the wet-day mean precipitation used the surface temperature as predictor.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. To execute this type:

```{r,eval=FALSE}
library(rmarkdown); render('eu-circle_case-study-3.Rmd',pdf_document())
```

Load the esd and ncdf packages.

```{r}
library(esd)
```

---------------------------------
```{r,eval=FALSE}
## Get MIDAS preciptation data that has laready been retrieved 
## from the MIDAS data base (UK Met Office)
source('readmidas.R')
```

Function to estimate probabilty for daily amount of exceeding a threshold x0 given that the distribution follows an exponential distribution and the records of the wet-day mean precipitation and frequency.

```{r}
PrexpPr <- function(mu,fw,x0=10) {
  Pr <- zoo(coredata(fw)*exp(-x0/coredata(mu)),order.by=year(fw))

  Pr.mu <- zoo(mean(coredata(fw))*exp(-x0/coredata(mu)),order.by=year(fw))
  Pr.fw <- zoo(coredata(fw)*exp(-x0/mean(coredata(mu))),order.by=year(fw))
  attr(Pr,'prob.f(mu)') <- Pr.mu
  attr(Pr,'prob.f(fw)') <- Pr.fw
  Pr <- attrcp(mu,Pr)
  attr(Pr,'variable') <- 'Pr'
  attr(Pr,'unit') <- 'probability'
  attr(Pr,'longname') <- paste('Probability of exceeding',x0)
  class(Pr) <- class(mu)
  invisible(Pr)
}
```

Load the data and weed out stations with lots of missing values

```{r}
load('pr.eu-circle-torbay.rda')

## First weeding of stations with many missing values
nok <- apply(coredata(pr),2,'nv')
x <- subset(pr,is=(nok> 15000))

## Estimate the annual wet-day mean precipitaiton
mu <- subset(annual(x,'wetmean',nmin=180),it=c(1961,2011))
## Estimate the annual wet-day mean frequency
fw <- subset(annual(x,'wetfreq',nmin=180),it=c(1961,2011))

## Second weeding of stations with many missing values in annual 
## aggregate
nok <- apply(coredata(mu),2,'nv')
mu <- subset(mu,is=(nok> 45))
fw<- subset(fw,is=(nok> 45))

## Check the records
diagnose(mu)

##Fill in the remaining gaps using a PCA-based interpolation scheme 
pcafill(mu) -> mu
pcafill(fw) -> fw
```

Examine the long-term trends in the wet-day mean precipitation

```{r}
plot(mu,new=FALSE); grid()
for (i in 1:dim(mu)[2]) lines(trend(mu[,i]))
```

```{r}
plot(fw); grid()
for (i in 1:dim(mu)[2]) lines(trend(mu[,i]))
```

```{r}
Pr.gt.10 <- PrexpPr(mu,fw,x0=10)
plot(Pr.gt.10); grid()
for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.10[,i]))
```

### 2016-04-07 KSS/dr-pr4KSS.Rmd
Results for Climatrans: precipitation and hot summer days

@@
 ```{r,eval=FALSE}
-library(rmarkdown); render('climatrans.Rmd',pdf_document())
+library(rmarkdown); render('dr-pr4KSS.Rmd',pdf_document())
 ```
@@
 } else load(fname.predictand)
+```
+
+```{r}
+## Show maps of trend
+map(mu.c,FUN='trend',new=FALSE,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1),rev=TRUE))
+title(expression(paste('Trend in ',mu,' (mm/day per decade) for cold months')))
+map(mu.w,FUN='trend',new=FALSE,colbar=list(pal='t2m',breaks=seq(-0.5,0.5,by=0.1),rev=TRUE))
+title(expression(paste('Trend in ',mu,' (mm/day per decade) for warm months')))
+map(100*fw.c,FUN='trend',new=FALSE,colbar=list(pal='t2m',breaks=seq(-2,2,by=0.1),rev=TRUE))
+title(expression(paste('Trend in ',f[w],' (% per decade) for cold months')))
+map(100*fw.w,FUN='trend',new=FALSE,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1),rev=TRUE))
+title(expression(paste('Trend in ',f[w],' (mm/day per decade) for cold months')))
+```
+
+Variance of an exponentially distributed variable is $\mu^2$:
+
+```{r}
+## Show maps of trend
+mu.c2 <- mu.c
+coredata(mu.c2) <- coredata(mu.c)^2
+map(mu.c2,FUN='trend',new=FALSE,colbar=list(pal='t2m',breaks=seq(-25,25,by=1),rev=TRUE))
+title(expression(paste('Trend in ',mu^2,' (variance per decade) for cold months')))
+mu.w2 <- mu.w
+coredata(mu.w2) <- coredata(mu.w)^2
+map(mu.w2,FUN='trend',new=FALSE,colbar=list(pal='t2m',breaks=seq(-10,10,by=1),rev=TRUE))
+title(expression(paste('Trend in ',mu^2,' (variance per decade) for warm months')))
 ```
@@
 ```{r}
-## Downscaling analysis
+## Downscaling analysis for the cold months
   ds.mu.c <- DS(pca.mu.c,eof.es.c)
@@
 
+```{r}
+## Downscaling analysis for the warm months
+  ds.mu.w <- DS(pca.mu.w,eof.es.w)
+  plot(ds.mu.w)
+```
+
+```{r}
+## Downscaling analysis for the cold months
+  ds.mu.C <- DS(pca.mu.C,eof.es.c)
+  plot(ds.mu.C)
+```
+
+```{r}
+## Downscaling analysis for the warm months
+  ds.mu.W <- DS(pca.mu.W,eof.es.w)
+  plot(ds.mu.W)
+```
+
+```{r}
+## Downscaling analysis for the cold months
+  ds.fw.c <- DS(pca.fw.c,eof.slp.c)
+  plot(ds.mu.c)
+```
+
+```{r}
+## Downscaling analysis for the warm months
+  ds.fw.w <- DS(pca.fw.w,eof.slp.w)
+  plot(ds.mu.w)
+```
+
+##**The residuals of the multiple regression analysis**
+
 We have now identified one connection between the large-scale conitions - vapour saturation pressure over the North Atlantic (a main moisture source) and the wet-day mean.
@@
 If the reminder looks onisy (e.g.
-'flat' eigenvalues), then the remainder may be assumed to be from local effects and can be represented by a noise model.
+'flat' eigenvalues), then the remainder may be assumed to be from local effects and can be represented by a noise model.dr
 
 ```{r}
-## Examine the PCA of residuals to see if there are any structures left
+## Examine the PCA of residuals for the cold monthsto see if there are any structures left
   mu.c.2 <- as.residual(as.station(ds.mu.c))
@@
 ```{r}
+## Examine the PCA of residuals for the warm months to see if there are any structures left
+  mu.w.2 <- as.residual(as.station(ds.mu.w))
+  mu.w.2 <- attrcp(mu.w,mu.w.2); class(mu.w.2) <- class(mu.w)
+  pca.mu.w.2 <- PCA(mu.w.2)
+  plot(pca.mu.w.2)
+```
+
+```{r}
+## Examine the PCA of residuals for the cold monthsto see if there are any structures left
+  fw.c.2 <- as.residual(as.station(ds.fw.c))
+  fw.c.2 <- attrcp(fw.c,fw.c.2); class(fw.c.2) <- class(fw.c)
+  pca.fw.c.2 <- PCA(fw.c.2)
+  plot(pca.fw.c.2)
+```
+
+```{r}
+## Examine the PCA of residuals for the warm months to see if there are any structures left
+  fw.w.2 <- as.residual(as.station(ds.fw.w))
+  fw.w.2 <- attrcp(fw.w,fw.w.2); class(fw.w.2) <- class(fw.w)
+  pca.fw.w.2 <- PCA(fw.w.2)
+  plot(pca.fw.w.2)
+```
+
+```{r}
+## Examine the time series of the residuals
+  plot.zoo(pca.mu.w.2,type='single')
+```
+
+```{r}
+## Examine the ACF for the leading PC of the residuals
+  plot(acf(pca.mu.w.2[,1]))
+```
+
+```{r}
+## Examine the power spectrum for the leading PC of the residuals
+  plot(spectrum(pca.mu.w.2[,1]))
+```
+
+```{r}
+## Examine the distribution of the leading PC of the residuals
+  plot(qqnorm(coredata(pca.mu.w.2[,1])))
+  qqline(coredata(pca.mu.w.2[,1])))
+```
+
+```{r}
+fndseres <- 'dse.nordic.precip.rda'
+if (!file.exists(fndseres)) {
 ## Downscaling analysis
@@
                          FUNX='C.C.eq',plot=TRUE,verbose=FALSE)
+  dse.fw.c <- DSensemble(pca.fw.c,predictor=slp.c,
+                         it=month.abb[c(1:3,10:12)],
+                         FUNX='C.C.eq',plot=TRUE,verbose=FALSE)
+  dse.mu.w <- DSensemble(pca.mu.w,predictor=es.w,
+                         it=month.abb[4:9],
+                         FUNX='C.C.eq',plot=TRUE,verbose=FALSE)
+  dse.fw.w <- DSensemble(pca.fw.w,predictor=slp.w,
+                         it=month.abb[4:9],
+                         FUNX='C.C.eq',plot=TRUE,verbose=FALSE)
+  save(file=fndseres,dse.mu.c,dse.fw.c,dse.mu.w,dse.fw.w)
+} else load(fndseres)
 

### 2016-04-07 CixPAG/paper58.Rmd
Update for Climatrans: precipitation and hot summer days

@@
 
-```{r}
+```{r,echo=TRUE}
 rm(list=ls())
@@
 library(esd)
+## Information about the system and session
+Sys.info()
+sessionInfo()
 ```

### 2016-04-08 EU-Circle/eu-circle_case-study-3.Rmd
Update for Climatrans: precipitation and hot summer days
paper58

@@
 
-```{r}
+```{r,echo=TRUE}  
+## Load the necessary R-package
 library(esd)
+## Information about the system and session
+Sys.info()
+sessionInfo()
 ```
@@
 ## from the MIDAS data base (UK Met Office)
-source('readmidas.R')
+
 ```

### 2016-04-08 paper58/paper58.Rmd
paper58

*Benestad et al - Upper-limit estimation of future precipitation return-values* 
-----------------------------------
title: "Analysis and calculations in manuscript"
author: "Rasmus Benestad"
date: "April 4, 2016"
output: pdf_document
---

##**R set-up**
First pre-amble that checks whether the esd-package is installed and installs it if needed. If it is not installed, install it from GitHub using devtools. Also install devtools if needed. This is only done once.

```{r,echo=TRUE}
rm(list=ls())
xlim <- c(0,35); ylim <- c(45,70)
readecad <- FALSE
figshare=TRUE
nmin=50

## Check if you need to get the esd-package:
install.esd <- ("esd" %in% rownames(installed.packages()) == FALSE)

if (install.esd) {
  print('Need to install the esd package')
  ## Need online access.
  ## Use the devtools-package for simple facilitation of installing.
  if ("devtools" %in% rownames(installed.packages()) == FALSE)
    install.packages('devtools')
  library(devtools)
  ## Install esd directly from github
  install_github('metno/esd')
  print('The latest version of esd has been installed')
}

## Start the esd-library:
library(esd)
## Information about the system and session
Sys.info()
sessionInfo()
```

Below are definition of a number of functions used in the data processing and analysis. The first are simple functions used in ploting.

**Simple functions which estimate various statistics**

  Functions to extract the numbers for mean,min,max,wettest month,driest month, estimate percentiles assuming an exponential distribution, skill-scores, and regression coefficients. These are used to make the R-code simpler, shorter, and improve clarity.

```{r}
## Return statistics for mean,min,max,wettest month,driest month
muclim <- function(x) {
  y <- coredata(x)
  iX <- mean((1:12)[is.element(y,max(y))])
  iN <- mean((1:12)[is.element(y,min(y))])
  stats <- c(mean(y,na.rm=TRUE),min(y,na.rm=TRUE),max(y,na.rm=TRUE),iX,iN)
  names(stats) <- c('mean','min','max','wettest month','driest month')
  stats
}
```

```{r}
## Produce a set of percentiles and their counterparts for the exponential distribution
qqexp <- function(x,x0=1) {
  x[x < x0] <- NA
  if (sum(is.finite(x))>0) {
    mu <- mean(x,na.rm=TRUE)
    qx <- quantile(x,probs=seq(0,1,length=101),na.rm=TRUE)
    qy <- -log(1-seq(0,1,length=101))*mu
  } else {
    qx <- rep(NA,101); qy <- rep(NA,101)
  }
  return(cbind(qx,qy))
}
```

```{r}
## Skill associated with predicting the wet-day mean mu
muskill <- function(x) {
  ## Estimate the skill of the calibration:
  r2 <- round(summary(lm(y ~ x,data=x))$r.squared,3)
  ## Negative slopes are not credible:
  if (summary(lm(y ~ x,data=x))$coefficients[2] < 0) r2 <- 0
  r2
}
```

```{r}
## A function for extracting the regression coefficients and their
## error terms.
beta <- function(x,verbose=FALSE) {
  wc.model <- lm(y ~ x, data=x)
  if (verbose) print(summary(wc.model))
  beta <- summary(wc.model)$coefficients[c(2,4)]
  return(beta)
}
```

**Data input and procesing**

Functions for reading the data and processing it so that it can readily be handelled in the analysis.

## Read the data from the CMIP5 GCMs
```{r}
readGCMs <- function(path='CMIP5.monthly/rcp45/',pattern='tas',
                     lon=c(-100,30),lat=c(0,40)) {
    ncfiles <- list.files(path=path,pattern=pattern,full.names=TRUE)
    n <- length(ncfiles)
    print(paste(n,'netCDF files'))
    X <- matrix(rep(NA,n*201),n,201)
    for (i in 1:n) {      print(paste(i,ncfiles[i]))
      ## Spatial average:                    
      gcm <- annual(spatial.avg.field(C.C.eq(retrieve(ncfiles[i],
                                                      lon=lon,lat=lat))))
      i1 <- is.element(1900:2100,year(gcm))
      i2 <- is.element(year(gcm),1900:2100)
      X[i,i1] <- coredata(gcm)[i2]
    }
    ## Extract the 5 & 95 percentile and the ensemble mean:
    print('Extract the 5 & 95 percentile and the ensemble mean')
    x <- apply(X,2,function(x) c(quantile(x,probs=c(0.05,0.95),na.rm=TRUE),mean(x,na.rm=TRUE)))
    print(dim(x))
    names(x) <- c('q05','q95','mean')
    x <- zoo(t(x),order.by=1900:2100)
    w2000 <- X[,is.element(1900:2100,2000)]
    w2050 <- X[,is.element(1900:2100,2020)]
    w2100 <- X[,is.element(1900:2100,2100)]
    plot(x,plot.type='single')
    attr(x,'path') <- path
    attr(x,'2000') <- w2000
    attr(x,'2050') <- w2050
    attr(x,'2100') <- w2100
    attr(x,'N') <- n
    attr(x,'region') <- paste(lon,lat,collapse=' ')
    return(x)
}
```

**Model calibration and regression analysis**

Main function that is used for calibrating the statistical model based on the mean seasonal cycle.

```{r}
## Calibrate a model for the wet-day mean mu using temperature as input
mucal <- function(x,pre=NULL,lon=c(-100,30),lat=c(0,40),
                  plot=FALSE,verbose=FALSE) {

## If no pre, use the crude NCEP-reanalysis provided in esd 
  if (is.null(pre)) {
    if (verbose) print('default predictor')
      t2m <- t2m.NCEP(lon=lon,lat=lat)
      pre <- spatial.avg.field(C.C.eq(t2m))
      if (plot) plot(EOF(t2m))
  } else
      if (is.character(pre))
          pre <- spatial.avg.field(C.C.eq(retrieve(ncfile=pre,lon=lon,lat=lat))) else
  if (inherits(pre,'field')) {
      if (is.T(pre)) pre <- spatial.avg.field(C.C.eq(pre)) else
      pre <- spatial.avg.field(pre)
  } else if (inherits(pre,'station')) pre <- pre
  z <- aggregate(pre,by=month,FUN='mean')

  cal <- data.frame(y=coredata(x),x=coredata(z))
  attr(cal,'standard.error') <- attr(x,'standard.error')
  stats <- cor.test(cal$y,cal$x)
  wc.model <- lm(y ~ x, data=cal)
  if (plot) {
      dev.new()
      par(bty='n',cex.sub=0.7,col.sub='grey40')
      ylim <- range(cal$y,na.rm=TRUE); xlim=range(cal$x,na.rm=TRUE)
      dy <- diff(ylim)/25
      plot(cal$x,cal$y,pch=19,cex=1.5,col='grey',
           ylab=expression(paste(mu,' (mm/day)')),
         xlab=expression(paste(e[s],' (Pa)')),
         ylim=ylim,xlim=xlim,
         main='"Worst-case" fit based on seasonal variations',
         sub=paste(loc(x),' (',round(lon(x),2),'E/',round(lat(x),2),'N; ',
             alt(x),'m.a.s.l.)',sep=''))
      segments(x0=cal$x,y0=cal$y,x1=cal$x,y1=cal$y+2*attr(x,'standard.error'),
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x,y1=cal$y-2*attr(x,'standard.error'),
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x+2*attr(z,'standard.error'),y1=cal$y,
               col='grey')
      segments(x0=cal$x,y0=cal$y,x1=cal$x-2*attr(z,'standard.error'),y1=cal$y,
               col='grey')
      points(cal$x,cal$y,pch=19,cex=1.5,col='grey')
      grid()
      abline(wc.model)
      text(xlim[1],ylim[2],paste('Correlation=',round(stats$estimate,2),
                                 '(','p-value=',
                                 100*round(stats$p.value,4),'%)'),
           pos=4,cex=0.7,col='grey')
      text(xlim[1],ylim[2]-dy,paste('Regression: y=',
                                    round(wc.model$coeff[1],4), '+',
                                    round(wc.model$coeff[2],4), 'x (R2=',
                                    round(summary(wc.model)$r.squared,2),')'),
           pos=4,cex=0.7,col='grey')
      par(new=TRUE,fig=c(0.5,0.97,0.1,0.5),yaxt='n',xaxt='n',xpd=TRUE,
          cex.axis=0.7,col.axis='grey')
      plot((cal$x - mean(cal$x))/sd(cal$x),type='l',lwd=2,
           ylab='',xlab='',col=rgb(0.6,0.3,0))
      lines((cal$y - mean(cal$y))/sd(cal$y),type='l',lwd=2,col=rgb(0,0.3,0.6))
      par(xaxt = "s")
      axis(1,at=1:12,labels=month.abb,col='grey')
  }
  invisible(cal)
}
```

Functions for projection and prediction for the future/past:

```{r}
## Projection based on the calibration with the annual cycle:
muproject <- function(x,gcm,verbose=FALSE,prct=TRUE) {
  #print('projection')
  wcmodel <- lm(y ~ x, data=x)
  if (verbose) {
    print(summary(wcmodel))
    print(dim(gcm))
  }
  pq05 <- data.frame(x=coredata(gcm[,1]))
  pq95 <- data.frame(x=coredata(gcm[,2]))
  pmea <- data.frame(x=coredata(gcm[,3]))
  y <- cbind(predict(wcmodel,newdata=pq05),
             predict(wcmodel,newdata=pq95),
             predict(wcmodel,newdata=pmea))
  if (verbose) print(dim(y))
  if (prct) {
    ii <- is.element(year(gcm),2000:2010)
    bline <- mean(y[ii,3])
    y <- 100*y/bline
  }
  y <- zoo(y,order.by=index(gcm))
  names(y) <- names(gcm)
  return(y)          
}
```

```{r}
## Predict values of the wet-day mean mu taking a given predictor
mupredict <- function(x,pre,verbose=FALSE,prct=TRUE) {
  if (verbose) print('projection')
  wcmodel <- lm(y ~ x, data=x)
  if (verbose) {
    print(summary(wcmodel))
  }
  eval <- data.frame(x=coredata(pre))
  y <- predict(wcmodel,newdata=eval)
  if (prct) {
    ii <- is.element(year(pre),1961:1990)
    bline <- mean(y[ii])
    if (verbose) print(length(bline))
    y <- 100*y/bline
  }
  y <- zoo(y,order.by=index(pre))
  names(y) <- 'mu'
  return(y)          
}
```

**Generation of graphics/maps**

The following functions are used for the presentation of the results

```{r}
## Generate a map for the PC weights for the different modes of the
## wet-day mean mu annual cycle.
mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
  r2 <- as.numeric(r2)
  colpc <- rev(colscal("t2m",n=100))
  cz <- round(100*abs(pca$v[,ipca])/quantile(abs(pca$v[,ipca]),0.95))
  cz[cz > 100] <- 100; cz[cz < 1] <- 1
  col <- colpc[cz]
  pch <- rep(19,length(r2))
  mo <- c(r2) > 0.4
  hi <- c(r2) > 0.6
  pch[!hi] <- 1
  pch[!mo] <- 4
  cex <- 1.5*c(r2) + 0.2

  map(mu, xlim=xlim,ylim=ylim,bg='grey80',new=FALSE)
  points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
  par(xpd=TRUE)
  text(10,73,paste('Annual cycle in PC',ipca,
      'with variance of',round(100*pca$d[ipca]^2/sum(pca$d^2)),'%'),pos=4)

  colbar(pretty(pca$v[,ipca],n=100),colpc,fig = c(0.05, 0.1, 0.05, 0.2))

  par(xaxt = "n", yaxt = "s", fig = c(0.05,0.25,0.80,0.95),
      mar = c(0, 1, 0, 0), new = TRUE, las = 1, cex.axis = 0.5,bty='n')
  plot(pca$u[,ipca],type='l',lwd=3,col="red")
  par(xaxt = "s")
  axis(1,at=1:12,labels=month.abb)
}
```

```{r}
## Plot shaded regions
shade <- function(x,col=rgb(0.5,0.5,0.5,0.3),border=NULL) {
  t <- index(x)
  if (is.null(border)) border <- col
  y <- coredata(x)
  polygon(c(t,rev(t)),c(y[,1],rev(y[,2])),col=col,border=border)
  lines(t,y[,3],lwd=5,col=col)
}
```

```{r}
## Estimate correlation between different sets of wet-day mean mu
## The function is designed to be used in apply for best efficiency
cormu <- function(x) {
  n <- length(x); nh <- n/2
  x1 <- x[1:nh]; x2 <- x[(nh+1):n]
  ok <- is.finite(x1) & is.finite(x2)
  return(cor(x1[ok],x2[ok]))
}
```

The definition of functions is followed by code that carry out the analysis based on these.  First get the data needed - if they are not stored locally, download from Figshare where they are stored.:

```{r}
##-----------------------------------------------------------------------------------------

## Need to obtain some of the data files - fetch from Figshare:

if (!file.exists("mu.worstcasemu.rda") & figshare) {
  download.file("http://files.figshare.com/2193033/mu.worstcasemu.rda",destfile="mu.worstcasemu.rda") 
}

if (!file.exists("pre.worstcasemu.rda") & figshare) {
  download.file("http://files.figshare.com/2193038/pre.worstcasemu.rda",destfile="pre.worstcasemu.rda") 
}

if (!file.exists("cmip5.rda") & figshare) {
  download.file("http://files.figshare.com/2193041/cmip5.rda",destfile="cmip5.rda") 
}
```

The data can also be refreshed or updated with ECA\&D data:

```{r}
## Preparations that only needs to be done once. 
if (readecad) {
## This section generates a processed data file from scratch using ECA&D data:
  pr <- station(src=c('metnod','ecad'),param='precip',nmin=nmin,it=c(1961,2010),lon=xlim,lat=ylim)
  pr <- subset(pr,it=c(1961,2014))
  nt <- apply(pr,2,FUN='nv')
  pr <- subset(pr,is=(nt >=19000))
  cpr <- coredata(pr)
  cpr[cpr > 250] <- NA
  coredata(pr) <- cpr
  save(file='pr.worstcasemu.rda',pr)
  file.remove('mu.worstcasemu.rda')
}

if (!file.exists('mu.worstcasemu.rda')) {
## pr.worstcasemu.rda is a huge file with daily precipdata based on ECA&D - generated above
  load('pr.worstcasemu.rda')

  ## Time series of the annual wet-day freq & mean - for evaluation
  ## Randomly sub-sample due to excessive volume:
  FW <- annual(pr,FUN='wetfreq',nmin=350)
  MU <- annual(pr,FUN='wetmean',nmin=350)
  fw <- aggregate(pr,month,FUN='wetfreq')
  mu <- aggregate(pr,month,FUN='wetmean')

  ## Strip away stations with a lot of missing data
  n <- apply(pr,2,FUN='nv')
  y1 <- subset(subset(pr,is=(n==max(n))),is=1)
  save(file='mu.worstcasemu.rda',mu,MU,fw,FW,y1,n)
} else load('mu.worstcasemu.rda')
```

The convention here is that the wet-day mean precipitation (precipitation intensity) is represented by symbol $\mu$ and the variable name 'mu' in the R-scripts. The wet-day frequency $f_w$ is referret to as 'fw'. Lower and upper case refer to the mean seasonal cycle and annually aggregated values respectively. There are some stations with many missing values and some with short series. Also exclude data records with unrealistic long-term trends (due to dubious data or short series).

```{r}
## Keep stations with no missing data and stations
## without suspect outlier trends
nok <- n
ok <- apply(mu,2,function(x) sum(!is.finite(x))==0) &
      apply(fw,2,function(x) sum(!is.finite(x))==0) &
      abs(apply(MU,2,FUN='trend.coef')) <= 1 &
      abs(apply(FW,2,FUN='trend.coef')) <= 0.02
mu <- subset(mu,is=ok) # REB 2015-03-01
MU <- subset(MU,is=ok) # REB 2015-03-01
fw <- subset(fw,is=ok)
FW <- subset(FW,is=ok)

## Remove stations with little data
nval <- apply(MU,2,nv)
mu <- subset(mu,is=(nval > 50))
MU <- subset(MU,is=(nval > 50))
fw <- subset(fw,is=(nval > 50))
FW <- subset(FW,is=(nval > 50))
nval <- apply(FW,2,nv)
mu <- subset(mu,is=(nval > 50))
MU <- subset(MU,is=(nval > 50))
fw <- subset(fw,is=(nval > 50))
FW <- subset(FW,is=(nval > 50))
```

Get and process the predictor data. Then the predictand data is processed: estimate annual mean aggregates and the mean seasonal cycle. Remove stations with large gaps of missing values.

```{r}
print("predictor")
if (!file.exists('pre.worstcasemu.rda')) {
  t2m <- retrieve('air.mon.mean.nc',lon=c(-100,30),lat=c(0,40))
  pre <- spatial.avg.field(C.C.eq(t2m))
  attr(pre,'region') <- '100W,30E/0N,40N'
  save(file='pre.worstcasemu.rda',pre)
} else load('pre.worstcasemu.rda')
```

Now the data is ready for the analysis. Calibrate the regression models and extract the regression coeffieients. Use 'apply' to speed up the process for multiple stations.

```{r}
## Extract the monthly aggregates for all stations
print('apply mucal')
V <- apply(mu,2,FUN='mucal',pre=pre)
Beta <- lapply(V,FUN='beta')

## Collect the R-squared statistics from lm(y ~ x) for each site
print('muskill')
r2 <- lapply(V,muskill)
```

**Plot figure 1**

Figure 1 illustrates how the mean seasonal cycle in $\mu$ and the area mean predictor compare and what the regression results for one example station.

```{r,echo=TRUE}
## Extract the mean, min, max, wettest month, and driest months in terms of mu
## for all stations:
print('muclim')
X <- apply(mu,2,FUN='muclim')
print(table(X[4,]))
wmns <- as.numeric(rownames(table(X[4,])))
nc <- max(wmns) - min(wmns) + 1
cols <- colscal(n=nc)
col1 <- cols[c(X[4,])]
cex <- 1.5*n/max(n)

## The relationship between mu and e_s for one station
## to show the calibration procedure
is <- (1:length(n))[X[4,]==8][1]
print(paste('plot mucal - is=',is))
mucal(subset(mu,is=is),pre=pre,verbose=TRUE,plot=TRUE)
figlab('Figure 1')
```

**Figure 2a+b**

Figure 2a+b show the principal component analysis (PCA) applied to the mean seasonal cycle for the different locations to show how it varies geographically.

```{r,echo=TRUE}
## Remove locations with missing values for PCA 
print('Mu - matrix for PCA')
Mu <- as.matrix(coredata(mu))

## Anomalies wrt the mean value at each location.
Mu <- apply(Mu,2,function(x) (x - mean(x)))
pca <- svd(Mu)

## Plot maps with PCs:
print('Maps with PCs')
mupcamap(mu,pca,1,xlim,ylim,r2)
figlab('Figure 2a')
mupcamap(mu,pca,2,xlim,ylim,r2)
figlab('Figure 2b')
```

**Mean seasonal cycle**

Examine how the components of the mean seasonal cycle correlates with the skill-score of the empirical models.

```{r,echo=TRUE}
print('Variance accounted for by the modes:')
print(round(100*pca$d**2/sum(pca$d**2),1))

## The sign of PCs is arbitrary...
print(cor.test(pca$v[,1],as.numeric(r2)))
print(cor.test(pca$v[,2],as.numeric(r2)))
print(cor.test(pca$v[,3],as.numeric(r2)))
```
**Trend analysis**

Estimate trend statistics and extract only the results for locations with a good match. The models calibrated here only seem to be valid for regions dominated by convective precipitation and are poor over regions where orographically forced precipitation dominate.

```{r,echo=TRUE}
## Estimate trend statistics for both fw and mu:
## Use subset - in the PCA there is something strange in the 1990s.
fw.trend <- 100*apply(subset(FW,it=c(1950,1990)),2,'trend.coef')/
                      apply(FW,2,'mean',na.rm=TRUE)
mu.trend <- 100*apply(subset(MU,it=c(1950,1990)),2,'trend.coef')/
                      apply(MU,2,'mean',na.rm=TRUE)

diagnose(MU)

## Select only the sites which have an R2 greater than 0.6:
is <- (1:length(r2))[as.numeric(r2)> 0.6]
mux <- subset(mu,is=is)
MUx <- subset(MU,is=is)
FWx <- subset(MU,is=is)
Xx <- X[,is]
Vx <- V[is]
```

**Aggregated results from the GCMs**

Need to collect aggregated results based on the ensembles of GCM projections.

```{r,echo=TRUE}
## Read the GCM ensembles andestimate the 5 and 95 percentiles as well as
## the ensemble mean.
## Get the annual mean temperature from GCM ensembles:
if (!file.exists('cmip5.rda')) {
  print('get CMIP5 RCP4.5')
  rcp4.5 <- readGCMs(path='CMIP5.monthly/rcp45/',pattern='tas_Amon_ens_rcp')
  print('get CMIP5 RCP8.5')
  rcp8.5 <- readGCMs(path='CMIP5.monthly/rcp85/',pattern='tas_Amon_ens_rcp')
  print('get CMIP5 RCP2.6')
  rcp2.6 <- readGCMs(path='CMIP5.monthly/rcp26/',pattern='tas_Amon_ens_rcp')
  save(file='cmip5.rda',rcp4.5,rcp8.5,rcp2.6)
} else load('cmip5.rda')

## derive time series for each location:
Z.rcp4.5 <- lapply(Vx,'muproject',rcp4.5)
Z.rcp8.5 <- lapply(Vx,'muproject',rcp8.5)
Z.rcp2.6 <- lapply(Vx,'muproject',rcp2.6)
t <- index(rcp4.5)
```
Extract the statistics of the different emission scenarios and time slices

```{r,echo=TRUE}
## Estimates for mu in 2010:
mu2010.rcp4.5 <- lapply(Z.rcp4.5,'window',start=2010,end=2010)
mu2010.rcp8.5 <- lapply(Z.rcp8.5,'window',start=2010,end=2010)
mu2010.rcp2.6 <- lapply(Z.rcp2.6,'window',start=2010,end=2010)
x2010 <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[3]]))
x2010u <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[2]]))
y2010 <- as.numeric(lapply(mu2010.rcp8.5,function(x) x[[3]]))
y2010u <- as.numeric(lapply(mu2010.rcp8.5,function(x) x[[2]]))
z2010 <- as.numeric(lapply(mu2010.rcp2.6,function(x) x[[3]]))
z2010u <- as.numeric(lapply(mu2010.rcp2.6,function(x) x[[2]]))

## Repeat for 2100:
mu2100.rcp4.5 <- lapply(Z.rcp4.5,'window',start=2100,end=2100)
mu2100.rcp8.5 <- lapply(Z.rcp8.5,'window',start=2100,end=2100)
mu2100.rcp2.6 <- lapply(Z.rcp2.6,'window',start=2100,end=2100)
x2100 <- as.numeric(lapply(mu2100.rcp4.5,function(x) x[[3]]))
x2100u <- as.numeric(lapply(mu2100.rcp4.5,function(x) x[[2]]))
y2100 <- as.numeric(lapply(mu2100.rcp8.5,function(x) x[[3]]))
y2100u <- as.numeric(lapply(mu2100.rcp8.5,function(x) x[[2]]))
z2100 <- as.numeric(lapply(mu2100.rcp2.6,function(x) x[[3]]))
z2100u <- as.numeric(lapply(mu2100.rcp2.6,function(x) x[[2]]))

## Data frams with changes in percentages:
mu.2100 <- data.frame(mean.RCP4.5=x2100 - x2010,q95.RCP4.5=x2100u - x2010u,
                      mean.RCP2.6=z2100 - z2010,q95.RCP4.5=z2100u - z2010u,
                      mean.RCP8.5=y2100 - y2010,q95.RCP8.5=y2100u - y2010u)
```

**Plot figure 3**

```{r,echo=TRUE}
## Fig 3
## Plot a map of projected values:
## Map showing RCP4.5 ensemble mean and the upper 95% change in the
## outer part of the symbol. Also an insert with box-plot diagram
## showing the other RCPs.

cols <- colscal(n=100,col='precip')
cx2100 <- round(x2100 - x2010)
cx2100u <- round(y2100u - y2010u)
cx2100[cx2100 < 1] <- 1; cx2100[cx2100 > 100] <- 100
cx2100u[cx2100u < 1] <- 1; cx2100u[cx2100u > 100] <- 100
colx <- cols[cx2100]
coly <- cols[cx2100u]
Cex <- 1.25

print("Plot a map of projected values:")
map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',gridlines=FALSE,
    colbar=list(col=cols,n=12,type="p",h=0.6,v=1))
points(lon(mux),lat(mux),pch=19,col=colx,cex=Cex)
points(lon(mux),lat(mux),pch=21,col=coly,cex=Cex,lwd=2)
par(xpd=TRUE)
text(20,73,'Wet-day mean: 2100')
legend(20,32,c(expression(bar(x)),expression(q[95])),
       pch=c(21,19),bty='n',col='grey',text.col='grey',horiz=TRUE)

colbar(pretty(c(x2100u- x2010u, x2100- x2010),n=15),cols,
       fig = c(0.05, 0.1, 0.05, 0.2))

par(new=TRUE,fig = c(0.05, 0.4, 0.75, 0.975),
    cex.axis=0.75,mar=c(1,1,0.1,0.1),xaxt='n')
boxplot(mu.2100,col=c(rep(rgb(0.5,0.5,0.5,0.3),2),
                      rep(rgb(0.5,1,0.5,0.3),2),
                      rep(rgb(1,0.5,0.5,0.3),2)))
par(xaxt='s')
axis(1,c(1,3,5),labels=c('RCP4.5','RCP2.6','RCP8.5'))
grid()

figlab('Figure 3')
```

**Supporting material/analysis (SM)**

Extract the essential information

```{r,echo=TRUE}
## Evaluation:
## Use the calibration strategy to predict the annual mu based on the
## predictor (t2m -> e_s)
print('evaluation: correlation')
mu.eval <- lapply(Vx,'mupredict',annual(pre),prct=FALSE)
m <- length(mu.eval); n <- length(mu.eval[[1]])
MUz <- matrix(unlist(mu.eval),n,m)
MUz <- zoo(MUz,order.by=index(mu.eval[[1]]))
MUz <- subset(MUz,it=MUx)
MUx <- subset(MUx,it=MUz)
FWx <- subset(FWx,it=MUz)
r.eval <- apply(rbind(coredata(MUz),coredata(MUx)),2,'cormu')
print(summary(r.eval))
```

##** ----------- Supporting Material ------------------------**

The supporting material includes a number of diagnostics and plots that chencks a number of assumptions made in this study.

##**figure SM2**

Presents the statistics of $R^2$ from the model calibration based on the different locations.

```{r,echo=TRUE}
par(xaxt='n',yaxt='n',bty='n')
plot(c(0,1),c(0,1),type='n',xlab='',ylab='')
text(0.5,0.5,'Supporting figures',cex=2,font=2)

## Plot the statistics of R2:
hist(100*as.numeric(r2),breaks=seq(0,100,by=5),lwd=2,col=rgb(0,0.3,0.5),
     xlab=expression(paste(R^2,' (%)')),freq=TRUE,
     main="Summary of regression scores")
grid()
figlab('Figure SM2')
```
##**Figure SM12**

A figure showing one example of projections for one location, showing all three emission scenarios.

```{r}
## Example of estimates for 2050:
print('Example plot - evolution')
par(bty='n')
i <- 1
N <- length(Z.rcp4.5)
plot(Z.rcp4.5[[i]],plot.type='single',lty=c(2,2,1),lwd=c(1,1,2),
     ylab='%',ylim=c(80,150),
     main=paste('Wet-day mean at',loc(subset(mux,is=i))))
shade(Z.rcp4.5[[i]],col=rgb(0.5,0.5,0.5,0.3))
shade(Z.rcp8.5[[i]],col=rgb(1,0.5,0.5,0.3))
shade(Z.rcp2.6[[i]],col=rgb(0.5,1,0.5,0.3))
grid()
figlab('Figure SM12')
```

Further processing for Fig SM3: extract information about trends in $\mu$.

```{r,echo=TRUE}
## Trend in projected wet-day mean
trendbeta <- unlist(lapply(Z.rcp4.5,function(x) trend.coef(x[,3])))
## Strange results:
print((1:N)[trendbeta < 0])

## test: Does the model predict observed trends?
print('evaluation: trends')

## Make sure to compare series with data for same times
mask <- !is.finite(coredata(MUx))
class(MUz) <- class(MUx)
muz <- coredata(MUz)
muz[mask] <- NA; dim(muz) <- dim(MUz)
coredata(MUz) <- muz

## Only look at stations with more than 50 years with data
ok <- (apply(coredata(MUx),2,nv) > 50)
MUz <- subset(MUz,is=ok)
MUx <- subset(MUx,is=ok)
FWx <- subset(FWx,is=ok)

trend.mux <- apply(MUx,2,'trend.coef')
trend.pre <- apply(MUz,2,'trend.coef')
trenderr.mux <- apply(MUx,2,'trend.err')
trenderr.pre <- apply(MUz,2,'trend.err')

## Need to get a picture whether the predictions gives a plausible
## upper limit.
trend.sense <- data.frame(x=c(-trend.mux,trend.mux),
                          y=c(-trend.pre,trend.pre))

xylim <- max(abs(c(trend.mux,trend.pre)))*c(-1,1)
```

##**Figure SM3**

Historical trends in the wet-day mean preciptation

```{r}
par(bty='n',col.sub='grey',cex.sub=0.8)
plot(trend.mux,trend.pre,pch=19,col=rgb(0.6,0.2,0,0.3),cex=1.5,
     main=expression(paste('Trends in ',mu,': observed and predicted upper limit')),
     xlab='Observed trend (mm/decade)',ylab='predicted trend (mm/decade)',
     xlim=xylim,ylim=xylim,
     sub=paste('Mean correlation for local year-to-year variations over t=[',
       start(MUx),',',end(MUx),
       '] is ',round(mean(r.eval),2),' (',round(quantile(r.eval,0.05),2),', ',
       round(quantile(r.eval,0.95),2),
       ')',sep='') )
grid()

polygon(c(xylim[1],xylim[2],xylim[1],xylim[1]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
        col=rgb(0.2,0.6,1,0.1),border=NA)
polygon(c(xylim[1],xylim[2],xylim[2],xylim[2]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
        col=rgb(1,0.2,0.2,0.1),border=NA)
points(trend.mux,trend.pre,pch=1,col=rgb(0,0,0,0.1),cex=1.5)

## Plot error bars 
apply(rbind(trend.mux,trend.pre,trenderr.mux,trenderr.pre),2,
      FUN=function(x) {lines(x[1]+c(-2,2)*x[3],x[2]+c(0,0),col=rgb(0.6,0.2,0,0.1))
                       lines(x[1]+c(0,0),x[2]+c(-2,2)*x[4],col=rgb(0.6,0.2,0,0.1))
                       lines(x[1]+c(-1,1)*0.01,x[2]+c(2,2)*x[4],col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(-1,1)*0.01,x[2]+c(-2,-2)*x[4],col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(-2,-2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
                       lines(x[1]+c(2,2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
                     })
figlab('Figure SM3')
```

##**Figure SM4**

```{r}
## Map showing trends in mu
map(MU,FUN='trend',colbar=list(breaks=seq(-1,1,length=21),rev=TRUE),cex=cex)
figlab('Figure SM4',ypos=0.999)
figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
```
##**Figure SM5**

```{r}
## 99-percentile wet-day mean and typical wet-day frequency
## The Tellus paper on specification approximately annual maximum.
## Prob(X>x) for annual maximum for 24hr precip
## Pr(X > x) = 1/365.25 = fw*(1-p): p = 1 - 1/(365.25*fw)
par(bty='n')
W <- apply(fw,2,FUN='muclim')
hist(100*(1-1/(W[1,]*365.25)),col='steelblue',lwd=2,breaks=seq(90,100,by=0.1),
     main='Wet-day percentile for annual maximum 24-precipitation',
     xlab='p (%)')
grid()
figlab('Figure SM5')
```

##**Figure SM6**

```{r}
## Statistics of trend in wet-day frequency
print('Wet-day frequency statistics')
hist(fw.trend,breaks=seq(-50,50,by=1),col='grey',lwd=2,
     main='Trend in wet-day frequency',
     xlab=expression(paste(f[w],' (%/decade)')))
grid()
figlab('Figure SM6')
```
##**Figure SM7**

```{r}
## Map showing trends in fw
#dev.new()
map(FW,FUN='trend',colbar=list(breaks=seq(-0.05,0.05,length=21),rev=TRUE),cex=cex)
figlab('Figure SM7',ypos=0.999)
figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
```

##**Figure SM1**

```{r}
## Figure SM1.
## test: See if the quantiles are consistent when the mean mu varies.
qtest <- aggregate(y1,year,FUN='qqexp')
qx <- c(coredata(qtest[,1:101]))
qy <- c(coredata(qtest[,102:202]))

par(bty='n')
plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
     main='Test: exponential distribution & changing mean',
     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
lines(c(0,40),c(0,40),col='grey')
figlab('Figure SM1')
```

##**Figure SM14: compare the mean seasonal variations in the different precipitation statistics**

```{r}
pr.mean <- aggregate(y1,by=month,FUN='mean')
pr.mu <- aggregate(y1,by=month,FUN='wetmean')
pr.fw <- aggregate(y1,by=month,FUN='wetfreq')
y1.l <- spell(y1,threshold=1)
pr.wet <- aggregate(subset(y1.l,is=1),by=month,FUN='mean')
pr.dry <- aggregate(subset(y1.l,is=2),by=month,FUN='mean')

par(bty='n',xaxt='n')
plot(merge(pr.mean,pr.mu,10*pr.fw,pr.wet,pr.dry),plot.type='single',
     col=c('steelblue','darkblue','grey','darkgreen','red'),
     lwd=c(3,3,1,1,1),ylab="",xlab="Calendar month",main=loc(y1))
grid()
par(yaxt='s',xaxt='s')
axis(1,at=1:12,labels=month.abb,cex.lab=0.7, col='grey')
axis(4,at=10*pretty(pr.fw),pretty(pr.fw),col='grey')

legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
               expression(bar(n[c*w*d])),expression(bar(n[c*d*d]))),bty='n',
       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1))
figlab('Figure SM9')
```

##**Figure SM10: Map of the future return values**

```{r}
print('Map of return values for 2100 with RCP4.5')
rv <- MUx
## x2100u gives the percentage of mu,
coredata(rv) <- t((x2100u/100)*apply(coredata(MUx),2,'mean')*log(365.25*t(coredata(FWx))))
cexr2 <- 1.5*c(as.numeric(r2)) + 0.2
map(rv,FUN='mean',cex=cexr2,colbar=list(breaks=seq(30,100,by=1)))
figlab('Figure SM10',ypos=0.999)
figlab('Return values for 2100 assuming RCP4.5',xpos=0.3,ypos=0.999)
```

##**Figure SM9**

```{r}
## Print the numbers:
print('--- Changes in 20% returnvalue in terms of % from 2010:')
print(lapply(mu.2100,summary))
print(paste('Summary for',length(mu.2100[[1]]),'locations'))

## Compare the regression coefficients derived from individual
## seasonal cycle with that derived from mean climatology.

data(mu.eq.f.tx)

col <- rgb(0.1,0.1,0.7,0.25)
mutx <- summary(mu.eq.f.tx)$coefficients[c(2,4)]
b1 <- as.numeric(lapply(Beta,function(x) x[1]))
e1 <- as.numeric(lapply(Beta,function(x) x[2]))

par(bty='n')
plot(b1,pch=19,col=col,cex=cexr2,
     main=expression(paste('Scaling coefficient for ',mu,' and ',e[s])),
     xlab='',ylab=expression(beta))
grid()
for (i in 1:length(b1)) {
  lines(rep(i,2),b1[i]+e1[i]*c(-2,2),col=col)
  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(-2,-2),col=col)
  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(2,2),col=col)
}
polygon(c(1,rep(length(b1),2),rep(1,2)),
        mutx[1]+mutx[2]*c(-2,-2,2,2,-2),
        border=rgb(0.5,0.5,0.5,0.4),col=rgb(0.5,0.5,0.5,0.3))
lines(c(1,length(b1)),rep(mutx[1],2),lwd=3,col=rgb(0.5,0.5,0.5,0.3))
figlab('Figure SM9')
```

##**Figure SM13**

```{r}
## Show the predictor area:
X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
figlab('Figure SM13',xpos=0.8,ypos=0.999)
```

##**Analysis from other continents and with local temperature**

Additional analysis testing the connection between local temperature and wet-day mean precipitation

```{r}
## Regression analysis between the wet-day mean mu and the mean temperature.

corhalf <- function(x) {
  n <- length(x)
  x1 <- x[1:(n/2)]; x2 <- x[(n/2+1):n]
  ok <- is.finite(x1) & is.finite(x2)
  r <- cor(x1[ok],x2[ok])
  return(r)
}
```

```{r}
## CLARIS
load('claris.Tx.rda')
load('claris.Pr.rda')

Tx1 <- Tx
Pr1 <- Pr

## COST-VALUE
load('VALUE_ECA_86_v2/stationsVALUE-exp1a.rda')
Tx2 <- Tx
Pr2 <- Pr
```

```{r}
## Read North american data:
if (!file.exists('mut2m.GDCN.rda')) {
  source('readGDCN.R')

  gdcn <- list.files('/disk1/GDCN-data_disk2',pattern='dly', full.names = TRUE)
  finfo <- file.info(gdcn)
  fok <- (finfo$size > 200000)
  gdcn <- gdcn[fok]
  n <- length(gdcn)

  plot(c(-180,180),c(-90,90),type='n',xlab='',ylab='')
  data(geoborders)
  lines(geoborders)

  for (i in 1:n) {
    pr <- readGDCN(gdcn[i])
    tx <- readGDCN(gdcn[i],param="tmax")
    if ( (nv(pr) >  20000) & (nv(tx) >  20000)) {
      pr <- subset(pr,it=c(1945,2015))
      tx <- subset(tx,it=c(1945,2015))

      if (i==1) {
        mu <- annual(pr,FUN='wetmean')
        fw <- annual(pr,FUN='wetfreq')
        t2m <- annual(tx,FUN='mean',na.rm=TRUE)
      } else {
        mu <- combine(mu,annual(pr,FUN='wetmean'))
        fw <- combine(fw,annual(pr,FUN='wetfreq'))
        t2m <- combine(t2m,annual(tx,FUN='mean',na.rm=TRUE))
      }
      print(paste(i,' (',n,'): ',loc(pr),', ',cntr(pr),' #validdata=',nv(pr),
                  'Tx: ',round(mean(tx,na.rm=TRUE),2),
                  round(min(tx,na.rm=TRUE),2),
                  round(max(tx,na.rm=TRUE),2),lat(tx)))
      points(lon(pr),lat(pr),pch=19,col='darkgreen')
    }
  }
  save(file='mut2m.GDCN.rda',mu,fw,t2m)
} else load('mut2m.GDCN.rda')
```

```{r}
## Aggregate annual statistics based on the combined data sources:
MU <- combine(mu,annual(Pr1,FUN='wetmean'),annual(Pr2,FUN='wetmean'))
FW <- combine(fw,annual(Pr1,FUN='wetfreq'),annual(Pr2,FUN='wetfreq'))
T2M <- combine(t2m,annual(Tx1,FUN='mean'),annual(Tx2,FUN='mean'))
nval <- apply(coredata(MU),2,'nv')
attr(T2M,'variable') <- 't2m'
es <- C.C.eq(T2M)

#map(MU,FUN='mean',cex=0)
#map(MU,FUN='trend',cex=0,colbar=list(rev=TRUE,breaks=seq(-1,1,length=21)))
#map(FW,FUN='trend',cex=0,colbar=list(rev=TRUE,breaks=seq(-0.01,0.01,length=21)))
calmu <- data.frame(x=as.numeric(apply(coredata(es),2,'mean',na.rm=TRUE)),
                    y=as.numeric(apply(coredata(MU),2,'mean',na.rm=TRUE)),
                    fw=as.numeric(apply(coredata(FW),2,'mean',na.rm=TRUE)),
                    z=alt(MU),lat=lat(MU),lon=lon(MU),nval=nval)
premu <- calmu; premu$x[0] <- 0
model.mutx <- lm(y ~ x, weights=fw,data=calmu)
print(summary(model.mutx))
data(geoborders)
Fw <- apply(coredata(fw),2,'sum',na.rm=TRUE)
```

##** Figure SM8**

Assess the connection between temperature and the wet-day mean precipitation.

```{r}
par(bty='n')
col <- rgb((1+sin(pi*calmu$lat/180))/2,
           cos(pi*calmu$lon/180)^2,
           1-(1+sin(pi*calmu$lat/180))/2,0.15)
cex <- 1.5*Fw/max(Fw,na.rm=TRUE)
plot(calmu$x,calmu$y,pch=19,col=col,cex=cex,
     main='Wet-day mean precipitation temperature dependency',
     ylab=expression(mu*phantom(0)*(mm/day)),
     xlab=expression(e[s]*phantom(0)**(Pa)))
points(calmu$x,calmu$y,pch=21,col=rgb(0.5,0.5,0.5,0.2),cex=cex)
lines(calmu$x,predict(model.mutx),col=rgb(0.4,0.4,0.4,0.1),lwd=2)

par(new=TRUE,fig=c(0.15,0.45,0.7,0.9),mar=rep(0,4),xaxt="n",yaxt="n")
plot(calmu$lon,calmu$lat,pch=19,col=col,cex=0.3*cex)
lines(geoborders,col='grey')
points(calmu$lon,calmu$lat,pch=19,col=col,cex=0.3*cex)
figlab('Figure SM8')

mu.eq.f.tx <- model.mutx
attr(mu.eq.f.tx,'input') <- 'saturation water pressure e_s (Pa)'
attr(mu.eq.f.tx,'predictand') <- 'C.C.eq(tmax)' 
attr(mu.eq.f.tx,'output') <- 'wet-day mean precipitation (mm/day)'
attr(mu.eq.f.tx,'calibration') <- 'mean climatology'
attr(mu.eq.f.tx,'source script') <- 'mut2m.R'
attr(mu.eq.f.tx,'timestamp') <- date()
attr(mu.eq.f.tx,'calibration_data') <- calmu
save(file='esd/data/mu.eq.f.tx.rda',mu.eq.f.tx)
```

```{r}
## Estimate the year-by-year correlation in es and mu

X <- matchdate(MU,es)
Y <- matchdate(es,MU)
w <- apply(coredata(fw),2,'mean')
ok <- w > 0.25
Z <- rbind(coredata(X),coredata(Y))
r <- apply(Z,2,corhalf)
hist(r)
```

### 2016-04-08 CORDEX/ensemble.R

@@ -1,0 +1,54 @@
+library('esd')
+data("IPCC.AR5.Table.9.A.1")
+data("global.t2m.gcm")
+gcmnm <- attr(global.t2m.cmip5,"model")
+gcmnm <- toupper(substr(gcmnm,1,nchar(gcmnm)-2))
+gcmnm[is.element(substr(gcmnm,1,13),"CSIRO-MK3-6-0")] <- "CSIRO-MK3.6.0"
+gcmnm[is.element(gcmnm,"FGOALS_G2")] <- "FGOALS-G2"
+gcmnm[is.element(gcmnm,"FIO-ESM")] <- "FIO-ESM V1.0"
+gcmnm[is.element(gcmnm,"BCC-CSM1-1")] <- "BCC-CSM1.1"
+gcmnm[is.element(gcmnm,"BCC-CSM1-1-M")] <- "BCC-CSM1.1(M)"
+gcmnm[is.element(gcmnm,"CESM1-BGC")] <- "CESM1(BGC)"
+gcmnm[is.element(gcmnm,"CESM1-CAM5")] <- "CESM1(CAM5)"
+gcmnm[is.element(gcmnm,"CNRM-CM5")] <- "CNRM-CM51"
+gcmnm[is.element(gcmnm,"INMCM4")] <- "INM-CM4"
+gcmnm[is.element(substr(gcmnm,1,7),"EC-EART")] <- "EC-EARTH"
+n <- length(gcmnm); m <- 12
+X <- matrix(rep(NA,n*m),n,m)
+for (i in 1:n) {
+  ii <- is.element(toupper(IPCC.AR5.Table.9.A.1$Model.Name),gcmnm[i])
+  if (sum(ii)>0)
+    X[i,] <- c(as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name)[ii][1],
+               as.character(IPCC.AR5.Table.9.A.1$Horizontal.Grid)[ii][1],
+               IPCC.AR5.Table.9.A.1$Number.of.Vert.Levels[ii][1],
+               IPCC.AR5.Table.9.A.1$Grid.Top.hPa[ii][1],
+               as.character(IPCC.AR5.Table.9.A.1$Component.Name.or.type)[ii][1],
+               as.character(IPCC.AR5.Table.9.A.1$Atmos.Chemistry.Component.Name)[ii][1],
[…]

### 2016-04-08 paper58/paper58.R

@@ -1,0 +1,1 @@
+/home/rasmusb/Dropbox/Public/LaTeX/paper58/paper58.R

### 2016-04-11 ClimaTrans/climatrans.Rmd
new updates - Climatrans

@@
 
-```{r}
+```{r,echo=TRUE}
 library(esd)
 library(ncdf)
+## Information about the system and session
+sessionInfo()
 ```
@@
 ## time origin: minutes since 1901-01-01 00:00
+library(ncdf)
 ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/delhi.nc')
@@
 if (!exists("es")) {
-  t2m <- retrieve('data/ncep/air.mon.mean.nc',lon=c(60,100),lat=c(0,23))
+  t2m <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,23))
   es <- aggregate(tas2es(t2m,season=season),year,'mean')
@@
 if (!exists("slp")) {
-  slp <- retrieve('data/ncep/slp.mon.mean.nc',lon=c(60,100),lat=c(0,23))
+  slp <- retrieve('slp.mon.mean.nc',lon=c(60,100),lat=c(0,23))
   slp <- aggregate(subset(slp,it=month.abb[season]),year,'mean')
@@
   plot(as.4seasons(anomaly(climatrans.tx)))
+  for(i in 1:3) lines(trend(subset(as.4seasons(anomaly(climatrans.tx)),is=i)))
+  grid()
 ```
@@
 ```{r}
+## The predictor
   predictor <- retrieve('air.mon.mean.nc',param='air',lon=c(60,90),lat=c(10,35))
+  predictor <- aggregate(subset(predictor,it='jja'),year,FUN='mean')
+  index(predictor) <- year(predictor)
+  eof.t2m <- EOF(predictor)
+## The predictand
+  Tx <- subset(as.4seasons(climatrans.tx),it='jja')
+  pca.tx <- PCA(Tx)
+  class(pca.tx) <- c("pca","station", "annual", "zoo")
+  index(pca.tx) <- year(pca.tx)
+  ds.tx <- DS(pca.tx,eof.t2m)
+## Check the residuals
+  tx.2 <- as.residual(as.station(ds.tx))
+  tx.2 <- attrcp(Tx,tx.2); class(tx.2) <- class(Tx)
+  pca.tx.2 <- PCA(tx.2)
+  plot(pca.tx.2)
+```
 
+Apply the analysis of number of hot days
+
+```{r}
   trh <- c(40,30,32)

### 2016-04-19 EU-Circle/eu-circle_torbay-precip.Rmd
EU-Cricle Trobay results

---
title: "eu-circle-torbay-precip.Rmd"
author: "Rasmus Benestad"
date: "April 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

This script retreives precipitation for Southwestern England from the EOBS data set and estimates the wet-day mean precipitation and frequency.

The EOBS data was chosen here because the rain gauge data from stations (MIDAS) was intermittent and not readily available for an analysis that makes use of PCA in order to make the most use of the information embedded in many parallel data records.

Rain gauges provide a very tiny sample of the rainfall, and previous work suggests that downscaling can provide a better picture if it's applied to a group of stations rather than single stations (Benestad et al., 2015; doi:10.342/tellusa.v67.28326). Downscaling applied to products from principal component analysis (PCA) benefits from an emphasis on cohernt structure of variability in time and space, which can improve the signal-to-noise ratio. PCA requires no missing values, but the EOBS gridded data provides complete records.

```{r predictand}
library(esd)
pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(-6,-1),lat=c(50,52))
mu <- annual(pr.eobs,FUN='wetmean')
fw <- annual(pr.eobs,FUN='wetfreq')
```

## Past trends

Climate change has taken place for some time, and past trends may provide some indication about whether a chosen climate parameter such as the wet-day mean precipitation $\mu$ or wet-day frequency $f_w$ are sensitive to a global warming:

```{r trend}
## The area mean wet-day mean precipitation
plot(mu,new=FALSE)
## The area mean wet-day frequency
plot(fw,new=FALSE)
## The early part of the record looks suspect:
mu <- subset(mu,it=c(1958,2015))
fw <- subset(fw,it=c(1958,2015))
## Show maps of trends for the wet-day mean precipitation
map(mu,FUN='trend',
    colbar=list(breaks=seq(-0.12,0.12,by=0.02),pal='t2m',rev=TRUE))
## Map of trends in wet-day frequency:
map(fw,FUN='trend',
    colbar=list(breaks=seq(-0.01,0.01,by=0.001),pal='t2m',rev=TRUE))
```

The trend analysis for the wet-day frequency suggest a general increase in the number of wet-days over the  perdiod 1960--2015 while the intencity .

The predictors can be organised as EOFs to simplify the analysis and make use of properties as the products being orthogonal:

```{r EOFs-in-precip}
## The area mean wet-day mean precipitation
eof.mu <- EOF(mu)
## The area mean wet-day frequency
plot(eof.mu)
## Show maps of trends for the wet-day mean precipitation
eof.fw <- EOF(fw)
## Map of trends in wet-day frequency:
plot(eof.fw)
```

The diagnostics of the EOFs suggest that the leading mode explains a high proportion of the variability on an annual time scale.

The mean sea-level pressure (SLP) is used as predictor for the annual wet-day frequency, as this gives a good description of the circulation. A number of different spatial domains were tested, and for SLP and $f_w$, the results were fairly robust with respect to the choice of domain.

The predictor for the wet-day mean precipitation was taken to be the saturation vapour pressure over the North Atlantic, which is regarded as the main source for the atmospheric moisture over northern Europe.

## Predictor

```{r predictor}
## The mean sea-level pressure
slp <- retrieve('slp.mon.mean.nc',lon=c(-80,0),lat=c(20,70))
## The surface temperature
t2m <- retrieve('air.mon.mean.nc',lon=c(-80,0),lat=c(10,70))
## Saturation water pressure estimated from maritime temperatures
es <- C.C.eq(mask(t2m,land=TRUE))
eof.slp<- EOF(annual(slp))
eof.t2m <- EOF(annual(t2m))
eof.es <- EOF(annual(es,FUN='mean'))
```

## Cannical Correlation analysis

Carry out canonical correlation analysis (CCA) to explore relationships between the large-scale conditions such as SLP/temperature and the precipitation over soutwestern England:

```{r cca}
## The wet-day frequency and SLP
cca.fw <- CCA(eof.fw,eof.slp)
plot(cca.fw)
cca.mu <- CCA(eof.mu,eof.es)
plot(cca.mu)
```

The CCA identifies the two patterns in the data that have the highest correlation - this pattern may not necessarily be the ones with the highest variance.

## Downscaling analysis

The downscaling emplies a multiple regression applied to each of the principal components for the EOFs of $\mu$ and $f_w$.

```{r DS}
## The wet-day frequency and SLP
ds.fw <- DS(subset(eof.fw,pattern=1:4),eof.slp,eofs=1:5)
plot(ds.fw)
ds.mu <- DS(subset(eof.mu,pattern=1:4),eof.es,eofs=1:7)
plot(ds.mu)
```

The downscaling includes cross-validation analysis, and the results shown are for the leading EOF of the predictands. We do not expect that the downscaling will capture a large fraction of the wet-day mean precipitation; local processes are also likely influencing.

Extract the residual and examine to see if there is any temporal-stpatial structure left: a tell-tale sign for noise is a flat eigenspectrum

```{r residual for mu}
## The residual of the wet-day precipitation
z.mu <- as.field(ds.mu)
#res.mu <- as.residual(z.mu)
res.mu <- mu - z.mu
res.mu <- attrcp(mu,res.mu)
class(res.mu) <- class(mu)
eof.res.mu <- EOF(res.mu)
plot(eof.res.mu)
```

For comparison the original data look like this:
```{r EOF for mu}
## The residual of the wet-day precipitation
plot(EOF(mu))
```

```{r residual for fw}
## Repeat with the wet-day frequency for the benefit of comparison
z.fw <- as.field(ds.fw)
#res.fw <- as.residual(ds.fw)
## Another way to estimate the residuals:
res.fw <- fw - z.fw
res.fw <- attrcp(fw,res.fw)
class(res.fw) <- class(fw)
eof.res.fw <- EOF(res.fw)
plot(eof.res.fw)
```

For comparison the original data look like this:
```{r EOF for fw}
## The residual of the wet-day precipitation
plot(EOF(fw))
```

Hence, there is remaining spatio-temporal structure left in the residuals from the downscaling. The challenge is to identify further factors on which both the wet-day mean precipitation and the wet-day frequency depend. It's surprising to see such strong modes in the residual of the wet-day frequency still.

### 2016-04-19 EU-Circle/eu-circle_torbay-precip.Rmd
EU-Cricle Trobay results

@@
 
-The predictors can be organised as EOFs to simplify the analysis and make use of properties as the products being orthogonal:
+The predictors can be organised as EOFs to simplify the analysis and make use of properties as the products being orthogonal. Here are the EOFS for the annual wet-day mean precipitation $\mu$
 
-```{r EOFs-in-precip}
+```{r EOFs-in-precip-mu}
 ## The area mean wet-day mean precipitation
@@
 plot(eof.mu)
+
+```
+
+and the annual wet-day frequency $f_w$
+
+```{r EOFs-in-precip-fw}
 ## Show maps of trends for the wet-day mean precipitation

### 2016-04-21 EU-Circle/eu-circle_case-study-3-midas.Rmd
new precipdata and combined Rmd script

**EU-Circle case study 3
-----------------------------------
title: Coastal Flooding (surface water, highway, sewer and watercourse flooding) across Torbay, UK
author: "Rasmus Benestad"
date: "April 6, 2016"
output: pdf_document
    fig_width:  8
    fig_height: 8
---

Torbay is located in South Devon (UK) and covers an area of approximately 62 km2 . The area has suffered from flooding over many years from a number of different sources, including surface water run-off, highway flooding, sewer flooding, main river and ordinary watercourse flooding during intense rainfall events. In addition the coastal areas of Torbay  suffer coastal flooding due to overtopping of the sea defences during high tides that coincide with easterly winds. It should be noted that the surface water, highway, sewer, main river and watercourse flooding is exacerbated in the low lying areas around the coast of Torquay, Paignton and Brixham during high tidal cycles when the capacity of the surface water outfalls discharging to coastal waters is impeded. In addition to the property flooding, during all of these flood events numerous roads have flooded to some extent, with some of the roads having to be closed to traffic until the flood water has subsided.

The main coast road linking Torquay to Paignton and Brixham has to be closed on a regular basis due to overtopping of the sea wall during high tides that coincide with easterly wind conditions. These closures result in long traffic diversions and delays. In addition, as a result of surface water flooding from watercourses and main rivers during intense rainfall events the main inland route linking Torquay, Paignton and Brixham has also been closed for periods resulting in no major roads being available between the towns until flooding has subsided. During the last major flooding event, which occurred on 24 th October 1999, over 200 properties across all three towns were affected by flooding. Approximately 50% of these properties were commercial properties including shops, restaurants, hotels, bars and a cinema. In addition to the property flooding both the major roads linking Torquay to Paignton and Brixham were closed for a significant period making travel within the Bay extremely difficult and affecting emergency response.

Obviously climate change can affect local flood risk in several ways. Impacts will depend on local conditions and vulnerability. Wetter winters and more of this rain falling in wet spells will increase river and watercourse flooding. More intense rainfall causes more surface run-off, increasing localised flooding and erosion. In turn this will increase pressure on drains, sewers and water quality. Rising sea levels, as a result of climate change, will increase local flood risk both in coastal regions from increased risk of overtopping of the sea wall and inland from main rivers and watercourses due to the interaction with drains, sewers and smaller watercourses. As sea level is predicted to rise by over 1m in Torbay over the next 100 years the frequency and impact of overtopping of the sea defences will increase resulting in more infrastructure and properties being affected by flooding.

**Analysis** 
The downscaling of the wet-day mean precipitation used the surface temperature as predictor.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. To execute this type:

```{r,eval=FALSE}
library(rmarkdown); render('eu-circle_case-study-3.Rmd',pdf_document())
```

Load the esd and ncdf packages.

```{r,echo=TRUE}  
## Load the necessary R-package
library(esd)
## Information about the system and session
sessionInfo()
```

---------------------------------
```{r,eval=FALSE}
## Get MIDAS preciptation data that has laready been retrieved 
## from the MIDAS data base (UK Met Office)

```

Function to estimate probabilty for daily amount of exceeding a threshold x0 given that the distribution follows an exponential distribution and the records of the wet-day mean precipitation and frequency.

```{r}
PrexpPr <- function(mu,fw,x0=10) {
  Pr <- zoo(coredata(fw)*exp(-x0/coredata(mu)),order.by=year(fw))

  Pr.mu <- zoo(mean(coredata(fw))*exp(-x0/coredata(mu)),order.by=year(fw))
  Pr.fw <- zoo(coredata(fw)*exp(-x0/mean(coredata(mu))),order.by=year(fw))
  attr(Pr,'prob.f(mu)') <- Pr.mu
  attr(Pr,'prob.f(fw)') <- Pr.fw
  Pr <- attrcp(mu,Pr)
  attr(Pr,'variable') <- 'Pr'
  attr(Pr,'unit') <- 'probability'
  attr(Pr,'longname') <- paste('Probability of exceeding',x0)
  class(Pr) <- class(mu)
  invisible(Pr)
}
```

Load the data and weed out stations with lots of missing values

```{r}
load('pr.eu-circle-torbay.rda') 
pr <- subset(pr,it=c(1960,2013))

## First weeding of stations with many missing values
nok <- apply(coredata(pr),2,'nv')
x <- subset(pr,is=(nok> 15000))

## Estimate the annual wet-day mean precipitaiton
mu <- subset(annual(x,'wetmean',nmin=100),it=c(1961,2011))
## Estimate the annual wet-day mean frequency
fw <- subset(annual(x,'wetfreq',nmin=100),it=c(1961,2011))

## Second weeding of stations with many missing values in annual 
## aggregate
nok <- apply(coredata(mu),2,'nv')
mu <- subset(mu,is=(nok>= 44))
fw<- subset(fw,is=(nok>= 44))

## Check the records
diagnose(mu)

##Fill in the remaining gaps using a PCA-based interpolation scheme 
pcafill(mu) -> mu
pcafill(fw) -> fw
```

Examine the long-term trends in the wet-day mean precipitation

```{r}
plot(mu,new=FALSE); grid()
for (i in 1:dim(mu)[2]) lines(trend(mu[,i]))
```

```{r}
plot(fw); grid()
for (i in 1:dim(mu)[2]) lines(trend(mu[,i]))
```

```{r}
Pr.gt.10 <- PrexpPr(mu,fw,x0=10)
plot(Pr.gt.10); grid()
for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.10[,i]))
```

## Useful information in the mean seasonal cycle?

another attempt to search for systematic influences on the precipitation statistics ($\mu$) is to examine the mean seasonal cycle. This type of approach may

```{r}
## Estimate the mean seasonal cycle in the wet-day mean precipitaiton
MU <- aggregate(x,month,FUN='wetmean')
## Estimate the mean seasonal cycle in wet-day mean frequency
FW <- aggregate(x,month,FUN='wetfreq')
## Estimate the mean seasonal cycle in wet-day mean frequency
MP <- aggregate(x,month,FUN='mean')
```

```{r}
## Show the mean seasonal cycle in the wet-day mean precipitation
plot(MU)
```

There is some indication of a mean seasonal cycle in $\mu$, with minimum values in late winter spring and higher values in late summer and autumn.

```{r}
## Show the mean seasonal cycle in the wet-day frequency
plot(FW)
```

There is a clear seasonal cycle in the wet-day occurrence - more rainy days in the uatumn.

```{r}
## Show the mean seasonal cycle in the total precipitation
MP <- 30*MP
attr(MP,'unit') <- 'mm/month'
plot(MP)
```

Most of the rain falls during winter.

```{r}
## Estimate the PCA for the annual cycle to identify the general features:
pca.MU <- PCA(MU,n=10)
plot(pca.MU)
```

Most stations have the same features when it comes to a mean seasonal cycle in $\mu$, with the leading PCA accounting for 80\% of the variance.

## Downscaling based on mean seasonal variations

```{r}
if (!file.exists('dse.mu.eu-circle-worstcase.rda')) {
  dse.mu <- DSensemble.mu.worstcase(x,biascorrect=TRUE,mask=TRUE,
                                    lon=c(-80,20),lat=c(10,60),rel.cord=FALSE)
  save(file='dse.mu.eu-circle-worstcase.rda',dse.mu)
} else load('dse.mu.eu-circle-worstcase.rda')
plot(dse.mu[[1]])
```

## Sea-level

Analyse the sea-level trends and connection to global temperatures to assess future outcome. First the stations from the GLOSS dataset (global monthly data):

```{r}
gloss <- aggregate(station.gloss(),year,FUN='max',na.rm=TRUE)
gloss <- subset(gloss,is=list(lon=c(-20,20),lat=c(45,55)))
plot(gloss)
```

Single station with hourly data from Newlyn:

```{r}
newlyn <- station.newlyn()
newlyn <- aggregate(newlyn,year,FUN='max',na.rm=TRUE)
plot(newlyn)
```

French tidal stations along the northern/eastern coast (SONEL):

```{r}
sonel <- aggregate(station.sonel(),year,FUN='max',na.rm=TRUE)
plot(sonel)
```

Combine the records from different sources

```{r}
index(newlyn) <- year(newlyn)
index(sonel) <- year(sonal)
index(gloss) <- year(gloss)
Z <- combine.stations(newlyn,sonel,gloss)
nv <- apply(coredata(Z),2,FUN='nv')
Z <- subset(Z,is=(nv > 50))
Z <- subset(Z,it=c(1950,2015))
plot(Z); grid()
```

Check for data availability

```{r}
diagnose(Z)
```

Make a map of the trend in the local sea-level - mm/decade

```{r trend-map}
zc <- coredata(Z)
zc[is.infinite(zc)] <- NA
zc -> coredata(Z)
map(Z,FUN='trend',cex=-2)
```

## Wind

Estimate wind windspeed from SLP measurements. Make use of the gridded EOBS data as the documentation for MIDAS data is poor and it's difficult to read the data and organise it, due to very large ASCII files and the way the data is stored.

```{r trend-map}
slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg_v12.0.nc',lon=c(-6,0),lat=c(48,52))
slp.eobs <- 10000*slp.eobs ## The data have some funny scaling factor
attr(slp.eobs,'unit') <- 'Pa'
map(slp.eobs,FUN='trend')
```

Pick three grid points from which we can estiamte the geostropic wind

```{r trend-map}
triangle <- subset(Z,is=c(2,3,7))
## The stations are not situated quite where there is valid slp data in the EOBS dataset
## It's only their coordinates that is used for interpolating SLP for three stations
## Using station objects as is-object makes the results a station object
attr(triangle,'longitude') <- c(-4,-2,-1.25)
attr(triangle,'latitude') <- c(50.50,51,49)
pp.eobs <- regrid(slp.eobs,is=triangle)
## Use triagulation for stations to estimate the wind
uv <- geostrophicwind(pp.eobs)
plot(annual(uv,nmin=300),main='zonal and meridional wind components',
     sub='Estimated from triangulation of mean sea-level pressure from EOBS')
```

The estimated annual mean geostrophic wind reveals that something changed between 2006 and 2010. The change seems to have taken place over 4 years.

```{r trend-event-2006}
plot(annual(uv,nmin=300),main='zonal and meridional wind components',
     xlim=c(2000,2010))
```

This event also is seen in the upper tail of the wind speed distribution. There are four features concerning these wind speed estimates: a downward trend from moderate high wind speeds in the 1950s, two years with a large drop in values in the 1970s, a number of years with high values around year 2010, and a slight increasing trend in the upper wind speeds over the record.

```{r wind-speed-trend}
## Trend in upper wind speed
ff <- sqrt(uv[,1]^2 + uv[,2]^2)
plot(annual(ff,"quantile",probs=0.98),lwd=2,
     main='Wind speed: annual 98-percentile (m/s)')
lines(trend(annual(ff,"quantile",probs=0.98)),lty=2)
grid()
trend.coef(annual(ff,"quantile",probs=0.98))
trend.err(annual(ff,"quantile",probs=0.98))
```

The trend is not siginificant at the 5-percentage level.

```{r wind-direction-trend}
## Trend in upper wind direction
dd <- zoo(180/pi*atan2(coredata(uv)[,1],coredata(uv)[,2]),order.by=index(uv))
plot(annual(dd,FUN='mean'),lwd=2,ylim=c(-180,180),
     main='Annual mean wind direction (degrees)')
axis(4,at=c(-180,-90,0,90,180),labels=c('S','E','N','W','S'))
for (j in c(-180,-90,0,90,180)) lines(range(index(dd)),rep(j,2),lty=3)
```

The change in 2006-2010 is also visible in the wind direction. This could be an event caused by a change in circulation, but it cluld also be due to changes in instrumentation, data used in the gridding or observational practices. It is possible to examine whether there were changes in the criculation by using ESD to compare the windspeeds with SLP from reanalyses.

```{r zonal-wind-check-get-slp}
## Trend in upper wind direction
SLP <- retrieve('slp.mon.mean.nc',lon=c(-50,30),lat=c(40,60))
eof.slp<- EOF(annual(SLP))
```

Apply the downscaling to the meridional wind component:

```{r zonal-wind-check1}
## Trend in upper wind direction
uam <- annual(subset(uv,is=2),nmin=300)
ds.u1 <- DS(uam,eof.slp,eofs=1:20)
plot(ds.u1)
```

The results suggest a mismatch between the reanalysis and the EOBS data, and the change after 2005 is not corroberated by other sources. We can therefore say that this feature is suspect.

```{r zonal-wind-check2}
## Trend in upper wind direction
ds.u2 <- DS(subset(uam,it=c(1950,2005)),eof.slp,eofs=1:20)
plot(ds.u2)
```

When the test is repeated for the shorter period excluding data after 2005, there is a reasonably good match between the EOBS data and the reanalysis. The trend analysis can be repeated with this event excluded

```{r wind-speed-trend-2}
## Trend in upper wind speed
plot(subset(annual(ff,"quantile",probs=0.98),it=c(1950,2005)),lwd=2,
     main='Wind speed: annual 98-percentile (m/s)')
lines(trend(subset(annual(ff,"quantile",probs=0.98),it=c(1950,2005))),lty=2)
grid()
```

The exclusion of the suspicious event also affects the long-term trend estiamte - now it's zero with a slight reduction.

```{r wind-speed-distribution}
## 
require(MASS)
hist(coredata(subset(ff,it=c(1950,2005))),lwd=2,breaks=seq(0,70,by=1),
     main='Wind speed  (m/s)',freq=FALSE,col=colscal(71))
grid()
## Maximum-likelihood fitting of the Weibull distribution - two parameters: shape and scale
f<-fitdistr(ff[is.finite(ff)], 'weibull')
lines(dweibull(seq(0,70,by=1),shape=f$estimate[1],scale=f$estimate[2]),lwd=3,col=rgb(1,0,0,0.3))
```

```{r reanalysis-wind-speed}
## 
UV <- geostrophicwind(SLP)
```

```{r wind-rose}
## wind rose
windrose(uv)
```

### 2016-04-21 EU-Circle/eu-circle_case-study-3.Rmd
new updates - Climatrans
new precipdata and combined Rmd script

@@
 ## Information about the system and session
-Sys.info()
 sessionInfo()
@@
 load('pr.eu-circle-torbay.rda') 
+pr <- subset(pr,it=c(1960,2013))
 
@@
 ## Estimate the annual wet-day mean precipitaiton
-mu <- subset(annual(x,'wetmean',nmin=180),it=c(1961,2011))
+mu <- subset(annual(x,'wetmean',nmin=100),it=c(1961,2011))
 ## Estimate the annual wet-day mean frequency
-fw <- subset(annual(x,'wetfreq',nmin=180),it=c(1961,2011))
+fw <- subset(annual(x,'wetfreq',nmin=100),it=c(1961,2011))
 
@@
 nok <- apply(coredata(mu),2,'nv')
-mu <- subset(mu,is=(nok> 45))
-fw<- subset(fw,is=(nok> 45))
+mu <- subset(mu,is=(nok>= 44))
+fw<- subset(fw,is=(nok>= 44))
 
@@
 ```
+
+## Useful information in the mean seasonal cycle?
+
+another attempt to search for systematic influences on the precipitation statistics ($\mu$) is to examine the mean seasonal cycle. This type of approach may
+
+```{r}
+## Estimate the mean seasonal cycle in the wet-day mean precipitaiton
+MU <- aggregate(x,month,FUN='wetmean')
+## Estimate the mean seasonal cycle in wet-day mean frequency
+FW <- aggregate(x,month,FUN='wetfreq')
+## Estimate the mean seasonal cycle in wet-day mean frequency
+MP <- aggregate(x,month,FUN='mean')
+```
+
+```{r}
+## Show the mean seasonal cycle in the wet-day mean precipitation
+plot(MU)
+```
+
+There is some indication of a mean seasonal cycle in $\mu$, with minimum values in late winter spring and higher values in late summer and autumn.
+
+```{r}
+## Show the mean seasonal cycle in the wet-day frequency
+plot(FW)
+```
+
+There is a clear seasonal cycle in the wet-day occurrence - more rainy days in the uatumn.
+
+```{r}
+## Show the mean seasonal cycle in the total precipitation
+MP <- 30*MP
+attr(MP,'unit') <- 'mm/month'
+plot(MP)
+```
+
+Most of the rain falls during winter.
+
+```{r}
+## Estimate the PCA for the annual cycle to identify the general features:
+pca.MU <- PCA(MU,n=10)
+plot(pca.MU)
+```
+
+Most stations have the same features when it comes to a mean seasonal cycle in $\mu$, with the leading PCA accounting for 80\% of the variance.
+
+## Downscaling based on mean seasonal variations
+
+```{r}
+if (!file.exists('dse.mu.eu-circle-worstcase.rda')) {
+  dse.mu <- DSensemble.mu.worstcase(x,biascorrect=TRUE,mask=TRUE,
+                                    lon=c(-80,20),lat=c(10,60),rel.cord=FALSE)
+  save(file='dse.mu.eu-circle-worstcase.rda',dse.mu)
+} else load('dse.mu.eu-circle-worstcase.rda')
+plot(dse.mu[[1]])
+```
+
+## Sea-level
+
+Analyse the sea-level trends and connection to global temperatures to assess future outcome. First the stations from the GLOSS dataset (global monthly data):
+
+```{r}
+gloss <- aggregate(station.gloss(),year,FUN='max',na.rm=TRUE)
+gloss <- subset(gloss,is=list(lon=c(-20,20),lat=c(45,55)))
+plot(gloss)
+```
+
+Single station with hourly data from Newlyn:
+
+```{r}
+newlyn <- station.newlyn()
+newlyn <- aggregate(newlyn,year,FUN='max',na.rm=TRUE)
+plot(newlyn)
+```
+
+French tidal stations along the northern/eastern coast (SONEL):
+
+```{r}
+sonel <- aggregate(station.sonel(),year,FUN='max',na.rm=TRUE)
+plot(sonel)
+```
+
+Combine the records from different sources
+
+```{r}
+index(newlyn) <- year(newlyn)
+index(sonel) <- year(sonal)
+index(gloss) <- year(gloss)
+Z <- combine.stations(newlyn,sonel,gloss)
+nv <- apply(coredata(Z),2,FUN='nv')
+Z <- subset(Z,is=(nv > 50))
+Z <- subset(Z,it=c(1950,2015))
+plot(Z); grid()
+```
+
+Check for data availability
+
+```{r}
+diagnose(Z)
+```
+
+Make a map of the trend in the local sea-level - mm/decade
+
+```{r sl-trend-map}
+zc <- coredata(Z)
+zc[is.infinite(zc)] <- NA
+zc -> coredata(Z)
+map(Z,FUN='trend',cex=-2)
+```
+
+## Wind
+
+Estimate wind windspeed from SLP measurements.
+Make use of the gridded EOBS data as the documentation for MIDAS data is poor and it's difficult to read the data and organise it, due to very large ASCII files and the way the data is stored.
+
+```{r slp-trend-map}
+slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg_v12.0.nc',lon=c(-6,0),lat=c(48,52))
+slp.eobs <- 10000*slp.eobs ## The data have some funny scaling factor
+attr(slp.eobs,'unit') <- 'Pa'
+map(slp.eobs,FUN='trend')
+```
+
+Pick three grid points from which we can estiamte the geostropic wind
+
+```{r trend-map}
+triangle <- subset(Z,is=c(2,3,7))
+## The stations are not situated quite where there is valid slp data in the EOBS dataset
+## It's only their coordinates that is used for interpolating SLP for three stations
+## Using station objects as is-object makes the results a station object
+attr(triangle,'longitude') <- c(-4,-2,-1.25)
+attr(triangle,'latitude') <- c(50.50,51,49)
+pp.eobs <- regrid(slp.eobs,is=triangle)
+## Use triagulation for stations to estimate the wind
+uv <- geostrophicwind(pp.eobs)
+plot(annual(uv,nmin=300),main='zonal and meridional wind components',
+     sub='Estimated from triangulation of mean sea-level pressure from EOBS')
+```
+
+The estimated annual mean geostrophic wind reveals that something changed between 2006 and 2010. The change seems to have taken place over 4 years.
+
+```{r trend-event-2006}
+plot(annual(uv,nmin=300),main='zonal and meridional wind components',
+     xlim=c(2000,2010))
+```
+
+This event also is seen in the upper tail of the wind speed distribution.
+There are four features concerning these wind speed estimates: a downward trend from moderate high wind speeds in the 1950s, two years with a large drop in values in the 1970s, a number of years with high values around year 2010, and a slight increasing trend in the upper wind speeds over the record.
+
+```{r wind-speed-trend}
+## Trend in upper wind speed
+ff <- sqrt(uv[,1]^2 + uv[,2]^2)
+plot(annual(ff,"quantile",probs=0.98),lwd=2,
+     main='Wind speed: annual 98-percentile (m/s)')
+lines(trend(annual(ff,"quantile",probs=0.98)),lty=2)
+grid()
+trend.coef(annual(ff,"quantile",probs=0.98))
+trend.err(annual(ff,"quantile",probs=0.98))
+```
+
+The trend is not siginificant at the 5-percentage level.
+
+```{r wind-direction-trend}
+## Trend in upper wind direction
+dd <- zoo(180/pi*atan2(coredata(uv)[,1],coredata(uv)[,2]),order.by=index(uv))
+plot(annual(dd,FUN='mean'),lwd=2,ylim=c(-180,180),
+     main='Annual mean wind direction (degrees)')
+axis(4,at=c(-180,-90,0,90,180),labels=c('S','E','N','W','S'))
+for (j in c(-180,-90,0,90,180)) lines(range(index(dd)),rep(j,2),lty=3)
+```
+
+The change in 2006-2010 is also visible in the wind direction.
+This could be an event caused by a change in circulation, but it cluld also be due to changes in instrumentation, data used in the gridding or observational practices.
+It is possible to examine whether there were changes in the criculation by using ESD to compare the windspeeds with SLP from reanalyses.
+
+```{r zonal-wind-check-get-slp}
+## Trend in upper wind direction
+SLP <- retrieve('slp.mon.mean.nc',lon=c(-50,30),lat=c(40,60))
+eof.slp<- EOF(annual(SLP))
+```
+
+Apply the downscaling to the meridional wind component:
+
+```{r zonal-wind-check1}
+## Trend in upper wind direction
+uam <- annual(subset(uv,is=2),nmin=300)
+ds.u1 <- DS(uam,eof.slp,eofs=1:20)
+plot(ds.u1)
+```
+
+The results suggest a mismatch between the reanalysis and the EOBS data, and the change after 2005 is not corroberated by other sources. We can therefore say that this feature is suspect.
+
+```{r zonal-wind-check2}
+## Trend in upper wind direction
+ds.u2 <- DS(subset(uam,it=c(1950,2005)),eof.slp,eofs=1:20)
+plot(ds.u2)
+```
+
+When the test is repeated for the shorter period excluding data after 2005, there is a reasonably good match between the EOBS data and the reanalysis.
+The trend analysis can be repeated with this event excluded
+
+```{r wind-speed-trend-2}
+## Trend in upper wind speed
+plot(subset(annual(ff,"quantile",probs=0.98),it=c(1950,2005)),lwd=2,
+     main='Wind speed: annual 98-percentile (m/s)')
+lines(trend(subset(annual(ff,"quantile",probs=0.98),it=c(1950,2005))),lty=2)
+grid()
+```
+
+The exclusion of the suspicious event also affects the long-term trend estiamte - now it's zero with a slight reduction.
+
+```{r wind-speed-distribution}
+## 
+require(MASS)
+hist(coredata(subset(ff,it=c(1950,2005))),lwd=2,breaks=seq(0,70,by=1),
+     main='Wind speed  (m/s)',freq=FALSE,col=colscal(71))
+grid()
+## Maximum-likelihood fitting of the Weibull distribution - two parameters: shape and scale
+f<-fitdistr(ff[is.finite(ff)], 'weibull')
+lines(dweibull(seq(0,70,by=1),shape=f$estimate[1],scale=f$estimate[2]),lwd=3,col=rgb(1,0,0,0.3))
+```
+
+```{r reanalysis-wind-speed}
+## 
+UV <- geostrophicwind(SLP)
+```
+
+```{r wind-rose}
+## wind rose
+windrose(uv)
+```
+
+## EOBS gridded data
+
+This script retreives precipitation for Southwestern England from the EOBS data set and estimates the wet-day mean precipitation and frequency.
+
+The EOBS data was chosen here because the rain gauge data from stations (MIDAS) was intermittent and not readily available for an analysis that makes use of PCA in order to make the most use of the information embedded in many parallel data records.
+
+Rain gauges provide a very tiny sample of the rainfall, and previous work suggests that downscaling can provide a better picture if it's applied to a group of stations rather than single stations (Benestad et al., 2015; doi:10.342/tellusa.v67.28326).
+Downscaling applied to products from principal component analysis (PCA) benefits from an emphasis on cohernt structure of variability in time and space, which can improve the signal-to-noise ratio.
+PCA requires no missing values, but the EOBS gridded data provides complete records.
+
+```{r predictand}
+pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(-6,-1),lat=c(50,52))
+mu <- annual(pr.eobs,FUN='wetmean')
+fw <- annual(pr.eobs,FUN='wetfreq')
+```
+
+## Past trends
+
+Climate change has taken place for some time, and past trends may provide some indication about whether a chosen climate parameter such as the wet-day mean precipitation $\mu$ or wet-day frequency $f_w$ are sensitive to a global warming:
+
+```{r trend}
+## The area mean wet-day mean precipitation
+plot(mu,new=FALSE)
+## The area mean wet-day frequency
+plot(fw,new=FALSE)
+## The early part of the record looks suspect:
+mu <- subset(mu,it=c(1958,2015))
+fw <- subset(fw,it=c(1958,2015))
+## Show maps of trends for the wet-day mean precipitation
+map(mu,FUN='trend',
+    colbar=list(breaks=seq(-0.12,0.12,by=0.02),pal='t2m',rev=TRUE))
+## Map of trends in wet-day frequency:
+map(fw,FUN='trend',
+    colbar=list(breaks=seq(-0.01,0.01,by=0.001),pal='t2m',rev=TRUE))
+```
+
+The trend analysis for the wet-day frequency suggest a general increase in the number of wet-days over the  perdiod 1960--2015 while the intencity .
+
+The predictors can be organised as EOFs to simplify the analysis and make use of properties as the products being orthogonal. Here are the EOFS for the annual wet-day mean precipitation $\mu$
+
+```{r EOFs-in-precip-mu}
+## The area mean wet-day mean precipitation
+eof.mu <- EOF(mu)
+## The area mean wet-day frequency
+plot(eof.mu)
+
+```
+
+and the annual wet-day frequency $f_w$
+
+```{r EOFs-in-precip-fw}
+## Show maps of trends for the wet-day mean precipitation
+eof.fw <- EOF(fw)
+## Map of trends in wet-day frequency:
+plot(eof.fw)
+```
+
+The diagnostics of the EOFs suggest that the leading mode explains a high proportion of the variability on an annual time scale.
+
+The mean sea-level pressure (SLP) is used as predictor for the annual wet-day frequency, as this gives a good description of the circulation.
+A number of different spatial domains were tested, and for SLP and $f_w$, the results were fairly robust with respect to the choice of domain.
+
+The predictor for the wet-day mean precipitation was taken to be the saturation vapour pressure over the North Atlantic, which is regarded as the main source for the atmospheric moisture over northern Europe.
+
+## Predictor
+
+```{r predictor}
+## The mean sea-level pressure
+slp <- retrieve('slp.mon.mean.nc',lon=c(-80,0),lat=c(20,70))
+## The surface temperature
+t2m <- retrieve('air.mon.mean.nc',lon=c(-80,0),lat=c(10,70))
+## Saturation water pressure estimated from maritime temperatures
+es <- C.C.eq(mask(t2m,land=TRUE))
+eof.slp<- EOF(annual(slp))
+eof.t2m <- EOF(annual(t2m))
+eof.es <- EOF(annual(es,FUN='mean'))
+```
+
+## Cannical Correlation analysis
+
+Carry out canonical correlation analysis (CCA) to explore relationships between the large-scale conditions such as SLP/temperature and the precipitation over soutwestern England:
+
+```{r cca}
+## The wet-day frequency and SLP
+cca.fw <- CCA(eof.fw,eof.slp)
+plot(cca.fw)
+cca.mu <- CCA(eof.mu,eof.es)
+plot(cca.mu)
+```
+
+The CCA identifies the two patterns in the data that have the highest correlation - this pattern may not necessarily be the ones with the highest variance.
+
+## Downscaling analysis
+
+The downscaling emplies a multiple regression applied to each of the principal components for the EOFs of $\mu$ and $f_w$.
+
+```{r DS}
+## The wet-day frequency and SLP
+ds.fw <- DS(subset(eof.fw,pattern=1:4),eof.slp,eofs=1:5)
+plot(ds.fw)
+ds.mu <- DS(subset(eof.mu,pattern=1:4),eof.es,eofs=1:7)
+plot(ds.mu)
+```
+
+The downscaling includes cross-validation analysis, and the results shown are for the leading EOF of the predictands.
+We do not expect that the downscaling will capture a large fraction of the wet-day mean precipitation; local processes are also likely influencing.
+
+Extract the residual and examine to see if there is any temporal-stpatial structure left: a tell-tale sign for noise is a flat eigenspectrum
+
+```{r residual for mu}
+## The residual of the wet-day precipitation
+z.mu <- as.field(ds.mu)
+#res.mu <- as.residual(z.mu)
+res.mu <- mu - z.mu
+res.mu <- attrcp(mu,res.mu)
+class(res.mu) <- class(mu)
+eof.res.mu <- EOF(res.mu)
+plot(eof.res.mu)
+```
+
+For comparison the original data look like this:
+```{r EOF for mu}
+## The residual of the wet-day precipitation
+plot(EOF(mu))
+```
+
+```{r residual for fw}
+## Repeat with the wet-day frequency for the benefit of comparison
+z.fw <- as.field(ds.fw)
+#res.fw <- as.residual(ds.fw)
+## Another way to estimate the residuals:
+res.fw <- fw - z.fw
+res.fw <- attrcp(fw,res.fw)
+class(res.fw) <- class(fw)
+eof.res.fw <- EOF(res.fw)
+plot(eof.res.fw)
+```
+
+For comparison the original data look like this:
+```{r EOF for fw}
+## The residual of the wet-day precipitation
+plot(EOF(fw))
+```
+
+Hence, there is remaining spatio-temporal structure left in the residuals from the downscaling.
+The challenge is to identify further factors on which both the wet-day mean precipitation and the wet-day frequency depend.
+It's surprising to see such strong modes in the residual of the wet-day frequency still.

### 2016-04-28 CORDEX/ensembles.Rmd
ensemble example for CORDEX

---
title: "ensembles"
author: "Rasmus Benestad"
date: "April 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

## **Simulated global mean temperartue**

Combine the simulated global mean temperature with information about the model configurations taken from Table 9-a-1 in the IPCC WG 1 2013 report ("AR5):

```{r}
library('esd')
data("IPCC.AR5.Table.9.A.1")
data("global.t2m.gcm")
gcmnm <- unlist(lapply(attr(global.t2m.gcm$global.t2m.cmip5.rcp45,'meta'),function(x) x$GCM))
n <- length(gcmnm); m <- 12
X <- matrix(rep(NA,n*m),n,m)

IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name <- as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name)
IPCC.AR5.Table.9.A.1$Ocean.Component.Name <- as.character(IPCC.AR5.Table.9.A.1$Ocean.Component.Name)
```

Index some of the entries which do not contain useful information.

```{r}
iam <- is.element(substr(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name,1,8),'Included') |
       is.element(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name,'<NA>')
IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name[iam] <- IPCC.AR5.Table.9.A.1$Model.Name[iam]
iom <- is.element(IPCC.AR5.Table.9.A.1$Ocean.Component.Name,'Included')
IPCC.AR5.Table.9.A.1$Ocean.Component.Name[iom] <- IPCC.AR5.Table.9.A.1$Model.Name[iom]
```

There are many different similar settings with different names. Try to organise and sort the different types under fewer and simpler categories. Here is a mix of conventions and ways to refer to the resolution.

```{r}
fixagcmnm <- grep('Included',as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name))
IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name[fixagcmnm] <- IPCC.AR5.Table.9.A.1$Model.Name[fixagcmnm]
```

```{r}
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- as.character(IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub(' ','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('-','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub(',','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('longitude','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('latitude','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('Nominally','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('equivalentto','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('FiniteVolume','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('Averagedcellsize','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('in','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('and','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('TL','T',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
Tres<- c('T42','T31','T159','T63','T159','T959','T213','T319','N48','N49','C180','C360','R42','N96')
for (ires in Tres) {
  ii <- grep(ires,IPCC.AR5.Table.9.A.1$Horizontal.Grid)
  IPCC.AR5.Table.9.A.1$Horizontal.Grid[ii] <- ires
}
IPCC.AR5.Table.9.A.1$Horizontal.Grid <- as.factor(IPCC.AR5.Table.9.A.1$Horizontal.Grid)
```

Fix the GCM names:

```{r}
IPCC.AR5.Table.9.A.1$Model.Name <- as.character(IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name <- gsub('-','.',IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name <- gsub('_','.',IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name <- gsub(' v1.0','',IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name <- gsub('(m)','.m',IPCC.AR5.Table.9.A.1$Model.Name,perl=TRUE,fixed=TRUE)
cesm1 <- grep('CESM1',IPCC.AR5.Table.9.A.1$Model.Name)
IPCC.AR5.Table.9.A.1$Model.Name[cesm1] <- 'CESM1'
IPCC.AR5.Table.9.A.1$Model.Name <- as.factor(IPCC.AR5.Table.9.A.1$Model.Name)
gcmnm <- gsub('-','.',gcmnm)
gcmnm <- gsub('_','.',gcmnm)
cesm1 <- grep('CESM1',gcmnm)
gcmnm[cesm1] <- 'CESM1'
cnrm5 <- grep('CNRM.CM5',gcmnm)
gcmnm[cnrm5] <- 'CNRM.CM51'
inmcm4 <- grep('inmcm4',gcmnm)
gcmnm[inmcm4] <- 'INM.CM4'
```

Also need to tidy up some of the description concerning the ocean models

```{r}
IPCC.AR5.Table.9.A.1$z.Co.ord <- as.character(IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('-coordinate','',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('hybrid','',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('sigma-z','sigma',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub(' ','',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('Isopycnic','isopycnic',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('zisopycnic','isopycnic',IPCC.AR5.Table.9.A.1$z.Co.ord)
IPCC.AR5.Table.9.A.1$z.Co.ord <- as.factor(IPCC.AR5.Table.9.A.1$z.Co.ord)
```

Make a data.frame that organises the model information for each simulation.

```{r, echo=FALSE}
for (i in 1:n) {
  ii <- is.element(toupper(IPCC.AR5.Table.9.A.1$Model.Name),toupper(gcmnm[i]))
  if (sum(ii)>0) {
    x <- c(as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Horizontal.Grid)[ii][1],
               IPCC.AR5.Table.9.A.1$Number.of.Vert.Levels[ii][1],
               IPCC.AR5.Table.9.A.1$Grid.Top.hPa[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Component.Name.or.type)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Atmos.Chemistry.Component.Name)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Land.Surface.Component.Name)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Ocean.Component.Name)[ii][1],
               IPCC.AR5.Table.9.A.1$Number.of.Vertical.Levels[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$z.Co.ord)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Ocean.Biogeochemistry.Component.Name)[ii][1],
               as.character(IPCC.AR5.Table.9.A.1$Sea.Ice.Component.Name)[ii][1])
    print(x)
    X[i,] <- x
  }  else print(paste('no match for ',gcmnm[i]))
}
```

Set the column names

```{r}
X <- as.data.frame(X)
colnames(X) <- c('atm.model','atm.grid',
                 'atm.lev','atm.top',
                 'atm.type','Atm.chem',
                 'Land','oce.model',
                 'oce.lev','oce.z',
                 'oce.bio','Seaice')
X$atm.lev <- as.integer(X$atm.lev)
X$atm.top <- as.numeric(X$atm.top)
X$oce.lev <- as.integer(X$oce.lev)
```

Combine the simulated temperature change with corresponding information about the model configuration

```{r}
dT <- colMeans(window(zoo(global.t2m.gcm[[1]]),start=2070,end=2099))-
      colMeans(window(zoo(global.t2m.gcm[[1]]),start=1970,end=1999))
X <- as.data.frame(cbind(round(dT,2),X))
```

##**Results**

Is the simulated temperature change sensitive to how the atmospheric model is configured?

```{r}
fit.atm <- lm(dT ~ atm.grid + atm.lev + atm.top + oce.lev + oce.z,data=X)
print(summary(fit.atm))
```

### 2016-04-29 Tanzania/tmintest.Rmd
Minimum temperatures in Tanzania

---
title: "Habiba"
author: "Rasmus Benestad"
date: "April 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

Load the esd-package and the data:

```{r}
library(esd)
load('tmininTz.rda')
load('predictors.rda')
data(ferder)
```

This is what we expect to see with good correspondence between predictand and predictor

```{r}
djf.e <- subset(as.4seasons(ferder),it='Jan')
corfield(djf.e,DJF.E,colbar=list(rev=FALSE),new=FALSE)
```

Now compare the station data with the corresponding interpolated data from the reanalysis

```{r}
dje.ei <- regrid(DJF.E,is=djf.e)
plot(djf.e,new=FALSE)
lines(dje.ei,lwd=3)
```

##**Tanzania**

Now we have a look at some of the stations in Tanzania: the ones with the poorest and the best score for the season with the lowest score. The first one is for Mwanza December-February mean daily minimum temperature:

```{r}
y <- subset(Y,is='Mwanza')
djf <- subset(as.4seasons(y),it='Jan')
## fix a little bug
attr(djf,'unit') <- 'degC'
corfield(djf,DJF,colbar=list(rev=FALSE),new=FALSE)
```

We can interpolate temperatures for the same coordinate as for Mwanza, but when we do the comparison, we need to use standardised values as the predictor represents the mean of daily mean temperatures $\overline{T}$ whereas the station is the mean daily minimum $T[min]$:

```{r}
djf.i <- regrid(DJF,is=y)
plot(stand(djf),new=FALSE)
lines(stand(djf.i),lwd=3)     
```

Dodoma was the station with the best correspondence between predictor and predictand:

```{r}
y2 <- subset(Y,is='Dodoma')
## There were two entries for Dodoma - exclude the one from GHCN
y2 <- subset(y2,is=1)
djf2 <- subset(as.4seasons(y2),it='Jan')
## fix a little bug
attr(djf2,'unit') <- 'degC'
corfield(djf2,DJF,colbar=list(rev=FALSE),new=FALSE)
```

```{r}
djf.i2 <- regrid(DJF,is=y2)
plot(stand(djf2),new=FALSE)
lines(stand(djf.i2),lwd=3)     
```

### 2016-04-29 CORDEX/ensembles.Rmd

@@
                as.character(IPCC.AR5.Table.9.A.1$Sea.Ice.Component.Name)[ii][1])
-    print(x)
+    #print(x)
     X[i,] <- x

### 2016-04-29 EU-Circle/eu-circle_case-study-3.Rmd

@@
 ## from the MIDAS data base (UK Met Office)
-
+load('pr.eu-circle-torbay.rda')
 ```
@@
 fw<- subset(fw,is=(nok>= 44))
+```
 
+Fill in missing values, but since there are too many missing values, we need to get some more data from the region.
+This is ok since we are working with annual aggregated values, which may imply larer spatial scales.
+In any case, these are used to assist the interpolation for the MIDAS sations.
+
+```{r}
+ ss <- select.station(param='precip',src='ecad',cntr=c('United Kingdom','France'),lat=c(45,55),lon=c(-10,0),nmin=50)
+pr.ecad <- station(ss)
+mu.ecad <- subset(annual(pr.ecad,'wetmean',nmin=100),it=c(1961,2011))
+fw.ecad <- subset(annual(pr.ecad,'wetfreq',nmin=100),it=c(1961,2011))
+## Second weeding of stations with many missing values in annual 
+## aggregate
+nok <- apply(coredata(mu.ecad),2,'nv')
+mu.ecad <- subset(mu.ecad,is=(nok>= 50))
+fw.ecad<- subset(fw.ecad,is=(nok>= 50))
+diagnose(mu.ecad)
+```
+
+supplement these to the MIDAS data:
+
+```{r}
+mu <- combine.stations(mu,mu.ecad)
+fw <- combine.stations(fw,fw.ecad)
+```
+
+```{r}
+##Fill in the remaining gaps using a PCA-based interpolation scheme 
 ## Check the records
 diagnose(mu)
-
-##Fill in the remaining gaps using a PCA-based interpolation scheme 
 pcafill(mu) -> mu
@@
 
-Most stations have the same features when it comes to a mean seasonal cycle in $\mu$, with the leading PCA accounting for 80\% of the variance.
+Most stations have the same features when it comes to a mean seasonal cycle in $\mu$, with the leading PCA accounting for 80% of the variance.
 
 ## Downscaling based on mean seasonal variations
+
+Estimating upper-limits can be an additional approach to exploring the effect of temperature on the precipitation intensity.
+This is a cruder means for quantifying and may provide further information when more traditional methods give limited information.
 
@@
   dse.mu <- DSensemble.mu.worstcase(x,biascorrect=TRUE,mask=TRUE,
-                                    lon=c(-80,20),lat=c(10,60),rel.cord=FALSE)
+                                    lon=c(-80,20),lat=c(10,60),rel.cord=FALSE,verbose=TRUE)
   save(file='dse.mu.eu-circle-worstcase.rda',dse.mu)

### 2016-04-29 CORDEX/ceofs.Rmd
Common EOFs and PCA for stations testing

---
title: "Common EOFs"
author: "Rasmus Benestad"
date: "April 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comparing spatial-temporal structures of variability

Common principal component analysis (PCA) and empirical orthogonal functions (EOFs) can be used to compare and analyse co-variance structures in multi-dimensional data. This is an R Markdown document which provides a demonstration of the concept. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

##*Sample data*

Here we use the R-package 'esd' <https://github.com/metno/esd> and sample data which come with this analysis tool. Note that the data are filtered and only describe large-scale features. Hence, it is ok for demonstration purposes, but true results need to be based on the raw temperatures.

Here we want to compare characteristical patterns of variability in marine tsurface temperatures associated with the winter season.

```{r}
library(esd)
X1 <- mask(t2m.NCEP(lon=c(-60,40),lat=c(40,70)),land=TRUE)
X2 <- t2m.NorESM.M(lon=c(-60,40),lat=c(40,70))
X2 <- subset(X2,it=range(index(X1)))
```

We limit the demonstration to the winter season.

```{r}
X1 <- subset(as.4seasons(X1),it='Jan')
X2 <- subset(as.4seasons(X2),it='Jan')
```

Common EOFs (PCAs) are combuted by constructing a data matrix containing anomalies from different sources, interpolated to a common spatial grid.

```{r}
XX <- combine(X1,X2)
```

The EOFs are then computed as in traditional sense, but differ only by the input, which now is a matrix of combined data. The spatial pattern (EOF) now represents both the reanalysis and the GCM, and it is only in the principal components (PCs) where we will find the differences between the two.

```{r}
ceof <- EOF(XX)
plot(ceof)
```

The results suggest that there are two leading modes which account for most of the variance. We can also inspect the second mode.

```{r}
plot(ceof,pattern=2)
```

To study similarities and differences between the reanalysis and the GCM, we need compare the statistics in the PCs. Below is a a plot with historgrams for the leading PC for both data sets:

```{r}
breaks <- seq(-0.5,0.5,by=0.25)
gcm <- hist(attr(ceof,'appendix.1')[,1],breaks=breaks,plot=FALSE)
rea <- hist(ceof[,1],breaks=breaks,plot=FALSE)
plot(rea$mids,rea$density,type='l',lwd=2,ylim=range(rea$density,gcm$density))
lines(gcm$mids,gcm$density,lwd=2,col='red')
```

We can apply statistical tests, like the Students t-test.

```{r}
print(t.test(coredata(ceof[,1]),coredata(attr(ceof,'appendix.1')[,1])))
```

There is also a more fancy diagnostic infographics for common EOFs in esd, which indicates the differences in terms of mean, standard deviation, and autocorrelation. An example is given below where the colour of the symbols indicate the EOF rank (only the blue ones are important) and the symbol size the ratio of the auto-correlation (small ones indicate .

```{r}
plot(diagnose(ceof),new=FALSE)
```

### 2016-04-29 CORDEX/pca4stations.Rmd
Common EOFs and PCA for stations testing

---
title: "PCA4station-data-inspection"
author: "Rasmus Benestad"
date: "April 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quality of station data

We want to look at a group of station data and test a strategy for assessing data quality.  This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

##**Station data**

Get some data to play with.

```{r}
library(esd)
library(utils)
if (!file.exists('precip.norway.rda')) download.file('https://ndownloader.figshare.com/files/2155751','precip.norway.rda')
load('precip.norway.rda')
## The precipitation data is 'Pr'
```

Now inspect the data - check for missing values

```{r}
diagnose(Pr)
```

Estimate annual aggregates:

```{r}
mu <- annual(Pr,'wetmean')
fw <- annual(Pr,'wetfreq')
map(mu,FUN='trend',new=FALSE)
```

Plot the original daily data

```{r}
plot(Pr,new=FALSE)
```

The wet-day mean precipitation $\mu$

```{r}
plot(mu,new=FALSE)
```

The wet-day frequency $f_w$

```{r}
plot(fw,new=FALSE)
```

Apply a principal component analysis (PCA) to the annual aggregates:

```{r}
pca.mu <- PCA(mu)
pca.fw <- PCA(fw)
```

Plot the leading PCA mode

```{r}
plot(pca.mu,new=FALSE)
```

Take a copy of the original data and mess it up by introducing deliberate errors and artifacts to degrade the information embedded. The idea is to explore the ways such errors affect the end result.

```{r}
X <- Pr
## Degrade the original data by synthetically introducing of errors and problems
d <- dim(X)
## Indices with values set to zero
s20.1 <- sample(1:d[1],1000); s20.2 <- sample(1:d[2],10)
## Indices with values set to random
s2r.1 <- sample(1:d[1],1000); s2r.2 <- sample(1:d[2],10)
## Set a random sub-selection of data points to zero or random values
X[s20.1,s20.2] <- 0
X[s2r.1,s2r.2] <- rexp(10000,rate=0.001)
## Set some locations to bad values
S2R <- sample(1:d[2],3)
X[,S2R] <- rexp(d[1],rate=c(1,0.1,0.5))
```

Estimate synthetic annual aggregates for $\mu$ and $f_w$.

```{r}
mu.s <- annual(X,'wetmean')
fw.s <- annual(X,'wetfreq')
map(mu,FUN='trend',new=FALSE)
```

```{r}
plot(X,new=FALSE)
```

```{r}
plot(mu.s,new=FALSE)
```

```{r}
plot(fw.s,new=FALSE)
```

```{r}
pca.smu <- PCA(mu.s)
pca.sfw <- PCA(fw.s)
```

```{r}
plot(pca.smu,new=FALSE)
```

### 2016-04-29 Tanzania/tmintest.Rmd

@@
 djf.e <- subset(as.4seasons(ferder),it='Jan')
-corfield(djf.e,DJF.E,colbar=list(rev=FALSE),new=FALSE)
+corfield(djf.e,DJF.E,colbar=list(breaks=seq(-1,1,by=0.1),rev=FALSE),new=FALSE)
 ```
@@
 attr(djf,'unit') <- 'degC'
-corfield(djf,DJF,colbar=list(rev=FALSE),new=FALSE)
+corfield(djf,DJF,colbar=list(breaks=seq(-1,1,by=0.1),rev=FALSE),new=FALSE)
 ```
@@
 attr(djf2,'unit') <- 'degC'
-corfield(djf2,DJF,colbar=list(rev=FALSE),new=FALSE)
+corfield(djf2,DJF,colbar=list(breaks=seq(-1,1,by=0.1),rev=FALSE),new=FALSE)
 ```
@@
 ```
+
+Songea was the station with the best scores in the March-May season
+
+```{r}
+y3 <- subset(Y,is='Songea')
+y3 <- subset(y3,is=1)
+mam <- subset(as.4seasons(y3),it='Apr')
+## fix a little bug
+attr(mam,'unit') <- 'degC'
+corfield(mam,MAM,colbar=list(breaks=seq(-1,1,by=0.1),rev=FALSE),new=FALSE)
+```
+
+```{r}
+mam.i <- regrid(MAM,is=y3)
+plot(stand(mam),new=FALSE)
+lines(stand(mam.i),lwd=3)     
+```
+
+## Have a look at the group of stations
+
+The use of principal component analysis (PCA) can reveal problems with some stations, since it will estimate patterns of covariance. We can apply PCA to the MAM season aggregated minimum temperature:
+
+```{r}
+tmin <- subset(Y,is=1:13)
+tmin <- subset(as.4seasons(tmin,nmin=60),it='Jan')
+diagnose(tmin)
+tmin <- pcafill(tmin)
+```
+
+Here is a plot of the evolution of the MAM mean minimum temperature:
+
+```{r}
+plot(tmin,new=FALSE)
+```
+
+The PCA suggests that there is one main pattern of variability, explaining 76% of the variance.
+The temperatures tend to increase over time, except for Mwanza (represented with the dark blue symbols which indicates negative weights).
+The other stations have similar weights, with most pronounced variability/trend near the coast.
+
+```{r}
+plot(PCA(tmin),new=FALSE)
+```

### 2016-04-29 CORDEX/pca4stations.Rmd
more on PCA & stations

@@
 
-Plot the leading PCA mode
+##**Synthesise errors**
+
+Take a copy of the original data and mess it up by introducing deliberate errors and artifacts to degrade the information embedded.
+The idea is to explore the ways such errors affect the end result.
+We can experiment with different degree of degration by changing the parameters that control the number of bad/random numbers we want to insert:
 
 ```{r}
-plot(pca.mu,new=FALSE)
+## number of bad data point in time
+n.t <- 4000
+#3 Number of bad data points in space
+n.s <- 50
 ```
-
-Take a copy of the original data and mess it up by introducing deliberate errors and artifacts to degrade the information embedded. The idea is to explore the ways such errors affect the end result.
 
@@
 ## Indices with values set to zero
-s20.1 <- sample(1:d[1],1000); s20.2 <- sample(1:d[2],10)
+s20.1 <- sample(1:d[1],n.t); s20.2 <- sample(1:d[2],n.s)
 ## Indices with values set to random
-s2r.1 <- sample(1:d[1],1000); s2r.2 <- sample(1:d[2],10)
+s2r.1 <- sample(1:d[1],n.t); s2r.2 <- sample(1:d[2],n.s)
 ## Set a random sub-selection of data points to zero or random values
 X[s20.1,s20.2] <- 0
-X[s2r.1,s2r.2] <- rexp(10000,rate=0.001)
+X[s2r.1,s2r.2] <- rexp(n.t*n.s,rate=0.001)
 ## Set some locations to bad values
@@
 
+Plot the degraded data:
+
 ```{r}
@@
 ```
+
+The aggregated degraded data: $\mu$
 
@@
 
+and $f_w$
+
 ```{r}
@@
 ```
+
+##**analysis of the original and degraded data**
+
+Apply PCA to the degraded data:
 
@@
 
+Plot the leading PCA mode of the original $\mu$
+
 ```{r}
+plot(pca.mu,new=FALSE)
+```
+
+Plot the leading PCA mode of the degraded $\mu$
+
+```{r}
+
 plot(pca.smu,new=FALSE)
 ```
+
+Plot the leading PCA mode of the original $\mu$
+
+```{r}
+plot(pca.fw,new=FALSE)
+```
+
+Plot the leading PCA mode of the degraded $\mu$
+
+```{r}
+
+plot(pca.sfw,new=FALSE)
+```
+
+##**Second information source**
+
+Retrieve mean sea-level pressure (SLP) and use a regression analysis to examine the data, knowing that the SLP is related to $f_w$.
+
+```{r}
+#slp <- annual(retrieve('slp.mon.mean.nc',lon=c(-60,30),lat=c(50,70)))
+#eof.slp<- EOF(slp)
+#save(file='eof.slp.rda',eof.slp)
+load('eof.slp.rda')
+ds <- DS(subset(pca.fw,pattern=1:3),eof.slp)
+ds.s <- DS(subset(pca.sfw,pattern=1:3),eof.slp)
+```
+
+Compare the results:
+```{r}
+plot(ds,new=FALSE)
+plot(ds.s,new=FALSE)
+```

### 2016-04-29 CORDEX/pca4stations.Rmd
more on PCA & stations

@@
 fw <- annual(Pr,'wetfreq')
-map(mu,FUN='trend',new=FALSE)
+map(mu,FUN='trend',new=FALSE,colbar=list(breaks=seq(-0.5,0.5,by=0.05),pal='t2m',rev=TRUE))
 ```
 
-Plot the original daily data
+The map suggests that the original data contains some randomness, as the trend in the wet-day mean precipitation $\mu$ does not indicate clear large-scale features, but contains siginicant random small-scale variations.
+This is not surprising, as $\mu$ seems to be more strongly influenced by local factors (e.g.
+convective rain) than e.g.
+the wet-day frequency $f_w$.
+
+```{r}
+map(mu,FUN='trend',new=FALSE,colbar=list(breaks=seq(-0.35,0.35,by=0.01),pal='t2m',rev=TRUE))
+```
+
+There are some trend outliers even for $f_w$, which look questionable.
+Due to capture/sampling limitations?
+Rain gauges only have diameters of the order of cm whereas the phenomenon they sample has scales of the order of km.
+
+We can plot the original daily data to get some idea of what it looks like
 
@@
 plot(Pr,new=FALSE)
+grid()
 ```
 
-The wet-day mean precipitation $\mu$
+The daily data contains a lot of information and it's hard to see relevant information.
+We can see large diffrences between the different stations and that there are some events with much higher amounts than others.
+similar plot of the annual wet-day mean precipitation $\mu$ can give a clearer picture of the long-term changes
 
 ```{r}
-plot(mu,new=FALSE)
+plot(mu,errorbar=FALSE,new=FALSE)
+grid()
 ```
 
-The wet-day frequency $f_w$
+The annual wet-day frequency $f_w$
 
@@
 plot(fw,new=FALSE)
+grid()
 ```
@@
 #3 Number of bad data points in space
-n.s <- 50
+n.s <- 25
 ```
+
+Apply the degrading:
 
@@
 
-Estimate synthetic annual aggregates for $\mu$ and $f_w$.
+Now, estimate synthetic annual aggregates for $\mu$ and $f_w$.
+Also make a new map of the trends in $\mu$ where the data has been degraded.
+The results from the trend analysis may not be so sensitive to the degraded data when only a selection of random numbers have been altered.
 
@@
 fw.s <- annual(X,'wetfreq')
-map(mu,FUN='trend',new=FALSE)
+map(mu.s,FUN='trend',new=FALSE,colbar=list(breaks=seq(-0.5,0.5,by=0.05),pal='t2m',rev=TRUE))
 ```
@@
 
-Apply PCA to the degraded data:
+We apply PCA to the degraded data to get a better picture of the spatial and temporal structures - we expect the covariance to be reduced and there should be less coherence between the different stations.
 
@@
 
-Plot the leading PCA mode of the original $\mu$
+Plot the leading PCA mode of the original $\mu$.
+The eigen-value spectrum (upper right panel) can provide an indication about the data quality.
+There are some coherent and pronounced variations when the leading modes are associated with a high proportion of the variance, and this is what we expect from parameters aggregated over reasonably long time scales.
+Hence, high variance associated with the leading mode suggests that there is some real information embedded in the data.
 
@@
 
-Plot the leading PCA mode of the degraded $\mu$
+We can compare the original results with leading PCA mode of the degraded $\mu$.
+Now the eigen-value spectrum is flatter, and the leading mode accounts for less of the variance.
+The spatial pattern is also more complex, without features that are associated with known geography.
+Also the time series contains less autocorrelation.
 
@@
 
-Plot the leading PCA mode of the original $\mu$
+Plot the leading PCA mode of the original $f_w$
 
@@
 
-Plot the leading PCA mode of the degraded $\mu$
+Plot the leading PCA mode of the degraded $f_w$
 

### 2016-05-02 EU-Circle/eu-circle_case-study-1.Rmd
EU-Circle: start scripts for Case 1 & 5

---
title: "eu-circle_case-study-1"
author: "Rasmus Benestad"
date: "May 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Extreme Dryness and forest fires on electricity and transport networks

Simultaneous forest fires ignite in the Bouche du Rhone department near Aix en Porvence, and in the Alpes Maritimes Department, at the French/Italian frontier, near a highway (A8) used by thousands of tourists. Due to the important smoke production, visibility is strongly reduced so that highway has to be closed, leading to an important traffic on the secondary road networks. The crossing of the frontier by car being completely impossible, traffic between France and Italy has to be diverted. Tourists are confined on highway rest areas. People, blocked on the roads, have difficulties to breathe because of the smoke and leave their cars, scattering into the nature. Additional accidents are caused because of the panic of people. Due to aerial firefighting, electricity lines have to be cut. Due to the fire smoke, aerial traffic in Nice airport has to be stopped. Numerous dwellings are without electricity and of course without telephones. Other emergency operations are disturbed because of the large delay of alert, major dispersion of means, decrease of available means.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

**Precipitation**

***Station data***

Represent the station data with lower-case $x$

```{r,echo=FALSE,message=FALSE}
library(esd)
ss <- select.station(param='precip',lon=c(5,10),lat=c(42,45),nmin=50)
map(ss)
```

```{r}
if (!file.exists('eu-circle_case-study-1_precio.rda')) {
  x <- station(ss)
  x <- subset(X,is=list(alt=-400))
  save(file='eu-circle_case-study-1_precio.rda',x)
} else load('eu-circle_case-study-1_precio.rda')
diagnose(x)
```

The raw daily precipitation data. The colour of the curves correspond to the colour marker of the location. There have been some instances with very large precipitation amounts in the past.

```{r}
plot(x,new=FALSE)
```

Examine the trend in the wet-day mean precipitation $\mu$

```{r}
mu <- annual(x,'wetmean')
wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
q95 <- annual(x,'wq95')
plot(mu,new=FALSE)
```

For many purposes, it may be useful to assume that the daily precipitation amount is approximately exponentially distributed for the days when it rains and that the probability of the precipitation exceedin a threshold value $x_0$ can be estimated according to $Pr(X>x_0) = f_w e^{-x_0/\mu}$ where $f_w$ is the wet-day frequency. Below is a test for whether the amount is close to exponential for a wet day.

```{r}
## check the relationship between wet-day 95-percentile and wet-day mean
plot(-log(0.05)*zoo(mu[,2]),zoo(q95[,2]),pch=19,xlim=c(0,120),ylim=c(0,120),
     main='Wet-day mean v.s. 95th wet percentile',
     xlab=expression(paste(-ln(1-0.95)*mu,' (mm/day)')), ylab=expression(paste(q[95],' (mm/day)')))
points(-log(0.05)*zoo(mu[,1]),zoo(q95[,1]),pch=19,col=rgb(1,0,0,0.2))
points(-log(0.05)*zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
points(-log(0.05)*zoo(mu[,4]),zoo(q95[,4]),pch=19,col=rgb(0,1,0,0.2))
points(-log(0.05)*zoo(mu[,5]),zoo(q95[,5]),pch=19,col=rgb(0,0.7,0.7,0.2))
points(-log(0.05)*zoo(mu[,6]),zoo(q95[,6]),pch=19,col=rgb(0.7,0,0.7,0.2))
points(-log(0.05)*zoo(mu[,7]),zoo(q95[,7]),pch=19,col=rgb(0.5,0.5,0.5,0.2))
grid()
legend(0,120,loc(climatrans.pr),pch=19,col=c(rgb(1,0,0,0.2),'black',rgb(0,0,1,0.2)),bty='n')
```

The scatter plot suggests that the points are clustered around the diagonal and hence the 24-hr precipitation amoun is approximately exponentially distributed.

We also need to examine the trend in the wet-day frequency

```{r}
fw <- annual(x,'wetfreq')
plot(mu,new=FALSE)
```

```{r}
z <- subset(x,it=c(1900,2016),is=1)
ncd <- spell(z,threshold=1)
plot(ncd)
```

How close is the duration of wet/dry spells to the geometric distribution?

```{r}
hist(ncd)
```

The mean value if the parameter that describes the shape of the geometric distribution. The comparison suggests that it does not give a bad representation of the observed charactereistics. Are there any trends in the annual mean durations?

```{r}
amncd <- annual(ncd)
plot(amncd,col=c(rgb(0,0,1),rgb(1,0.3,0)))
grid()
lines(trend(subset(amncd,is=1)),lty=2)
lines(trend(subset(amncd,is=2)),lty=2)
```

There is a

Represent gridded precipitation with upper-case $X$

***Gridded precipitation***

Gridded data are not ideal when it comes to extremes, as the gridding process makes the data spatially inhomogeneous. But

```{r}
X <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(4,8),lat=c(43,46))
map(X)
```

### 2016-05-02 EU-Circle/eu-circle_case-study-5.Rmd
EU-Circle: start scripts for Case 1 & 5

---
title: "EU-Circle Case study 5"
author: "Rasmus Benestad"
date: "May 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rapid Winter Flooding (melting ice, narrow mountain streams, flooding) around Dresden, Germany

Dresden is the largest city in the Eastern part of Germany, Saxony, near the Czech border. It is crossed by the large river Elbe (its width is around 110m in Dresden) which comes from the Czech Republic and flows through Magdeburg and Hamburg into the North Sea. The region between Dresden and the Czech border, but also in the near surroundings in the Eastern and South-Western directions, is quite hilly with mountains up to ca. 1200 m. In the recent past (especially notable are the big floodings in 2002 and 2013) there has been quite some floodings caused by intense and long rain which led to floods of the Elbe.However, there is another threat which is happening more often and hence causing more damage because of the currently ongoing climate changes. In the mountains are lots of small mountain streams which are really small on a normal day (a typical one has a width of 1.5m and a depth of 0.5m). But in cases of heavy rains these small streams rapidly grow quite large (1-2h is not uncommon), especially compared to their normal size (sometimes they grow to their hundredfold size in respect of the amount of water they carry). This causes a lot of damage, especially as prediction and short-term prevention mechanisms are not really possible because of the small timescale. The damages caused by such events are quite local and often not related to the flooding of big rivers. In these  regions there are often roads going through valley which are vital in a sense that if they are  not available any more, quite long bypasses have to be taken. Additionally railways,  especially railway bridges, are often effected as well. These two aspects show the impacts on the transport network.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

### 2016-05-13 CORDEX/ensembles.Rmd
Added a set with new Rmd scripts

@@
 ```{r}
-fit.atm <- lm(dT ~ atm.grid + atm.lev + atm.top + oce.lev + oce.z,data=X)
+fit.atm <- lm(dT ~ atm.grid + atm.lev + atm.top + oce.lev + oce.z +
+                   atm.grid*atm.lev + atm.grid*atm.top +       # interactions
+                   atm.grid*oce.lev + atm.grid*oce.z,data=X)
 print(summary(fit.atm))

### 2016-05-13 CORDEX/pca4predictands.Rmd
Added a set with new Rmd scripts

---
title: "PCA 4 predictands"
author: "Rasmus Benestad"
date: "May 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Demonstration

This R-markdown script provides some of the material presented in <http://www.icrc-cordex2016.org/images/pdf/Programme/Abstracts/Orals/Parallel_B/B3_5_Benestad.pdf> at the ICRC-CORDEX conference <http://www.icrc-cordex2016.org/index.php> in Stockholm, May 2016.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r,warnings=FALSE}
## Script to plot PCA of stations
library(esd)
```

Get the R-package esd (source: <https://github.com/metno/esd>)

```{r}
## Demonstration using real data from the ECA&D: Europe
Y0 <- station(lon=c(-20,20),lat=c(40,60),nmin=70,src='ecad')
map(Y0,FUN='mean')
## Do the analysis for the winter season only (DJF)
Y <- subset(as.4seasons(Y0),it='Jan')

## Need to remove period with little data and stations with little data
Y <- subset(Y,it=c(1950,2015))
nv <- apply(coredata(Y),2,'nv')
Y <- subset(Y,is=nv > 50)
## Show the data availability
diagnose(Y)

## Fill in missing data
Y <- pcafill(Y)
diagnose(Y)

## compute PCA
pca <- PCA(Y)
plot(pca,new=FALSE)
pca <- subset(pca,pattern=1:4)

## Get a predictor - need to download the file because OpenDap doesn't seem to work
tas <- retrieve('ftp://ftp.cdc.noaa.gov/Datasets/ncep.reanalysis.derived/surface/air.mon.mean.nc',lon=c(-30,30),lat=c(35,65))
tas <- subset(as.4seasons(tas),it='Jan')
eof <- EOF(tas)

## Example of downscaling
ds <- DS(pca,eof)
plot(ds)
```

### 2016-05-30 Tanzania/dse.tz-ncep.Rmd
Add DSensemble results

---
title: "Tanzania"
author: "Rasmus Benestad"
date: "May 10, 2016"
output: html_document
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r,message=FALSE}
## Script to downscale temperature from Tanzania.
## Data from Habiba.

library(esd)
```

# Functions

Prepare the predictands

```{r,message=FALSE}
predictands <- function(param='tmin',incl.ghcn=FALSE) {
  ## Prepare the predictand data
  #print("predictands")
  fname <- paste(param,'.tz.rda',sep='')
  path <- paste('Habiba/',param,sep='')
  if (!file.exists(fname)) {
  files <- list.files(path,pattern='rda',full.names=TRUE)
  locs <- list.files(path,pattern='rda')
  dse <- grep('dse.tm',files)
  files <- files[-dse]; locs <- locs[-dse]
  print(files)
  locs <- substr(locs,1,nchar(locs)-4)
  Y <- NULL
  for (i in 1:length(files)) {
    load(files[i])
    y1 <- eval(parse(text=locs[i]))
    if (is.null(Y)) Y <- y1 else
                    Y <- combine.stations(Y,y1)
  }
  save(file=fname,Y)
  plot(Y)
} else load(fname)
  #print('GHCN')
  if (incl.ghcn) {
    xstations <- switch(param,'tmin'='Habiba/tn.eafrica.rda',
                            'tmax'='Habiba/tx.eafrica.rda')
    load(xstations)
    if (param=='tmin') y <- tn.eafrica else y <- tx.eafrica
    y <- subset(y,it=c(1950,2015))
    Y <- combine(Y,y)
  }
  map(Y,FUN='mean',cex=-2)
  invisible(Y)
}
```

Function that applies the downscaling to a combination of predictands in terms of principal component analysis (PCA) results and predictors that involve large multi-model ensembles.

```{r,message=FALSE}
downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
                      period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
                      lon=c(0,100),lat=c(65,90),eofs=1:6,n=10,rel.cord=FALSE,select=NULL) {

  print('downscale')

## Use a time and space window:  
  Y <- subset(Y,it=period)
  Y <- subset(Y,is=list(lat=lat,lon=lon))

## Estimate seasonal means & weed out stations with little data
  Y4 <- subset(as.4seasons(Y,FUN=FUN,nmin=60),it=it)
  ok <- apply(coredata(Y4),1,nv)
  Y4 <- subset(Y4,it=ok>0)
  nok <- apply(coredata(Y4),2,nv)
  Y4 <- subset(Y4,is=nok>15)

  print(paste(round(100*sum(!is.finite(Y4))/length(!is.finite(Y4))),'% missing',sep=''))
  if (plot) map(Y,FUN=FUN,cex=-2)

  nmiss <- round(100*sum(!is.finite(Y4))/length(!is.finite(Y4)))
  print(paste(nmiss,'% missing',sep=''))

## Fill missing data using PCA-based regression
  Z <- pcafill(Y4)

  pca <- PCA(Z,n=n)
  if (plot) plot(pca)

## Downscale results
  print('DSensemble')
  dse.pca <- DSensemble(pca,predictor=predictor,FUNX=FUNX,verbose=verbose,
                        biascorrect=TRUE,rcp=rcp,eofs=eofs,select=select,
                        lon=lon,lat=lat,rel.cord=rel.cord,it=it)

  attr(dse.pca,'N.missing') <- nmiss
  invisible(dse.pca)
}
```

Function that extracts results for a given year

```{r}
distill <- function(x,it=2050) {
  y <- subset(x,it=it)
  attributes(y) <- NULL
  y <- mean(y)
  return(y)
}
```

Function that grids the results

```{r,message=FALSE}
gridmap <- function(dse.results,param=NULL,
                    fast=TRUE,it=c(2070,2099),it0=c(1979,2012),verbose=FALSE,
                    breaks=seq(0,7,length.out=15),pal='warm',rev=FALSE) {
  if (verbose) print('gridmap')
  library(LatticeKrig)

  if (is.null(param)) param <- varid(dse.results) 
  if (verbose) print ('as.station')
  if (inherits(dse.results,'pca')) dse.station <- as.station(dse.results)  else
                                   dse.station <- dse.results

  if (verbose) print('distill')
  Z0 <- lapply(dse.station,distill,it=it0)
  Z <- lapply(dse.station,distill,it=it)
  lon <- unlist(lapply(dse.station,lon))
  lat <- unlist(lapply(dse.station,lat))
  alt <- unlist(lapply(dse.station,function(x) alt(attr(x,'station'))))
  X <- attr(dse.station[[1]],'station')
  z0 <- matrix(unlist(Z0),1,length(dse.station))
  z <- matrix(unlist(Z),1,length(dse.station))
  data(etopo5)
  etopo5 <- subset(etopo5,is=list(lon=range(lon),lat=range(lat)))
  etopo5[etopo5<=0] <- NA

  if (verbose) print('grid')
  ## Set the grid to be the same as that of etopo5:
  grid <- structure(list(x=lon(etopo5),y=lat(etopo5)),class='gridList')

  #3 Flag dubplicated stations:
  ok <- !(duplicated(lon) & duplicated(lat))

  ## Spread in the  90-percente interval changing
  obj <- LatticeKrig( x=cbind(lon[ok],lat[ok]),
                      y=z[1,ok] - z0[1,ok],Z=alt[ok] )

  ##  obj <- LatticeKrig( x=cbind(lon[ok],lat[ok]), y=z[2,ok],Z=alt[ok])
  w <- predictSurface(obj, grid.list = grid,Z=etopo5)
  w$z[is.na(etopo5)] <- NA

  ## Get rid of packages that have functions of same name:
  detach("package:LatticeKrig")
  detach("package:fields")
  detach("package:spam")
  detach("package:grid")
  detach("package:maps")

  ## Convert the results from LatticeKrig to esd:
  W <- w$z
  attr(W,'variable') <- param
  attr(W,'unit') <- 'degC' 
  attr(W,'longitude') <- w$x
  attr(W,'latitude') <- w$y
  class(W) <- class(etopo5)

  ## Make a projection that zooms in on the Barents region
  #dev.new()

  colbar <- list(breaks=round(breaks,2),rev=rev,pal=pal)
  if (verbose) print(colbar)
  if (is.null(rev)) rev <- switch(param,'t2m'=FALSE,'pr'=TRUE)
  Wx <- max(abs(W),na.rm=TRUE)
  if (is.null(breaks)) breaks <- round(seq(0,Wx,length=31),2)
  if (fast)
    map(W,xlim=range(lon(W)),ylim=range(lat(W)),
        colbar=colbar,new=FALSE)
  else {
    attr(W,'variable') <- NULL
    attr(W,'unit') <- NULL
    map(W,xlim=range(lon(W)),ylim=range(lat(W)),projection='sphere',
        colbar=colbar,verbose=verbose)
  }
  figlab(paste(it,collapse='-'),ypos=0.99)
}

```

Function that applies tests to theresults through applying PCA

```{r,warning=FALSE}
pcaasses <- function(Y) {
  ## Cut away blocks with missing data:

  nv <- apply(coredata(Y),2,nv)
  Y <- subset(Y,is=nv > 12000)
  Y4s <- as.4seasons(Y,nmin=60)
  y1 <- subset(Y4s,it='Jan')
  y2 <- subset(Y4s,it='Apr')
  y3 <- subset(Y4s,it='Jul')
  y4 <- subset(Y4s,it='Oct')

  ## Fill in missing
  y1 <- pcafill(y1)
  y2 <- pcafill(y2)
  y3 <- pcafill(y3)
  y4 <- pcafill(y4)

  ## Apply PCA:
  pca.djf <- PCA(y1)
  pca.mam <- PCA(y2)
  pca.jja <- PCA(y3)
  pca.son <- PCA(y4)

  ## Examine the results:
  print('DJF - PCA')
  plot(pca.djf,new=FALSE)
  print('MAM - PCA')
  plot(pca.mam,new=FALSE)
  print('JJA - PCA')
  plot(pca.jja,new=FALSE)
  print('SON - PCA')
  plot(pca.son,new=FALSE)

  ## compare the station data with reanalysis - based on CCA
  ## Read reanalysis:
  X <- retrieve('air.mon.mean.nc',lon=range(lon(Y))+c(-3,3),lat=range(lat(Y))+c(-3,3))
  X4s <- as.4seasons(X)

  eof.djf <- EOF(subset(X4s,it='Jan'))
  eof.mam <- EOF(subset(X4s,it='Apr'))
  eof.jja <- EOF(subset(X4s,it='Jul'))
  eof.son <- EOF(subset(X4s,it='Oct'))

  cca.djf <- CCA(pca.djf,eof.djf)
  cca.mam <- CCA(pca.mam,eof.mam)
  cca.jja <- CCA(pca.jja,eof.jja)
  cca.son <- CCA(pca.son,eof.son)

  print('DJF - CCA')
  plot(cca.djf,new=FALSE)
  print('MAM - CCA')
  plot(cca.mam,new=FALSE)
  print('JJA - CCA')
  plot(cca.jja,new=FALSE)
  print('SON - CCA')
  plot(cca.son,new=FALSE)

## The quick and effiecint way to downscale all the stations:
  ds.djf <- DS(subset(pca.djf,pattern=1:4),eof.djf)
  ds.mam <- DS(subset(pca.mam,pattern=1:4),eof.mam)
  ds.jja <- DS(subset(pca.jja,pattern=1:4),eof.jja)
  ds.son <- DS(subset(pca.son,pattern=1:4),eof.son)

  print('DJF - DS')
  plot(ds.djf,new=FALSE)
  #dev.new()
  print('MAM - DS')
  plot(ds.mam,new=FALSE)
  #dev.new()
  print('JJA - DS')
  plot(ds.jja,new=FALSE)
  #dev.new()
  print('SON - DS')
  plot(ds.son,new=FALSE)

  return(list(pca.djf=pca.djf,pca.mam=pca.mam,pca.jja=pca.jja,
              pca.son=pca.son,y.djf=y1,y.mam=y2,y.jja=y3,y.son=y4))
}
```

#Maximum temperature

The code below shows how the functions are used and applied to the data for the daily maximum $T_x$. The first chunck of code carries out some preliminary analysis, the second shows how the results are derived, and the third show how the plots are made.

```{r,message=FALSE,warning=FALSE}
#-----------------------------------------------------------------------
# Define season and parameter

param <- 'tmax'
FUN <- 'mean'
reanalysis <- 'air.mon.mean.nc'
#reanalysis <- 'ERAINT_t2m_mon.nc'
FUNX <- 'mean'
data.path <- 'dse.Tz.NCEP/'

predictands(param) -> Y

Y <- subset(Y,it=c(1961,2011))
lon=range(lon(Y))+c(-10,10); lat=range(lat(Y))+c(-15,15)

## Check the predictands:
X <- pcaasses(Y)
```

The PCA suggest that most of the variance in the station data for daily maximum temperature $T_x$ can be represented by a small number of leading modes. For the SON season, however, the second mode is more prominent, suggesting more complicated structure where spatial variability has less coherency between the stations.

The CCA suggests dissimilar spatial temperature pattern with the higest temporal correlation in the predictand and predictor. The closest spatial similarity is seen in SON.

The DS analysis applied to the reanalysis suggests strongest match in MAM and weakest in DJF. The spatial patterns differ to some extent and there is not a great match between the predictor and predictand.

```{r,message=FALSE,warning=FALSE}
## Get the predictand -> Y

for (it in c('djf','mam','jja','son')) {

  ## Get the large-scale predictor:
  if (!exists('predictor')) {
    T2M <- retrieve(reanalysis,lon=lon,lat=lat)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)
  } else if (length(month(subset(predictor,it=it)))==0)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)

  #print(paste('Generating dse.',param,'.tz.rcp45.',it,'.rda',sep=''))

  ## Carry out the downscaling:
  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp45.',it,'.rda',sep=''))) {
    #dev.new()
    dse.t2m.tz.rcp45 <- downscale(Y,predictor,it,param,FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp45.',it,'.rda',sep=''),dse.t2m.tz.rcp45)
  }

  #print(paste('Generating dse.',param,'.tz.rcp26.',it,'.rda',sep=''))

  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp26.',it,'.rda',sep=''))) {
    dse.t2m.tz.rcp26 <- downscale(Y,predictor,it,param,rcp='rcp26',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp26.',it,'.rda',sep=''),dse.t2m.tz.rcp26)
  }

   #print(paste('Generating dse.',param,'.tz.rcp85.',it,'.rda',sep=''))

  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp85.',it,'.rda',sep=''))) {
    dse.t2m.tz.rcp85 <- downscale(Y,predictor,it,param,rcp='rcp85',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp85.',it,'.rda',sep=''),dse.t2m.tz.rcp85)
  }
}
#print('--- Completed downscaling ---')
```

##The March-May season.

```{r,message=FALSE,warning=FALSE}
## analysis:

load(paste(data.path,'dse.',param,'.tz.rcp85.mam.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
plot(subset(z,is='Arusha'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('figure4g.',param,'.pdf'))

plot(subset(z,is='Musoma'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('figure4h.',param,'.pdf'))

gridmap(z)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.pdf',sep=''))

validate(z)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.diagnose.pdf',sep=''))
```

The plume plots look OK for Arusha and Musoma, and the observations fit reasonably well within the envelope of downscaled ensemble (RCP8.5). The gridded map suggests strongest warming in the north (~5C increase in 2099).

The rank test (Wilcox) indicates reasonable scores, although the rank distribution is not entirely flat. The diagnostic (target plot) for the diagnosis suggest that the trend in the ensemble is higher than in observation (horizontal axis) and that the inter-annual variance (vertical axis) is smaller in the ESD results.

##The December-February season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.djf.rda',sep=''))  
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.pdf',sep=''))
validate(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.validate.pdf',sep=''))
diagnose(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.diagnose.pdf',sep=''))
```

A weaker warming (~2C) is projected in DJF, also with strongest response in the north.  But the rank scores are weak.

##The June-August season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.jja.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.pdf',sep=''))
validate(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.validate.pdf',sep=''))
diagnose(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.diagnose.pdf',sep=''))
```

The assessment suggest the JJA season is better downscaled, with reasonable rank scores. The projected warming is more uniform over Tanzania, albeit with some local variations.

##The September-November season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.pdf',sep=''))
validate(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.validate.pdf',sep=''))
diagnose(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))

```

The SON season is associated with strongest warming in the north, and displays a clear north-south gradient. The warming is projected to be 0.7--1.3C by 2099, and the rank scores are moderately good - the distribution of the ranks had a U-shape rather than being flat.

##Check a random station: is is well-represented by the PCA?

```{r,message=FALSE,warning=FALSE}
## Check the PCA-results:
z.djf <- as.station(X$pca.djf)
y.djf <- X$y.djf
par(new=FALSE)
plot.zoo(subset(z.djf,is=1),subset(y.djf,is=1),
         main=paste('Test: Winter temperature at',loc(subset(y.djf,is=1))))
plot(subset(z.djf,is=1),lwd=3,col='grey70')
lines(subset(y.djf,is=1))
```

The single station data could be well-recovered from the PCA product.

#Implementation - Minimum temperature

Get and prepare the predictand data for the daily minimum temperature $T_n$.

```{r,message=FALSE,warning=FALSE}
#-----------------------------------------------------------------------
# Define season and parameter

param <- 'tmin'

predictands(param) -> Y

Y <- subset(Y,it=c(1961,2011))
lon=range(lon(Y))+c(-10,10); lat=range(lat(Y))+c(-15,15)

## Check the predictands:
X <- pcaasses(Y)
```

Apply the ESD models to implement thedownscaling

```{r,message=FALSE,warning=FALSE}
## Get the predictand -> Y

for (it in c('djf','mam','jja','son')) {

  ## Get the large-scale predictor:
  if (!exists('predictor')) {
    T2M <- retrieve(reanalysis,lon=lon,lat=lat)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)
  } else if (length(month(subset(predictor,it=it)))==0)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)

  print(paste('Generating dse.',param,'.tz.rcp45.',it,'.rda',sep=''))

  ## Carry out the downscaling:
  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp45.',it,'.rda',sep=''))) {
    #dev.new()
    dse.t2m.tz.rcp45 <- downscale(Y,predictor,it,param,FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp45.',it,'.rda',sep=''),dse.t2m.tz.rcp45)
  }

  print(paste('Generating dse.',param,'.tz.rcp26.',it,'.rda',sep=''))

  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp26.',it,'.rda',sep=''))) {
    dse.t2m.tz.rcp26 <- downscale(Y,predictor,it,param,rcp='rcp26',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp26.',it,'.rda',sep=''),dse.t2m.tz.rcp26)
  }

   print(paste('Generating dse.',param,'.tz.rcp85.',it,'.rda',sep=''))

  if (!file.exists(paste(data.path,'dse.',param,'.tz.rcp85.',it,'.rda',sep=''))) {
    dse.t2m.tz.rcp85 <- downscale(Y,predictor,it,param,rcp='rcp85',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste(data.path,'dse.',param,'.tz.rcp85.',it,'.rda',sep=''),dse.t2m.tz.rcp85)
  }
}
print('--- Completed downscaling ---')
```

The PCA suggest that most of the variance in the station data for daily maximum temperature $T_n$ can be represented by a small number of leading modes. For the SON season, however, the second mode is more prominent, suggesting more complicated structure where spatial variability has less coherency between the stations.

The CCA suggests dissimilar spatial temperature pattern with the higest temporal correlation in the predictand and predictor.

The DS analysis applied to the reanalysis suggests strongest match in MAM and weakest in DJF. The spatial patterns differ to some extent and there is not a great match between the predictor and predictand.

##The March-May season

```{r,message=FALSE,warning=FALSE}
## analysis:

load(paste(data.path,'dse.',param,'.tz.rcp85.mam.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
plot(subset(z,is='Arusha'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('figure4g.',param,'.pdf'))

plot(subset(z,is='Musoma'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('figure4h.',param,'.pdf'))

gridmap(z)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.pdf',sep=''))

validate(z)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.validate.pdf',sep=''))

diagnose(z,new=FALSE)
figlab('MAM')
#dev.copy2pdf(file=paste('map.',param,'.mam.diagnose.pdf',sep=''))
```

The plume plots look good for Arusha and Musoma, and strongest warming (3.5-4C) is projected for the eastern part of Tanzania. The evaluation of the quality gave moderate results - some discrepancies in terms of the rank scores.

##The December-February season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.djf.rda',sep=''))  
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.pdf',sep=''))
validate(z)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('DJF')
#dev.copy2pdf(file=paste('map.',param,'.djf.diagnose.pdf',sep=''))
```

Strongest warming projected for the southeast part of the country (~3C), and the evaluation suggest moderate scores. Again, still some discrepancies.

##The June-August season

```{r,message=FALSE,warning=FALSE}
load(paste(data.path,'dse.',param,'.tz.rcp85.jja.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.pdf',sep=''))
validate(z)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('JJA')
#dev.copy2pdf(file=paste('map.',param,'.jja.diagnose.pdf',sep=''))
```

Most pronounced warming is projected for the souther part (~3C), and the assessment indicates moderate quality with some discrepancies. The ensemble results tend to account for lower inter-annual variability than in the observations and weaker trend (except for one station with higher trend).

##The September-November season

```{r,message=FALSE,warning=FALSE}
#load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
load(paste(data.path,'dse.',param,'.tz.rcp45.son.rda',sep=''))
z <- as.station(dse.t2m.tz.rcp85)
gridmap(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.pdf',sep=''))
validate(z)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.validate.pdf',sep=''))
diagnose(z,new=FALSE)
figlab('SON')
#dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))

z <- as.station(dse.t2m.tz.rcp85)
plot(subset(z,is='Arusha'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('SON')

plot(subset(z,is='Musoma'),target.show=FALSE,map.show=FALSE,legend=FALSE,xlim=c(1900,2100),new=FALSE)
figlab('SON')
```

Strongest projected warming in the south (~3C) and moderate quality.

## Check a random station: is is well-represented by the PCA?

```{r,message=FALSE,warning=FALSE}
## Check the PCA-results:
z.djf <- as.station(X$pca.djf)
y.djf <- X$y.djf
plot.zoo(subset(z.djf,is=1),subset(y.djf,is=1),
         main=paste('Test: Winter temperature at',loc(subset(y.djf,is=1))))
plot(subset(z.djf,is=1),lwd=3,col='grey70')
lines(subset(y.djf,is=1))
```

Single station records can be recovered from the PCA.

#Summary

There are problems with some stations, which is visible in the PCA, CCA and DS carried out with just reanalysis. The spatial patterns indicate non-smooth and incoherent structures, suggesting problems with some of the stations and a mismatch with the reanalysis. Furthermore, the cross-validation reveal moderate scores (r typically less than 0.7), and the evaluations based on trends, range of inter-annual variability and rank score reveal some discrepancies.

The NCEP/NCAR reanalysis is probably better than ERAINT for ESD model calibration.

### 2016-05-30 Tanzania/dse.tz.Rmd
Add DSensemble results

--- 2016-05-30 Tanzania/dse.tz-ncep.Rmd
@@
 
-# R Markdown
+## R Markdown
 
@@
 
-The code below shows how the functions are used and applied to the data for the daily maximum $T_x$.
+The code below shows how the functions are used and applied to the data for the daily maximum temperature $T_x$.
 The first chunck of code carries out some preliminary analysis, the second shows how the results are derived, and the third show how the plots are made.
@@
 FUN <- 'mean'
-reanalysis <- 'air.mon.mean.nc'
-#reanalysis <- 'ERAINT_t2m_mon.nc'
+#reanalysis <- 'air.mon.mean.nc'
+reanalysis <- 'ERAINT_t2m_mon.nc'
 FUNX <- 'mean'
-data.path <- 'dse.Tz.NCEP/'
+data.path <- 'dse.Tz.ERAINT/'
 
@@
 The rank test (Wilcox) indicates reasonable scores, although the rank distribution is not entirely flat.
-The diagnostic (target plot) for the diagnosis suggest that the trend in the ensemble is higher than in observation (horizontal axis) and that the inter-annual variance (vertical axis) is smaller in the ESD results.
+The diagnostic (target plot) for the diagnosis suggest a reasonable reproduction of trend (horizontal axis) and variance (vertical axis), although four stations were associated with higher variance than suggested by the ensemble and the observed trends were weaker than suggested by the ensemble.
 
@@
 
-A weaker warming (~2C) is projected in DJF, also with strongest response in the north.  But the rank scores are weak.
+The projected warming was again hiher in the north (~3C), and the evaluation suggested reasonable skill, although not perfect.
 
@@
 
-The assessment suggest the JJA season is better downscaled, with reasonable rank scores. The projected warming is more uniform over Tanzania, albeit with some local variations.
+In June-August, the projected warming in $T_x$ was found to be greatest the west (~3C), but the assessment of the rank score suggested a poor match between observations and the downscaled ensemble.
 
@@
 #dev.copy2pdf(file=paste('map.',param,'.son.diagnose.pdf',sep=''))
-
 ```
 
-The SON season is associated with strongest warming in the north, and displays a clear north-south gradient.
-The warming is projected to be 0.7--1.3C by 2099, and the rank scores are moderately good - the distribution of the ranks had a U-shape rather than being flat.
+The results suggest strongest warming in the north, but the map is unrealistic with only a latitude-based gradient. The assessment of the ranked score also suggests a weak performance.
 
-##Check a random station: is is well-represented by the PCA?
+## Test of PCA
+
+Check a random station: is it well-represented by the PCA?
 
@@
 
-The single station data could be well-recovered from the PCA product.
+The data for Arusha was recovered from the PCA products.
 
-#Implementation - Minimum temperature
+#Minimum temperature
 
@@
 
-Apply the ESD models to implement thedownscaling
+The PCA suggests that most of the variance in $T_n$ was captured by a small number of leading modes.
+The CCA found that the patterns in the predictand and predictor with highest correlation were not so similar in terms of spatial structure.
+
+The DS applied to the reanalysis found highest scores for MAM, and again, the spatial patterns were dissimilar.
+
+## Apply the ESD models to implement the downscaling
 
@@
 
-The PCA suggest that most of the variance in the station data for daily maximum temperature $T_n$ can be represented by a small number of leading modes.
-For the SON season, however, the second mode is more prominent, suggesting more complicated structure where spatial variability has less coherency between the stations.
-
-The CCA suggests dissimilar spatial temperature pattern with the higest temporal correlation in the predictand and predictor.
-
-The DS analysis applied to the reanalysis suggests strongest match in MAM and weakest in DJF.
-The spatial patterns differ to some extent and there is not a great match between the predictor and predictand.
-
 ##The March-May season
@@
 
-The plume plots look good for Arusha and Musoma, and strongest warming (3.5-4C) is projected for the eastern part of Tanzania.
-The evaluation of the quality gave moderate results - some discrepancies in terms of the rank scores.
+The plume plots for Arusha and Musoma look ok.
+Strongest warming (~3C) is projected for the eastern part of the country.
+THe assessment of the downscaled results in terms of ranked score suggest that the observations have a different character to the downscaled results (cold bias).
 
@@
 
-Strongest warming projected for the southeast part of the country (~3C), and the evaluation suggest moderate scores. Again, still some discrepancies.
+The projections suggest strongest warming in the southeast (~2C), and again, the evaluation indicates a mismatch between observations and downscaled results. Poor performance.
 
@@
 
-Most pronounced warming is projected for the souther part (~3C), and the assessment indicates moderate quality with some discrepancies.
-The ensemble results tend to account for lower inter-annual variability than in the observations and weaker trend (except for one station with higher trend).
+Strongest warming in the southwest (~3C), and higher ranked scores than the two previous seasons, but the observations have a higher variance than accounted for by the downscaled ensemble and the the trends in the downscaled results are weaker than those observed.
 
@@
 ```{r,message=FALSE,warning=FALSE}
-#load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
-load(paste(data.path,'dse.',param,'.tz.rcp45.son.rda',sep=''))
+load(paste(data.path,'dse.',param,'.tz.rcp85.son.rda',sep=''))
 z <- as.station(dse.t2m.tz.rcp85)
@@
 
-Strongest projected warming in the south (~3C) and moderate quality.
+Greatest warming in the southwest, but these results look suspect. The plume-plot for Arusha exhibits a spread that increases over time that spans from cooling to warming.
 
-## Check a random station: is is well-represented by the PCA?
+##Check a random station: is it well-represented by the PCA?
 
@@
 
-Single station records can be recovered from the PCA.
+The original data is recovered from the PCA product.
 
@@
 
-The NCEP/NCAR reanalysis is probably better than ERAINT for ESD model calibration.
+The ERAINT is probably too short for proper model calibration.

### 2016-05-30 Tanzania/tmintest.Rmd
Add DSensemble results

@@
 
-We can interpolate temperatures for the same coordinate as for Mwanza, but when we do the comparison, we need to use standardised values as the predictor represents the mean of daily mean temperatures $\overline{T}$ whereas the station is the mean daily minimum $T[min]$:
+We can interpolate temperatures for the same coordinate as for Mwanza, but when we do the comparison, we need to use standardised values as the predictor represents the mean of daily mean temperatures $\bar{T}$ whereas the station is the mean daily minimum $T_{min}$:
 

### 2016-05-30 Tanzania/dse.tz-ncep.R

@@ -1,0 +1,135 @@
+
+## Script to downscale temperature from Tanzania.
+## Data from Habiba.
+
+library(esd)
+setwd('~/R/storeA')
+
+predictands <- function(param='tmin',incl.ghcn=FALSE) {
+  ## Prepare the predictand data
+  #print("predictands")
+  fname <- paste(param,'.tz.rda',sep='')
+  path <- paste('Habiba/',param,sep='')
+  if (!file.exists(fname)) {
+  files <- list.files(path,pattern='rda',full.names=TRUE)
+  locs <- list.files(path,pattern='rda')
+  dse <- grep('dse.tm',files)
+  files <- files[-dse]; locs <- locs[-dse]
+  print(files)
+  locs <- substr(locs,1,nchar(locs)-4)
+  Y <- NULL
+  for (i in 1:length(files)) {
+    load(files[i])
+    y1 <- eval(parse(text=locs[i]))
+    if (is.null(Y)) Y <- y1 else
+                    Y <- combine.stations(Y,y1)
+  }
+  save(file=fname,Y)
+  plot(Y)
+} else load(fname)
+  #print('GHCN')
+  if (incl.ghcn) {
+    xstations <- switch(param,'tmin'='Habiba/tn.eafrica.rda',
+                            'tmax'='Habiba/tx.eafrica.rda')
+    load(xstations)
+    if (param=='tmin') y <- tn.eafrica else y <- tx.eafrica
+    y <- subset(y,it=c(1950,2015))
+    Y <- combine(Y,y)
+  }
+  map(Y,FUN='mean',cex=-2)
+  invisible(Y)
+}
+
+downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
+                      period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
+                      lon=c(0,100),lat=c(65,90),eofs=1:5,n=3,rel.cord=FALSE,select=1:5) {
+
+  print('downscale')
+  
+## Use a time and space window:  
+  Y <- subset(Y,it=period)
+  Y <- subset(Y,is=list(lat=lat,lon=lon))
+
+## Estimate seasonal means & weed out stations with little data
+  Y4 <- subset(as.4seasons(Y,FUN=FUN,nmin=60),it=it)
[…]

### 2016-05-30 Tanzania/dse.tz.R

@@ -1,0 +1,360 @@
+## Script to downscale temperature from Tanzania.
+## Data from Habiba.
+#!/user/bin/env Rscript
+#args=commandArgs (trailingOnly=TRUE)
+#print(paste('Switch working directory to',args[1]))
+#setwd(args[1])
+
+library(esd)
+setwd('/lustre/storeA/users/rasmusb')
+
+predictands <- function(param='tmin') {
+  ## Prepare the predictand data
+  print("predictands")
+  fname <- paste(param,'.tz.rda',sep='')
+  path <- paste('Habiba/',param,sep='')
+  if (!file.exists(fname)) {
+  files <- list.files(path,pattern='rda',full.names=TRUE)
+  locs <- list.files(path,pattern='rda')
+  dse <- grep('dse.tm',files)
+  files <- files[-dse]; locs <- locs[-dse]
+  print(files)
+  locs <- substr(locs,1,nchar(locs)-4)
+  Y <- NULL
+  for (i in 1:length(files)) {
+    load(files[i])
+    y1 <- eval(parse(text=locs[i]))
+    if (is.null(Y)) Y <- y1 else
+                    Y <- combine.stations(Y,y1)
+  }
+  save(file=fname,Y)
+  plot(Y)
+} else load(fname)
+  print('GHCN')
+  xstations <- switch(param,'tmin'='Habiba/tn.eafrica.rda',
+                      'tmax'='Habiba/tx.eafrica.rda')
+  load(xstations)
+  if (param=='tmin') y <- tn.eafrica else y <- tx.eafrica
+  y <- subset(y,it=c(1950,2015))
+  Y <- combine(Y,y)
+  map(Y)
+  invisible(Y)
+}
+
+downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
+                      period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
+                      lon=c(0,100),lat=c(65,90),eofs=1:6,n=10,rel.cord=FALSE,select=NULL) {
+
+  print('downscale')
+  
+## Use a time and space window:  
+  Y <- subset(Y,it=period)
+  Y <- subset(Y,is=list(lat=lat,lon=lon))
+
[…]

### 2016-06-13 CixPAG/CixPAG-aphrodite.Rmd

--- 2016-04-28 CORDEX/ensembles.Rmd
@@
 ---
-title: "ensembles"
+title: "CixPAG-aphrodites"
 author: "Rasmus Benestad"
-date: "April 27, 2016"
+date: "May 11, 2016"
 output: html_document
@@
 
-## **Simulated global mean temperartue**
+```{r}
+library(esd)
+```
 
-Combine the simulated global mean temperature with information about the model configurations taken from Table 9-a-1 in the IPCC WG 1 2013 report ("AR5):
+## Including Plots
 
 ```{r}
-library('esd')
-data("IPCC.AR5.Table.9.A.1")
-data("global.t2m.gcm")
-gcmnm <- unlist(lapply(attr(global.t2m.gcm$global.t2m.cmip5.rcp45,'meta'),function(x) x$GCM))
-n <- length(gcmnm); m <- 12
-X <- matrix(rep(NA,n*m),n,m)
-
-IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name <- as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name)
-IPCC.AR5.Table.9.A.1$Ocean.Component.Name <- as.character(IPCC.AR5.Table.9.A.1$Ocean.Component.Name)
+odurl <- 'http://iridl.ldeo.columbia.edu/SOURCES/.RIHN/.aphrodite/.V1101/.APHRO_MA_050deg/.precip/dods'
+con <- nc_open(url)
+print.ncdf4(con)
 ```
-
-Index some of the entries which do not contain useful information.
-
-```{r}
-iam <- is.element(substr(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name,1,8),'Included') |
-       is.element(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name,'<NA>')
-IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name[iam] <- IPCC.AR5.Table.9.A.1$Model.Name[iam]
-iom <- is.element(IPCC.AR5.Table.9.A.1$Ocean.Component.Name,'Included')
-IPCC.AR5.Table.9.A.1$Ocean.Component.Name[iom] <- IPCC.AR5.Table.9.A.1$Model.Name[iom]
-```
-
-There are many different similar settings with different names.
-Try to organise and sort the different types under fewer and simpler categories.
-Here is a mix of conventions and ways to refer to the resolution.
-
-```{r}
-fixagcmnm <- grep('Included',as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name))
-IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name[fixagcmnm] <- IPCC.AR5.Table.9.A.1$Model.Name[fixagcmnm]
-```
-
-```{r}
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- as.character(IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub(' ','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('-','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub(',','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('longitude','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('latitude','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('Nominally','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('equivalentto','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('FiniteVolume','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('Averagedcellsize','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('in','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('and','',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- gsub('TL','T',IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-Tres<- c('T42','T31','T159','T63','T159','T959','T213','T319','N48','N49','C180','C360','R42','N96')
-for (ires in Tres) {
-  ii <- grep(ires,IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-  IPCC.AR5.Table.9.A.1$Horizontal.Grid[ii] <- ires
-}
-IPCC.AR5.Table.9.A.1$Horizontal.Grid <- as.factor(IPCC.AR5.Table.9.A.1$Horizontal.Grid)
-```
-
-Fix the GCM names:
-
-```{r}
-IPCC.AR5.Table.9.A.1$Model.Name <- as.character(IPCC.AR5.Table.9.A.1$Model.Name)
-IPCC.AR5.Table.9.A.1$Model.Name <- gsub('-','.',IPCC.AR5.Table.9.A.1$Model.Name)
-IPCC.AR5.Table.9.A.1$Model.Name <- gsub('_','.',IPCC.AR5.Table.9.A.1$Model.Name)
-IPCC.AR5.Table.9.A.1$Model.Name <- gsub(' v1.0','',IPCC.AR5.Table.9.A.1$Model.Name)
-IPCC.AR5.Table.9.A.1$Model.Name <- gsub('(m)','.m',IPCC.AR5.Table.9.A.1$Model.Name,perl=TRUE,fixed=TRUE)
-cesm1 <- grep('CESM1',IPCC.AR5.Table.9.A.1$Model.Name)
-IPCC.AR5.Table.9.A.1$Model.Name[cesm1] <- 'CESM1'
-IPCC.AR5.Table.9.A.1$Model.Name <- as.factor(IPCC.AR5.Table.9.A.1$Model.Name)
-gcmnm <- gsub('-','.',gcmnm)
-gcmnm <- gsub('_','.',gcmnm)
-cesm1 <- grep('CESM1',gcmnm)
-gcmnm[cesm1] <- 'CESM1'
-cnrm5 <- grep('CNRM.CM5',gcmnm)
-gcmnm[cnrm5] <- 'CNRM.CM51'
-inmcm4 <- grep('inmcm4',gcmnm)
-gcmnm[inmcm4] <- 'INM.CM4'
-```
-
-Also need to tidy up some of the description concerning the ocean models
-
-```{r}
-IPCC.AR5.Table.9.A.1$z.Co.ord <- as.character(IPCC.AR5.Table.9.A.1$z.Co.ord)
-IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('-coordinate','',IPCC.AR5.Table.9.A.1$z.Co.ord)
-IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('hybrid','',IPCC.AR5.Table.9.A.1$z.Co.ord)
-IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('sigma-z','sigma',IPCC.AR5.Table.9.A.1$z.Co.ord)
-IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub(' ','',IPCC.AR5.Table.9.A.1$z.Co.ord)
-IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('Isopycnic','isopycnic',IPCC.AR5.Table.9.A.1$z.Co.ord)
-IPCC.AR5.Table.9.A.1$z.Co.ord <- gsub('zisopycnic','isopycnic',IPCC.AR5.Table.9.A.1$z.Co.ord)
-IPCC.AR5.Table.9.A.1$z.Co.ord <- as.factor(IPCC.AR5.Table.9.A.1$z.Co.ord)
-```
-
-Make a data.frame that organises the model information for each simulation.
-
-```{r, echo=FALSE}
-for (i in 1:n) {
-  ii <- is.element(toupper(IPCC.AR5.Table.9.A.1$Model.Name),toupper(gcmnm[i]))
-  if (sum(ii)>0) {
-    x <- c(as.character(IPCC.AR5.Table.9.A.1$Atmosphere.Component.Name)[ii][1],
-               as.character(IPCC.AR5.Table.9.A.1$Horizontal.Grid)[ii][1],
-               IPCC.AR5.Table.9.A.1$Number.of.Vert.Levels[ii][1],
-               IPCC.AR5.Table.9.A.1$Grid.Top.hPa[ii][1],
-               as.character(IPCC.AR5.Table.9.A.1$Component.Name.or.type)[ii][1],
-               as.character(IPCC.AR5.Table.9.A.1$Atmos.Chemistry.Component.Name)[ii][1],
-               as.character(IPCC.AR5.Table.9.A.1$Land.Surface.Component.Name)[ii][1],
-               as.character(IPCC.AR5.Table.9.A.1$Ocean.Component.Name)[ii][1],
-               IPCC.AR5.Table.9.A.1$Number.of.Vertical.Levels[ii][1],
-               as.character(IPCC.AR5.Table.9.A.1$z.Co.ord)[ii][1],
-               as.character(IPCC.AR5.Table.9.A.1$Ocean.Biogeochemistry.Component.Name)[ii][1],
-               as.character(IPCC.AR5.Table.9.A.1$Sea.Ice.Component.Name)[ii][1])
-    print(x)
-    X[i,] <- x
-  }  else print(paste('no match for ',gcmnm[i]))
-}
-```
-
-Set the column names
-
-```{r}
-X <- as.data.frame(X)
-colnames(X) <- c('atm.model','atm.grid',
-                 'atm.lev','atm.top',
-                 'atm.type','Atm.chem',
-                 'Land','oce.model',
-                 'oce.lev','oce.z',
-                 'oce.bio','Seaice')
-X$atm.lev <- as.integer(X$atm.lev)
-X$atm.top <- as.numeric(X$atm.top)
-X$oce.lev <- as.integer(X$oce.lev)
-```
-
-Combine the simulated temperature change with corresponding information about the model configuration
-
-```{r}
-dT <- colMeans(window(zoo(global.t2m.gcm[[1]]),start=2070,end=2099))-
-      colMeans(window(zoo(global.t2m.gcm[[1]]),start=1970,end=1999))
-X <- as.data.frame(cbind(round(dT,2),X))
-```
-
-##**Results**
-
-Is the simulated temperature change sensitive to how the atmospheric model is configured?
-
-```{r}
-fit.atm <- lm(dT ~ atm.grid + atm.lev + atm.top + oce.lev + oce.z,data=X)
-print(summary(fit.atm))
-```

### 2016-06-13 ClimaTrans/climatrans.Rmd

@@
 ```{r}
+wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
 season<-5:10
@@
 mu <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetmean')
-q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'quantile',probs=0.95)
+q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wq95')
 ```

### 2016-06-13 EU-Circle/eu-circle_case-study-1.Rmd

@@
 output: html_document
+bibliography: bibliography.bib
 ---
@@
 
-```{r,echo=FALSE,message=FALSE}
+```{r,message=FALSE}
 library(esd)
-ss <- select.station(param='precip',lon=c(5,10),lat=c(42,45),nmin=50)
-map(ss)
 ```
 
-```{r}
+```{r,warning=FALSE}
 if (!file.exists('eu-circle_case-study-1_precio.rda')) {
+  ss <- select.station(param='precip',lon=c(5,10),lat=c(42,45),nmin=50,src=c('ecad','ghcnd'))
   x <- station(ss)
-  x <- subset(X,is=list(alt=-400))
+  x <- subset(x,is=list(alt=-400))
   save(file='eu-circle_case-study-1_precio.rda',x)
 } else load('eu-circle_case-study-1_precio.rda')
+## Only keep the ECAD data
+x <- subset(x,is=is.element(src(x),'ECAD'))
+map(x,FUN='q95',new=TRUE)
 diagnose(x)
@@
 
-```{r}
+```{r,warning=FALSE}
 plot(x,new=FALSE)
+grid()
 ```
@@
 
-```{r}
-mu <- annual(x,'wetmean')
+```{r,warning=FALSE}
+mu <- annual(x,'wetmean',nmin=250)
 wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
@@
 plot(mu,new=FALSE)
+for (i in 1:dim(mu)[2]) lines(trend(subset(mu,is=i)),lty=2)
+grid()
+
 ```
+
+There appears to be a slight increasing trend in $\mu$ for most stations, according to a linear regression analysis against time.
+There are also pronounced year-to-year variations in $\mu$ with high "spikes" in some years.
+
+Need to check to see if $\mu$ is correlated with number of rainy days, e.g. can be high due to smaller samples and higher sampling fluctuations:
+
+```{r}
+x1 <- zoo(subset(annual(x,'count',threshold=1),is=1))
+y1 <- zoo(subset(mu,is=1))
+plot(x1,y1,xlab=expression(n[X>1*mm]),ylab=expression(mu),xlim=c(30,150),ylim=c(5,30),
+     main='Test: wet-day mean precipitation dependency on sample size')
+for (i in 2:dim(mu)[2]) {
+  x1 <- zoo(subset(annual(x,'count',threshold=1),is=i))
+  y1 <- zoo(subset(mu,is=i))
+  points(x1,y1)
+}
+grid()
+```
+
+There is no indication suggesting that high value of annual $\mu$ is connected with a low number of rain events.
 
@@
 
-```{r}
+```{r,warning=FALSE}
 ## check the relationship between wet-day 95-percentile and wet-day mean
-plot(-log(0.05)*zoo(mu[,2]),zoo(q95[,2]),pch=19,xlim=c(0,120),ylim=c(0,120),
+plot(-log(0.05)*zoo(mu[,2]),zoo(q95[,1]),pch=19,xlim=c(0,120),ylim=c(0,120),
      main='Wet-day mean v.s. 95th wet percentile',
      xlab=expression(paste(-ln(1-0.95)*mu,' (mm/day)')), ylab=expression(paste(q[95],' (mm/day)')))
-points(-log(0.05)*zoo(mu[,1]),zoo(q95[,1]),pch=19,col=rgb(1,0,0,0.2))
+points(-log(0.05)*zoo(mu[,1]),zoo(q95[,2]),pch=19,col=rgb(1,0,0,0.2))
 points(-log(0.05)*zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
@@
 grid()
-legend(0,120,loc(climatrans.pr),pch=19,col=c(rgb(1,0,0,0.2),'black',rgb(0,0,1,0.2)),bty='n')
+legend(0,120,loc(mu),pch=19,col=c('black',rgb(1,0,0,0.2),rgb(0,0,1,0.2),rgb(0,1,0,0.2),rgb(0,0.7,0.7,0.2),rgb(0.7,0,0.7,0.2),rgb(0.5,0.5,0.5,0.2)),bty='n',cex=0.5)
 ```
 
-The scatter plot suggests that the points are clustered around the diagonal and hence the 24-hr precipitation amoun is approximately exponentially distributed.
+The scatter plot suggests that the points are mainly clustered around the diagonal and hence the 24-hr precipitation amoun is approximately exponentially distributed.
+The Marseille data seem to be less similar to the exponential distribution than the other sites.
 
-We also need to examine the trend in the wet-day frequency
+We also need to examine the trend in the wet-day frequency:
 
-```{r}
-fw <- annual(x,'wetfreq')
-plot(mu,new=FALSE)
+```{r,warning=FALSE}
+fw <- annual(x,'wetfreq',nmin=250)
+plot(fw,new=FALSE)
+for (i in 1:dim(fw)[2]) lines(trend(subset(fw,is=i)),lty=2)
+grid()
 ```
 
-```{r}
+The rain gauge data suggests decreasing trends in $f_w$ for most locations, except for the highest station (360 m a.s.l.) further inland (Faverges-de-la-Tour) and with a short record.
+
+Estimate the evolution in the probability for heavy precipitation: $Pr(X > x) = f_w e^{-x_0/\mu}$.
+
+```{r,warning=FALSE}
+x0 <- 100 #mm
+Pr <- zoo(1-pbinom(0,size=365,prob=coredata(fw)*exp(-x0/coredata(mu))),order.by=year(fw))
+class(Pr) <- class(mu)
+Pr <- attrcp(mu,Pr)
+attr(Pr,'variable') <- 'Pr'
+attr(Pr,'unit') <- 'probability'
+attr(Pr,'longname') <- paste('probability of at least one day with more than',x0,'mm') 
+plot(Pr,ylab=expression(Pr(n>0)),xlab='',new=FALSE,errorbar=FALSE,
+          main=paste('Probability of at least one day with more than',x0,'mm'))
+for (i in 1:dim(Pr)[2]) lines(trend(subset(Pr,is=i)),lty=2)
+grid()
+```
+
+The results indicate large annual variations in the probability of extreme precipitation ($X > 100mm$), suggesting a high sensitivity to both $\mu$ and $f_w$.
+This sensitivity is also seen in the different estimates for different locations, even if the difference between the mean $\mu$ is not as dramatic.
+
+One relevant parameter is the duration of dry and wet spells (number of dry/wet consecutive days).
+The grapghic below shows the length of dry (red) and wet (blue) intervals (dry are plotted with negative sign for clarity).
+
+```{r,warning=FALSE}
 z <- subset(x,it=c(1900,2016),is=1)
@@
 plot(ncd)
+grid()
 ```
 
-How close is the duration of wet/dry spells to the geometric distribution?
+The dry runs usually last longer than the wet runs.
+some of the longest dry spells have lasted for ~100 days, but it has rarely rained more than 5 days in a row.
+One question is whether the dry/wet spell statistics can be considered to be a result of random/stochastic processes, or if they contain more structure.
 
-```{r}
+If there is only a fixed probability for wet/dry day that detemines the outcome, then we can use the geometric distribution to represent the statistics of the wet/dry spells:  <https://en.wikipedia.org/wiki/Geometric_distribution>.
+The question then is how closely the duration of wet/dry spells follow the geometric distribution.
+The histogram below compares the data to the pdfs for the geometric distribution and suggest a good match.
+
+```{r,warning=FALSE}
 hist(ncd)
@@
 
-The mean value if the parameter that describes the shape of the geometric distribution.
-The comparison suggests that it does not give a bad representation of the observed charactereistics.
-Are there any trends in the annual mean durations?
+The mean value if the parameter that describes the shape of the geometric distribution. The comparison suggests that it does not give a bad representation of the observed charactereistics.
 
-```{r}
+The second question is then: Are there any trends in the annual mean durations?
+
+```{r,warning=FALSE}
 amncd <- annual(ncd)
-plot(amncd,col=c(rgb(0,0,1),rgb(1,0.3,0)))
+plot(amncd,col=c(rgb(0,0,1),rgb(1,0.3,0)),new=FALSE)
 grid()
 lines(trend(subset(amncd,is=1)),lty=2)
-lines(trend(subset(amncd,is=2)),lty=2)
+mcdd <- trend(subset(amncd,is=2))
+lines(mcdd,lty=1)
+print(attr(mcdd,'coefficients'))
 ```
 
-There is a
+There is a slight increasing trend in the annual mean duration of dry spells, which also has a consequence for the probability of long-lasting meteorological droughts.
 
-Represent gridded precipitation with upper-case $X$
+```{r,warning=FALSE}
+plot(1-pgeom(1:100,1/mean(mcdd)),type='l',xlab='consecutive dry days',ylab=expression(Pr(X > x)))
+plot(index(mcdd),1-pgeom(30, 1/mcdd),type='l',xlab='',ylab=expression(Pr(X > 30*days)),
+     main='Estimated probability for 30-day dry spell',ylim=c(0,0.05))
+for (i in 1:dim(Pr)[2]) lines(trend(subset(Pr,is=i)),lty=2)
+grid()
+```
+
+Linear trend analysis applied to both the duration of dry spells (solid line for one site in figure above) and the probability of more than 100 mm rain per day (dashed line showing several sites) both exhibit long-term increases over the last ~100 years.
+This tendency can be interpreted as the risk of long dry periods increases while the amount that falls as rain gets more extreme for the wet days.
+In other words, there has been a growing risk for both droughts and flash floods.
 
@@
 
-Gridded data are not ideal when it comes to extremes, as the gridding process makes the data spatially inhomogeneous. But
+The geographical pattern of trend in dry spell duration can be analysed using gridded data from EOBs (based on the ECA&D station records) Here gridded precipitation is represented with upper-case $X$.
+Gridded data are not ideal when it comes to extremes, as the gridding process makes the data spatially inhomogeneous.
+There are geographical variations in the climatological precipitation, with more precipitation over the higher elevations and the southern Alps.
+
+```{r,warning=FALSE}
+X <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(4,8),lat=c(43,46))
+map(X,new=FALSE)
+```
+
+A similar analysis of the mean dry spell duration indictes that the dry periods typically last longer near the coast.
 
 ```{r}
-X <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(4,8),lat=c(43,46))
-map(X)
+mean.dry <- function(x,na.rm=TRUE) mean(subset(spell(x,threshold=1),is=2),na.rm=na.rm)
+map(X,FUN='mean.dry',new=FALSE,colbar=list(pal='warm'))
 ```
+
+It is also along the coast where there has been a strongest trend in the mean duration of the dry spells, which has an implication for the probability of long dry periods according to the geometric distribution.
+
+```{r}
+trend.dry <- function(x,na.rm=TRUE) trend.coef(subset(spell(x,threshold=1),is=2))
+map(X,FUN='trend.dry',new=FALSE,colbar=list(pal='warm'))
+```
+
+## Temperature
+
+For heat waves, use a similar approach as described in the supporting material of where a GLM was used to estimate the number of events with a basis of the 'Poisson' distribution.

### 2016-06-30 SPECS/specs-decade-mos.Rmd
SPECS decadal

---
title: "SPECS-decadal-MOS"
author: "Rasmus Benestad"
date: "June 22, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supporting material for Using statistical downscaling in decadal predictions

This document provides the source code (R) and the embedded plots for the investigation into the question whether model output statistics (MOS) can improve the skill of decadal forecasts. The term 'decadal' is loosely used for time scales between 1 and 10 years.

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

## The settings:

```{r initialise}
## Script that applies MOS to SPECS decadal forecasts

library(esd)
library(ncdf)
setwd("~/R")

downloadfromunican <- FALSE   # Preparations - get the data
downscaledreanalysis <- FALSE # SM: Assessment of large-small scale link
looexperiments <- FALSE        # Cross-validation MOS on forecasts
diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts

source("~/Dropbox/Public/SPECS/R/udg2esd_met.R") ## Upload udg2esd function

## This variable is needed because CDO did not fix the times properly
## for the predictands (EOBS)...
time = c(364, 729, 1095, 1460, 1825, 2190, 2556, 2921, 3286, 3651, 4017, 4382, 
    4747, 5112, 5478, 5843, 6208, 6573, 6939, 7304, 7669, 8034, 8400, 8765, 
    9130, 9495, 9861, 10226, 10591, 10956, 11322, 11687, 12052, 12417, 12783, 
    13148, 13513, 13878, 14244, 14609, 14974, 15339, 15705, 16070, 16435, 
    16800, 17166, 17531, 17896, 18261, 18627, 18992, 19357, 19722, 20088, 
    20453, 20818, 21183, 21549, 21914, 22279, 22644, 23010, 23375, 23740, 
    23921) ;

```

## Define the functions

Formatting and pre-processing the data

```{r functions}
#--- functions ---

fcst2field <- function(x,it0=1,im=1,verbose=FALSE) {

  if (verbose) print(paste('fcst2field',it0,im))
  lon <- x$xyCoords$x; nx <- length(lon)
  lat <- x$xyCoords$y; ny <- length(lat)
  index <- x$Dates$start; nt <- length(index)

## The data is scrambled with unconvntional x-y-order - generate unscramble index:
  usi <- t(matrix(1:(nx*ny),ny,nx))
  param <- x$Variable$varName
  unit <- attr(x$Variable,'units')
  if (verbose) print(paste(param,unit))

## one field
  df <- matrix(x$Data[it0,im,,,],nt,nx*ny)
  y <- zoo(df[,c(usi)],order.by=as.Date(index))
  Y <- as.field(y,param=param,unit=unit,lon=lon,lat=lat,
              info=paste('member',im ,'initialisation',it0))
  if (verbose) print('---')
  invisible(Y)
}

```

Extracting the correct predictor information from the data

```{r xpredictor}

## Function to extract a subset and organise the predictor data
## for all ensemble members
Xpredictor <- function(x,m=1:3,it=NULL,is=NULL,mask=FALSE,
                       FUN='mean',verbose=FALSE) {
  if (verbose) print('Xpredictor')
  ## x is the data, m is the forecast interval (years), and it
  ## specifies e.g. months
  ## Get the dimensions of the data structure
  d <- dim(x$Data)
  if (verbose) {print(summary(x)); print(d)}
  ## initialisations members times nx ny

  ## Prepare one member per loop
  for (im in 1:d[2]) {
    index <- x$Dates$start; nt <- length(index); jj <- 1
    for (ii in m) {
      if (verbose) print(paste('member',im,'initialisation',ii))
      ## Read the entire field
      y1 <- fcst2field(x,it0=10-ii,im=im,verbose=verbose)
      ## Extract a sub-region
      y1 <- subset(y1,is=is)
      ## If it!=NULL select subset of the year 
      y1 <- aggregate(subset(y1,it=it),year,FUN)
      ## If mask, remove the land component
      if (mask) y1 <- mask(y1,land=TRUE)
      if (verbose) {print(index(y1)); print(index(lag(zoo(y1),k=-ii)))}
      ## Let Y1 contain the end results: sum the different lead times
      if (jj==1)  Y1 <- zoo(y1) else
                  Y1 <- zoo(Y1) + zoo(lag(y1,k=-ii))
      jj <- jj + 1
    }
    ## Divide by the number of initialisations to get the forecast interval mean
    coredata(Y1) <- coredata(Y1)/length(m)
    ## Set the appropriate attributes and class
    if (verbose) {print('Set the appropriate attributes'); print(names(attributes(y1)))}
    Y1 <- attrcp(y1,Y1)
    if (verbose) print(paste('Set the appropriate class',class(y1),collapse='-'))
    class(Y1) <- class(y1)
    if (verbose) print(class(y1))

    ## The index is set by the year and the calendar month/day reflects
    ## member number
    if (d[2]<=12) index(Y1) <- as.Date(paste(year(Y1),im,'01',sep='-'),
                                       origin = as.Date("1950-01-01")) else
                  index(Y1) <- as.Date(paste(year(Y1),'01-01',sep='-')+im,
                                       origin = as.Date("1950-01-01"))
    if (verbose) print(index(Y1))

    ## Outside the loop for each member: combine the different members
    if (im==1) Y <- zoo(Y1) else
               Y <- c(zoo(Y),zoo(Y1))
  }

  Y <- attrcp(y1,Y,ignore='dimnames')
  attr(Y,'dimnames') <- NULL
  ## Need to subsample every m year
  it <- is.element(year(Y),seq(min(year(Y)),max(year(Y)),by=max(m)))
  Y <- subset(Y,it=it)
  class(Y) <- class(y1)
  if (verbose) {
    print(paste(max(m),'-year intervals for ',d[2],' ensemble members',sep=''))
    print(index(Y))
  }
  attr(Y,'longitude') <- lon(y1)
  attr(Y,'latitude') <- lat(y1)
  attr(Y,'dimensions') <- c(length(lon(y1)),length(lat(y1)),length(index(Y)))
  ## Keep track of how the members are represented
  attr(Y,'n_members') <- d[2]
  if (d[2]<=12) attr(Y,'member_by') <- 'month' else
                attr(Y,'member_by') <- 'day'
  ## The end-result contains a field object with
  if (verbose) print('Xpredictor successfully completed!')
  invisible(Y)
}

```

Aggregate daily data from the EOBS data, using CDO (because the data size is too big for R).

```{r aggregate}
aggregatedailyfield <- function(fname,cline='cdo monmean',
                                output='mu_0.25deg_reg_v12.0.nc') {
  print('This function only works if you have CDO installed')
  print(paste(cline,fname,output))
  system(paste(cline,fname,output))
  x <- retrieve(output)
  invisible(x)
}

```

Estimating the wet-day mean from the EOBS data using CDO (because the data size is too big for R).

```{r wetdaymean}
wetdaymean <- function(ifile,cline='cdo yearmean',fname='data/wetdaymean.nc') {
  print('calculate mask')
  system(paste('cdo gtc,1',ifile,'data/mask.nc'))
  ## her blir ofile en mask med 1 i alle tidsskritt med data over konstanten (satt til 0),
  ## og verdi 0 i alle andre
  print('apply mask')
  system(paste('cdo ifthen data/mask.nc',ifile, fname))
  ## her appliseres masken på ifile og lagres i ofile. (Øvrige punkter får verdien NA.)

  ##Disse går å kombinere til en rad:
  ##  print(paste('cdo ifthen -gtc,0',ifile,ifile,fname))
  ##  system(paste('cdo ifthen -gtc,0',ifile,ifile,fname))
  X <- aggregatedailyfield(fname=fname,cline=cline)
  Y <- aggregatedailyfield(fname=,'data/mask.nc',cline=cline,output='fw_0.25deg_reg_v12.0.nc')
  file.remove(fname); file.remove('data/mask.nc')
}

```

function setting up the experiment:

```{r D53exp}
## LOO experimtent: X is the predictor and Y is the predictand
D53Exp <- function(X,Y,m=c(1,3),it=NULL,is=NULL,mask=FALSE,verbose=TRUE) {
  if (verbose) print('D53Exp: Prepare predictor')
  ## Prepare the predictor:
  X <- Xpredictor(X,m=m,it=it,is=is,mask=mask,verbose=verbose)
  ## Prepare the predictand: repeat the same precitand to match the
  ## ensemble members - give same time stamp/date:
  if (verbose) print(paste('Predictand',varid(Y)))
  n <- attr(X,'n_members')
  Y0 <- Y

  ## Aggregate the predictand over the forecast interval as done in Xpredictor
  if (verbose) print(paste('Aggregate over years ',m,collapse='-'))
  for (j in 1:length(m)) {
    if (j==1) y <- zoo(Y) else
              y <- y + lag(zoo(Y),k=-m[j])
  }
  coredata(y) <- coredata(y)/length(m)
  if (verbose) {print('dim after aggragation');print(dim(Y0)); print(dim(y))}

  ## repeat the observations with corrsponding time stamps as the ensemble members of the forecast
  if (verbose) print('repeat predictand')
  for (i in 1:n) {
    if (verbose) print(paste('predictand',i))
    if (attr(X,'member_by')=='month')
      index(y) <- as.Date(paste(year(y),i,'01',sep='-'),
                          origin = as.Date("1950-01-01")) else
      index(y) <- as.Date(paste(year(y),'01-01',sep='-')+i,
                          origin = as.Date("1950-01-01"))
    if (verbose) print(index(y))
    if (i==1) Y <- y else Y <- c(Y,y)
  }
  ## Copy all attributes etc.
  Y <- attrcp(Y0,Y)
  class(Y) <- c('field','day','zoo')
  attr(Y,'dimensions') <- c(length(lon(Y0)),length(lat(Y0)),length(index(Y)))
  if (verbose) print('predictand sucessfully processed')
  ## The experiment
  xeof <- EOF(mask(X,land=TRUE))
  yrs <- year(X)[is.element(year(X),year(Y))]
  yrs <- yrs[!duplicated(yrs)]
  if (verbose) {print(yrs); print('LOO x-val...')}
  ## LOO
  for (it in 1:length(yrs)) {
    ## set index to exclude all members for the same evaluation interval
    iloo <- !is.element(year(Y),yrs[it])
    yloo <- subset(Y,it=iloo)
    if (verbose) {
      print(paste('LOO: it=',it,'years:'))
      print(table(year(yloo)))
    }
    class(yloo) <- class(X)
    yeof <- EOF(yloo,n=4)
    ## Some hacking needed to ensure that subset doesn't change the index type to year.
    class(yeof)[3] <- 'day';  class(xeof)[3] <- 'day';
    ds.loo <- DS(yeof,xeof,verbose=verbose)
    ## Extract the forecast for the evaluation interval:
    if (verbose) print('reconstruct field')
    jloo <- is.element(year(xeof),yrs[it])
    xval <- subset(as.field(predict(ds.loo,newdata=xeof)),it=jloo)
    ## combine each LOO result into one series:
    if (it==1) XVAL <- xval else
               XVAL <- c(XVAL,xval)
  }
  if (verbose) print('LOO completed')
  XVAL <- attrcp(Y0,XVAL); class(XVAL) <- class(Y)
  attr(XVAL,'dimensions') <- c(attr(Y,'dimensions')[1:2],length(index(XVAL)))
  ## Plot the correlation map between LOO results and original:
  rmap <- corfield(aggregate(XVAL,year,'mean'),aggregate(Y,year,'mean'),plot=FALSE)
  map(rmap,colbar=list(breaks=seq(-1,1,by=0.1),pal='t2m',rev=TRUE),type='fill',new=FALSE)
  invisible(XVAL)
}

```

Extract the {\em m}-year annual mean MOS-based forecast.

```{r mannualfcst}
mannualfcst <- function(x,l=1,it=NULL,is=NULL,mask=FALSE,verbose=FALSE) {
  ## A function that extract l-year aggregated data from the forecasts
  ## It uses Xpredictor to represent each ensemble member as a different
  ## calendar month, and needs to seperate these months and aggregate them
  ## seperately before synthesising a new corresponding field object with
  ## multi-year aggregated values.

  X <- Xpredictor(x,m=l,it=it,is=is,mask=mask,verbose=verbose)
  n <- attr(X,'n_members')
  if (attr(X,'member_by')=='month') {
    if (verbose) print('separate ensemble members by calendar month')
    m <- as.integer(rownames(table(month(X))))
    j <- 1
    for (i in m) {
      y <- pentad(subset(X,it=month.abb[i]),l=l)
      index(y) <- as.Date(paste(year(y),i,'01',sep='-'))
      if (j==1) Y <- zoo(y) else
                Y <- c(Y,zoo(y))
      j <- j+1
    }
  }
  Y <- attrcp(X,Y)
  class(Y) <- class(X)
  attr(Y,'dimnames') <- NULL
  invisible(Y)
}

```

Compare the {\em m}-year annual mean MOS-based forecast with corresponding EOBS statistics

```{r compare}
comparemannual <- function(tas_DECA,reanalysis='air.mon.mean.nc',mask=TRUE,l= 3,eofs=NULL) {

  X <- mannualfcst(tas_DECA,l=l,mask=mask,verbose=TRUE)
  Y <- pentad(retrieve(reanalysis,lon=range(lon(X)),lat=range(lat(X))),l=l)

  if (is.T(X)) varid <- 'tas' else {
    varid <- 'psl'
    coredata(Y) <- 100*coredata(Y)
    attr(X,'unit') <- 'Pa'
    attr(Y,'unit') <- 'Pa'
  }
  if (mask) land <- '' else land <- '-nomask'

## Compare the multi-annual scale ensemble mean forecast with the NCEP/NCAL reanalysis
  Xm <- aggregate(X,year,'mean')
  index(Xm) <- year(Xm)
  class(Y) <- class(Xm)

  ## Apply an EOF-filter to emphasise only the large scale anomalies:
  if (!is.null(eofs)) {
    print(paste('Only keep the large-scale structures; EOFs:',paste(eofs,collapse=' ,')))
    Xm.eofs <- subset(EOF(Xm),pattern=eofs)
    Xm <- eof2field(Xm.eofs)
    land <- paste(land,'eof_',paste(eofs,collapse='+'))
  }

  rmap <- corfield(Xm,Y,new=FALSE)
  if (l==1) figlab(paste(l,'year'),ypos=0.999) else figlab(paste(l,'years'),ypos=0.999)

## Generate a common EOF

  if (mask) Z <- mask(Y,land=TRUE) else Z <- Y
  Z <- subset(Z,it=range(year(X)))
  for (i in 1:10) {
    z <- subset(X,it=month.abb[i])
    index(z) <- year(z)
    Z <- combine(Z,z)
  }
  ceof <- EOF(Z)

  plot(ceof,type='fill',new=FALSE)
  figlab(paste(l,'years'),ypos=0.999)
  invisible(rmap)
## Generate a common EOF
}

```

Get the diagnostics of the MOS forecasts

```{r diagnose}
diagnoseMOS <- function(results,y,l=1) {
  print(paste('diagnoseMOS',results))
  load(results)  
  eval(parse(text=paste('zmos <-',substr(results,14,nchar(results)-4))))
  attr(zmos,'variable') <- varid(y)
  print(range(year(zmos)))
  y <- pentad(y,it0=min(year(zmos)),l=l)
  print(index(y))
  print(index(zmos))
  print('ensemble mean: aggregate(.)')
  ## Different ensemble emmbers are represented as calendar months
  Z <- aggregate(zmos,year,'mean')
  index(Z) <- year(Z)
  print('correlation: ensemble mean FCST & reanalysis')
  r <- corfield(Z,y,plot=FALSE)
  map(r,colbar=list(breaks=seq(-1,1,by=0.1),pal='t2m',rev=TRUE),type='fill',new=FALSE)
  figlab('EOBS+ensemble mean MOS',ypos=0.999)

  print('Common EOFs')
  y <- subset(y,it=range(year(Z)))
  ceof <- EOF(combine(y,Z))

  plot(ceof,type='fill',new=FALSE)
  figlab('EOBS+MOS',xpos=0.8,ypos=0.999)
  figlab('ensemble mean MOS (red) + observed (black)')

  plot(diagnose(ceof),new=FALSE)
  figlab('EOBS+MOS')

  print('return correlation statistics')
  invisible(quantile(c(r),na.rm=TRUE))
}

```

Check the space-time scale dependency.

```{r scales}
scaledependency <- function(x) {
  print('scaledependency')
  ## compare time and spatial scales
  stopifnot(inherits(x,"field"))
  d <- attr(x,'dimensions')
  ## convert zoo's 2D to 3D data
  y <- t(x); dim(y) <- d
  nd <- trunc(min(d[1:2])/2)
  scales <- matrix(rep(nd*trunc(d[3]/2)),nd,trunc(d[3]/2))
  timeseries <- matrix(rep(nd*d[3]),nd,d[3])
  print(dim(scales))
  for (i in 1:nd) {
    ix <- nd + -i:i; iy <- nd + -i:i
    ix[ix < 1] <- 1; ix[ix > d[1]] <- d[1]
    iy[iy < 1] <- 1; iy[iy > d[2]] <- d[2]
    print(c(i,range(ix),range(iy)))
    z <- coredata(y[ix,iy,])
    dim(z) <- c((1+2*i)^2,d[3])
    z <- colMeans(z)
    f <- spectrum(z,method="pgram",plot=FALSE)$spec
    ## Save smoothed filter and 
    #scales[i,] <- filter(rev(f[1:trunc(d[3]/2)]),rep(1,3)/3)
    scales[i,] <- rev(f[1:trunc(d[3]/2)])
    timeseries[i,] <- z
  }
  attr(scales,'timescales') <- rev(1/(12*spectrum(z,plot=FALSE)$freq[1:trunc(d[3]/2)]))
  attr(scales,'unit_time') <- 'year'
  attr(scales,'gridboxes') <- (1+2*(1:nd))^2
  attr(scales,'unit_gridboxes') <- 'number_of_boxes'
  attr(scales,'timeseries') <- zoo(t(timeseries),order.by=index(x))
  image(attr(scales,'timescales'),attr(scales,'gridboxes'),t(scales))
  invisible(t(scales))
}

##------------------------------------------------------------------------
```

## Implementation

```{r implement}
## Get the predictand: the gridded wet-day mean preciptation from the EOBS data

## Get the data:

if (downloadfromunican) {
  print('Get data from oceano.macc.unican.es')
  datadir <- 'SPECS-decadal'
  if (!file.exists('SPECS-decadal')) {
    dir.create('SPECS-decadal')
    system(paste('rsync -rtuv --progress ',datadir,'lpcaron@oceano.macc.unican.es:~/'))
  }
}

print('Get the predictands: mu & fw from EOBS')
if (!file.exists('mu_0.25deg_reg_v12.0.nc'))
  mu.eobs <-  wetdaymean('data.ECAD/rr_0.25deg_reg_v12.0.nc') else
  mu.eobs <- retrieve('mu_0.25deg_reg_v12.0.nc')
fw.eobs <- retrieve('fw_0.25deg_reg_v12.0.nc')
attr(mu.eobs,'variable') <- 'mu'
attr(fw.eobs,'variable') <- 'fw'
attr(fw.eobs,'unit') <- 'fraction'
index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))

```

```{r ds}

## Display the downscaled reanalysis data for assessing link between large and small scales  - in the shape of EOFs
if (downscaledreanalysis) {
  mu.eof <- EOF(mu.eobs)
  plot(mu.eof,colbar=list(pal='t2m',breaks=seq(-0.02,0.02,by=0.001),type='fill',rev=TRUE),new=FALSE)
  figlab('EOBS',ypos=0.999)

## The wet-day frequency:

  fw.eof <- EOF(fw.eobs)
  plot(fw.eof,colbar=list(pal='t2m',breaks=seq(-0.01,0.015,by=0.001),type='fill',rev=TRUE),new=FALSE)
  figlab('EOBS',ypos=0.999)

## Test to examine the link between large-scale temperature and wet-day mean precip
## Derive the saturation vapour pressure from temperature
## retrieve did not get the right time stamp:
  es <- annual(C.C.eq(retrieve('air.mon.mean.nc',lon=c(-60,40),lat=c(20,70))))
  es.eof <- EOF(mask(es,land=TRUE))

## Also get the mean sea-level pressure for the wet-day frequency 
  slp <- annual(retrieve('slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
  slp.eof <- EOF(slp)

## Additional diagnostics: canonical correlation analysis:
  cca <- CCA(es.eof,slp.eof)
  plot(cca,colbar1=list(breaks=seq(-300,300,by=10)),colbar2=list(breaks=seq(-20,20,by=1)),new=FALSE)
  figlab(expression(paste('NCEP/NCAR: ',e[s](T),'-slp')))

  class(mu.eof) <- class(es.eof)
  mu.ds <- DS(mu.eof,es.eof)
  class(fw.eof) <- class(slp.eof)
  fw.ds <- DS(fw.eof,slp.eof)

  plot(mu.ds,colbar1=list(breaks=seq(-0.025,0.025,by=0.005),rev=TRUE),
       colbar2=list(breaks=seq(-350,350,by=25)),type='fill',new=FALSE,)
  figlab('ESD(NCEP/NCAR)',ypos=0.999)

  plot(fw.ds,colbar1=list(breaks=seq(0.0,0.02,by=0.001),rev=TRUE),
       colbar2=list(breaks=seq(-20,20,by=2.5)),type='fill',new=FALSE)
  figlab('ESD(NCEP/NCAR)',ypos=0.999)
}

```

```{r data}

## The predictor: Convert from UDG to field.object:
## psl_DECA
## tas_DECA
##slp <- udg2esd(psl_DECA)
##t2m <- udg2esd(tas_DECA)
##> summary(psl_DECA)
##                    Length   Class  Mode     
##Variable                   2 -none- list     
##Data                79704000 -none- numeric  
##xyCoords                   2 -none- list     
##Dates                      2 -none- list     
##InitializationDates        9 -none- list     
##Members                   10 -none- character
##
##> dim(psl_DECA$Data)
## [1]   9  10 480  41  45

print('Get the decadal predictors saved locally in "data/gfdl_DECA.rda"')
load('data/gfdl_DECA.rda')

```

Leave-one-out cross-validation experiments

```{r loo}

if (looexperiments) {
## Carry out a LOO cross-validation experiment
  print('Check results for LOO experiment - if absent do it')
  if (!file.exists('specsdecadal.z.fw.y1.rda')) {
    print('do specsdecadal.z.fw.y1.rda')
    z.fw.y1 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.rda',z.fw.y1)
    rm("z.fw.y1"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.rda')) {
    print('do specsdecadal.z.mu.y1.rda')
    z.mu.y1 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.rda',z.mu.y1)
    rm("z.mu.y1"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.3.rda')) {
    print('do specsdecadal.z.fw.y1.3.rda')
    z.fw.y1.3 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:3,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.3.rda',z.fw.y1.3)
    rm("z.fw.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.3.rda')) {
    print('do specsdecadal.z.mu.y1.3.rda')
    z.mu.y1.3 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:3,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.3.rda',z.mu.y1.3)
    rm("z.mu.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.5.rda')) {
    print('do specsdecadal.z.fw.y1.5.rda')
    z.fw.y1.5 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:5,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.5.rda',z.fw.y1.5)
    rm("z.fw.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.5.rda')) {
    print('do specsdecadal.z.mu.y1.5.rda')
    z.mu.y1.5 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:5,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.5.rda',z.mu.y1.5)
    rm("z.mu.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.9.rda')) {
    print('do specsdecadal.z.fw.y1.9.rda')
     z.fw.y1.9 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:9,it=NULL,is=NULL,mask=FALSE)
     save(file='specsdecadal.z.fw.y1.9.rda',z.fw.y1.9)
     rm("z.fw.y1.9"); gc(reset=TRUE)
   }
  if (!file.exists('specsdecadal.z.mu.y1.9.rda')) {
    print('do specsdecadal.z.mu.y1.9.rda')
    z.mu.y1.9 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:9,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.9.rda',z.mu.y1.9)
    rm("z.mu.y1.9"); gc(reset=TRUE)
  }

## Diagnostics:
  print('diagnose the MOS results')
  load('specsdecadal.z.mu.y1.3.rda')
  it0 <- range(year(z.mu.y1.3))
  print(paste('time period',paste(it0,collapse='-')))
  score.mu.y1 <- diagnoseMOS('specsdecadal.z.mu.y1.rda',mu.eobs)
  score.fw.y1 <- diagnoseMOS('specsdecadal.z.fw.y1.rda',fw.eobs)
  score.mu.y1.3 <- diagnoseMOS('specsdecadal.z.mu.y1.3.rda',mu.eobs,l=3)
  score.fw.y1.3 <- diagnoseMOS('specsdecadal.z.fw.y1.3.rda',fw.eobs,l=3)
  score.mu.y1.5 <- diagnoseMOS('specsdecadal.z.mu.y1.5.rda',mu.eobs,l=5)
  score.fw.y1.5 <- diagnoseMOS('specsdecadal.z.fw.y1.5.rda',fw.eobs,l=5)
  score.mu.y1.9 <- diagnoseMOS('specsdecadal.z.mu.y1.9.rda',mu.eobs,l=9)
  score.fw.y1.9 <- diagnoseMOS('specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)
  while (dev.cur()>1) dev.off()
## Repeat and make a table of the results...

}

```

Diagnose the forecasts

```{r diagnosefcst}
## Common EOFs to diagnose the results?

if (diagnoseFCST) {
  comparemannual(tas_DECA,mask=TRUE,l= 1)
  comparemannual(tas_DECA,mask=TRUE,l= 2)
  comparemannual(tas_DECA,mask=TRUE,l= 4)
  comparemannual(tas_DECA,mask=TRUE,l= 5)
  comparemannual(tas_DECA,mask=TRUE,l= 6)
  comparemannual(tas_DECA,mask=TRUE,l= 7)
  comparemannual(tas_DECA,mask=TRUE,l= 8)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 1)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 2)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 3)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 4)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 5)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 6)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 7)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 8)
  rmap.1.10.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:10)
  rmap.1.10.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:10)
}
```

```{r rmaps}
  ## Test the forecasts of the larg-scale anomalies/gravest EOFs
  rmap.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3)
  rmap.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3)
  rmap.1.3.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:3)
  rmap.1.3.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:3)
  rmap.1.5.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:5)
  rmap.1.2.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:2)
  rmap.1.9.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:9)
```

```{r rdiff}
  rdiff <- rmap.1.3.nomask.3 - rmap.nomask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-3 vs full field',ypos=0.99)

  rdiff <- rmap.1.3.mask.3 - rmap.mask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-3 vs full field',ypos=0.99)

  rdiff <- rmap.1.5.nomask.3 - rmap.nomask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-5 vs full field',ypos=0.99)

  rdiff <- rmap.1.2.nomask.3 - rmap.nomask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-2 vs full field',ypos=0.99)

  rdiff <- rmap.1.9.nomask.3 - rmap.nomask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-9 vs full field',ypos=0.99)
```

```{r t2m scales}
## Scaling dependency: time - space
t2m <- retrieve('~/data/ncep/air.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
t2m <- anomaly(t2m)
zst <- scaledependency(t2m)
#zst[,1] <- NA
d <- dim(zst)

image(attr(zst,'timescales'),attr(zst,'gridboxes'),t(t(zst)/colMeans(zst)),
      main='Surface temperature: spectral power',
      xlab='Temporal scale (years)',ylab='spatial scale (grid boxes)')
contour(attr(zst,'timescales'),attr(zst,'gridboxes'),zst,add=TRUE)

plot(attr(zst,'timeseries'),plot.type='single',col=colscal(20),main='T(2m)')

```

## Repeat for mean sea-level pressure (SLP)

```{r slp}

## Repeat for SLP
slp <- retrieve('~/data/ncep/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
slp <- anomaly(slp)
yst <- scaledependency(slp)
image(attr(yst,'timescales'),attr(yst,'gridboxes'),t(t(yst)/colMeans(yst)),
      main='Mean sea-level pressure: spectral power',
      xlab='Temporal scale (years)',ylab='spatial scale (grid boxes)')
contour(attr(yst,'timescales'),attr(yst,'gridboxes'),yst,add=TRUE)

plot(attr(yst,'timeseries'),plot.type='single',col=colscal(20),main='SLP')

## Distributions? Ensemble mean and spread -> dnorm()?

## Scores: SPECS libraries?

## The dependency of number of heavy precip event on frequency/sample size

ns <- trunc(10*1.1^seq(1,100,by=1))
nx <- rep(NA,length(ns))
for (i in 1:length(ns)) nx[i] <- sum(rexp(ns[i],rate=1/7) > 50)
plot(ns,nx,main='Number of extreme cases (x>50)',
     xlab='sample size',ylab=expression(n[x>50]),
     sub=expression(paste('exponential distribution: ',mu,'=7')))

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2016-06-30 SPECS/specs_decadal-exp-d52.3_esd-loocv.R

@@ -1,0 +1,643 @@
+## Script that applies MOS to SPECS decadal forecasts
+
+library(esd)
+library(ncdf)
+
+downloadfromunican <- FALSE   # Preparations - get the data
+downscaledreanalysis <- FALSE # SM: Assessment of large-small scale link
+looexperiments <- FALSE        # Cross-validation MOS on forecasts
+diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts
+
+source("~/Dropbox/Public/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
+
+## This variable is needed because CDO did not fix the times properly
+## for the predictands (EOBS)...
+time = c(364, 729, 1095, 1460, 1825, 2190, 2556, 2921, 3286, 3651, 4017, 4382, 
+    4747, 5112, 5478, 5843, 6208, 6573, 6939, 7304, 7669, 8034, 8400, 8765, 
+    9130, 9495, 9861, 10226, 10591, 10956, 11322, 11687, 12052, 12417, 12783, 
+    13148, 13513, 13878, 14244, 14609, 14974, 15339, 15705, 16070, 16435, 
+    16800, 17166, 17531, 17896, 18261, 18627, 18992, 19357, 19722, 20088, 
+    20453, 20818, 21183, 21549, 21914, 22279, 22644, 23010, 23375, 23740, 
+    23921) ;
+
+
+
+#--- functions ---
+
+fcst2field <- function(x,it0=1,im=1,verbose=FALSE) {
+
+  if (verbose) print(paste('fcst2field',it0,im))
+  lon <- x$xyCoords$x; nx <- length(lon)
+  lat <- x$xyCoords$y; ny <- length(lat)
+  index <- x$Dates$start; nt <- length(index)
+
+## The data is scrambled with unconvntional x-y-order - generate unscramble index:
+  usi <- t(matrix(1:(nx*ny),ny,nx))
+  param <- x$Variable$varName
+  unit <- attr(x$Variable,'units')
+  if (verbose) print(paste(param,unit))
+
+## one field
+  df <- matrix(x$Data[it0,im,,,],nt,nx*ny)
+  y <- zoo(df[,c(usi)],order.by=as.Date(index))
+  Y <- as.field(y,param=param,unit=unit,lon=lon,lat=lat,
+              info=paste('member',im ,'initialisation',it0))
+  if (verbose) print('---')
+  invisible(Y)
+}
+
+
[…]

### 2016-07-29 ClimaTrans/climatrans.Rmd
ClimaTrans added material

@@
 ```
+
+Below is a listing of the function used to extract the raw precipitation results averaged over the Indian sub-continent.
+
+```{r}
+RMSE <- function(x,y) sqrt(sum((x-y)^2))/length(x)
+
+precipcmip <- function(predictor='data/ERAINT/era-int-precip-mon.nc',
+                    path="CMIP5.monthly/",rcp='rcp45',pattern="pr_Amon_ens_",
+                    type='ncdf',it=c(1950,2015),plot=TRUE,verbose=FALSE,
+                    lon=c(70,90),lat=c(7,30),select=NULL,xlim=NULL,fname=NULL) {
+
+  if (is.null(fname)) fname <- paste('precipcmip.',paste(lon,lat,sep='.',collapse='-'),'.rda',sep='')
+
+  print(paste('precipcmip: results (will be) stored in',fname))
+
+  if (file.exists(fname)) {
+    load(fname)
+    return(results)
+  }
+
+  ## Diagnose the common EOFs and effect of bias-adjustment
+
+  if (verbose) print(paste('getcmip, it=',it))
+
+  if (verbose) print('predictor')
+  if (is.character(predictor))
+    ## Unit: "mm"
+    rea <- retrieve(ncfile=predictor,lon=lon,lat=lat,
+                    type=type,verbose=verbose) else
+  if (inherits(predictor,'field'))
+    rea <- subset(predictor,is=list(lon=lon,lat=lat))
+  if (!is.null(it)) {
+    rea <- subset(rea,it=it)
+    it <- c(min(year(rea)),max(year(rea)))
+  }
+
+  ## The reanalysis
+  if (verbose) print('aggregate the reanalysis')
+  Rea <- aggregate.area(rea,FUN='mean')
+  X <- aggregate(Rea,month,FUN='mean')
+  Xt <- annual(Rea,FUN='sum')
+
+  if (plot) plot(X,lwd=3,col='black',main='Climatology',ylim=c(0,2*round(max(X))),
+                 ylab=ylab(X),xlab='Calendar month',map.type='rectangle')
+
+  # Ensemble GCMs
+  path <- file.path(path,rcp,fsep = .Platform$file.sep)
+  ncfiles <- list.files(path=path,pattern=pattern,full.name=TRUE)
+  N <- length(ncfiles)
+
+  if (is.null(select)) select <- 1:N else
+                       N <- length(select)
+  if (verbose) {print('GCMs:'); print(path); print(ncfiles[select])}
+
+  Z <- matrix(rep(NA,N*12),N,12)
+  Zt <- matrix(rep(NA,N*201),N,201)
+  gcmnm <- rep("",N); rmse <- rep(NA,N); r <- rep(NA,N)
+
+  ## Set up a list environment to keep all the results
+  if (verbose) print("loop...") 
+  for (i in 1:N) {
+    if (verbose) print(ncfiles[select[i]])
+
+    ## unit: "kg m-2 s-1"
+    gcm <- retrieve(ncfile = ncfiles[select[i]],type=type,
+                          lon=range(lon(rea)),
+                          lat=range(lat(rea)),verbose=verbose)
+    GCM <- aggregate.area(gcm,FUN='mean')
+    ## Annual total precipitation
+    zt <- annual(GCM,FUN='sum')
+    i1 <- is.element(year(zt),1900:2100)
+    i2 <- is.element(1900:2100,year(zt))
+    Zt[i,i2] <- coredata(zt)[i1]
+
+    if (!is.null(it)) {
+      if (verbose) print('Extract some months ot a time period')
+      if (verbose) print(it)
+      gcm <- subset(gcm,it=it)
+    }
+    ## The mean annual cycle
+    ## The units of the GCMs are mm/day - multiply by 30
+    z <- 30*aggregate(GCM,month,FUN='mean')
+    Z[i,] <- coredata(z)
+    #gcmnm[i] <- attr(gcm,'model_id')
+    gcmnm.i <- paste(attr(gcm,'model_id'),attr(gcm,'realization'),sep="-r")
+    gcmnm.i <- gsub(' ','',gcmnm.i)
+    gcmnm.i <- gsub('-','.',gcmnm.i)
+    rmse[i] <- round(RMSE(coredata(X),coredata(z)),2)
+    r[i] <- round(cor(coredata(X),coredata(z)),2)
+    gcmnm[i] <- gcmnm.i
+    print(paste(gcmnm.i,unit(GCM),unit(Rea),rmse[i],r[i]))
+    if (plot) lines(z,col=rgb(i/N,0,0.5+0.5*r[i],0.1),lwd=3)
+  }
+  attr(Z,'GCMs') <- gcmnm
+   if (plot) {
+     lines(X,col='black',lwd=3)
+     grid()
+     figlab(paste(start(Rea),end(Rea)))
+   }
+  results <- list(reanalysis=X,GCMs=z,
+                  obs=Xt,gcm=zoo(t(Zt),order.by=1900:2100),rmse=rmse,cor=r)
+  dev.copy2pdf(file=paste('precipcmip.',
+                 paste(lon,lat,sep='.',collapse='-'),'.pdf',sep=''))
+  save(file=fname,results)
+  return(results)
+}
+```
+
 ##*Precipitation*
+
+We look at various aspects of the precipitation in order to get an idea about what is happening over the sub-continent of India and at the three Indian megacities.
+First, we present features from the past, such as the mean annual cycle, historical long-term trends, and statistical distribution for different aspects connected to precipitation: wet-day frequency and the wet-day mean (precipitation intencity).
 
@@
 
-## Analysis
+### Analysis of rain gauge data
 
@@
 plot(aggregate(climatrans.pr,month,'wetfreq'),new=FALSE)
+grid()
 ```
+
+The monsoon season is clearly visible in the wet-day frequency $f_w$, being a period when it rains  almost every day in Mumbay (Bombay) to somewhat less than every other day in the two other cities New Delhi and Bangalore.
+The duration of the Monsoon season is longest in Bangalore (~Apr-Oct) and shortest in New Delhi (~Jun-Sep).
 
@@
 plot(aggregate(climatrans.pr,month,'wetmean'),new=FALSE)
+grid()
 ```
+
+There are less pronounced difference in the precipitation intencity $\mu$ between the Monsoon season and the other Calendar months, although the most intense rains tend to take place in Mumbay during June and later in the season in New Delhi (Sep) and Bangalore (Sep-Oct).
+
+Both the frequency and intensity of the rains are typically higher in Mumbay than in New Delhi and Bangalore, and Mumbay has at outset a more challenging situation in terms of coping with extreme rains than the two other megacities.
+This difference is also clearly visible in the typical monthly precipitation totals, which are much higher for Mumbay and peaks in July when both $f_w$ and $\mu$ tend to be high.
 
@@
 plot(aggregate(climatrans.pr,month,'sum'),new=FALSE)
+grid()
 ```
 
-Estimate annual aggregated statistics for the months May-October (the wet/Monsoon season):
+### Testing the precipitation statistics.
+
+The idea that the wet-day precipitation approcimately follows the exponential distribution enbles a simplified analysis and interpreation of the reainfall statistics and the prediction/projection of heavy precipitation events.
+
+We concentrate on the monsoon season and estimate annual aggregated statistics for the months May-October:
 
@@
 
-Compare the data to the exponential distribution - if the data is exponentially distributed, then the wet-day mean precipitation $\mu$ and the 95-percentile $q_{95}$ are propportional by $q_p = -ln(1-p)\mu$.
+A simple test to see if the data follows the exponential distribution is to compare the data to the exponential distribution - if the data is exponentially distributed, then the wet-day mean precipitation $\mu$ and the 95-percentile $q_{95}$ are propportional by $q_p = -ln(1-p)\mu$.
 
@@
 
-Compare the histogram of the preciptation data with an exponential distribution
+another way to visualise this is to compare the histogram of the preciptation data with an exponential distribution
+
 ```{r}
@@
 lines(seq(0,250,by=1),dexp(seq(0,250,by=1),rate=1/mean(z)),lwd=3,col='red')
+grid()
 ```
 
-##**Historical long-term trends**
+###**Historical long-term trends in precipitation**
 
@@
 for(i in 1:3) lines(trend(subset(fw,is=i)))
+grid()
+print(trend(fw,result='pval')) 
 ```
+
+The long-term trends only hint to potentially slight changes, and only the trend in New Delhi appears to be statistically significant at the 5% level.
 
@@
 for(i in 1:3) lines(trend(subset(mu,is=i)))
+grid()
+print(trend(mu,result='pval'))
 ```
 
-If the 24-hr preciptation amounts is exponential (when it rains), then the probability of exceeding a threshold is $Pr(X>x_0) = f_w e^{-x_0/\mu}$. Since we know both
+For $\mu$, only the record for Mumbay suggest a significant long-term trend at the 5%-level.
+There were a couple of years during the 1950s which stand out as unique seasons with particularly intense rainfall in Mumbay.
+
+If the 24-hr preciptation amounts is exponential (when it rains), then the probability of exceeding a threshold is $Pr(X>x_0) = f_w e^{-x_0/\mu}$.
+Since we know both, we can approximately quantify the probability for days receiving more than 50 mm of rain.
 
@@
 for(i in 1:3) lines(trend(subset(Pr,is=i)))
+grid()
 ```
@@
 ## Aggregate to mean seasonal cycle and annual mean values
-plot(annual(climatrans.pr,'sum'),new=FALSE)
+pt <- annual(climatrans.pr,'sum')
+attr(pt,'unit') <- 'mm/season'
+plot(pt,new=FALSE)
 for(i in 1:3) lines(trend(subset(annual(climatrans.pr,'wetfreq'),is=i)))
+grid()
 ```
 
-##**Predictors**
+##**Predictors for downscaling**
 
@@
 The aggregation of the temperature over the wet-season may involve the average, minimum or the maximum values.
-Trial anderror to see what works and what doesn't:
+Trial and error to see what works and what doesn't: the results suggest a reasonable choise was May-October season mean.
 
@@
 
-The mean seal-level pressure (SLP) is usually a promising predictor for the wet-day frequency.
+The mean seal-level pressure (SLP) is usually a promising predictor for the wet-day frequency for the May-October season.
 
@@
 y <- subset(mu,is=2)
-corfield(y,es,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)))
+corfield(y,es,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)),new=FALSE)
 ```
@@
 z <- subset(fw,is=2)
-corfield(z,slp,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)))
+corfield(z,slp,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)),new=FALSE)
 ```
@@
 pca.fw <-PCA(fw)
+plot(pca.mu,new=FALSE)
 ```
 
-### **Estimate EOFs for the predictands**
+The leading mode for the seasonal wet-day mean precipitation is dominated by Mumbay with almost zero covariance with Bangalore.
+
+```{r}
+plot(pca.fw,new=FALSE)
+```
+
+The PCA pattern has more similar spatial weights for $f_w$ over the three megacities, even though there is a more equal share of variance between the modes.
+
+### **Estimate EOFs for the predictors**
+
 ```{r}
@@
 
+The dominant mode of variability in the May-October mean $e_s$ featured largest magnitudes east of south Indian coast.
+
 ### **Plot the EOFs for the temperature-based predictor**
@@
 ```
+
+The main May-October SLP mode was associated with the strength of the Monsoon winds.
 
@@
 
-### 
-```{r,echo=FALSE, fig.width=6, fig.height=6}
+### The wet-day mean precipitation
+
+The predictor for the wet-day mean precipitation was taken to be the saturation vapour pressure $e_s$ estimated from maritime surface air taken from the NCEP/NCAR reanalysis 1.
+This choice was motivated by physics, expecting the intensity to be influenced by the amount of moisture in the air - which again has ita main source from the evaopration over the oceans.
+
+```{r fig.width=6, fig.height=6}
 ## The function DS uses multiple regression by default
 ds.mu <- DS(pca.mu,eof.es,eofs=1:20)
+plot(ds.mu,colbar1=list(breaks=seq(0,1,by=0.1)),
+           colbar2=list(breaks=seq(-500,500,by=25)),new=FALSE)
+```
+
+The results for the wet-daty mean $\mu$ were problematic in the sense that the evaluation of the downscaling model (cross-validation) revealed low scores, in particular associated with a small number of extreme seasons with exceptionally high $\mu$.
+
+## The wet-day frequency
+
+Mean sea level pressure was used as a predictor to downscale the wet-day frequency $f_w$, which too is motivated from physical reasoning: the wind speed/direction and transport of moisture influence the probability for precipitation.
+
+```{r}
 ds.fw <- DS(pca.fw,eof.slp,eofs=1:20)
-plot(ds.mu,colbar1=list(breaks=seq(0,1,by=0.1),rev=2),
-     colbar2=list(breaks=seq(-500,500,by=25),rev=TRUE),new=FALSE)
 plot(ds.fw,new=FALSE)
 ```
+
+The downscaled results for $f_w$ were associated with high skill (cross-validation = 0.71), and the associated predictor pattern implies that the rain occurrence is associated with large-scale southwesterly wind field, typically associated with the monsoon.
 
@@
 
-```{r}
-## Examine the PCA of residuals for the cold monthsto see if there are any structures left
-  mu.2 <- as.residual(as.station(ds.mu))
-  mu.2 <- attrcp(mu,mu.2); class(mu.2) <- class(mu)
+```{r pca-of-residual-mu}
+## Examine the PCA of residuals for the cold months to see if there are any structures left
+  mu.2 <- as.station(as.residual(ds.mu))
   pca.mu.2 <- PCA(mu.2)
-  plot(pca.mu.2)
+  plot(pca.mu.2,new=FALSE)
+  grid()
+  str(mu.2)
 ```
@@
 
-```{r}
+```{r pca-of-residual-fw}
 ## Examine the PCA of residuals for the cold monthsto see if there are any structures left
-  fw.2 <- as.residual(as.station(ds.fw))
-  fw.2 <- attrcp(fw,fw.2); class(fw.2) <- class(fw)
+  fw.2 <- as.station(as.residual(ds.fw))
   pca.fw.2 <- PCA(fw.2)
-  plot(pca.fw.2)
+  plot(pca.fw.2,new=FALSE)
+  grid()
 ```
@@
 
-Here include the argument xfuns='tas2es' to use the option annual(tas2es(gcm)) rather than the option annual(gcm,FUN='tas2es') which will fail. Also
+Here include the argument xfuns='tas2es' to use the option annual(tas2es(gcm)) rather than the option annual(gcm,FUN='tas2es') which will fail.
+Also, we apply the downscaling to the mean sea level pressure (MSLP).
 
@@
 
-#### *Maximum temperature*
+*Delhi* 
+Validation of the downscaled ensembles underestimated both the observed interannual variations and trend in the May-October mean $\mu$.
+The projections do not indicate much change in the future, however, the analysis does not capture well all factors that may influence the precipitation intencity.
+It may be more sensitive to small-scale conditions and processes (e.g.
+mesoscale convection).
+
+The donscaled ensemble provides an inadequate description of $f_w$ for Delhi, although both models and observations agree on near-zero trends.
+There are uncounted sources for interannual variability, which implies and underestimation of dry and wet seasons.
+
+*Mumbay*
+The downscaled results provide an inadequate description of both $\mu$ and $f_w$ in terms of interannual variability and trend ($\mu$).
+The results do not suggest future change in $\mu$, however, there are some indication of a slight increase in $f_w$.
+
+*Bangalore*
+The downscaled results provide an inadequate description of both $\mu$ and $f_w$ in terms of interannual variability and trend, and suggest future no change in $\mu$ or $f_w$.
+
+## GCM raw results for precipitation over the Indian sub-continent:
+
+As an extra check, a plot is made of the raw CMIP5 RCP4.5 GCM results for the precipitation averaged over the Indian continent (70-90E/7-30N).
+Here, the annual mean precipitation $\overline{x}$ is shown rather than May-October $f_w$ and $\mu$, however, $\overline{x}= f_w \mu$ and most of the rainfall typically falls within the May-October period.
+Hence, any change in the annual mean precipitation is mostly due to the rainy monsoon season.
+
+The results suggest a modest increase simulated for precipitation.
+The corresponding estimate from the NCAR/NCEP reanalyses suggests an increasing trend, however, such results should be viewed with caution doe to potential inhomogeneities (<https://climatedataguide.ucar.edu/climate-data/atmospheric-reanalysis-overview-comparison-tables>)
+
+```{r}
+## India:
+Z2 <- precipcmip(lon=c(70,90),lat=c(7,30))
+par(bty='n')
+X <- zoo(x=apply(coredata(Z2$gcm),2,function(x) 100*(x/mean(x,na.rm=TRUE))),order.by=index(Z2$gcm))
+plot(X,lwd=3,col=rgb(0,0.5,1,0.2),plot.type='single',ylab='Annual precip (%)')
+lines(100*Z2$obs/mean(Z2$obs,na.rm=TRUE),lwd=3)
+grid()
+```
+
+## * Analysis of maximum temperature*
 
@@
   load('climatrans.tx.rda')
-  plot(as.4seasons(anomaly(climatrans.tx)))
+  plot(as.4seasons(anomaly(climatrans.tx)),new=FALSE)
   for(i in 1:3) lines(trend(subset(as.4seasons(anomaly(climatrans.tx)),is=i)))
+  grid()
+```
+
+## Are the hot events behaving like Poisson distribution?
+
+Take hot events to be any day warmer than 30C:
+
+```{r}
+nhot <- annual(climatrans.tx,FUN='count',threshold=35)
+plot(nhot,new=FALSE)
+for(i in 1:3) lines(trend(subset(nhot,is=i)))
+grid()
+print(trend(nhot,result='pval')) 
+```
+
+```{r}
+print(paste('Mean number=',mean(nhot[,1],na.rm=TRUE),'Variance=',var(nhot[,1],na.rm=TRUE),
+            'should equal',mean(nhot[,1],na.rm=TRUE),'for Poisson process'))
+hist(coredata(nhot[,1]),freq=FALSE,col='grey',
+     main=loc(nhot)[1])
+lines(dpois(x=seq(0,max(nhot[,1]),by=1),lambda=mean(nhot[,1],na.rm=TRUE)),lwd=2,col='red')
 grid()
@@
   eof.t2m <- EOF(predictor)
+```
+
+Organise the predictand
+
+```{r Tx-predictand}
 ## The predictand
@@
   index(pca.tx) <- year(pca.tx)
+```
+
+Apply the downscaling to maximum temperature and check the residuals.
+
+```{r}
   ds.tx <- DS(pca.tx,eof.t2m)
+  plot(ds.tx,new=FALSE)
+```
+
+### Check the residuals of the regression analysis
+
+```{r}
 ## Check the residuals
-  tx.2 <- as.residual(as.station(ds.tx))
-  tx.2 <- attrcp(Tx,tx.2); class(tx.2) <- class(Tx)
+  tx.2 <- as.station(as.residual(ds.tx))
   pca.tx.2 <- PCA(tx.2)
-  plot(pca.tx.2)
+  plot(pca.tx.2,new=FALSE)
 ```
+
+The eigenvalue spectrum is not flat, and the time series hints of some temporal structure, however, the spatial weights suggests little spatial covariance.
+
+## Hot day statistics
 
@@
     ## the assumption that the temperature is close to being normally distributed
-    hw <- hotsummerdays(x=y,dse=dse.tx,threshold=trh[i],plot=FALSE)
-    plot(hw,new=FALSE)
+    hw <- hotsummerdays(x=y,dse=dse.tx,threshold=trh[i],plot=TRUE,new=FALSE)
+    #plot(hw)
   }
 ```
+
+The general linear models provide a reasonable description of the number of hot events given the summer mean temperature.
+The tails of the distribution do not quite conform to the normal distribution, however.
+The downscaled model summer mean temperatures are comparable with observed  summer mean temperatures.
+
+### Check link between the mean and 95-percentile
 
@@
 
-Check whether the maximum temperature is close to being normally distributed
-
-```{r}
-## New Delhi:	
-  new.delhi.tx <- subset(climatrans.tx,is=1)
-## Bangalore:
-  bangalore.tx <- subset(climatrans.tx,is=3)
-## Bombay:     
-  bombay.tx <- subset(climatrans.tx,is=2)
-
-  y <- new.delhi.tx
-  par(mfcol=c(2,2))
-  qqnorm(coredata(subset(y,it='djf')),col='grey',main='DJF')
-  qqline(coredata(subset(y,it='djf')))
-  qqnorm(coredata(subset(y,it='mam')),col='grey',main='MAM')
-  qqline(coredata(subset(y,it='mam')))
-  qqnorm(coredata(subset(y,it='jja')),col='grey',main='JJA')
-  qqline(coredata(subset(y,it='jja')))
-  qqnorm(coredata(subset(y,it='son')),col='grey',main='SON')
-  qqline(coredata(subset(y,it='son')))
-```
-
 Check the long-term change: historic trend in the seasonal spread (standard deviation) - the above estimates assumed that the spread is constant.
@@
 ```{r}
-  plot(as.4seasons(anomaly(climatrans.tx),FUN='sd'),new=FALSE)
+  tx.sd <- as.4seasons(anomaly(climatrans.tx),FUN='sd')
+  plot(subset(tx.sd,it='jja'),new=FALSE)
+  for(i in 1:3) lines(trend(subset(subset(tx.sd,it='jja'),is=i)))
+  grid()
+  print(trend(subset(tx.sd,it='jja'),result='pval'))
+  print(colMeans(climatrans.tx))
 ```
+
+The interannual variabilty in the seasonal daily max temperature variance of ~2 degrees is small compared to the temperature level (especially in degrees Kelvin).
+
+### Check if the variance in standard deviation conforms to randomness: normally distributed
+
+```{r}
+for (i in 1:3) {
+  qqnorm(coredata(subset(subset(tx.sd,it='jja'),is=i)),main=loc(tx.sd)[i])
+}
+```

### 2016-07-29 ClimaTrans/climatrans.Rmd
Climatrans

@@
 ## Information about the system and session
-sessionInfo()
 ```
@@
   return(results)
+}
+```
+
+The function `hotsummerdays` is provided in `esd`, and the listing of the code is provided below:
+
+```{r, eval=FALSE}
+hotsummerdays <- function (x, y = NULL, dse = NULL, it = "jja", threshold = 30, 
+    verbose = FALSE, plot = TRUE, nmin = 90, new = TRUE, ...) 
+{
+    if (verbose) 
+        print("mildwinterdays")
+    stopifnot(inherits(x, "station"))
+    if (is.null(y)) 
+        y <- x
+    djf <- subset(x, it = it)
+    djfy <- subset(y, it = it)
+    nwd1 <- annual(djfy, FUN = "count", threshold = threshold, 
+        nmin = nmin)
+    mwd1 <- annual(djf, FUN = "mean", nmin = nmin)
+    cal <- data.frame(x = c(coredata(mwd1)), y = c(coredata(nwd1)))
+    dfit <- glm(y ~ x, family = "poisson", data = cal)
+    if (plot) {
+        if (new) 
+            dev.new()
+        par(bty = "n")
+        plot(cal, pch = 19, ylim = c(0, 90), xlab = expression(paste("mean temperature ", 
+            (degree * C))), ylab = "number of hot days", main = loc(x))
+        pre <- data.frame(x = seq(min(cal$x, na.rm = TRUE) - 
+            1, max(cal$x, na.rm = TRUE) + 5, by = 0.1))
+        lines(pre$x, exp(predict(dfit, newdata = pre)), col = rgb(1, 
+            0, 0, 0.3), lwd = 3)
+        if (new) 
+            dev.new()
+        djf.sd <- sd(coredata(djf), na.rm = TRUE)
+        qqnorm(coredata(djf))
+        qqline(coredata(coredata(djf)), col = "red")
+        grid()
+    }
+    if (is.null(dse)) 
+        dse <- DSensemble.t2m(x, biascorrect = TRUE, verbose = verbose, 
+            plot = plot)
+    djf.dse <- subset(dse, it = "djf")
+    index(djf.dse) <- year(djf.dse)
+    ovl <- window(djf.dse, start = year(start(x)), end = year(end(x)))
+    djf.dse <- djf.dse - mean(coredata(ovl), na.rm = TRUE) + 
+        mean(coredata(mwd1), na.rm = TRUE)
+    q1 <- data.frame(x = apply(coredata(djf.dse), 1, quantile, 
+        probs = 0.05, na.rm = TRUE))
+    q2 <- data.frame(x = apply(coredata(djf.dse), 1, quantile, 
+        probs = 0.95, na.rm = TRUE))
+    qm <- data.frame(x = apply(coredata(djf.dse), 1, mean, na.rm = TRUE))
+    obs <- data.frame(x = coredata(mwd1))
+    t <- year(index(djf.dse))
+    preq1 <- exp(predict(dfit, newdata = q1))
+    preq1[preq1 > 90] <- NA
+    tr1 <- predict(lm(preq1 ~ t + I(t^2) + I(t^3) + I(t^4) + 
+        I(t^5)))
+    tr1[!is.finite(preq1)] <- NA
+    preq2 <- exp(predict(dfit, newdata = q2))
+    preq2[preq2 > 90] <- NA
+    tr2 <- predict(lm(preq2 ~ t + I(t^2) + I(t^3) + I(t^4) + 
+        I(t^5)))
+    tr2[!is.finite(preq2)] <- NA
+    prem <- exp(predict(dfit, newdata = qm))
+    prem[prem > 90] <- NA
+    tr3 <- predict(lm(prem ~ t + I(t^2) + I(t^3) + I(t^4) + I(t^5)))
+    tr3[!is.finite(prem)] <- NA
+    Nwd <- zoo(cbind(preq1, preq2, prem, tr1, tr2, tr3), order.by = t)
+    nwd.pre <- zoo(exp(predict(dfit, newdata = obs)), order.by = year(mwd1))
+    if (plot) {
+        if (new) 
+            dev.new()
+        par(bty = "n")
+        plot(zoo(djf.dse, order.by = year(djf.dse)), plot.type = "single", 
+            col = rgb(0.5, 0.5, 0.5, 0.2), ylab = expression(paste("mean temperature", 
+                (degree * C))), xlab = "", main = loc(x))
+        points(mwd1, pch = 19)
+        grid()
+        if (new) 
+            dev.new()
+        par(bty = "n")
+        plot(Nwd, plot.type = "single", lwd = 5, main = loc(x), 
+            ylim = c(0, 90), xlab = "", ylab = paste("number of hot days: T(2m) > ", 
+                threshold, unit(x)), col = c(rgb(0.5, 0.5, 0.7, 
+                0.5), rgb(0.8, 0.5, 0.5, 0.5), rgb(0.8, 0.5, 
+                0.8, 0.5), rgb(0.3, 0.3, 0.6, 0.5), rgb(0.6, 
+                0.3, 0.3, 0.5), rgb(0.6, 0.3, 0.6, 0.5)), ...)
+        grid()
+        points(nwd1, pch = 19)
+        lines(nwd.pre, col = rgb(0.5, 0.5, 0.5, 0.5))
+    }
+    Nwd <- attrcp(x, Nwd)
+    attr(Nwd, "unit") <- "days"
+    attr(Nwd, "info") <- paste("number of hot days: t2m > ", 
+        threshold)
+    attr(Nwd, "observation") <- nwd1
+    attr(Nwd, "nwd.pre") <- nwd.pre
+    index(Nwd) <- t
+    class(Nwd) <- c("nevents", "zoo")
+    invisible(Nwd)
 }
@@
 
-## Are the hot events behaving like Poisson distribution?
-
-Take hot events to be any day warmer than 30C:
-
-```{r}
-nhot <- annual(climatrans.tx,FUN='count',threshold=35)
-plot(nhot,new=FALSE)
-for(i in 1:3) lines(trend(subset(nhot,is=i)))
-grid()
-print(trend(nhot,result='pval')) 
-```
-
-```{r}
-print(paste('Mean number=',mean(nhot[,1],na.rm=TRUE),'Variance=',var(nhot[,1],na.rm=TRUE),
-            'should equal',mean(nhot[,1],na.rm=TRUE),'for Poisson process'))
-hist(coredata(nhot[,1]),freq=FALSE,col='grey',
-     main=loc(nhot)[1])
-lines(dpois(x=seq(0,max(nhot[,1]),by=1),lambda=mean(nhot[,1],na.rm=TRUE)),lwd=2,col='red')
-grid()
-```
-
 Downscale the seasonal seasonal mean maximum temperature and then use the mean temperature to estimate the number of hot days assuming that the daily distribution is approximately Gaussian for a given season.
@@
 ```{r}
-  trh <- c(40,30,32)
+  trh <- c(40,35,35)
   for (i in 1:3) {
@@
     ## the assumption that the temperature is close to being normally distributed
-    hw <- hotsummerdays(x=y,dse=dse.tx,threshold=trh[i],plot=TRUE,new=FALSE)
+    dse.tx[is.element(year(dse.tx),2100),] <- NA   # some of the results for 2100 are suspect 
+    hw <- hotsummerdays(x=y,dse=dse.tx,threshold=trh[i],it=NULL,plot=TRUE,new=FALSE)
     #plot(hw)
@@
 ### Check link between the mean and 95-percentile
+
+More traditional approaches of downscaling typically involve either downscaling daily data and estimate a probability density function (pdf) from these results or downscale the pdf directly in terms of the parameters describing the shape (<http://climatescience.oxfordre.com/view/10.1093/acrefore/9780190228620.001.0001/acrefore-9780190228620-e-27>).
+The probability of hot days may be estimated based on the pdf, and it is important to check the consistency between the data and the relevant parametric distribution.
 
@@
 ```
+
+## Are the hot events behaving like Poisson distribution?
+
+Take hot events to be any day warmer than 40C:
+
+```{r}
+for(i in 1:3) {
+  nhot1 <- annual(subset(climatrans.tx,is=i),FUN='count',threshold=trh[i])
+  if (i==1) nhot <- nhot1 else nhot <- combine(nhot,nhot1)
+}
+plot(nhot,new=FALSE)
+for(i in 1:3) lines(trend(subset(nhot,is=i)))
+grid()
+print(trend(nhot,result='pval')) 
+```
+
+```{r}
+print(paste('Mean number=',mean(nhot[,1],na.rm=TRUE),'Variance=',var(nhot[,1],na.rm=TRUE),
+            'should equal',mean(nhot[,1],na.rm=TRUE),'for Poisson process'))
+hist(coredata(nhot[,1]),freq=FALSE,col='grey',
+     main=loc(nhot)[1])
+lines(dpois(x=seq(0,max(nhot[,1]),by=1),lambda=mean(nhot[,1],na.rm=TRUE)),lwd=2,col='red')
+grid()
+```

### 2016-07-29 ClimaTrans/climatrans.Rmd
update

@@
 mu <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetmean')
+nd <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'nv')
 q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wq95')
@@
 
+```{r}
+print(summary(coredata(pt)))
+print(summary(coredata(fw*nd)))
+print(summary(coredata(mu)))
+```
+
 ##**Predictors for downscaling**
@@
 Trial and error to see what works and what doesn't: the results suggest a reasonable choise was May-October season mean.
+
+```{r}
+## Retrieve the reanalysis from FTP
+if (!file.exists("air.mon.mean.nc"))
+  download.file('ftp://ftp.cdc.noaa.gov/Datasets/ncep.reanalysis.derived/surface/air.mon.mean.nc',
+              destfile = 'air.mon.mean.nc')
+if (!file.exists("slp.mon.mean.nc"))
+  download.file('ftp://ftp.cdc.noaa.gov/Datasets/ncep.reanalysis.derived/surface/slp.mon.mean.nc',
+              destfile = 'slp.mon.mean.nc')
+```
 
@@
   grid()
+  print(summary(coredata(climatrans.tx)))
 ```

### 2016-07-29 gmst-outlook/gmst-outlook.Rmd
update

---
title: "gmst-outlook"
author: "Rasmus Benestad"
date: "July 28, 2016"
output: pdf_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r}
library(esd)
download.file('ftp://ftp.cdc.noaa.gov/Datasets/ncep.reanalysis.derived/surface/air.mon.mean.nc',
              destfile = 'air.mon.mean.nc')
```

Get the NCEP/NCAR reanalysis 1 from FTP and estimate the global mean anomaly

```{r}
t2m <- retrieve('air.mon.mean.nc')
gmst <- aggregate.area(t2m,FUN='mean')
gmsta <- anomaly(gmst,ref=1961:1990)
```

Find the most recent complete month and set up a regression model

```{r}
t <- index(gmst)
mon <- month(t[length(t)])
Y <- annual(gmsta)
x <- aggregate(subset(gmsta,it=1:mon),year,FUN='mean')
z <- lm(coredata(Y) ~ coredata(x))
y <- predict(z,newdata=x)
n <- length(y)
```

Plot the analysis

```{r}
ylim<- range(c(coredata(Y),coredata(y)),na.rm=TRUE) + 0.25*c(-1,1)
plot(zoo(x),zoo(Y),ylim=ylim,
     xlab=paste('Jan-',month.abb[mon],' mean',sep=''),
     ylab='annual mean', sub='base line: 1961-1990; source NCEP/NCAR reanalysis 1',
     main='Global mean temperature',pch=19,cex=1.5)
lines(rep(coredata(x)[n],2),ylim,lwd=7,col='grey90')
arrows(coredata(x)[n],coredata(y)[n],min(x,na.rm=TRUE),coredata(y)[n],lty=2)
points(coredata(x)[n],coredata(y)[n],cex=1.5,lwd=2)
text(coredata(x)[n],min(y),max(year(x)),sub=1)
grid()
```

### 2016-07-29 dipper.R

@@ -1,0 +1,86 @@
+## R-script to analyse dipper population
+
+library(esd)
+library(ncdf4)
+
+examine <- FALSE
+
+dipper <- read.table('~/Desktop/dipper.csv',header=TRUE)
+
+dipper <- zoo(x=dipper[[2]],order.by=dipper[[1]])
+y <- as.station(dipper,loc='Lyngdalselva',
+                lon=7,lat=58.5,alt=520,
+                param='population',unit='count',
+                reference='Marlène Gamelon')
+
+predictor <- retrieve("air.mon.mean.nc",lon=c(-1,17),lat=c(57,62))
+## Only use the winter temperature:
+predictor <- aggregate(subset(predictor,it='djf'),year,FUN='mean')
+
+if (examine) {
+  ## Check the distribution
+  n <- seq(0,150,by=10); mu <- mean(dipper)
+  hist(coredata(dipper),breaks=n,freq=FALSE,col='grey')
+  lines(x,dpois(x,mu),lwd=2,col='red')
+
+  ## Single ESD for inspection
+  X <- EOF(predictor)
+  z <- DS(y,X)
+  plot(z)
+}
+
+## Downscale the dipper population directly based on the large-scale
+## annual mean temperature
+
+dev.new()
+if (!file.exists('dipper.Z.rcp45.rda')) {
+  Z.rcp45 <- DSensemble.annual(y,biascorrect=TRUE,
+                               predictor=predictor,
+                               lon=c(-1,17),lat=c(57,62),
+                               abscoords=TRUE)
+  save(file='dipper.Z.rcp45.rda',Z.rcp45)
+} else load('dipper.Z.rcp45.rda')
+
+if (!file.exists('dipper.Z.rcp85.rda')) {
+  Z.rcp85 <- DSensemble.annual(y,biascorrect=TRUE,
+                               rcp="rcp85",predictor=predictor,
+                               lon=c(-1,17),lat=c(57,62),
+                               abscoords=TRUE)
+  save(file='dipper.Z.rcp85.rda',Z.rcp85)
+} else load('dipper.Z.rcp85.rda')
+
+if (!file.exists('dipper.Z.rcp26.rda')) {
+  Z.rcp26 <- DSensemble.annual(y,biascorrect=TRUE,
[…]

### 2016-08-02 ClimaTrans/climatrans.Rmd

@@
 library(ncdf)
+#rm(list=ls())
 ## Information about the system and session
@@
 str(climatrans.pr)
+plot(climatrans.pr,new=FALSE)
 ```
+
 ##**Sensitivity tests**
@@
 nd <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'nv')
+ndhr <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'count',threshold=100)
 q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wq95')
@@
 ```{r, fig.width=6, fig.height=6}
-## Wet-day mean
+## Probability of exceedin 50 mm/day
 Pr <- PrexpPr(mu,fw,x0=50) 
@@
 grid()
+print(trend(Pr,result='pval'))
 ```
@@
 ```{r, fig.width=6, fig.height=6}
-## Aggregate to mean seasonal cycle and annual mean values
+## Aggregate to annual totals
 pt <- annual(climatrans.pr,'sum')
@@
 plot(pt,new=FALSE)
-for(i in 1:3) lines(trend(subset(annual(climatrans.pr,'wetfreq'),is=i)))
+for(i in 1:3) lines(trend(subset(pt,is=i)))
+grid()
+print(trend(pt,result='pval'))
+```
+
+Also, it's interesting to see how the number of days with heave rainfall changes over time:
+
+```{r, fig.width=6, fig.height=6}
+plot(ndhr,new=FALSE)
 grid()
@@
 eof.slp <- EOF(slp)
+eof.t2m <- EOF(aggregate(subset(t2m,is=list(lon=c(65,80),lat=c(10,20))),year,'mean'))
 index(eof.es) <- year(eof.es)
 index(eof.slp) <- year(eof.slp)
+index(eof.t2m) <- year(eof.t2m)
 ```
@@
 plot(eof.slp,new=FALSE)
+```
+
+The main May-October T(2m) mode was associated with the strength of the Monsoon winds.
+
+### **Plot the EOFs for the SLP**
+```{r, fig.width=6, fig.height=6}
+plot(eof.t2m,new=FALSE)
 ```
@@
 
-## The wet-day frequency
-
-Mean sea level pressure was used as a predictor to downscale the wet-day frequency $f_w$, which too is motivated from physical reasoning: the wind speed/direction and transport of moisture influence the probability for precipitation.
-
-```{r}
-ds.fw <- DS(pca.fw,eof.slp,eofs=1:20)
-plot(ds.fw,new=FALSE)
-```
-
-The downscaled results for $f_w$ were associated with high skill (cross-validation = 0.71), and the associated predictor pattern implies that the rain occurrence is associated with large-scale southwesterly wind field, typically associated with the monsoon.
-
-## **Examine the residuals from the downscaling**
+### Examine the residuals from the downscaling for the wet-day mean
 
@@
 
-The residuals for the wet-day frequency:
+### Alternative approach: downscale the number of days with heavy rainfall
+
+Model the number with heavy rainfall (more than 100 mm/day) in Mumbay.
+
+```{r fig.width=6, fig.height=6}
+ndhr.cal <- data.frame(y = coredata(subset(subset(ndhr,is=2),it=c(1949,2004))),
+                       X = coredata(subset(eof.es,it=c(1949,2004))))
+ds.ndhr <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr.cal,family="poisson")
+plot(1949:2004,ndhr.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[2])
+grid()
+lines(1949:2004,exp(predict(ds.ndhr)),col='red')
+```
+
+Check if the number of days with heavy precipitation is Poisson-distributed
+
+```{r}
+print(paste('Mean number=',mean(ndhr[,2],na.rm=TRUE),'Variance=',var(ndhr[,2],na.rm=TRUE),
+            'should equal',mean(ndhr[,2],na.rm=TRUE),'for Poisson process'))
+h <- hist(coredata(ndhr[,2]),freq=FALSE,col='grey',
+     main=loc(ndhr)[2])
+lines(h$mids-0.5,dpois(x=h$mids-0.5,lambda=mean(ndhr[,2],na.rm=TRUE)),lwd=2,col='red')
+grid()
+```
+
+## The wet-day frequency
+
+Mean sea level pressure was used as a predictor to downscale the wet-day frequency $f_w$, which too is motivated from physical reasoning: the wind speed/direction and transport of moisture influence the probability for precipitation.
+
+```{r}
+ds.fw <- DS(pca.fw,eof.slp,eofs=1:20)
+plot(ds.fw,new=FALSE)
+```
+
+The downscaled results for $f_w$ were associated with high skill (cross-validation = 0.71), and the associated predictor pattern implies that the rain occurrence is associated with large-scale southwesterly wind field, typically associated with the monsoon.
+
+### The residuals for the wet-day frequency:
 
@@
 
-#### **Do the downscaling based on the CMIP5 RCP4.5 ensemble**
+## **Do the downscaling based on the CMIP5 RCP4.5 ensemble**
 
@@
   qqnorm(coredata(subset(subset(tx.sd,it='jja'),is=i)),main=loc(tx.sd)[i])
+  qqline(coredata(subset(subset(tx.sd,it='jja'),is=i)),main=loc(tx.sd)[i])
 }

### 2016-08-03 ClimaTrans/climatrans.Rmd
ClimaTrans

@@
+---
+output: pdf_document
+---
 *Climatrans results: Work package 3: Outline of the development trends and climate change impact up to 2050* 

### 2016-09-08 ClimaTrans/climatrans.Rmd
Added R-scripts used for the ERL paper cited in AACA

--- 2016-08-02 ClimaTrans/climatrans.Rmd
@@
-*Climatrans results: Work package 3: Outline of the development trends and climate change impact up to 2050* 
+---
+output: pdf_document
+---
+* Supporting Material for 'Predicting events: Hot days and heavy precipitation'*
+
 -----------------------------------
@@
 **Background**
+
+***Climatrans results: Work package 3: Outline of the development trends and climate change impact up to 2050***
 The main objectives of the project are: 1) Assess climate change and environmental impacts in urban areas in India related to the transport sector.
@@
 ```
+
+## The highest values expected is proportional to the sample size
+
+The actual sample size for precipitation amount is $N=n_{days} f_w$, and for a variable with an particular exponential distribution, we can demonstrate that the highest value is a function of the sample size $N$:
+
+```{r}
+N = round(10^seq(1,5,length=300))
+qx <- rep(NA,length(N))
+for (i in 1:length(N)) qx[i] <- max(rexp(N[i]))
+plot(N,qx,log='x',main='Maximum value vs sample size',ylab='Maximum value')
+```
+
+## Some analyses to examine the data and its quality
+
+A simple ideal demonstration that simulates the gridding process where a grid-point data is a weighted sum of four nearby measurements (similar to a bi-linear interpolation).
+
+```{r}
+## Test weighted sum of daily rainfall.
+## Simulate gridding
+
+N <- 100000
+iw <- runif(4*N)
+x <- rexp(4*N)
+x[iw < 0.6] <- 0
+W <- runif(4)
+W <- W/sum(W)
+dim(x) <- c(N,4)
+y <- x %*% W
+
+breaks <- seq(-1,20,by=0.1)
+hx1 <- hist(x[,1],plot=FALSE,breaks=breaks)
+hx2 <- hist(x[,2],plot=FALSE,breaks=breaks)
+hx3 <- hist(x[,3],plot=FALSE,breaks=breaks)
+hx4 <- hist(x[,4],plot=FALSE,breaks=breaks)
+maxx <- max(hx1$counts,hx2$counts,hx3$counts,hx4$counts)
+hy <- hist(y,breaks=breaks,xlim=c(0,5),ylim=c(0,maxx),col="grey",border="grey",
+           main='Test gridded precip',xlab='amount',
+           sub="'gridded' result = weighted sum of 4 stations")
+lines(hx1$mids,hx1$counts,col='red',lwd=3)
+lines(hx2$mids,hx2$counts,col='red',lwd=3)
+lines(hx3$mids,hx3$counts,col='red',lwd=3)
+lines(hx4$mids,hx4$counts,col='red',lwd=3)
+legend(3,maxx,c('Individual simulation','Weighted sum'),
+      col=c('red','grey'),lwd=c(3,10),bty='n')
+```
+
+## Example of station record
+
+The format of the station records bears a structure from the past where it was more important to take up little space rather than being human readable.
+
+```{r eval=FALSE}
+102197208 491010005000000000000000000000000000000000000000001090000000002820249 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 49102000000000000000000000000000300000000000000000000000000000000     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 491110000000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 4911200000000000000000000000000000000000000000000000000000000         1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 491210000000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 49122000000000000000000000000000000000000000000000000000000000000     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500110000000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 50012000000000000000000000000000000000000000000000000000000000000     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500210000000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 50022000000000000000000000000000000000000000000000000                 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500310000000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 50032000000000000000000000000000000000000000000000000000000000000     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500410000000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 5004200000000000000000000000000000000000000000000000000000000         1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500510000000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 50052000000000000000000000000000000080003000500000000000000000000     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500610000000000000000000300050000000000000000004304750051001301470053 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 5006201220089005800790076000000200000000000000010001800380000         1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500710005004100280000002000150513154901420030039108410729012200760051 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 50072004600080457074403560066006101070269046007320330001300710058     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500810028000300410147006100710081027701420152010700130015000000050005 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 50082000000250000000000030000000000000000000000000000005600030175     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 500910470107207921283027400580208007900080000000003811201065800030000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 5009200030003002003350043000000380000000000000000000000000000         1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 501010028000000000000000000000000009900000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 50102000000000000000000000000000000000000000000000000000000000000     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 501110000000000000000000000000000000000000000000000000000000000280000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 5011200000000000000000000000000000000000000000000000000000000         1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 501210000000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 50122000000000000000000000000000000000000000000000000000000000000     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 51011    000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 51012000000000000000000000000000000000000000000000000000000000000     1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 51021    000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 51022000000000000000000000000000000000000000000000000                 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+102197208 51031    000000000000000000000000000000000000000000000000000000000000 1207  BOMBAY CITY     BOMBAY SANTACRUZOBSY   19
+```
+
+The way to decode such files has been provided, but the way the information is stored makes it prone to errors and mistakes when reading the data:
+
+```{r eval=FALSE}
+Format Specifications
+Column                                           Data
+A                                                Catchment No. - 3 Digits
+B                                                Latitude - 2 Digits
+C                                                Longitudes - 2 Digits
+D                                                Station No. - 2 Digits
+E                                                Blank
+F                                                Year - 2 Digits
+G                                                Month - 2 Digits
+H                                                Card No. - 1 Digit
+I to W                                           Rainfall in 4 Digits (Without Decimal)
+                                                 If Col H has a value 1 - Rainfall from
+Dates 1 to 15; value 2 - Rainfall from Dates 17 to 31
+X                                                If Col H has value 1 - Rainfall for
+the date 16; value 2 - Rainfall for the month - (5 Digits)
+Y                                                State No. - 2 Digits
+Z                                                District No. - 2 Digits
+AA                                               STN No. - 2 Digits
+AB                                               District Name - 16 Characters
+AC                                               Station Name - 16 Characters
+AD                                               Class - 5 Characters
+AE                                               Secret Classification - 2 Digits
+AF                                               Century Year - 19 or 20
+```

### 2016-09-08 AACA/aacads.R

@@ -1,0 +1,15 @@
+library(esd)
+
+dayswithnosnow <- function(x) {
+  x <- -x
+  y <- annual(x,FUN='count',threshold=-1)
+  attr(y,'unit') <- 'days < 1 cm'
+  invisible(y)
+}
+
+load('sd.aaca.rda')
+y <- dayswithnosnow(sd.aaca)
+nok <- apply(coredata(y),2,nv)
+y <- subset(y,is=nok > 40)
+map(y,FUN='trend',cex=-2,colbar=list(breaks=seq(-22,22,by=2)))
+

### 2016-09-08 AACA/dsaaca.R

@@ -1,0 +1,101 @@
+### Downscale scenarios for AACA - the Barents region
+### temperature and precipitation.
+### Based on PCA-representation for the predictand.
+### Use 'pcafill' to fill in missing values.
+### Rasmus E. Benestad, Oslo, November, 2015
+
+library(esd)
+
+
+downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
+                      period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
+                      lon=c(0,100),lat=c(65,90),eofs=1:6,n=10,rel.cord=FALSE,select=NULL) {
+
+  print('downscale')
+  
+## Use a time and space window:  
+  Y <- subset(Y,it=period)
+  Y <- subset(Y,is=list(lat=lat,lon=lon))
+
+## Estimate seasonal means & weed out stations with little data
+  Y4 <- subset(as.4seasons(Y,FUN=FUN),it=it)
+  ok <- apply(coredata(Y4),1,nv)
+  Y4 <- subset(Y4,it=ok>0)
+  nok <- apply(coredata(Y4),2,nv)
+  Y4 <- subset(Y4,is=nok>15)
+
+  print(paste(round(100*sum(!is.finite(Y4))/length(!is.finite(Y4))),'% missing',sep=''))
+  if (plot) map(Y,FUN=FUN,cex=-2)
+
+  nmiss <- round(100*sum(!is.finite(Y4))/length(!is.finite(Y4)))
+  #print(paste(nmiss,'% missing',sep=''))
+  
+## Fill missing data using PCA-based regression
+  Z <- pcafill(Y4)
+
+  pca <- PCA(Z,n=n)
+  if (plot) plot(pca)
+
+## Downscale results
+  print('DSensemble')
+  dse.pca <- DSensemble(pca,predictor=predictor,FUNX=FUNX,verbose=verbose,
+                        biascorrect=TRUE,rcp=rcp,eofs=eofs,select=select,
+                        lon=lon,lat=lat,rel.cord=rel.cord,it=it)
+
+  attr(dse.pca,'N.missing') <- nmiss
+  invisible(dse.pca)
+}
+
+
+
+#-----------------------------------------------------------------------
+# Define season and parameter
+
[…]

### 2016-09-08 AACA/esd4aaca.R

@@ -1,0 +1,193 @@
+# Read the data from AARI/EALAT and store as 'esd' station data.
+
+library(esd)
+
+## The Russian data are stored in a annoying and cumbersome format
+## without a consistent standard set-up
+readEALAT <- function(fname,param='t2m',verbose=FALSE) {
+  skip=0; eoh <- FALSE
+  ## Check the length of the header
+  if (verbose) {print(fname); print(param)}
+  while (!eoh) {
+    skip=skip + 1
+    z <- readLines(fname,n=skip)
+    z <- z[length(z)]
+    eoh <- length(grep('/',z)>0)
+  }
+  ## Clean the data:
+  h <- readLines(fname,n=skip)
+  ZZ <- readLines(fname)
+  zz <- ZZ
+  for (i in 1:length(zz)) zz[i] <- gsub('*',' ',ZZ[i],fixed=TRUE,perl=TRUE)
+  for (i in 1:length(zz)) zz[i] <- gsub('?',' ',ZZ[i],fixed=TRUE,perl=TRUE)
+  for (i in 1:length(zz)) zz[i] <- gsub('+',' ',ZZ[i],fixed=TRUE,perl=TRUE)
+  for (i in 1:length(zz))
+    zz[i] <- gsub('9999.9','    NA',zz[i],fixed=TRUE,perl=TRUE)
+  for (i in 1:length(zz))
+    zz[i] <- gsub('999.9','   NA',zz[i],fixed=TRUE,perl=TRUE)
+  for (i in 1:length(zz))
+    zz[i] <- gsub(' 999','  NA',zz[i],fixed=TRUE,perl=TRUE)
+
+  width <- switch(tolower(param),
+                  't2m'=c(5,6,3,3,6,6,6),
+                  'precip'=c(5,6,3,3,6),
+                  'sd'=c(4,3,3,4),
+                  'tx'=c(5,6,3,3,6,6,6),
+                  'tn'=c(5,6,3,3,6,6,6))
+  iyr <- switch(tolower(param),
+                  't2m'=2,'precip'=2,'sd'=1,
+                  'tx'=2,'tn'=2)
+  nc <- nchar(zz[(skip+1):length(zz)])
+  ok <- is.element(nc,max(nc)+c(-1,0))
+  if (file.exists('readEALAT.txt')) file.remove('readEALAT.txt')
+  writeLines(zz[(skip+1):length(zz)][ok],con='readEALAT.txt')
+  if (verbose) print(zz[1:skip])
[…]

### 2016-09-08 AACA/juntunen.R

@@ -1,0 +1,59 @@
+## R script to downscale temperature etc for kittila for the AACA report.
+
+library(esd)
+
+#https://no.wikipedia.org/wiki/Kittil%C3%A4
+#Koordinater: 67°38′48,678″N 24°54′21,622″Ø
+
+if (!file.exists('kittila.rda')) {
+  kittila <- read.table('~/Dropbox/Public/AACA/kittila_timeseries.csv',sep=';',
+                        as.is=TRUE,col.names=c('date','precip','t2m','SWE'))
+  t2m <- zoo(kittila$t2m,order.by=as.Date(kittila$date))
+  precip <- zoo(kittila$precip,order.by=as.Date(kittila$date))
+  SWE <- zoo(kittila$SWE,order.by=as.Date(kittila$date))
+
+  t2m <- as.station(t2m,loc='Kittila',lon=24.90,lat=67.63,alt=196,
+                    unit='deg C',param='temperature')
+  precip <- as.station(precip,loc='Kittila',lon=24.90,lat=67.63,alt=196,
+                       unit='mm',param='precipitation')
+  SWE <- as.station(SWE,loc='Kittila',lon=24.90,lat=67.63,alt=196,
+                    unit='mm',param='Snow depth water equivalent')
+  save(file='kittila.rda',t2m,precip,SWE)
+} else load('kittila.rda')
+
+## Projections for kittila temperature
+if (!file.exists('dse.kittila.rda')) {
+  dse.kittila <- DSensemble.t2m(t2m,biascorrect=TRUE,verbose=FALSE)
+  save(file='dse.kittila.rda',dse.kittila)
+} else load('dse.kittila.rda')
+
+
+## Historic trends in temperature, precip (mu and fw), and SWE.
+
+mu <- annual(precip,FUN='wetmean')
+fw <- annual(precip,FUN='wetfreq')
+
+plot(annual(t2m))
+lines(trend(annual(t2m)))
+
+plot(mu)
+lines(trend(mu))
+
+plot(fw)
+lines(trend(fw))
+
+## Annual maximum snow depth
+plot(annual(SWE,FUN='max'))
+lines(trend(annual(SWE,FUN='max')))
+
+## Number of days with snow
+plot(annual(SWE,FUN='count',threshold=1))
+lines(trend(annual(SWE,FUN='count',threshold=1)))
+
+## Use seasonal variations in mu or variations in geographic vaiations in
[…]

### 2016-09-08 AACA/paper59SM.R

@@ -1,0 +1,38 @@
+## Rasmus E. Benestad
+## Test of pcafill
+
+library(esd)
+param <- 't2m'
+FUN <- 'mean'
+lon=c(0,100); lat=c(65,90)
+reanalysis <- 'air.mon.mean.nc'
+period=c(1950,2015)
+
+if (param=='precip') FUNX <- 'C.C.eq' else FUNX <- 'mean'
+
+## Get the predictand -> Y
+if (!exists('Y')) {
+  load(paste(param,'.aaca.rda',sep=''))
+  Y <- eval(parse(text=paste(param,'.aaca',sep='')))
+}
+Y <- subset(Y,is=list(lat=lat,lon=lon))
+
+## Use a time and space window:  
+for (it in c('djf','mam','jja','son')) {
+  y <- subset(Y,it=period)
+
+## Estimate seasonal means & weed out stations with little data
+  Y4 <- subset(as.4seasons(y,FUN=FUN),it=it)
+  ok <- apply(coredata(Y4),1,nv)
+  Y4 <- subset(Y4,it=ok>0)
+  nok <- apply(coredata(Y4),2,nv)
+  Y4 <- subset(Y4,is=nok>15)
+
+  map(y,FUN=FUN,cex=-1.5)
+  dev.copy2pdf(file=paste('paper59-stations.mean',param,'.',it,'pdf',sep=''))
+
+## Fill missing data using PCA-based regression
+  dev.new()
+  Z <- pcafill.test(Y4,max.miss=50)
+  dev.copy2pdf(file=paste('paper59-pcafill.test.',param,'.',it,'.pdf',sep=''))
+}

### 2016-09-08 AACA/paper59_Atlas.R

@@ -1,0 +1,873 @@
+### Present maps of downscaled scenarios for AACA - the Barents region
+### temperature and precipitation.
+### Based on PCA-representation for the predictand.
+### Use 'pcafill' to fill in missing values.
+### Rasmus E. Benestad, Oslo, November, 2015
+
+library(esd)
+library(ncdf4)
+
+distill <- function(x,it=2050,probs=c(0.05,0.95)) {
+  y <- subset(x,it=it)
+  attributes(y) <- NULL
+  y <- quantile(y,probs=probs,na.rm=TRUE)
+  return(y)
+}
+
+
+grid2ncdf <- function(x,fname='grid2ncdf.nc') {
+  print('grid2ncdf')
+  dimlon <- ncdim_def( "longitude", "degree_east", lon(x) )
+  dimlat <- ncdim_def( "latitude", "degree_north", lat(x) )
+  x4nc <- ncvar_def(varid(x)[1], unit(x)[1], list(dimlon,dimlat), -1, 
+                    longname=attr(x,'longname'), prec="integer")
+     
+     # Create a netCDF file with this variable
+  ncnew <- nc_create( fname, x4nc )
+
+  y <- coredata(x); attributes(y) <- NULL
+  meany <- mean(y,na.rm=TRUE)
+  y <- round(100*(y-meany))
+  y[!is.finite(y)] <- -99
+  # Write some values to this variable on disk.
+  ncvar_put( ncnew, x4nc, round(y) )
+  ncatt_put( ncnew, x4nc, "add_offset", meany, prec="float" )
+  ncatt_put( ncnew, x4nc, "scale_factor", 0.01, prec="float" ) 
+  ncatt_put( ncnew, x4nc, "_FillValue", -99, prec="float" ) 
+  ncatt_put( ncnew, x4nc, "missing_value", -99, prec="float" ) 
+  ncatt_put( ncnew, 0, "description", 
+             "Gridded downscaled results using LaticeKrig")
+  nc_close(ncnew)
+}
+
+propchange <- function(X,it0=c(1986,2015)) {
[…]

### 2016-09-08 AACA/paper59cca.R

@@ -1,0 +1,17 @@
+library(esd)
+
+it <- month.abb[5:9]
+
+load('pr.aaca.rda')
+load('t2m.aaca.rda')
+
+Y <- aggregate(subset(pr.aaca,it=it),year,'wetmean')
+Y <- PCA(pcafill(Y))
+X <- aggregate(subset(t2m.aaca,it=it),year,'mean')
+X <- PCA(pcafill(X))
+
+cca <- CCA(X,Y)
+plot(cca)
+
+
+

### 2016-09-08 AACA/paper59diagnoseESD.R

@@ -1,0 +1,294 @@
+## R-script to diagnose the downscaled (ESD) results based on the scores
+## generated by the downscaling process.
+## Rasmus Benestad
+
+library(esd)
+
+balls <- function(x,y=NULL) {
+  for (i in 1:20) points(x,y,cex=seq(2,0.1,length=20)[i],
+  col=rgb(i/20,i/20,i/20))
+}
+
+show <-  function(x,N,i,gcmnm.i,add=FALSE,xlim=NULL,ylim=NULL,verbose=FALSE,...) {
+  n <- length(x$mean.diff)
+  j <- 1:n
+  col <- rgb(j/n,abs(sin(pi*j/n)),(1-j/n),0.3)
+  if (is.null(xlim)) xlim <- c(-1,1)
+  if (is.null(ylim)) ylim <- c(1,N)
+  
+  if (!add) {
+    dev.new(width=7,height=10)
+    par(bty="n")
+    par0 <- par()
+    plot(xlim,ylim,type="n",
+         ylab="GCM",xlab=expression(1- sigma[p*r*e]/sigma[r*e*f]),
+         main=paste("Diagnostics: common EOFs",attr(x,'variable')),
+         xlim=xlim,ylim=ylim,col="grey")
+    grid()
+    #legend(xlim[1],ylim[2],c("same sign","different sign"),
+    #       pch=c(19,21),bty="n",col="grey")
+    par(xpd=TRUE)
+    text(xlim[2],ylim[2],'EOF #',col='grey40',cex=0.8,pos=3)
+
+    par(new=TRUE,fig=c(0.85,0.95,0.70,0.85),mar=c(0,3,0,0),
+        cex.axis=0.7,yaxt="s",xaxt="n",las=1)
+     colbar <- rbind(1:n,1:n)
+    image(1:2,1:n,colbar,col=col)
+    par(fig=par0$fig,mar=par0$mar,cex.axis=par0$cex.axis,
+        yaxt=par0$yaxt,xaxt=par0$xaxt,las=par0$las,new=TRUE)
+    plot(xlim,ylim,type="n",xlim=xlim,ylim=ylim,
+         xlab='',ylab='',main='',sub='')
+    par(par0$new)
+  }
+  cex <- abs(0.1 - x$mean.diff)/0.1
+  pch <- rep(19,n); pch[cex < 0] <- 21
+  cex <- abs(cex); cex[cex > 2] <- 2
+  if (verbose) {
+     print('Mean difference:');print(x$mean.diff)
[…]

### 2016-09-08 AACA/paper59dsmufw.R

@@ -1,0 +1,87 @@
+## Downscaling exercise to examine the link between large-scale predictors and the local
+## precipitation parameters.
+
+library(esd)
+
+param <- 'pr'
+FUN <- 'wetfreq'
+#it <- NULL
+#it <- month.abb[4:9]
+it <- month.abb[c(1:3,10:12)]
+period <- c(1950,2015)
+eofs <- 1:4
+n <- 10
+verbose <- FALSE
+plot <- FALSE
+
+if (FUN=='wetmean') {
+  FUNX <- 'C.C.eq'
+  pattern <- "tas_Amon_ens_"
+  reanalysis <- 'air.mon.mean.nc'
+  prop=TRUE
+  lon <- c(-50,10); lat <- c(40,75)
+} else {
+  FUNX <- 'mean'
+  pattern <- "psl_Amon_ens_"
+  reanalysis <- 'slp.mon.mean.nc'
+  prop=FALSE
+  lon <- c(-40,70); lat <- c(50,85)
+}
+
+## The predictor:
+
+  predictor <- retrieve(reanalysis,lon=lon,lat=lat)
+  predictor <- subset(predictor,it=it)
+  if (FUNX=='C.C.eq') predictor <- C.C.eq(predictor)
+  eof <- EOF(annual(predictor,nmin=6))
+
+## Get the predictand -> Y
+
+  load(paste(param,'.aaca.rda',sep=''))
+  Y <- eval(parse(text=paste(param,'.aaca',sep='')))
+  Y <- subset(Y,it=period)
+
+  ## Extract summer or winter
+  if (!is.null(it)) {
+    print(paste('Extract',paste(it,collapse='-')))
+    Y <- subset(Y,it=it)
+    nmin <- 28*length(it)
+  } else nmin=360
+  if (verbose) {print(paste('nmin=',nmin)); print(dim(Y))}
+  ## Remove suspect data:
+  cY <- coredata(Y); cY[cY > 500] <- NA; cY -> coredata(Y)
+
+
+## Estimate seasonal means & weed out stations with little data
+  Y4 <- annual(Y,FUN=FUN,nmin=6*28)
+  counts <- zoo(annual(Y,FUN='count',nmin=6*28))
+  ok <- apply(coredata(Y4),1,nv)
+  Y4 <- subset(Y4,it=ok>0)
+  nok <- apply(coredata(Y4),2,nv)
[…]

### 2016-09-08 AACA/weedaaca.R

@@ -1,0 +1,66 @@
+## Need to iterate a couple of times before all duplicates are removed.
+
+library(esd)
+
+#load('t2m.aaca.rda')
+load('pr.aaca.rda')
+t2m.aaca <- pr.aaca
+
+lonlat <- paste(round(lon(t2m.aaca),3),round(lat(t2m.aaca),3))
+n <- length(lonlat)
+dlonlat <- rownames(table(lonlat[duplicated(lonlat)]))
+
+## Index for stations to keep 
+keep <- rep(TRUE,n)
+
+for (i in 1:length(dlonlat)) {
+  is <- is.element(lonlat,dlonlat[i])
+  y <- subset(t2m.aaca,is=is)
+  nv <- apply(coredata(y),2,nv)
+  print(c(sum(is),loc(y),nv))
+  remove <- (1:n)[is]
+  if (sum(nv==max(nv))>1) {
+     remove <- remove[is.element(nv,max(nv))][-1]
+   } else
+     remove <- remove[nv==max(nv)]
+  keep[remove] <- FALSE
+  plot(y)
+}
+
+print(sum(keep))
+t2m.aaca <- subset(t2m.aaca,is=keep)
+t2m.aaca <- subset(t2m.aaca,it=c(1900,2015))
+esd::map(t2m.aaca,FUN='mean',cex=-2)
+
+#save(file='t2m.aaca.rda',t2m.aaca)
+save(file='pr.aaca.rda',pr.aaca)
+
+
+if (TRUE) {
+## Remove duplicated records:
+dy <- duplicated(stid(Y))
+d2 <- (1:dim(Y)[2])[dy]
+for (i in 1:length(d2))
+  print(apply(subset(t2m.aaca,is=c(d2[i],d2[i]-1)),2,nv))
+Y <- subset(Y,is=!dy)
+eval(parse(text=paste('Y -> ',param,'.aaca',sep='')))
+eval(parse(text=paste('save(file="',param,'.aaca.rda",',param,'.aaca)',sep='')))
+}
+
+param <- 'pr'
+FUN <- 'mean'
+it <- 'son'
+reanalysis <- 'air.mon.mean.nc'
+ 
+if (param=='precip') FUNX <- 'C.C.eq' else FUNX <- 'mean'
+
+
+## Remove unrealistically high amounts.
+## Get the predictand -> Y
+if (!exists('Y')) {
[…]

### 2016-09-19 ClimaTrans/climatrans.Rmd
Added run-off and dipper examples

@@
 
+### Function to evaluate spatio-temporal variance structure for SLP
+
+```{r}
+showceofvar <-  function(x,N,i,gcmnm.i,add=FALSE,xlim=NULL,ylim=NULL,verbose=FALSE,...) {
+  n <- length(x$mean.diff)
+  j <- 1:n
+  col <- rgb(j/n,abs(sin(pi*j/n)),(1-j/n),0.3)
+  if (is.null(xlim)) xlim <- c(-1,1)
+  if (is.null(ylim)) ylim <- c(1,N)
+
+  if (!add) {
+    par(bty="n")
+    par0 <- par()
+    plot(xlim,ylim,type="n",
+         ylab="GCM",xlab=expression(1- sigma[p*r*e]/sigma[r*e*f]),
+         main=paste("Diagnostics: common EOFs",attr(x,'variable')),
+         xlim=xlim,ylim=ylim,col="grey")
+    grid()
+    #legend(xlim[1],ylim[2],c("same sign","different sign"),
+    #       pch=c(19,21),bty="n",col="grey")
+    par(xpd=TRUE)
+    text(xlim[2],ylim[2],'EOF #',col='grey40',cex=0.8,pos=3)
+
+    par(new=TRUE,fig=c(0.85,0.95,0.70,0.85),mar=c(0,3,0,0),
+        cex.axis=0.7,yaxt="s",xaxt="n",las=1)
+     colbar <- rbind(1:n,1:n)
+    image(1:2,1:n,colbar,col=col)
+    par(fig=par0$fig,mar=par0$mar,cex.axis=par0$cex.axis,
+        yaxt=par0$yaxt,xaxt=par0$xaxt,las=par0$las,new=TRUE)
+    plot(xlim,ylim,type="n",xlim=xlim,ylim=ylim,
+         xlab='',ylab='',main='',sub='')
+    par(par0$new)
+  }
+  cex <- abs(0.1 - x$mean.diff)/0.1
+  pch <- rep(19,n); pch[cex < 0] <- 21
+  cex <- abs(cex); cex[cex > 2] <- 2
+  if (verbose) {
+     print('Mean difference:');print(x$mean.diff)
+     print('Ration of standard deviation');print(x$sd.ratio)
+     print('Size');print(cex)
+     print('col');print(col)
+     #points(x$mean.diff,1-x$sd.ratio,pch=pch,col='grey75',cex=1)
+  }
+
+  points(1-x$sd.ratio,rep(i,n),pch=pch,col=col,cex=cex)
+  text(xlim[1],i,gcmnm.i,pos=4,cex=0.7)
+}
+
+biasdiag <- function(predictor,param='slp',
+                    path="CMIP5.monthly/",pattern="slp_Amon_ens_",type='ncdf',
+                    period=c(1950,2015),rcp='rcp45',verbose=TRUE,
+                    lon=c(0,100),lat=c(65,90),select=NULL,xlim=NULL) {
+
+  ## Diagnose the common EOFs and effect of bias-adjustment
+
+  it <- month.abb[as.numeric(rownames(table(month(predictor))))]
+  if (verbose) print(paste('biasdiag, it=',it))
+
+  # Ensemble GCMs
+  path <- file.path(path,rcp,fsep = .Platform$file.sep)
+  ncfiles <- list.files(path=path,pattern=pattern,full.name=TRUE)
+  N <- length(ncfiles)
+
+  if (is.null(select)) select <- 1:N else
+                       N <- length(select)
+  if (verbose) {print('GCMs:'); print(path); print(ncfiles[select])}
+
+  d.y <- 6 # Number of leading EOFs to evalue
+  years <- 1900:2100
+  m <- length(years)
+  months <- rep(month(y)[1],m)
+  X <- matrix(rep(NA,N*m*d.y),N,m*d.y)
+  dim(X) <- c(d.y[2],N,m)
+  gcmnm <- rep("",N)
+  scorestats <- matrix(rep(NA,N*9),N,9)
+  colnames(scorestats) <- c("1-r.xval","mean.diff","sd.ratio","autocorr.ratio",
+                            "res.trend","res.K-S","res.ar1",'amplitude.ration',
+                            "1-R2")
+
+  t <- as.Date(paste(years,months,'01',sep='-'))
+
+  cols <- rgb(seq(1,0,length=100),rep(0,100),seq(0,1,length=100),0.15)
+
+  flog <- file("DSensemble.ceof-log.txt","at")
+
+  ## Set up a list environment to keep all the results
+  results <- list() 
+  if (verbose) print("loop...") 
+  for (i in 1:N) {
+    if (verbose) print(ncfiles[select[i]])
+    gcm <- retrieve(ncfile = ncfiles[select[i]],type=type,
+                          lon=range(lon(predictor)),
+                          lat=range(lat(predictor)),verbose=verbose)
+    if (!is.null(it)) {
+      if (verbose) print('Extract some months ot a time period')
+      if (verbose) print(it)
+      gcm <- subset(gcm,it=it)
+    }
+    #gcmnm[i] <- attr(gcm,'model_id')
+    gcmnm.i <- paste(attr(gcm,'model_id'),attr(gcm,'realization'),sep="-r")
+    gcmnm.i <- gsub(' ','',gcmnm.i)
+    gcmnm.i <- gsub('-','.',gcmnm.i)
+    print(gcmnm.i)
+    GCM <- aggregate(subset(gcm,it=month.abb[month(y)[1]]),year,FUN='mean') 
+    SLPGCM <- combine(predictor,GCM)
+    Z <- try(EOF(SLPGCM))
+    diag <- diagnose(Z)
+    showceofvar(diag,N,i,gcmnm.i,add=(i!=1),xlim=xlim,verbose=verbose)
+    eval(parse(text=paste('results$',it,'.',gcmnm.i,' <- diag',sep='')))
+  }
+  return(results)
+}
+
+```
+
 ##*Precipitation*
@@
 nd <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'nv')
-ndhr <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'count',threshold=100)
+ndhr100 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'count',threshold=100)
+ndhr50 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'count',threshold=50)
 q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wq95')
@@
 ```{r, fig.width=6, fig.height=6}
-plot(ndhr,new=FALSE)
+plot(ndhr100,new=FALSE)
+grid()
+```
+
+And more moderate events
+
+```{r, fig.width=6, fig.height=6}
+plot(ndhr50,new=FALSE)
 grid()
@@
 ```{r fig.width=6, fig.height=6}
-ndhr.cal <- data.frame(y = coredata(subset(subset(ndhr,is=2),it=c(1949,2004))),
+ndhr.cal <- data.frame(y = coredata(subset(subset(ndhr100,is=2),it=c(1949,2004))),
                        X = coredata(subset(eof.es,it=c(1949,2004))))
@@
 lines(1949:2004,exp(predict(ds.ndhr)),col='red')
+```
+
+Model the number with heavy rainfall (more than 50 mm/day) in Delhi and Bangalore.
+
+```{r fig.width=6, fig.height=6}
+## Delhi
+ndhr50.1.cal <- data.frame(y = coredata(subset(subset(ndhr50,is=1),it=c(1949,2004))),
+                       X = coredata(subset(eof.es,it=c(1949,2004))))
+ds.ndhr50.1 <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr50.1.cal,family="poisson")
+plot(1949:2004,ndhr50.1.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[1])
+grid()
+lines(1949:2004,exp(predict(ds.ndhr50.1)),col='red')
+
+# Bangalore
+ndhr50.3.cal <- data.frame(y = coredata(subset(subset(ndhr50,is=3),it=c(1949,2004))),
+                       X = coredata(subset(eof.es,it=c(1949,2004))))
+ds.ndhr50.3 <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr50.3.cal,family="poisson")
+plot(1949:2004,ndhr50.3.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[3])
+grid()
+lines(1949:2004,exp(predict(ds.ndhr50.3)),col='red')
 ```
@@
       col=c('red','grey'),lwd=c(3,10),bty='n')
+```
+
+## Examine the common EOFs for SLP over the Indian sub-continent
+
+```{r}
+print('Compare common EOFs')
+diag.slp <- biasdiag(slp)
 ```

### 2016-09-19 ClimaTrans/dipper.Rmd
Added run-off and dipper examples

---
title: "Dipper downscaling"
author: "Rasmus Benestad"
date: "August 22, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example of downscaling a population of species

We can use the assumption of Poisson-type distribution for a count of e.g. population or events such as the bird "dipper" that lives in a colony in southern Norway. The dipper is a type of bird that lives in water falls. Its population is affected by the temperature and weather the water freezes during the winter.

This analysis was carried out for M. Gamelon, who provided the data on the dipper population

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r dipper}
## R-script to analyse dipper population

library(esd)
library(ncdf4)

examine <- FALSE

dipper <- read.table('~/data/dipper.csv',sep=',',header=TRUE)

dipper <- zoo(x=dipper[[2]],order.by=dipper[[1]])
y <- as.station(dipper,loc='Lyngdalselva',
                lon=7,lat=58.5,alt=520,
                param='population',unit='count',
                reference='Marlène Gamelon')

predictor <- retrieve("air.mon.mean.nc",lon=c(-1,17),lat=c(57,62))
## Only use the winter temperature:
predictor <- aggregate(subset(predictor,it='djf'),year,FUN='mean')

if (examine) {
  ## Check the distribution
  n <- seq(0,150,by=10); mu <- mean(dipper)
  hist(coredata(dipper),breaks=n,freq=FALSE,col='grey')
  lines(x,dpois(x,mu),lwd=2,col='red')

  ## Single ESD for inspection
  X <- EOF(predictor)
  z <- DS(y,X)
  plot(z,new=FALSE)
}
```

```{r}
## Downscale the dipper population directly based on the large-scale
## annual mean temperature

if (!file.exists('dipper.Z.rcp45.rda')) {
  Z.rcp45 <- DSensemble.annual(y,biascorrect=TRUE,
                               predictor=predictor,
                               lon=c(-1,17),lat=c(57,62),
                               abscoords=TRUE)
  save(file='dipper.Z.rcp45.rda',Z.rcp45)
} else load('dipper.Z.rcp45.rda')

if (!file.exists('dipper.Z.rcp85.rda')) {
  Z.rcp85 <- DSensemble.annual(y,biascorrect=TRUE,
                               rcp="rcp85",predictor=predictor,
                               lon=c(-1,17),lat=c(57,62),
                               abscoords=TRUE)
  save(file='dipper.Z.rcp85.rda',Z.rcp85)
} else load('dipper.Z.rcp85.rda')

if (!file.exists('dipper.Z.rcp26.rda')) {
  Z.rcp26 <- DSensemble.annual(y,biascorrect=TRUE,
                               rcp="rcp26",predictor=predictor,
                               lon=c(-1,17),lat=c(57,62),
                               abscoords=TRUE)
  save(file='dipper.Z.rcp26.rda',Z.rcp26)
} else load('dipper.Z.rcp26.rda')

year <- year(Z.rcp45)
ci90.rcp45 <- apply(coredata(Z.rcp45),1,quantile,
                    probs=c(0.05,0.95),na.rm=TRUE)
ci90.rcp26 <- apply(coredata(Z.rcp26),1,quantile,
                    probs=c(0.05,0.95),na.rm=TRUE)
ci90.rcp85 <- apply(coredata(Z.rcp85),1,quantile,
                    probs=c(0.05,0.95),na.rm=TRUE)

par(bty='n')
plot(range(year),range(ci90.rcp45,ci90.rcp26,ci90.rcp85,y),
     type='n',xlab='',ylab='Population',main='Dipper',
     sub=loc(y))
grid()
polygon(c(year,rev(year)),c(ci90.rcp85[1,],rev(ci90.rcp85[2,])),
                            col=rgb(1,0.5,0,0.3),border=rgb(1,0.5,0))
polygon(c(year,rev(year)),c(ci90.rcp45[1,],rev(ci90.rcp45[2,])),
                            col=rgb(0.5,1,0,0.3),border=rgb(0.5,1,0))
polygon(c(year,rev(year)),c(ci90.rcp26[1,],rev(ci90.rcp26[2,])),
                            col=rgb(0,0.5,1,0.3),border=rgb(0,0.5,1))
lines(year(y),coredata(y),lwd=4,pch=19,type='b')

legend(1900,200,c('RCP2.6','RCP4.5','RCP8.5'),
       col=c(rgb(0,0.5,1),rgb(0.5,1,0),rgb(1,0.5,0)),
       lwd=7,lty=1,bty='n')
legend(1905,170,'observed',lwd=4,pch=19,lty=1,bty='n')

```

### 2016-09-22 ClimaTrans/climatrans.Rmd
ClimaTrans updates

@@
 library(ncdf)
+setwd('~/R')
 #rm(list=ls())
@@
     if (plot) {
-        if (new) 
-            dev.new()
         par(bty = "n")
@@
             0, 0, 0.3), lwd = 3)
-        if (new) 
-            dev.new()
         djf.sd <- sd(coredata(djf), na.rm = TRUE)
@@
         grid()
-        if (new) 
-            dev.new()
         par(bty = "n")
@@
   points(1-x$sd.ratio,rep(i,n),pch=pch,col=col,cex=cex)
-  text(xlim[1],i,gcmnm.i,pos=4,cex=0.7)
-}
-
-biasdiag <- function(predictor,param='slp',
-                    path="CMIP5.monthly/",pattern="slp_Amon_ens_",type='ncdf',
+  #print(paste(' > ',i,i,gcmnm.i[i]))
+  text(xlim[1],i,gcmnm.i[i],pos=4,cex=0.6)
+}
+
+biasdiag <- function(predictor,param='slp',it=NULL,
+                    path="CMIP5.monthly/",pattern="psl_Amon_ens_",type='ncdf4',
                     period=c(1950,2015),rcp='rcp45',verbose=TRUE,
@@
 
-  it <- month.abb[as.numeric(rownames(table(month(predictor))))]
   if (verbose) print(paste('biasdiag, it=',it))
@@
 
-  d.y <- 6 # Number of leading EOFs to evalue
+  nceofs <- 6 # Number of leading EOFs to evalue
   years <- 1900:2100
   m <- length(years)
-  months <- rep(month(y)[1],m)
-  X <- matrix(rep(NA,N*m*d.y),N,m*d.y)
-  dim(X) <- c(d.y[2],N,m)
-  gcmnm <- rep("",N)
+  months <- rep(month(nceofs),m)
+  X <- matrix(rep(NA,N*m*nceofs),N,m*nceofs)
+  dim(X) <- c(nceofs,N,m)
+  gcmnm.i <- rep("",N)
   scorestats <- matrix(rep(NA,N*9),N,9)
@@
   results <- list() 
+  if (!file.exists('diag.ceof.slp.rda')) {
     if (verbose) print("loop...") 
     for (i in 1:N) {
+
       if (verbose) print(ncfiles[select[i]])
@@
         gcm <- subset(gcm,it=it)
-    }
-    #gcmnm[i] <- attr(gcm,'model_id')
-    gcmnm.i <- paste(attr(gcm,'model_id'),attr(gcm,'realization'),sep="-r")
-    gcmnm.i <- gsub(' ','',gcmnm.i)
-    gcmnm.i <- gsub('-','.',gcmnm.i)
-    print(gcmnm.i)
-    GCM <- aggregate(subset(gcm,it=month.abb[month(y)[1]]),year,FUN='mean') 
+        gcm <- subset(gcm,it=range(year(predictor)))
+      }
+      gcmnm <- paste(attr(gcm,'model_id'),attr(gcm,'realization'),sep="-r")
+      gcmnm <- gsub(' ','',gcmnm)
+      gcmnm <- gsub('-','.',gcmnm)
+      gcmnm.i[i] <- gcmnm
+      print(gcmnm.i[i])
+      GCM <- aggregate(gcm,year,FUN='mean') 
       SLPGCM <- combine(predictor,GCM)
@@
       showceofvar(diag,N,i,gcmnm.i,add=(i!=1),xlim=xlim,verbose=verbose)
-    eval(parse(text=paste('results$',it,'.',gcmnm.i,' <- diag',sep='')))
-  }
+      eval(parse(text=paste('results$member.',i,'.',gcmnm.i[i],' <- diag',sep='')))
+    } 
+    save(file='diag.ceof.slp.rda',results,gcmnm.i)
+  } else {
+    load('diag.ceof.slp.rda')
+    N <- length(gcmnm.i)
+    for (i in 1:N) {
+      #print(paste(' < ', i,gcmnm.i[i]))
+      showceofvar(results[[i]],N,i,gcmnm.i,add=(i!=1),xlim=xlim,verbose=verbose)
+    }
+  }
+  dev.copy2pdf(file=paste('diag.ceof.slp.india.pdf',sep=''))
   return(results)
@@
 ds.ndhr <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr.cal,family="poisson")
-plot(1949:2004,ndhr.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[2])
+plot(1949:2004,ndhr.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr100)[2])
 grid()
@@
 ds.ndhr50.1 <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr50.1.cal,family="poisson")
-plot(1949:2004,ndhr50.1.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[1])
+plot(1949:2004,ndhr50.1.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr50)[1])
 grid()
@@
 ds.ndhr50.3 <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr50.3.cal,family="poisson")
-plot(1949:2004,ndhr50.3.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[3])
+plot(1949:2004,ndhr50.3.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr50)[3])
 grid()
@@
 ```{r}
-print(paste('Mean number=',mean(ndhr[,2],na.rm=TRUE),'Variance=',var(ndhr[,2],na.rm=TRUE),
-            'should equal',mean(ndhr[,2],na.rm=TRUE),'for Poisson process'))
-h <- hist(coredata(ndhr[,2]),freq=FALSE,col='grey',
-     main=loc(ndhr)[2])
-lines(h$mids-0.5,dpois(x=h$mids-0.5,lambda=mean(ndhr[,2],na.rm=TRUE)),lwd=2,col='red')
+print(paste('Mean number=',mean(ndhr100[,2],na.rm=TRUE),'Variance=',var(ndhr100[,2],na.rm=TRUE),
+            'should equal',mean(ndhr100[,2],na.rm=TRUE),'for Poisson process'))
+h <- hist(coredata(ndhr100[,2]),freq=FALSE,col='grey',
+     main=loc(ndhr100)[2])
+lines(h$mids-0.5,dpois(x=h$mids-0.5,lambda=mean(ndhr100[,2],na.rm=TRUE)),lwd=2,col='red')
 grid()
@@
 
+Common EOFs can be used to assess the GCMs' ability to reproduce the spatio-temporal structure in covariance in cinnection to mean sea-level pressure (SLP) variations over the Indian sub-continent.
+In orther words, whether the GCMs reproduce the observed circulation features such as wind patterns associated with the monsoon.
+The code below compares the statistical character of the odel results.
+
 ```{r}
 print('Compare common EOFs')
-diag.slp <- biasdiag(slp)
+diag.slp <- biasdiag(slp,it=month.abb[season],verbose=FALSE)
 ```

### 2016-09-27 CixPAG/CixPAG-imd.Rmd

--- 2016-09-19 ClimaTrans/dipper.Rmd
@@
 ---
-title: "Dipper downscaling"
+title: "CixPAG"
 author: "Rasmus Benestad"
-date: "August 22, 2016"
+date: "August 3, 2016"
 output: pdf_document
@@
 
-## Example of downscaling a population of species
-
-We can use the assumption of Poisson-type distribution for a count of e.g.
-population or events such as the bird "dipper" that lives in a colony in southern Norway.
-The dipper is a type of bird that lives in water falls.
-Its population is affected by the temperature and weather the water freezes during the winter.
-
-This analysis was carried out for M. Gamelon, who provided the data on the dipper population
-
 ## R Markdown
@@
 
-```{r dipper}
-## R-script to analyse dipper population
-
+```{r esd}
 library(esd)
-library(ncdf4)
-
-examine <- FALSE
-
-dipper <- read.table('~/data/dipper.csv',sep=',',header=TRUE)
-
-dipper <- zoo(x=dipper[[2]],order.by=dipper[[1]])
-y <- as.station(dipper,loc='Lyngdalselva',
-                lon=7,lat=58.5,alt=520,
-                param='population',unit='count',
-                reference='Marlène Gamelon')
-
-predictor <- retrieve("air.mon.mean.nc",lon=c(-1,17),lat=c(57,62))
-## Only use the winter temperature:
-predictor <- aggregate(subset(predictor,it='djf'),year,FUN='mean')
-
-if (examine) {
-  ## Check the distribution
-  n <- seq(0,150,by=10); mu <- mean(dipper)
-  hist(coredata(dipper),breaks=n,freq=FALSE,col='grey')
-  lines(x,dpois(x,mu),lwd=2,col='red')
-
-  ## Single ESD for inspection
-  X <- EOF(predictor)
-  z <- DS(y,X)
-  plot(z,new=FALSE)
-}
 ```
 
-```{r}
-## Downscale the dipper population directly based on the large-scale
-## annual mean temperature
+## Reading data predictand
 
-if (!file.exists('dipper.Z.rcp45.rda')) {
-  Z.rcp45 <- DSensemble.annual(y,biascorrect=TRUE,
-                               predictor=predictor,
-                               lon=c(-1,17),lat=c(57,62),
-                               abscoords=TRUE)
-  save(file='dipper.Z.rcp45.rda',Z.rcp45)
-} else load('dipper.Z.rcp45.rda')
-
-if (!file.exists('dipper.Z.rcp85.rda')) {
-  Z.rcp85 <- DSensemble.annual(y,biascorrect=TRUE,
-                               rcp="rcp85",predictor=predictor,
-                               lon=c(-1,17),lat=c(57,62),
-                               abscoords=TRUE)
-  save(file='dipper.Z.rcp85.rda',Z.rcp85)
-} else load('dipper.Z.rcp85.rda')
-
-if (!file.exists('dipper.Z.rcp26.rda')) {
-  Z.rcp26 <- DSensemble.annual(y,biascorrect=TRUE,
-                               rcp="rcp26",predictor=predictor,
-                               lon=c(-1,17),lat=c(57,62),
-                               abscoords=TRUE)
-  save(file='dipper.Z.rcp26.rda',Z.rcp26)
-} else load('dipper.Z.rcp26.rda')
-
-year <- year(Z.rcp45)
-ci90.rcp45 <- apply(coredata(Z.rcp45),1,quantile,
-                    probs=c(0.05,0.95),na.rm=TRUE)
-ci90.rcp26 <- apply(coredata(Z.rcp26),1,quantile,
-                    probs=c(0.05,0.95),na.rm=TRUE)
-ci90.rcp85 <- apply(coredata(Z.rcp85),1,quantile,
-                    probs=c(0.05,0.95),na.rm=TRUE)
-
-par(bty='n')
-plot(range(year),range(ci90.rcp45,ci90.rcp26,ci90.rcp85,y),
-     type='n',xlab='',ylab='Population',main='Dipper',
-     sub=loc(y))
-grid()
-polygon(c(year,rev(year)),c(ci90.rcp85[1,],rev(ci90.rcp85[2,])),
-                            col=rgb(1,0.5,0,0.3),border=rgb(1,0.5,0))
-polygon(c(year,rev(year)),c(ci90.rcp45[1,],rev(ci90.rcp45[2,])),
-                            col=rgb(0.5,1,0,0.3),border=rgb(0.5,1,0))
-polygon(c(year,rev(year)),c(ci90.rcp26[1,],rev(ci90.rcp26[2,])),
-                            col=rgb(0,0.5,1,0.3),border=rgb(0,0.5,1))
-lines(year(y),coredata(y),lwd=4,pch=19,type='b')
-
-legend(1900,200,c('RCP2.6','RCP4.5','RCP8.5'),
-       col=c(rgb(0,0.5,1),rgb(0.5,1,0),rgb(1,0.5,0)),
-       lwd=7,lty=1,bty='n')
-legend(1905,170,'observed',lwd=4,pch=19,lty=1,bty='n')
-
+```{r pressure, echo=FALSE}
+pt <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
+tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
+map(pt)
 ```

### 2016-09-27 ClimaTrans/climatrans.Rmd

@@
 
-Model the number with heavy rainfall (more than 50 mm/day) in Delhi and Bangalore.
+Model the number with heavy rainfall (more than 50 mm/day) in Delhi and Bangalore. The results for Delhi and Bangalore also indicate a failure to capture the number of events.
 
@@
 ds.ndhr50.1 <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr50.1.cal,family="poisson")
-plot(1949:2004,ndhr50.1.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr50)[1])
+plot(1949:2004,ndhr50.1.cal$y,pch=19,ylab=expression(N(X > 50*mm)),xlab='',main=loc(ndhr50)[1])
 grid()
@@
 ds.ndhr50.3 <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr50.3.cal,family="poisson")
-plot(1949:2004,ndhr50.3.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr50)[3])
+plot(1949:2004,ndhr50.3.cal$y,pch=19,ylab=expression(N(X > 50*mm)),xlab='',main=loc(ndhr50)[3])
 grid()
@@
 ```{r}
-print(paste('Mean number=',mean(ndhr100[,2],na.rm=TRUE),'Variance=',var(ndhr100[,2],na.rm=TRUE),
-            'should equal',mean(ndhr100[,2],na.rm=TRUE),'for Poisson process'))
+print(paste('Mean number=',round(mean(ndhr100[,2],na.rm=TRUE),1),
+            'Variance=',round(var(ndhr100[,2],na.rm=TRUE),1),
+            'should equal',round(mean(ndhr100[,2],na.rm=TRUE),1),'for Poisson process'))
 h <- hist(coredata(ndhr100[,2]),freq=FALSE,col='grey',
@@
 lines(h$mids-0.5,dpois(x=h$mids-0.5,lambda=mean(ndhr100[,2],na.rm=TRUE)),lwd=2,col='red')
+grid()
+```
+
+Also check whether the number of events in Delhi and Bangalore approximately follow a Poisson distribution.
+
+```{r}
+# Delhi
+print(paste('Mean number=',round(mean(ndhr50[,1],na.rm=TRUE),1),
+            'Variance=',round(var(ndhr50[,1],na.rm=TRUE),1),
+            'should equal',round(mean(ndhr50[,1],na.rm=TRUE),1),'for Poisson process'))
+h <- hist(coredata(ndhr50[,1]),freq=FALSE,col='grey',
+     main=loc(ndhr50)[1])
+lines(h$mids-0.5,dpois(x=h$mids-0.5,lambda=mean(ndhr50[,1],na.rm=TRUE)),lwd=2,col='red')
+grid()
+
+# Bangalore
+print(paste('Mean number=',round(mean(ndhr50[,3],na.rm=TRUE),1),
+            'Variance=',round(var(ndhr50[,3],na.rm=TRUE),1),
+            'should equal',round(mean(ndhr50[,3],na.rm=TRUE),1),'for Poisson process'))
+h <- hist(coredata(ndhr50[,3]),freq=FALSE,col='grey',
+     main=loc(ndhr50)[3])
+lines(h$mids-0.5,dpois(x=h$mids-0.5,lambda=mean(ndhr50[,3],na.rm=TRUE)),lwd=2,col='red')
+grid()
+```
+
+The number of days with heavy rain is deviating somewhat from the Poisson distribution for the Indian data.
+We can compare the assumption of Poisson distribution with Norwegian data (all year) with known high quality (we have the metadata), and the results suggest a closer fit, although this data too is over-dispersed.
+
+```{r}
+data(bjornholt)
+ndhr30 <- aggregate(bjornholt,year,'count',threshold=30)
+print(paste('Mean number=',round(mean(ndhr30,na.rm=TRUE),1),'Variance=',round(var(ndhr30,na.rm=TRUE),1),
+            'should equal',round(mean(ndhr30,na.rm=TRUE),1),'for Poisson process'))
+h <- hist(coredata(ndhr30),freq=FALSE,col='grey',
+     main=loc(ndhr30))
+lines(h$mids-0.5,dpois(x=h$mids,lambda=mean(ndhr30,na.rm=TRUE)),lwd=2,col='red')
 grid()
@@
 
-```{r}
+```{r fig.width=12, fig.height=20}
 print('Compare common EOFs')
@@
 
+This test shows a comparison of the standard deviation of the principal components (PCs) associated with the reanalysis SLP and the GCM SLP over the sub-indian continent (region shown in the maps above) and the year-to-year variability in the monsoon season mean SLP.
+A good reproduction of the spatio-temporal variance structure for the SLP is expected to yield leading EOFs with similar variance and hence the blue symbols should be close to zero for $1 - \rho_{gcm}/\rho_{ref}$ along the x-axis.
+The results suggest otherwise, with large differences in the standard deviation connected to the PCs, and indicate that all GCMs have difficulties reproducing the monsoon features in the monsoon SLP.
+
 ## Example of station record
@@
 The format of the station records bears a structure from the past where it was more important to take up little space rather than being human readable.
+This cryptic and non-human readable way of recording data is associated several risks of error introduction, as it is easy to mistype some of the '0' entries and it is hard to spot obvious mistakes (e.g.
+unreasonable and unphysical values).
+Furthermore, reading data of this type of format also is prone to errors, as the data is stored in a way that may vary from time to time.
+There is also a high risk of the introduction of errors when reading such data.
 

### 2016-09-27 ClimaTrans/runoff-vikedal.Rmd

---
title: "Runoff-vikedal"
author: "Rasmus Benestad"
date: "September 5, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example of downscaling high runoff events

This example shows how the number of events with high runoff (more than 30 cubic meters per second) is connected to precipitation statistics that can be downscaled through empirical-statistical means. Such a connection indicates that it may be possible to downscale number of high flow events through a hybrid scheme based on parameter downscaling (of the wet-day mean precipitation and wet day frequency) and number-of-event modelling.

The example here is for Vikedal, a small catchment in western Norway.

The hydrological data was provided by Yan Lin from NIVA (<http://www.niva.no>).

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r getdata}
library(esd)
# VIKEDAL (met station no. 46850, LAT 59.5558N, LON 5.9955E
hydroIO <- read.table("~/Dropbox/data/Vikedal.csv",sep=";",header=TRUE)
loc <- 'Vikedal'
lon <- 5.9955
lat <- 59.5558
alt <- 159
stid <- 46850

hydroIO$Date <- as.character(hydroIO$Date)      # Fix
hydroIO$Date <- paste(substr(hydroIO$Date,7,10),substr(hydroIO$Date,4,5),substr(hydroIO$Date,1,2),sep='-')
hydroIO$Temperature.C <- hydroIO$Temperatrue.C  # Fix
P <- zoo(x=hydroIO$Precipitation..mm,order.by=as.Date(hydroIO$Date))
t2m <- zoo(x=hydroIO$Temperature.C,order.by=as.Date(hydroIO$Date))
u <- zoo(x=hydroIO$Flow..m3.s,order.by=as.Date(hydroIO$Date))
```

```{r plotdata}
par(bty="n")
plot(merge(P,u,as.anomaly(t2m)),plot.type='single',col=c("steelblue","black","red"),lwd=c(3,1,2),new=FALSE)
#dev2bitmap('hydroex-daily.png',res=150)

Pmm <- aggregate(P,as.yearmon,FUN='mean')
Tmm <- aggregate(t2m,as.yearmon,FUN='mean')
umm <- aggregate(u,as.yearmon,FUN='mean')

#plot(coredata(P),coredata(u))

cal <- data.frame(u=coredata(umm),t2m=coredata(Tmm),Pmm=coredata(Pmm))
fit <- lm(u ~ t2m + Pmm,data=cal)
print(summary(fit))

par(bty="n")
plot(coredata(umm),predict(fit),
     main='Monthly river flow predicted from monthly T and P',
     ylim=range(coredata(Pmm)),
     xlab=expression(u[hydro]),ylab=expression(u==f(T[2*m],P)))
lines(range(coredata(Pmm)),range(coredata(Pmm)),col="red",lty=2)
grid()
#dev2bitmap(file='hydroex.png',res=150)

temp <- as.station(t2m,lon=lon,lat=lat,alt=alt,unit='deg C',param='t2m',loc=loc,stid=stid)

if (FALSE) {
z.t2m <- try(DSensemble.t2m(temp,biascorrect=TRUE,plot=TRUE))

#dev2bitmap('hydroex-dse-z.t2m.png',res=150)
save(file='hydroex-dse-z.t2m.rda',z.t2m)
}

precip <- as.station(P,lon=lon,lat=lat,alt=alt,unit='mm/day',param='precip',loc=loc,stid=stid)
if (FALSE) {
z.pre <- try(DSensemble.precip(precip,biascorrect=TRUE,plot=TRUE))
#dev2bitmap('hydroex-dse-z.pre.png',res=150)
save(file='hydroex-dse-z.pre.rda',z.pre)
}

flow <- as.station(u,lon=lon,lat=lat,alt=alt,unit='m^3/s',param='u',loc=loc,stid=stid)

muam <- annual(precip,FUN='exceedance',threshold=1)
fwam <- annual(precip,FUN='wetfreq',threshold=1)
uamx <- annual(flow,FUN='count',threshold=30)
attr(uamx,'variable') <- 'n(X > m^3/s)'
attr(uamx,'unit') <- ' '

am <- data.frame(u=coredata(uamx),mu=coredata(muam),fw=coredata(fwam))
emu <- glm(u ~ mu + fw, data=am, family='poisson')

plot(uamx,lwd=3,ylim=c(0,70),col='red',new=FALSE)
lines(zoo(exp(predict(emu)),order.by=index(uamx)))

text(1980,70,pos=4,'Number of annual events: u > 30 m³/s')
legend(1980,65,c("hydrologic model",expression(f(mu,f[w]))),
       lty=1,col=c("red","black"),lwd=c(3,1),bty="n")
#dev2bitmap('hydroex-u.gt.30.png',res=150)
```

### 2016-10-17 EU-Circle/eu-circle_case-study-3.Rmd
moved ABC4CDE to own repository

@@
 ---------------------------------
-```{r,eval=FALSE}
-## Get MIDAS preciptation data that has laready been retrieved 
-## from the MIDAS data base (UK Met Office)
-load('pr.eu-circle-torbay.rda')
-```
 
@@
 
-```{r}
+```{r function-probability}
 PrexpPr <- function(mu,fw,x0=10) {
@@
   attr(Pr,'unit') <- 'probability'
-  attr(Pr,'longname') <- paste('Probability of exceeding',x0)
+  attr(Pr,'longname') <- paste('Probability of exceeding',x0,'mm')
   class(Pr) <- class(mu)
@@
 
-```{r}
+```{r read-precip-data}
+## Get MIDAS preciptation data that has laready been retrieved 
+## from the MIDAS data base (UK Met Office)
 load('pr.eu-circle-torbay.rda') 
@@
 x <- subset(pr,is=(nok> 15000))
+
+## Quality check: remove outliers days with more than 300 mm - set to NA
+xc <- coredata(x); xc[xc > 300] <- NA; xc -> coredata(x); rm('xc')
 
@@
 
-```{r}
+```{r ecad-data}
  ss <- select.station(param='precip',src='ecad',cntr=c('United Kingdom','France'),lat=c(45,55),lon=c(-10,0),nmin=50)
@@
 
-```{r}
+```{r combine-stations}
 mu <- combine.stations(mu,mu.ecad)
@@
 
-```{r}
+```{r fill in missing data}
 ##Fill in the remaining gaps using a PCA-based interpolation scheme 
@@
 
-```{r}
-Pr.gt.10 <- PrexpPr(mu,fw,x0=10)
-plot(Pr.gt.10); grid()
-for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.10[,i]))
+```{r probabilities}
+Pr.gt.30 <- PrexpPr(mu,fw,x0=30)
+plot(Pr.gt.30); grid()
+for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.30[,i]))
 ```
@@
 
-```{r}
+```{r aggregate}
 ## Estimate the mean seasonal cycle in the wet-day mean precipitaiton
@@
 
-```{r}
+```{r precipitation-intensity}
 ## Show the mean seasonal cycle in the wet-day mean precipitation
@@
 
-```{r}
+```{r wet-day frequency}
 ## Show the mean seasonal cycle in the wet-day frequency
@@
 
-```{r}
+```{r sesonal-precipitation}
 ## Show the mean seasonal cycle in the total precipitation
@@
 
-```{r}
+```{r winter-rain}
 ## Estimate the PCA for the annual cycle to identify the general features:
@@
 
-```{r}
+```{r downscale-ensemble}
 if (!file.exists('dse.mu.eu-circle-worstcase.rda')) {
   dse.mu <- DSensemble.mu.worstcase(x,biascorrect=TRUE,mask=TRUE,
-                                    lon=c(-80,20),lat=c(10,60),rel.cord=FALSE,verbose=TRUE)
+                                lon=c(-80,20),lat=c(10,60),
+                                rel.cord=FALSE,verbose=TRUE)
   save(file='dse.mu.eu-circle-worstcase.rda',dse.mu)
 } else load('dse.mu.eu-circle-worstcase.rda')
-plot(dse.mu[[1]])
+#plot(dse.mu[[1]])
+plot.zoo(dse.mu[[1]],plot.type='single',col=rgb(1,0,0,0.05))
+grid()
 ```
@@
 index(newlyn) <- year(newlyn)
-index(sonel) <- year(sonal)
+index(sonel) <- year(sonel)
 index(gloss) <- year(gloss)
@@
 ```{r slp-trend-map}
-slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg_v12.0.nc',lon=c(-6,0),lat=c(48,52))
+slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg_v13.1.nc',lon=c(-6,0),lat=c(48,52))
 slp.eobs <- 10000*slp.eobs ## The data have some funny scaling factor
@@
 ```{r predictand}
-pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(-6,-1),lat=c(50,52))
+pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v13.0.nc',lon=c(-6,-1),lat=c(50,52))
 mu <- annual(pr.eobs,FUN='wetmean')

### 2016-10-17 KSS/ds-pr4KSS.Rmd
moved ABC4CDE to own repository

--- 2016-04-07 KSS/dr-pr4KSS.Rmd
@@
 ## Represent the predictors in terms of EOFs
-eof.t2m.w <- EOF(t2m.w)
+eof.t2m.w <- EOF(t2m.w)d
 eof.t2m.c <- EOF(t2m.c)

### 2016-10-17 KSS/ds-t2m4KSS.Rmd
moved ABC4CDE to own repository

---
title: "Downscale t2m for KSS"
author: "Rasmus Benestad"
date: "October 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r esd}
library(esd)
```

## Function definitions:

```{r}
downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
                      period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
                      lon=c(-20,40),lat=c(50,80),ip=1:6,n=6,
                      rel.cord=FALSE,select=NULL) {

  print('downscale')

## Use a time and space window:  
  Y <- subset(Y,it=period)
  Y <- subset(Y,is=list(lat=lat,lon=lon))

## Estimate seasonal means & weed out stations with little data
  Y4 <- subset(as.4seasons(Y,FUN=FUN),it=it)
  ok <- apply(coredata(Y4),1,nv)
  Y4 <- subset(Y4,it=ok>0)
  nok <- apply(coredata(Y4),2,nv)
  Y4 <- subset(Y4,is=nok>15)

  print(paste(round(100*sum(!is.finite(Y4))/length(!is.finite(Y4))),'% missing',
  sep=''))
  if (plot) map(Y,FUN=FUN,cex=-2)

  nmiss <- round(100*sum(!is.finite(Y4))/length(!is.finite(Y4)))
  #print(paste(nmiss,'% missing',sep=''))

## Fill missing data using PCA-based regression
  Z <- pcafill(Y4)

  pca <- PCA(Z,n=n)
  if (plot) plot(pca)

## Downscale results
  print('DSensemble')
  dse.pca <- DSensemble(pca,predictor=predictor,FUNX=FUNX,verbose=verbose,
                        biascorrect=TRUE,rcp=rcp,ip=ip,select=select,
                        lon=lon,lat=lat,rel.cord=rel.cord,it=it)

  attr(dse.pca,'N.missing') <- nmiss
  invisible(dse.pca)
}
```

## Retrieving the data

Get the predictand data: daily temperature from ECAD:

```{r predictand-data}
if (!file.exists('t2m.nordic.rda')) {
  ss <- select.station(param='t2m',lon=c(5,30),lat=c(55,72),src='ecad',nmin=50)
  X <- station(ss)
  save(X,file='t2m.nordic.rda')
} else load('t2m.nordic.rda')
```

```{r}
## Only use the most recent 66 years 
X <- subset(X,it=c(1950,2015))
## Remove records with many missing datapoints
nv <- apply(coredata(X),2,'nv')
```

```{r}
#-----------------------------------------------------------------------
# Define season and parameter

param <- 't2m'
FUN <- 'mean'
reanalysis <- 'air.mon.mean.nc'
lon=c(0,100); lat=c(65,90)

if (param=='precip') FUNX <- 'C.C.eq' else FUNX <- 'mean'

## Get the predictand -> Y
if (!exists('Y')) {
  load(paste(param,'.nordic.rda',sep=''))
}

for (it in c('djf','mam','jja','son')) {

  ## Get the large-scale predictor:
  if (!exists('predictor')) {
    T2M <- retrieve(reanalysis,lon=lon,lat=lat)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)
  } else if (length(month(subset(predictor,it=it)))==0)
    predictor <- subset(as.4seasons(T2M,FUNX=FUNX),it=it)

  print(paste('Generating dse.kss.t2m.rcp45.',it,'.rda',sep=''))

  ## Carry out the downscaling:
  if (!file.exists(paste('dse.kss.t2m.rcp45.',it,'.rda',sep=''))) {
    Z <- downscale(Y,predictor,it,param,FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste('dse.kss.t2m.rcp45.',it,'.rda',sep=''),Z)
  }

  print(paste('Generating dse.kss.t2m.rcp26.',it,'.rda',sep=''))

  if (!file.exists(paste('dse.kss.t2m.rcp26.',it,'.rda',sep=''))) {
    Z <- downscale(Y,predictor,it,param,rcp='rcp26',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste('dse.kss.t2m.rcp26.',it,'.rda',sep=''),Z)
  }

   print(paste('Generating dse.kss.t2m.rcp85.',it,'.rda',sep=''))

  if (!file.exists(paste('dse.kss.t2m.rcp85.',it,'.rda',sep=''))) {
    Z <- downscale(Y,predictor,it,param,rcp='rcp85',
                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
    save(file=paste('dse.kss.t2m.rcp85.',it,'.rda',sep=''),Z)
  }
}
print('--- Completed downscaling ---')
```

## Post-processing

```{r}
## source the functions for gridding
demo(pca2eof,ask=FALSE)
load('dse.kss.t2m.rcp45.djf.rda')
## Filter and reduce data size by reteining only the 4 leading PCA
Z <- subset(Z,ip=1:4)
## Convert the PCA object to EOF 
Z <- as.eof.dsensemble.pca(Z)
save(Z,file='dse.KSS.t2m.rcp45.djf.rda')
map(Z,it=c(2040,2060))
```

### 2016-10-17 EU-Circle/eu-circle_case-study-3.Rmd
Added DS scripts for temperature

@@
 ```{r trend-map}
-triangle <- subset(Z,is=c(2,3,7))
+## Fudge it - pick some windstations, but do not use their data. Reset the
+## coordinates to select a triangle of SLP from the EOBS data.
+ss <- select.station(param='slp',src='ecad',cntr=c('France','UNITED KINGDOM'))
+fx <- station(ss)
+triangle <- subset(fx,is=1:3)
 ## The stations are not situated quite where there is valid slp data in the EOBS dataset
@@
 plot(annual(uv,nmin=300),main='zonal and meridional wind components',
-     sub='Estimated from triangulation of mean sea-level pressure from EOBS')
+     sub='Estimated from triangulation of mean sea-level pressure from EOBS',zoom=7)
 ```

### 2016-10-17 CixPAG/CixPAG-imd.Rmd
Something new

@@
 tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
-map(pt)
+map(tx)
 ```
+
+```{r}
+Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
+eof.tx <- EOF(Tx.maxjja)
+
+Tx.nhd <- annual(tx,FUN='count',threshold=40)
+```

### 2016-10-18 KSS/ds-t2m4KSS.Rmd
update from VGL

@@
 if (!file.exists('t2m.nordic.rda')) {
-  ss <- select.station(param='t2m',lon=c(5,30),lat=c(55,72),src='ecad',nmin=50)
-  X <- station(ss)
-  save(X,file='t2m.nordic.rda')
+  ss <- select.station(param='tg',lon=c(5,30),lat=c(55,72),src='ecad',nmin=50)
+  Y <- station(ss)
+  save(Y,file='t2m.nordic.rda')
 } else load('t2m.nordic.rda')
@@
 ## Only use the most recent 66 years 
-X <- subset(X,it=c(1950,2015))
+Y <- subset(Y,it=c(1960,2015))
 ## Remove records with many missing datapoints
-nv <- apply(coredata(X),2,'nv')
+nv <- apply(coredata(Y),2,'nv')
+Y <- subset(Y,is=nv > 14000)
 ```
@@
 
-param <- 't2m'
 FUN <- 'mean'
 reanalysis <- 'air.mon.mean.nc'
-lon=c(0,100); lat=c(65,90)
-
-if (param=='precip') FUNX <- 'C.C.eq' else FUNX <- 'mean'
-
-## Get the predictand -> Y
-if (!exists('Y')) {
-  load(paste(param,'.nordic.rda',sep=''))
-}
+lon=c(-30,30); lat=c(50,75)
+FUNX <- 'mean'
+verbose=TRUE
 
@@
   if (!file.exists(paste('dse.kss.t2m.rcp45.',it,'.rda',sep=''))) {
-    Z <- downscale(Y,predictor,it,param,FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
+    Z <- downscale(Y,predictor,it,param,FUN=FUN,FUNX=FUNX,lon=lon,lat=lat,verbose=verbose)
     save(file=paste('dse.kss.t2m.rcp45.',it,'.rda',sep=''),Z)
@@
     Z <- downscale(Y,predictor,it,param,rcp='rcp26',
-                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
+                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat,verbose=verbose)
     save(file=paste('dse.kss.t2m.rcp26.',it,'.rda',sep=''),Z)
@@
     Z <- downscale(Y,predictor,it,param,rcp='rcp85',
-                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat)
+                                    FUN=FUN,FUNX=FUNX,lon=lon,lat=lat,verbose=verbose)
     save(file=paste('dse.kss.t2m.rcp85.',it,'.rda',sep=''),Z)

### 2016-10-19 EU-Circle/eu-circle_case-study-3.Rmd

@@
+---
+output: pdf_document
+---
 **EU-Circle case study 3
@@
 ```{r slp-trend-map}
-slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg_v13.1.nc',lon=c(-6,0),lat=c(48,52))
+slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg.nc',lon=c(-6,0),lat=c(48,52))
 slp.eobs <- 10000*slp.eobs ## The data have some funny scaling factor
@@
 ```{r predictand}
-pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v13.0.nc',lon=c(-6,-1),lat=c(50,52))
+pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(-6,-1),lat=c(50,52))
 mu <- annual(pr.eobs,FUN='wetmean')

### 2016-10-19 EU-Circle/eu-circle_case-study-3.Rmd
Update

@@
 ## The wet-day frequency and SLP
-ds.fw <- DS(subset(eof.fw,pattern=1:4),eof.slp,eofs=1:5)
+ds.fw <- DS(subset(eof.fw,ip=1:4),eof.slp,eofs=1:5)
 plot(ds.fw)
-ds.mu <- DS(subset(eof.mu,pattern=1:4),eof.es,eofs=1:7)
+ds.mu <- DS(subset(eof.mu,ip=1:4),eof.es,eofs=1:7)
 plot(ds.mu)

### 2016-10-19 CixPAG/CixPAG-imd.Rmd
update from VGL

--- 2016-09-27 CixPAG/CixPAG-imd.Rmd
@@
 
-## Reading data predictand
+## Reading data predictand: maximum temperature
 
-```{r pressure, echo=FALSE}
-pt <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
+```{r IMD maxmum temperature}
 tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
-map(pt)
+map(tx)
 ```
+
+Some initial exploration of the data - the mean summer maximum daily maximum temperature
+
+```{r}
+Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
+eof.tx <- EOF(Tx.maxjja)
+plot(eof.tx,new=FALSE)
+```
+
+Trend in the men summer maximum daily maximum temperature
+
+```{r}
+map(Tx.maxjja,FUN='trend',new=FALSE)
+```
+
+The mean number of days with daily maximum temperature exceeding 40C.
+
+```{r}
+Tx.nhd40 <- annual(tx,FUN='count',threshold=40)
+map(Tx.nhd40,new=FALSE)
+```
+
+## Precipitation
+
+```{r IMD maxmum temperature}
+rr <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
+```
+
+Monsoon rain statistics for the past - any interesting trends?
+
+```{r}
+mu <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetmean')
+MU <- annual(rr,FUN='wetmean')  ## Use capital letters for theentire year
+fw <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetfreq')
+FW <- annual(rr,FUN='wetfreq')
+pt <- annual(rr,FUN='sum')
+map(pt,new=FALSE)
+```
+
+The precipitation intencity during the monsoon months
+
+```{r}
+map(mu,new=FALSE)
+map(mu,FUN='trend',new=FALSE)
+```
+
+The precipitation frequency during the monsoon months
+
+```{r}
+map(fw,new=FALSE)
+map(fw,FUN='trend',new=FALSE)
+```

### 2016-10-20 KSS/dse4KSS.R
KSS: R-jobs for parralel queing

@@ -1,0 +1,124 @@
+#!/usr/bin/env Rscript
+args=commandArgs(trainingOnly=TRUE)
+
+## R-job for downscaling Nordic temperatures for KSS 
+
+library(esd)
+
+
+## Function definitions:
+
+downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
+                      period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
+                      lon=c(-20,40),lat=c(50,80),ip=1:6,n=6,
+                      pattern='tas_Amon_ens',
+                      rel.cord=FALSE,select=NULL) {
+  
+  print('downscale')
+  
+  ## Use a time and space window:  
+  Y <- subset(Y,it=period)
+  Y <- subset(Y,is=list(lat=lat,lon=lon))
+  
+  ## Estimate seasonal means & weed out stations with little data
+  Y4 <- subset(as.4seasons(Y,FUN=FUN),it=it)
+  ok <- apply(coredata(Y4),1,nv)
+  Y4 <- subset(Y4,it=ok>0)
+  nok <- apply(coredata(Y4),2,nv)
+  Y4 <- subset(Y4,is=nok>15)
+  
+  print(paste(round(100*sum(!is.finite(Y4))/length(!is.finite(Y4))),'% missing',
+              sep=''))
+  if (plot) map(Y,FUN=FUN,cex=-2)
+  
+  nmiss <- round(100*sum(!is.finite(Y4))/length(!is.finite(Y4)))
+  #print(paste(nmiss,'% missing',sep=''))
+  
+  ## Fill missing data using PCA-based regression
+  Z <- pcafill(Y4)
+  
+  pca <- PCA(Z,n=n)
+  if (plot) plot(pca)
+  
+  ## Downscale results
+  print('DSensemble')
+  dse.pca <- DSensemble(pca,predictor=predictor,FUNX=FUNX,verbose=verbose,
+                        biascorrect=TRUE,rcp=rcp,ip=ip,select=select,
+                        pattern=pattern,
+                        lon=lon,lat=lat,rel.cord=rel.cord,it=it,plot=plot)
+  
+  attr(dse.pca,'N.missing') <- nmiss
+  invisible(dse.pca)
+}
+
+##-----------------------------------------------------------------------
[…]

### 2016-10-20 KSS/dse4KSS.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+echo "Submit downscaling jobs for KSS as parallel tasks"
+#for PARAM in t2m mu fw
+for PARAM in t2m #TEST
+do  
+  #for SEASON in djf mam jja son 
+  for SEASON in djf # TEST
+  do
+    #for RCP in rcp45 rcp85 rcp26
+    for RCP in rcp45
+    do
+      echo dse4KSS.$PARAM.$SEASON.$RCP.job
+      sed 's|param|'$PARAM'|;
+           s|season|'$SEASON'|;
+           s|rcp|'$RCP'|' dse4KSS.job > dse4KSS.$PARAM.$SEASON.$RCP.job
+      qsub dse4KSS.$PARAM.$SEASON.$RCP.job
+    done
+  done
+done 
+
+

### 2016-10-20 KSS/dse4KSS.R
dse4KSS updated & fixed

@@ -1,10 +1,11 @@
 #!/usr/bin/env Rscript
-args=commandArgs(trainingOnly=TRUE)
+args=commandArgs(trailingOnly=TRUE)
 
 ## R-job for downscaling Nordic temperatures for KSS 
+print('--- R-job for downscaling Nordic temperatures for KSS ---')
+print(args)
 
 library(esd)
-
 
 ## Function definitions:
 
@@ -54,6 +55,7 @@ downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
 ##-----------------------------------------------------------------------
 ## Define season and parameter: passed from the parameters
 
+print('Settings')
 param <- args[1]  # parameter
 rcp <- args[2]
 it <- args[3]
@@ -89,7 +91,9 @@ verbose=TRUE
 
 ## Retrieving the data
 ## Get the predictand data: daily temperature from ECAD:
-  
+
+print('Retrieve data')
+
 if (!file.exists(predictand)) {
   ss <- select.station(param=varid,lon=c(5,30),lat=c(55,72),src='ecad',nmin=50)
   Y <- station(ss)
@@ -102,7 +106,8 @@ Y <- subset(Y,it=c(1960,2015))
 nv <- apply(coredata(Y),2,'nv')
 Y <- subset(Y,is=nv > 14000)
 
-  
+print('predictand')
+
   ## Get the large-scale predictor:
   if (!exists('predictor')) {
     T2M <- retrieve(reanalysis,lon=lon,lat=lat)

### 2016-10-20 CixPAG/CixPAG-imd.Rmd
CixPAG downscaling

@@
 
-```{r IMD maxmum temperature}
+```{r IMD_tx}
 tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
+attr(tx,'unit') <- 'degC'
 map(tx)
@@
 
-```{r}
+```{r max_summer_temps}
 Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
-eof.tx <- EOF(Tx.maxjja)
-plot(eof.tx,new=FALSE)
+index(Tx.maxjja) <- year(Tx.maxjja)
+eof.Y <- EOF(Tx.maxjja,n=5)
+plot(eof.Y,new=FALSE)
 ```
@@
 
-```{r}
+```{r number_hot_days}
 Tx.nhd40 <- annual(tx,FUN='count',threshold=40)
@@
 
+## Downscaling
+
+```{r}
+predictor <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,25))
+predictor <- annual(subset(predictor,it='jja'),nmin=3)
+eof.X <- EOF(predictor,n=10)
+plot(eof.X,new=FALSE)
+```
+
+Check the link between large and small scales
+
+```{r}
+class(eof.Y) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+class(eof.X) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+index(eof.Y) <- as.Date(paste(year(eof.Y),'-07-01',sep=''))
+index(eof.X) <- as.Date(paste(year(eof.X),'-07-01',sep=''))
+ds.test <- DS(eof.Y,eof.X,ip=1:6)
+plot(ds.test,new=FALSE)
+```
+
+### Apply the downscaling to the CMIP5 ensemble RCP4.5
+
+Carry out the downscaling once and save the results.
+
+```{r}
+if (!file.exists('dse.Tx.jja.cixpag.rda')) {
+    dse.Tx <- DSensemble.eof(eof.Y,eof.X,ip=1:6,it='jja',
+                             predictor='air.mon.mean.nc',verbose=FALSE)
+    save(dse.Tx,file='dse.Tx.jja.cixpag.rda')
+} else load('dse.Tx.jja.cixpag.rda')
+```
+
 ## Precipitation
 
-```{r IMD maxmum temperature}
+```{r IMD_precip}
 rr <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
@@
 
-```{r}
+```{r aggregated_precip}
 mu <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetmean')

### 2016-10-20 CixPAG/CixPAG-imd.R
Fixed dse4KSS jobs

@@ -1,0 +1,50 @@
+library(esd)
+ 
+## Reading data predictand: maximum temperature
+
+tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
+attr(tx,'unit') <- 'degC'
+map(tx)
+ 
+## Some initial exploration of the data - the mean summer maximum daily maximum temperature
+
+Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
+index(Tx.maxjja) <- year(Tx.maxjja)
+eof.Y <- EOF(Tx.maxjja,n=5)
+plot(eof.Y,new=FALSE)
+ 
+## Trend in the men summer maximum daily maximum temperature
+
+map(Tx.maxjja,FUN='trend',new=FALSE)
+ 
+## The mean number of days with daily maximum temperature exceeding 40C. 
+
+Tx.nhd40 <- annual(tx,FUN='count',threshold=40)
+map(Tx.nhd40,new=FALSE)
+ 
+## Downscaling
+
+predictor <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,25))
+predictor <- annual(subset(predictor,it='jja'),nmin=3)
+eof.X <- EOF(predictor,n=10)
+plot(eof.X,new=FALSE)
+ 
+## Check the link between large and small scales
+
+class(eof.Y) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+class(eof.X) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+index(eof.Y) <- as.Date(paste(year(eof.Y),'-07-01',sep=''))
+index(eof.X) <- as.Date(paste(year(eof.X),'-07-01',sep=''))
+ds.test <- DS(eof.Y,eof.X,ip=1:6)
+plot(ds.test,new=FALSE)
+ 
+
+### Apply the downscaling to the CMIP5 ensemble RCP4.5 
+
+## Carry out the downscaling once and save the results. 
+
+if (!file.exists('dse.Tx.jja.cixpag.rda')) {
+  dse.Tx <- DSensemble.eof(eof.Y,eof.X,ip=1:6,it='jja',
+                           predictor='air.mon.mean.nc',verbose=FALSE)
+  save(dse.Tx,file='dse.Tx.jja.cixpag.rda')
+} else load('dse.Tx.jja.cixpag.rda')

### 2016-10-20 KSS/dse4KSS.R

@@ -55,6 +55,8 @@ downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
 ##-----------------------------------------------------------------------
 ## Define season and parameter: passed from the parameters
 
+path <- 
+
 print('Settings')
 param <- args[1]  # parameter
 rcp <- args[2]
@@ -64,7 +66,7 @@ if (param=='t2m') {
   FUN='mean'
   FUNX='mean'
   pattern='tas_Amon_ens'
-  predictand <- '/lustre/t2m.nordic.rda/t2m.nordic.rda'
+  predictand <- '/lustre/storeB/users/rasmusb/data/t2m.nordic.rda'
   varid <- 'tg'
 }
 if (param=='mu') {
@@ -72,7 +74,7 @@ if (param=='mu') {
   FUN='wetmean'
   FUNX='meanx'
   pattern='tas_Amon_ens'
-  predictand <- '/lustre/t2m.nordic.rda/rr.nordic.rda'
+  predictand <- '/lustre/storeB/users/rasmusb/data/rr.nordic.rda'
   varid <- 'precip'
 }
 if (param=='fw') {
@@ -80,14 +82,16 @@ if (param=='fw') {
   FUN='wetfreq'
   FUNX='mean'
   pattern='psl_Amon_ens'
-  predictand <- '/lustre/t2m.nordic.rda/rr.nordic.rda'
+  predictand <- '/lustre/storeB/users/rasmusb/data/rr.nordic.rda'
   varid <- 'precip'
 }
 
 outdir <- '/lustre/storeB/users/rasmusb/dse4KSS/'
+if (!file.exists(outdir)) dir.create(outdir)
+
 lon=c(-30,30); lat=c(50,75)
 
-verbose=TRUE
+verbose=FALSE
 
 ## Retrieving the data
 ## Get the predictand data: daily temperature from ECAD:

### 2016-10-20 KSS/dse4KSS.sh

@@ -1,14 +1,14 @@
 #!/bin/bash
 
 echo "Submit downscaling jobs for KSS as parallel tasks"
-#for PARAM in t2m mu fw
-for PARAM in t2m #TEST
+for PARAM in t2m mu fw
+#for PARAM in t2m #TEST
 do  
-  #for SEASON in djf mam jja son 
-  for SEASON in djf # TEST
+  for SEASON in djf mam jja son 
+  #for SEASON in djf # TEST
   do
-    #for RCP in rcp45 rcp85 rcp26
-    for RCP in rcp45
+    for RCP in rcp45 rcp85 rcp26
+    #for RCP in rcp45
     do
       echo dse4KSS.$PARAM.$SEASON.$RCP.job
       sed 's|param|'$PARAM'|;

### 2016-10-20 CixPAG/CixPAG-imd.R
Computed dsensemble results for CixPag

@@ -48,3 +48,4 @@ if (!file.exists('dse.Tx.jja.cixpag.rda')) {
                            predictor='air.mon.mean.nc',verbose=FALSE)
   save(dse.Tx,file='dse.Tx.jja.cixpag.rda')
 } else load('dse.Tx.jja.cixpag.rda')
+

### 2016-10-20 CixPAG/CixPAG-imd-precip.Rmd
minor fixes

--- 2016-10-20 CixPAG/CixPAG-imd.Rmd
@@
 ---
-title: "CixPAG"
+title: "CixPAG-IMD precip"
 author: "Rasmus Benestad"
-date: "August 3, 2016"
+date: "October 20, 2016"
 output: pdf_document
@@
 This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-```{r esd}
-library(esd)
-```
-
-## Reading data predictand: maximum temperature
-
-```{r IMD_tx}
-tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
-attr(tx,'unit') <- 'degC'
-map(tx)
-```
-
-Some initial exploration of the data - the mean summer maximum daily maximum temperature
-
-```{r max_summer_temps}
-Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
-index(Tx.maxjja) <- year(Tx.maxjja)
-eof.Y <- EOF(Tx.maxjja,n=5)
-plot(eof.Y,new=FALSE)
-```
-
-Trend in the men summer maximum daily maximum temperature
-
-```{r}
-map(Tx.maxjja,FUN='trend',new=FALSE)
-```
-
-The mean number of days with daily maximum temperature exceeding 40C.
-
-```{r number_hot_days}
-Tx.nhd40 <- annual(tx,FUN='count',threshold=40)
-map(Tx.nhd40,new=FALSE)
-```
-
-## Downscaling
-
-```{r}
-predictor <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,25))
-predictor <- annual(subset(predictor,it='jja'),nmin=3)
-eof.X <- EOF(predictor,n=10)
-plot(eof.X,new=FALSE)
-```
-
-Check the link between large and small scales
-
-```{r}
-class(eof.Y) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
-class(eof.X) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
-index(eof.Y) <- as.Date(paste(year(eof.Y),'-07-01',sep=''))
-index(eof.X) <- as.Date(paste(year(eof.X),'-07-01',sep=''))
-ds.test <- DS(eof.Y,eof.X,ip=1:6)
-plot(ds.test,new=FALSE)
-```
-
-### Apply the downscaling to the CMIP5 ensemble RCP4.5
-
-Carry out the downscaling once and save the results.
-
-```{r}
-if (!file.exists('dse.Tx.jja.cixpag.rda')) {
-    dse.Tx <- DSensemble.eof(eof.Y,eof.X,ip=1:6,it='jja',
-                             predictor='air.mon.mean.nc',verbose=FALSE)
-    save(dse.Tx,file='dse.Tx.jja.cixpag.rda')
-} else load('dse.Tx.jja.cixpag.rda')
-```
 
@@
 ```
+
+## Downscaling
+
+```{r max_summer_temps}
+mu.jjas <- annual(subset(rr,it=month.abb[6:9]),nmin=4)
+index(mu.jjas) <- year(mu.jjas)
+eof.Y <- EOF(mu.jjas,n=5)
+plot(eof.Y,new=FALSE)
+```
+
+```{r}
+predictor <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,25))
+predictor <- maskland(predictor,land=TRUE)
+predictor <- C.C.eq(predictor)
+predictor <- annual(subset(predictor,it='jja'),nmin=3)
+eof.X <- EOF(predictor,n=10)
+plot(eof.X,new=FALSE)
+```
+
+Check the link between large and small scales
+
+```{r}
+class(eof.Y) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+class(eof.X) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+index(eof.Y) <- as.Date(paste(year(eof.Y),'-07-01',sep=''))
+index(eof.X) <- as.Date(paste(year(eof.X),'-07-01',sep=''))
+ds.test <- DS(eof.Y,eof.X,ip=1:6)
+plot(ds.test,new=FALSE)
+```

### 2016-10-20 CixPAG/CixPAG-imd.Rmd
minor fixes

@@
 ---
-title: "CixPAG"
+title: "CixPAG-IMD-Tx"
 author: "Rasmus Benestad"
@@
 
-## Precipitation
-
-```{r IMD_precip}
-rr <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
+```{r}
+## Present annual maximum daily maximum temperature from RCP4.5 ensemble mean
+z0 <- map(dse.Tx,it=2010,new=FALSE)
+## annual maximum daily maximum temperature for 2050 based on RCP4.5 ensemble mean
+z1 <- map(dse.Tx,it=2050,new=FALSE)
+## Projected change in annual maximum daily maximum temperature
+dz <- z1-z0
+dz <- attrcp(z1,dz)
+attr(dz,'dimensions') <- attr(z1,'dimensions')
+class(dz) <- class(z1)
+map(dz,new=FALSE)
 ```
-
-Monsoon rain statistics for the past - any interesting trends?
-
-```{r aggregated_precip}
-mu <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetmean')
-MU <- annual(rr,FUN='wetmean')  ## Use capital letters for theentire year
-fw <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetfreq')
-FW <- annual(rr,FUN='wetfreq')
-pt <- annual(rr,FUN='sum')
-map(pt,new=FALSE)
-```
-
-The precipitation intencity during the monsoon months
-
-```{r}
-map(mu,new=FALSE)
-map(mu,FUN='trend',new=FALSE)
-```
-
-The precipitation frequency during the monsoon months
-
-```{r}
-map(fw,new=FALSE)
-map(fw,FUN='trend',new=FALSE)
-```

### 2016-10-20 CixPAG/CixPAG-imd.Rmd
small fix

@@
 
-Trend in the men summer maximum daily maximum temperature
+Trend in the max summer maximum daily maximum temperature
 
-```{r}
+```{r map_trend_tx}
 map(Tx.maxjja,FUN='trend',new=FALSE)
@@
 
-```{r}
+```{r predictor}
 predictor <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,25))
@@
 
-```{r}
+```{r test_downscaling}
 class(eof.Y) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
@@
 
-```{r}
+```{r downscale_ensemble}
 if (!file.exists('dse.Tx.jja.cixpag.rda')) {
@@
 
-```{r}
+```{r show_maps}
 ## Present annual maximum daily maximum temperature from RCP4.5 ensemble mean
-z0 <- map(dse.Tx,it=2010,new=FALSE)
+z0 <- map(dse.Tx,it=2010,new=FALSE,plot=FALSE)
 ## annual maximum daily maximum temperature for 2050 based on RCP4.5 ensemble mean
-z1 <- map(dse.Tx,it=2050,new=FALSE)
+z1 <- map(dse.Tx,it=2050,new=FALSE,plot=FALSE)
 ## Projected change in annual maximum daily maximum temperature

### 2016-10-20 KSS/dse4KSS.R

@@ -55,8 +55,6 @@ downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
 ##-----------------------------------------------------------------------
 ## Define season and parameter: passed from the parameters
 
-path <- 
-
 print('Settings')
 param <- args[1]  # parameter
 rcp <- args[2]

### 2016-11-09 CixPAG/meanspell.R
CixPAG meanspell

@@ -1,0 +1,47 @@
+## Script that reads European temperature data and explores the connection between 
+## the seasonal mean temperature and the mean length of the warm/cold spells 
+
+it <- 'djf'
+cold <- TRUE
+if (cold) is <- 2 else is <- 1
+threshold <- 0
+
+library(esd)
+ss <- select.station(src='ecad',param='tg',nmin=75)
+d <- dim(ss)
+x <- rep(NA,d[1]); y <- x
+for (is in 1:d[1]) {
+  Tx <- station(ss[is,])
+  print(loc(Tx))
+  z <- subset(Tx,is=is)
+  ## Make sure that there are values above and below the given threshold - otherwise
+  ## spell will not work.
+  if ( (sum(z > threshold,na.rm=TRUE)>100) & (sum(z < threshold,na.rm=TRUE)> 100) ) {
+    s <- spell(z,threshold=threshold)
+    y[is] <- mean(subset(subset(s,is=2),it=it),na.rm=TRUE)
+    x[is] <- mean(subset(z,it=it),na.rm=TRUE)
+    plot(x,y,main='Mean temperature & mean length of freezing spells',
+        sub='source: ECA&D',pch=19,col=rgb(0.5,0,0,0.3),
+        xlab=expression(bar(T)),ylab=expression(bar(tau[T < T0])))
+    calfit <- data.frame(x=max(x,na.rm=TRUE) - x[is.finite(x) & is.finite(y)],
+                         y=y[is.finite(x) & is.finite(y)])
+    attr(calfit,'max(x)') <- max(x,na.rm=TRUE)
+    #fit <- lm(y ~ I(x) + I(x^2),data=calfit)
+    #fit <- glm(y ~ x,data=calfit,family='poisson')
+    #if (is > 10) lines(calfit$x[order(cal.fit$x)]-max(x,na.rm=TRUE),
+    #                   predict(fit)[order(calfit$x)],col='red')
+  }
+  grid()
+}
+
+attr(x,'description') <- paste(it,'mean temperature (degC)')
+if (cold) attr(y,'description') <- 'mean cold spell length (days)' else
+          attr(y,'description') <- 'mean warm spell length (days)'
+attr(y,'threshold') <- treshold
+attr(x,'label') <- expression(bar(T))
+attr(y,'label') <- xpression(bar(tau[T < T0]))
+meanspell <- data.frame(meanT=x,meanL=y)
+attr(meanspell,'fit') <- fit
[…]

### 2016-11-09 KSS/dse4KSS.R

@@ -11,7 +11,7 @@ library(esd)
 
 downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
                       period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
-                      lon=c(-20,40),lat=c(50,80),ip=1:6,n=6,
+                      lon=c(-20,40),lat=c(50,80),ip=1:2,n=3,
                       pattern='tas_Amon_ens',
                       rel.cord=FALSE,select=NULL) {
   
@@ -19,7 +19,6 @@ downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
   
   ## Use a time and space window:  
   Y <- subset(Y,it=period)
-  Y <- subset(Y,is=list(lat=lat,lon=lon))
   
   ## Estimate seasonal means & weed out stations with little data
   Y4 <- subset(as.4seasons(Y,FUN=FUN),it=it)
@@ -45,8 +44,7 @@ downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
   print('DSensemble')
   dse.pca <- DSensemble(pca,predictor=predictor,FUNX=FUNX,verbose=verbose,
                         biascorrect=TRUE,rcp=rcp,ip=ip,select=select,
-                        pattern=pattern,
-                        lon=lon,lat=lat,rel.cord=rel.cord,it=it,plot=plot)
+                        nmin=1,pattern=pattern,lon=lon,lat=lat,rel.cord=rel.cord,it=it,plot=plot)
   
   attr(dse.pca,'N.missing') <- nmiss
   invisible(dse.pca)
@@ -55,43 +53,48 @@ downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
 ##-----------------------------------------------------------------------
 ## Define season and parameter: passed from the parameters
 
-path <- 
-
 print('Settings')
 param <- args[1]  # parameter
 rcp <- args[2]
 it <- args[3]
+landmask <- FALSE
+
 if (param=='t2m') {
   reanalysis <- 'air.mon.mean.nc'
[…]

### 2016-11-09 KSS/dse4KSS.sh

@@ -2,9 +2,10 @@
 
 echo "Submit downscaling jobs for KSS as parallel tasks"
 for PARAM in t2m mu fw
+#for PARAM in mu fw
 #for PARAM in t2m #TEST
 do  
-  for SEASON in djf mam jja son 
+  for SEASON in djf mam jja son ondjfm amjjas
   #for SEASON in djf # TEST
   do
     for RCP in rcp45 rcp85 rcp26

### 2016-11-09 KSS/testesd.R

@@ -1,0 +1,97 @@
+param <- 't2m'
+it <- 'djf'
+period <- c(1950,2015)
+n <- 3
+ip <- 1:2
+landmask <- FALSE
+
+if (param=='t2m') {
+  #reanalysis <- 'air.mon.mean.nc'
+  reanalysis <- 'ERA40_t2m_mon.nc'
+  FUN='mean'
+  FUNX='mean'
+  pattern='tas_Amon_ens'
+  predictand <- '/lustre/storeB/users/rasmusb/data/t2m.nordic.rda'
+  varid <- 'tg'
+  #lon=c(-10,40); lat=c(50,75)
+  lon=c(-5,40); lat=c(55,72)
+}
+if (param=='mu') {
+  reanalysis <- 'air.mon.mean.nc'
+  FUN='wetmean'
+  FUNX='C.C.eq'
+  pattern='tas_Amon_ens'
+  predictand <- '/lustre/storeB/users/rasmusb/data/rr.nordic.rda'
+  varid <- 'precip'
+  lon=c(-90,10); lat=c(25,75)
+  landmask <- TRUE
+}
+if (param=='fw') {
+  reanalysis <- 'slp.mon.mean.nc'
+  FUN='wetfreq'
+  FUNX='mean'
+  pattern='psl_Amon_ens'
+  predictand <- '/lustre/storeB/users/rasmusb/data/rr.nordic.rda'
+  varid <- 'precip'
+  lon=c(-30,30); lat=c(50,75)
+}
+
+outdir <- '/lustre/storeB/users/rasmusb/dse4KSS/'
+if (!file.exists(outdir)) dir.create(outdir)
+
+verbose=FALSE
+
+## Retrieving the data
+## Get the predictand data: daily temperature from ECAD:
+
+print('Retrieve data')
+
+if (!file.exists(predictand)) {
+  ss <- select.station(param=varid,lon=c(5,30),lat=c(55,72),src='ecad',nmin=50)
+  Y <- station(ss)
+  save(Y,file=predictand)
+} else load(predictand)
+
+## Only use the most recent 66 years 
+Y <- subset(Y,it=c(1960,2015))
+## Remove records with many missing datapoints
+nv <- apply(coredata(Y),2,'nv')
+Y <- subset(Y,is=nv > 14000)
+
+print('predictor')
+
+## Get the large-scale predictor:
+T2M <- retrieve(reanalysis,lon=lon,lat=lat)
[…]

### 2016-11-09 CixPAG/test.Rmd
EU-Circle case 3

---
title: "test"
author: "Rasmus Benestad"
date: "October 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r esd}
library(esd)
```

## Reading data predictand: maximum temperature

```{r IMD_tx, warning=FALSE}
tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
attr(tx,'unit') <- 'degC'
map(tx)
```

Some initial exploration of the data - the mean summer maximum daily maximum temperature

```{r max_summer_temps, warning=FALSE}
Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
index(Tx.maxjja) <- year(Tx.maxjja)
eof.Y <- EOF(Tx.maxjja,n=5)
plot(eof.Y,new=FALSE)
```

Trend in the max summer maximum daily maximum temperature

```{r map_trend_tx, warning=FALSE}
map(Tx.maxjja,FUN='trend',new=FALSE)
```

The mean number of days with daily maximum temperature exceeding 40C.

```{r number_hot_days, warning=FALSE}
Tx.nhd40 <- annual(tx,FUN='count',threshold=40)
map(Tx.nhd40,new=FALSE)
```

## Downscaling

```{r predictor, warning=FALSE}
predictor <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,25))
predictor <- annual(subset(predictor,it='jja'),nmin=3)
eof.X <- EOF(predictor,n=10)
plot(eof.X,new=FALSE)
```

Check the link between large and small scales

```{r test_downscaling, warning=FALSE}
class(eof.Y) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
class(eof.X) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
index(eof.Y) <- as.Date(paste(year(eof.Y),'-07-01',sep=''))
index(eof.X) <- as.Date(paste(year(eof.X),'-07-01',sep=''))
ds.test <- DS(eof.Y,eof.X,ip=1:6)
plot(ds.test,new=FALSE)
```

### Apply the downscaling to the CMIP5 ensemble RCP4.5

Carry out the downscaling once and save the results.

```{r downscale_ensemble, warning=FALSE}
if (!file.exists('dse.Tx.jja.cixpag.rda')) {
    dse.Tx <- DSensemble.eof(eof.Y,eof.X,ip=1:6,it='jja',
                             predictor='air.mon.mean.nc',verbose=FALSE)
    save(dse.Tx,file='dse.Tx.jja.cixpag.rda')
} else load('dse.Tx.jja.cixpag.rda')
```

### 2016-11-09 EU-Circle/eu-circle_case-study-3.Rmd
EU-Circle case 3

@@
 
-```{r function-probability}
+```{r function-probability, warning=FALSE, message=FALSE}
 PrexpPr <- function(mu,fw,x0=10) {
@@
 
-```{r read-precip-data}
+```{r read-precip-data, warning=FALSE, message=FALSE}
 ## Get MIDAS preciptation data that has laready been retrieved 
@@
 
-```{r ecad-data}
+```{r ecad-data, warning=FALSE, message=FALSE}
  ss <- select.station(param='precip',src='ecad',cntr=c('United Kingdom','France'),lat=c(45,55),lon=c(-10,0),nmin=50)
@@
 
-```{r combine-stations}
+```{r combine-stations, warning=FALSE, message=FALSE}
 mu <- combine.stations(mu,mu.ecad)
@@
 
-```{r fill in missing data}
+```{r fill in missing data, warning=FALSE, message=FALSE}
 ##Fill in the remaining gaps using a PCA-based interpolation scheme 
@@
 
-```{r}
+```{r long-term-mu, warning=FALSE, message=FALSE}
 plot(mu,new=FALSE); grid()
@@
 
-```{r probabilities}
+```{r probabilities, warning=FALSE, message=FALSE}
 Pr.gt.30 <- PrexpPr(mu,fw,x0=30)
@@
 
-```{r aggregate}
+```{r aggregate, warning=FALSE, message=FALSE}
 ## Estimate the mean seasonal cycle in the wet-day mean precipitaiton
@@
 
-```{r precipitation-intensity}
+```{r precipitation-intensity, warning=FALSE, message=FALSE}
 ## Show the mean seasonal cycle in the wet-day mean precipitation
@@
 
-```{r wet-day frequency}
+```{r wet-day frequency, warning=FALSE, message=FALSE}
 ## Show the mean seasonal cycle in the wet-day frequency
@@
 
-```{r sesonal-precipitation}
+```{r sesonal-precipitation, warning=FALSE, message=FALSE}
 ## Show the mean seasonal cycle in the total precipitation
@@
 
-```{r winter-rain}
+```{r winter-rain, warning=FALSE, message=FALSE}
 ## Estimate the PCA for the annual cycle to identify the general features:
@@
 
-```{r downscale-ensemble}
+```{r downscale-ensemble, warning=FALSE, message=FALSE}
 if (!file.exists('dse.mu.eu-circle-worstcase.rda')) {
@@
 
-```{r}
+```{r sea-level, results="hide", warning=FALSE, message=FALSE}
 gloss <- aggregate(station.gloss(),year,FUN='max',na.rm=TRUE)
 gloss <- subset(gloss,is=list(lon=c(-20,20),lat=c(45,55)))
+```
+
+Plot the GLOSS tidal data:
+
+```{r}
 plot(gloss)
@@
 
-```{r}
+```{r newlyn, results="hide", warning=FALSE, message=FALSE}
 newlyn <- station.newlyn()
 newlyn <- aggregate(newlyn,year,FUN='max',na.rm=TRUE)
+```
+
+Plot the Newlyn tidal data:
+
+```{r}
 plot(newlyn)
@@
 
-```{r}
+```{r sonel, warning=FALSE, message=FALSE}
 sonel <- aggregate(station.sonel(),year,FUN='max',na.rm=TRUE)
@@
 
-```{r}
+```{r combine-sea-level, warning=FALSE, message=FALSE}
 index(newlyn) <- year(newlyn)
@@
 
-```{r sl-trend-map}
+```{r sl-trend-map, warning=FALSE, message=FALSE}
 zc <- coredata(Z)
@@
 
-```{r slp-trend-map}
+```{r slp-trend-map, warning=FALSE, message=FALSE}
 slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg.nc',lon=c(-6,0),lat=c(48,52))
@@
 
-```{r trend-map}
+```{r trend-map, warning=FALSE, message=FALSE}
 ## Fudge it - pick some windstations, but do not use their data. Reset the
@@
 
-```{r trend-event-2006}
+```{r trend-event-2006, warning=FALSE, message=FALSE}
 plot(annual(uv,nmin=300),main='zonal and meridional wind components',
@@
 
-```{r wind-speed-trend}
+```{r wind-speed-trend, warning=FALSE, message=FALSE}
 ## Trend in upper wind speed
@@
 
-```{r wind-direction-trend}
+```{r wind-direction-trend, warning=FALSE, message=FALSE}
 ## Trend in upper wind direction
@@
 
-```{r zonal-wind-check-get-slp}
+```{r zonal-wind-check-get-slp, warning=FALSE, message=FALSE}
 ## Trend in upper wind direction
@@
 
-```{r zonal-wind-check1}
+```{r zonal-wind-check1, warning=FALSE, message=FALSE}
 ## Trend in upper wind direction
@@
 
-```{r zonal-wind-check2}
+```{r zonal-wind-check2, warning=FALSE, message=FALSE}
 ## Trend in upper wind direction
@@
 
-```{r wind-speed-trend-2}
+```{r wind-speed-trend-2, warning=FALSE, message=FALSE}
 ## Trend in upper wind speed
@@
 
-```{r wind-speed-distribution}
+```{r wind-speed-distribution, warning=FALSE, message=FALSE}
 ## 
@@
 
-```{r reanalysis-wind-speed}
+```{r reanalysis-wind-speed, warning=FALSE, message=FALSE}
 ## 
@@
 
-```{r predictand}
+```{r predictand, warning=FALSE, message=FALSE}
 pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(-6,-1),lat=c(50,52))
@@
 
-```{r trend}
+```{r trend, warning=FALSE, message=FALSE}
 ## The area mean wet-day mean precipitation
@@
 
-```{r EOFs-in-precip-mu}
+```{r EOFs-in-precip-mu, warning=FALSE, message=FALSE}
 ## The area mean wet-day mean precipitation
@@
 
-```{r EOFs-in-precip-fw}
+```{r EOFs-in-precip-fw, warning=FALSE, message=FALSE}
 ## Show maps of trends for the wet-day mean precipitation
@@
 
-```{r predictor}
+```{r predictor, warning=FALSE, message=FALSE}
 ## The mean sea-level pressure
@@
 
-```{r cca}
+```{r cca, warning=FALSE, message=FALSE}
 ## The wet-day frequency and SLP
@@
 
-```{r DS}
+```{r DS, warning=FALSE, message=FALSE}
 ## The wet-day frequency and SLP
@@
 
-```{r residual for mu}
+```{r residual-for-mu, warning=FALSE, message=FALSE}
 ## The residual of the wet-day precipitation
@@
 For comparison the original data look like this:
-```{r EOF for mu}
+```{r EOF-for-mu, warning=FALSE, message=FALSE}
 ## The residual of the wet-day precipitation
@@
 
-```{r residual for fw}
+```{r residual-for-fw, warning=FALSE, message=FALSE}
 ## Repeat with the wet-day frequency for the benefit of comparison
@@
 For comparison the original data look like this:
-```{r EOF for fw}
+```{r EOF-for-fw, warning=FALSE, message=FALSE}
 ## The residual of the wet-day precipitation

### 2016-11-09 dipper/dipper.R

@@ -5,7 +5,7 @@ library(ncdf4)
 
 examine <- FALSE
 
-dipper <- read.table('~/Desktop/dipper.csv',header=TRUE)
+dipper <- read.table('~/Dropbox/data/dipper.csv',header=TRUE)
 
 dipper <- zoo(x=dipper[[2]],order.by=dipper[[1]])
 y <- as.station(dipper,loc='Lyngdalselva',

### 2016-11-10 CixPAG/meanspell.R
CixPAG spells and means

@@ -3,27 +3,28 @@
 
 it <- 'djf'
 cold <- TRUE
-if (cold) is <- 2 else is <- 1
 threshold <- 0
+if (cold) is <- 2 else is <- 1
 
 library(esd)
 ss <- select.station(src='ecad',param='tg',nmin=75)
 d <- dim(ss)
 x <- rep(NA,d[1]); y <- x
-for (is in 1:d[1]) {
-  Tx <- station(ss[is,])
+for (i in 1:d[1]) {
+  Tx <- station(ss[i,])
   print(loc(Tx))
-  z <- subset(Tx,is=is)
+  z <- subset(Tx,is=i)
   ## Make sure that there are values above and below the given threshold - otherwise
   ## spell will not work.
-  if ( (sum(z > threshold,na.rm=TRUE)>100) & (sum(z < threshold,na.rm=TRUE)> 100) ) {
+  if ( (sum(z > threshold,na.rm=TRUE)>1000) & (sum(z < threshold,na.rm=TRUE)> 1000) ) {
     s <- spell(z,threshold=threshold)
-    y[is] <- mean(subset(subset(s,is=2),it=it),na.rm=TRUE)
-    x[is] <- mean(subset(z,it=it),na.rm=TRUE)
-    plot(x,y,main='Mean temperature & mean length of freezing spells',
+    y[i] <- mean(subset(subset(s,is=is),it=it),na.rm=TRUE)
+    x[i] <- mean(subset(z,it=it),na.rm=TRUE)
+    plot(x,y,main=paste('Mean',it,'temperature & mean length of intervals',
+                        c('below','above')[c(cold,!cold)],threshold,'C'),
         sub='source: ECA&D',pch=19,col=rgb(0.5,0,0,0.3),
-        xlab=expression(bar(T)),ylab=expression(bar(tau[T < T0])))
-    calfit <- data.frame(x=max(x,na.rm=TRUE) - x[is.finite(x) & is.finite(y)],
+        xlab=expression(bar(T)),ylab=expression(bar(tau)))
+    calfit <- data.frame(x=x[is.finite(x) & is.finite(y)],
                          y=y[is.finite(x) & is.finite(y)])
     attr(calfit,'max(x)') <- max(x,na.rm=TRUE)
     #fit <- lm(y ~ I(x) + I(x^2),data=calfit)
@@ -37,9 +38,9 @@ for (is in 1:d[1]) {
 attr(x,'description') <- paste(it,'mean temperature (degC)')
[…]

### 2016-11-16 CixPAG/meanspell.R
CixPAG

@@ -1,48 +1,100 @@
 ## Script that reads European temperature data and explores the connection between 
 ## the seasonal mean temperature and the mean length of the warm/cold spells 
 
-it <- 'djf'
-cold <- TRUE
-threshold <- 0
-if (cold) is <- 2 else is <- 1
+it <- 'jja'
+if (it =='djf') {
+  cold <- TRUE
+  threshold <- 0
+  is <- 2
+} else {
+  cold <- FALSE
+  threshold <- 20
+  is <- 1
+}
+
+AR <- function(n,mean=1,sd=1,a1=0.8) {
+  rn <- rnorm(n,mean=mean,sd=sd)
+  for (i in 2:n) rn[i] <- (a1*rn[i-1] + (1-a1)*rn[i])
+  invisible(rn)
+}
 
 library(esd)
 ss <- select.station(src='ecad',param='tg',nmin=75)
 d <- dim(ss)
-x <- rep(NA,d[1]); y <- x
+x <- rep(NA,d[1]); y <- x; std <- y
+q.spell <- rep(NA,d[1]*10); dim(q.spell) <- c(d[1],10); q.geom <- q.spell
+
 for (i in 1:d[1]) {
-  Tx <- station(ss[i,])
-  print(loc(Tx))
-  z <- subset(Tx,is=i)
+  z <- station(ss[i,])
+  print(loc(z))
   ## Make sure that there are values above and below the given threshold - otherwise
   ## spell will not work.
   if ( (sum(z > threshold,na.rm=TRUE)>1000) & (sum(z < threshold,na.rm=TRUE)> 1000) ) {
     s <- spell(z,threshold=threshold)
     y[i] <- mean(subset(subset(s,is=is),it=it),na.rm=TRUE)
+    
+    ## Compare the spell-distribution with a geometric distribution
+    q.spell[i,] <- quantile(subset(subset(s,is=is),it=it),probs=seq(0.05,0.95,by=0.1),na.rm=TRUE)
+    q.geom[i,] <- qgeom(p=seq(0.05,0.95,by=0.1),prob=1/y[i])
+    std[i] <- sd(subset(z,it=it),na.rm=TRUE)
     x[i] <- mean(subset(z,it=it),na.rm=TRUE)
     plot(x,y,main=paste('Mean',it,'temperature & mean length of intervals',
                         c('below','above')[c(cold,!cold)],threshold,'C'),
[…]

### 2016-11-17 EU-Circle/eu-circle_case-study-3.Rmd
EU-Circle case study 3

@@
 plot(MU)
+grid()
 ```
@@
 plot(FW)
+grid()
 ```
@@
 plot(MP)
+grid()
 ```
@@
 
-```{r winter-rain, warning=FALSE, message=FALSE}
+```{r winter-rain, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
 ## Estimate the PCA for the annual cycle to identify the general features:
@@
 
+## EOBS gridded data
+
+This script retreives precipitation for Southwestern England from the EOBS data set and estimates the wet-day mean precipitation and frequency.
+
+The EOBS data was chosen here because the rain gauge data from stations (MIDAS) was intermittent and not readily available for an analysis that makes use of PCA in order to make the most use of the information embedded in many parallel data records.
+
+Rain gauges provide a very tiny sample of the rainfall, and previous work suggests that downscaling can provide a better picture if it's applied to a group of stations rather than single stations (Benestad et al., 2015; doi:10.342/tellusa.v67.28326).
+Downscaling applied to products from principal component analysis (PCA) benefits from an emphasis on cohernt structure of variability in time and space, which can improve the signal-to-noise ratio.
+PCA requires no missing values, but the EOBS gridded data provides complete records.
+
+```{r predictand, warning=FALSE, message=FALSE}
+pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(-6,-1),lat=c(50,52))
+mu <- annual(pr.eobs,FUN='wetmean')
+fw <- annual(pr.eobs,FUN='wetfreq')
+```
+
+## Past trends
+
+Climate change has taken place for some time, and past trends may provide some indication about whether a chosen climate parameter such as the wet-day mean precipitation $\mu$ or wet-day frequency $f_w$ are sensitive to a global warming:
+
+```{r trend, warning=FALSE, message=FALSE}
+## The area mean wet-day mean precipitation
+plot(mu,new=FALSE)
+## The area mean wet-day frequency
+plot(fw,new=FALSE)
+## The early part of the record looks suspect:
+mu <- subset(mu,it=c(1958,2015))
+fw <- subset(fw,it=c(1958,2015))
+## Show maps of trends for the wet-day mean precipitation
+map(mu,FUN='trend',
+    colbar=list(breaks=seq(-0.12,0.12,by=0.02),pal='t2m',rev=TRUE))
+## Map of trends in wet-day frequency:
+map(fw,FUN='trend',
+    colbar=list(breaks=seq(-0.01,0.01,by=0.001),pal='t2m',rev=TRUE))
+```
+
+The trend analysis for the wet-day frequency suggest a general increase in the number of wet-days over the  perdiod 1960--2015 while the intencity .
+
+The predictors can be organised as EOFs to simplify the analysis and make use of properties as the products being orthogonal. Here are the EOFS for the annual wet-day mean precipitation $\mu$
+
+```{r EOFs-in-precip-mu, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+## The area mean wet-day mean precipitation
+eof.mu <- EOF(mu)
+## The area mean wet-day frequency
+plot(eof.mu)
+
+```
+
+and the annual wet-day frequency $f_w$
+
+```{r EOFs-in-precip-fw, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+## Show maps of trends for the wet-day mean precipitation
+eof.fw <- EOF(fw)
+## Map of trends in wet-day frequency:
+plot(eof.fw)
+```
+
+The diagnostics of the EOFs suggest that the leading mode explains a high proportion of the variability on an annual time scale.
+
+The mean sea-level pressure (SLP) is used as predictor for the annual wet-day frequency, as this gives a good description of the circulation.
+A number of different spatial domains were tested, and for SLP and $f_w$, the results were fairly robust with respect to the choice of domain.
+
+The predictor for the wet-day mean precipitation was taken to be the saturation vapour pressure over the North Atlantic, which is regarded as the main source for the atmospheric moisture over northern Europe.
+
+## Predictor
+
+```{r predictor, warning=FALSE, message=FALSE}
+## The mean sea-level pressure
+slp <- retrieve('slp.mon.mean.nc',lon=c(-80,0),lat=c(20,70))
+## The surface temperature
+t2m <- retrieve('air.mon.mean.nc',lon=c(-80,0),lat=c(10,70))
+## Saturation water pressure estimated from maritime temperatures
+es <- C.C.eq(mask(t2m,land=TRUE))
+eof.slp<- EOF(annual(slp))
+eof.t2m <- EOF(annual(t2m))
+eof.es <- EOF(annual(es,FUN='mean'))
+```
+
+## Cannical Correlation analysis
+
+Carry out canonical correlation analysis (CCA) to explore relationships between the large-scale conditions such as SLP/temperature and the precipitation over soutwestern England:
+
+```{r cca, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+## The wet-day frequency and SLP
+cca.fw <- CCA(eof.fw,eof.slp)
+plot(cca.fw,new=FALSE)
+cca.mu <- CCA(eof.mu,eof.es)
+plot(cca.mu,new=FALSE)
+```
+
+The CCA identifies the two patterns in the data that have the highest correlation - this pattern may not necessarily be the ones with the highest variance.
+
+## Downscaling analysis
+
+The downscaling emplies a multiple regression applied to each of the principal components for the EOFs of $\mu$ and $f_w$.
+
+```{r DS, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+## The wet-day frequency and SLP
+ds.fw <- DS(subset(eof.fw,ip=1:4),eof.slp,eofs=1:5)
+plot(ds.fw,new=FALSE)
+ds.mu <- DS(subset(eof.mu,ip=1:4),eof.es,eofs=1:7)
+plot(ds.mu,new=FALSE)
+```
+
+The downscaling includes cross-validation analysis, and the results shown are for the leading EOF of the predictands.
+We do not expect that the downscaling will capture a large fraction of the wet-day mean precipitation; local processes are also likely influencing.
+
+Extract the residual and examine to see if there is any temporal-stpatial structure left: a tell-tale sign for noise is a flat eigenspectrum.
+
+```{r residual-for-mu, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+## The residual of the wet-day precipitation
+z.mu <- as.field(ds.mu)
+#res.mu <- as.residual(z.mu)
+res.mu <- mu - z.mu
+res.mu <- attrcp(mu,res.mu)
+class(res.mu) <- class(mu)
+eof.res.mu <- EOF(res.mu)
+plot(eof.res.mu)
+```
+
+The residuals from the downscaling of $\mu$ still give eigenvalues which indicate that there are differnt coherent modes present associated with different variance.
+The spatial pattern is similar to the spatial patterns in the original data.
+This exercise suggests that the anaysis failed to capture any strong link between the predictor (vapour pressure estimated over the North Atlantic) and the predictand.
+One reason may be a shifting source region for moisture at different times of the year and from year to year.
+
+For comparison the original data look like this:
+```{r EOF-for-mu, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+## The original of the wet-day precipitation
+plot(EOF(mu))
+```
+
+```{r residual-for-fw, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+## Repeat with the wet-day frequency for the benefit of comparison
+z.fw <- as.field(ds.fw)
+#res.fw <- as.residual(ds.fw)
+## Another way to estimate the residuals:
+res.fw <- fw - z.fw
+res.fw <- attrcp(fw,res.fw)
+class(res.fw) <- class(fw)
+eof.res.fw <- EOF(res.fw)
+plot(eof.res.fw)
+```
+
+The residual for the wet-day frequency ...
+
+For comparison the original data look like this:
+```{r EOF-for-fw, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+## The original of the wet-day precipitation
+plot(EOF(fw))
+```
+
+Hence, there is remaining spatio-temporal structure left in the residuals from the downscaling.
+The challenge is to identify further factors on which both the wet-day mean precipitation and the wet-day frequency depend.
+It's surprising to see such strong modes in the residual of the wet-day frequency still.
+
 ## Sea-level
@@
 plot(gloss)
+grid()
 ```
@@
 plot(newlyn)
+grid()
 ```
@@
 plot(sonel)
+grid()
 ```
@@
      sub='Estimated from triangulation of mean sea-level pressure from EOBS',zoom=7)
+grid()
 ```
@@
 plot(annual(uv,nmin=300),main='zonal and meridional wind components',
-     xlim=c(2000,2010))
+     xlim=c(2000,2010),map.show=FALSE)
+grid()
 ```
@@
 
-```{r zonal-wind-check1, warning=FALSE, message=FALSE}
+```{r zonal-wind-check1, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
 ## Trend in upper wind direction
@@
 
-```{r zonal-wind-check2, warning=FALSE, message=FALSE}
+```{r zonal-wind-check2, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
 ## Trend in upper wind direction
@@
 ```{r reanalysis-wind-speed, warning=FALSE, message=FALSE}
-## 
-UV <- geostrophicwind(SLP)
+## Estimate grostrophich wind
+UV <- geostrophicwind(SLP,progressbar=FALSE)
 ```
@@
 ```
-
-## EOBS gridded data
-
-This script retreives precipitation for Southwestern England from the EOBS data set and estimates the wet-day mean precipitation and frequency.
-
-The EOBS data was chosen here because the rain gauge data from stations (MIDAS) was intermittent and not readily available for an analysis that makes use of PCA in order to make the most use of the information embedded in many parallel data records.
-
-Rain gauges provide a very tiny sample of the rainfall, and previous work suggests that downscaling can provide a better picture if it's applied to a group of stations rather than single stations (Benestad et al., 2015; doi:10.342/tellusa.v67.28326).
-Downscaling applied to products from principal component analysis (PCA) benefits from an emphasis on cohernt structure of variability in time and space, which can improve the signal-to-noise ratio.
-PCA requires no missing values, but the EOBS gridded data provides complete records.
-
-```{r predictand, warning=FALSE, message=FALSE}
-pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(-6,-1),lat=c(50,52))
-mu <- annual(pr.eobs,FUN='wetmean')
-fw <- annual(pr.eobs,FUN='wetfreq')
-```
-
-## Past trends
-
-Climate change has taken place for some time, and past trends may provide some indication about whether a chosen climate parameter such as the wet-day mean precipitation $\mu$ or wet-day frequency $f_w$ are sensitive to a global warming:
-
-```{r trend, warning=FALSE, message=FALSE}
-## The area mean wet-day mean precipitation
-plot(mu,new=FALSE)
-## The area mean wet-day frequency
-plot(fw,new=FALSE)
-## The early part of the record looks suspect:
-mu <- subset(mu,it=c(1958,2015))
-fw <- subset(fw,it=c(1958,2015))
-## Show maps of trends for the wet-day mean precipitation
-map(mu,FUN='trend',
-    colbar=list(breaks=seq(-0.12,0.12,by=0.02),pal='t2m',rev=TRUE))
-## Map of trends in wet-day frequency:
-map(fw,FUN='trend',
-    colbar=list(breaks=seq(-0.01,0.01,by=0.001),pal='t2m',rev=TRUE))
-```
-
-The trend analysis for the wet-day frequency suggest a general increase in the number of wet-days over the  perdiod 1960--2015 while the intencity .
-
-The predictors can be organised as EOFs to simplify the analysis and make use of properties as the products being orthogonal. Here are the EOFS for the annual wet-day mean precipitation $\mu$
-
-```{r EOFs-in-precip-mu, warning=FALSE, message=FALSE}
-## The area mean wet-day mean precipitation
-eof.mu <- EOF(mu)
-## The area mean wet-day frequency
-plot(eof.mu)
-
-```
-
-and the annual wet-day frequency $f_w$
-
-```{r EOFs-in-precip-fw, warning=FALSE, message=FALSE}
-## Show maps of trends for the wet-day mean precipitation
-eof.fw <- EOF(fw)
-## Map of trends in wet-day frequency:
-plot(eof.fw)
-```
-
-The diagnostics of the EOFs suggest that the leading mode explains a high proportion of the variability on an annual time scale.
-
-The mean sea-level pressure (SLP) is used as predictor for the annual wet-day frequency, as this gives a good description of the circulation.
-A number of different spatial domains were tested, and for SLP and $f_w$, the results were fairly robust with respect to the choice of domain.
-
-The predictor for the wet-day mean precipitation was taken to be the saturation vapour pressure over the North Atlantic, which is regarded as the main source for the atmospheric moisture over northern Europe.
-
-## Predictor
-
-```{r predictor, warning=FALSE, message=FALSE}
-## The mean sea-level pressure
-slp <- retrieve('slp.mon.mean.nc',lon=c(-80,0),lat=c(20,70))
-## The surface temperature
-t2m <- retrieve('air.mon.mean.nc',lon=c(-80,0),lat=c(10,70))
-## Saturation water pressure estimated from maritime temperatures
-es <- C.C.eq(mask(t2m,land=TRUE))
-eof.slp<- EOF(annual(slp))
-eof.t2m <- EOF(annual(t2m))
-eof.es <- EOF(annual(es,FUN='mean'))
-```
-
-## Cannical Correlation analysis
-
-Carry out canonical correlation analysis (CCA) to explore relationships between the large-scale conditions such as SLP/temperature and the precipitation over soutwestern England:
-
-```{r cca, warning=FALSE, message=FALSE}
-## The wet-day frequency and SLP
-cca.fw <- CCA(eof.fw,eof.slp)
-plot(cca.fw)
-cca.mu <- CCA(eof.mu,eof.es)
-plot(cca.mu)
-```
-
-The CCA identifies the two patterns in the data that have the highest correlation - this pattern may not necessarily be the ones with the highest variance.
-
-## Downscaling analysis
-
-The downscaling emplies a multiple regression applied to each of the principal components for the EOFs of $\mu$ and $f_w$.
-
-```{r DS, warning=FALSE, message=FALSE}
-## The wet-day frequency and SLP
-ds.fw <- DS(subset(eof.fw,ip=1:4),eof.slp,eofs=1:5)
-plot(ds.fw)
-ds.mu <- DS(subset(eof.mu,ip=1:4),eof.es,eofs=1:7)
-plot(ds.mu)
-```
-
-The downscaling includes cross-validation analysis, and the results shown are for the leading EOF of the predictands.
-We do not expect that the downscaling will capture a large fraction of the wet-day mean precipitation; local processes are also likely influencing.
-
-Extract the residual and examine to see if there is any temporal-stpatial structure left: a tell-tale sign for noise is a flat eigenspectrum
-
-```{r residual-for-mu, warning=FALSE, message=FALSE}
-## The residual of the wet-day precipitation
-z.mu <- as.field(ds.mu)
-#res.mu <- as.residual(z.mu)
-res.mu <- mu - z.mu
-res.mu <- attrcp(mu,res.mu)
-class(res.mu) <- class(mu)
-eof.res.mu <- EOF(res.mu)
-plot(eof.res.mu)
-```
-
-For comparison the original data look like this:
-```{r EOF-for-mu, warning=FALSE, message=FALSE}
-## The residual of the wet-day precipitation
-plot(EOF(mu))
-```
-
-```{r residual-for-fw, warning=FALSE, message=FALSE}
-## Repeat with the wet-day frequency for the benefit of comparison
-z.fw <- as.field(ds.fw)
-#res.fw <- as.residual(ds.fw)
-## Another way to estimate the residuals:
-res.fw <- fw - z.fw
-res.fw <- attrcp(fw,res.fw)
-class(res.fw) <- class(fw)
-eof.res.fw <- EOF(res.fw)
-plot(eof.res.fw)
-```
-
-For comparison the original data look like this:
-```{r EOF-for-fw, warning=FALSE, message=FALSE}
-## The residual of the wet-day precipitation
-plot(EOF(fw))
-```
-
-Hence, there is remaining spatio-temporal structure left in the residuals from the downscaling.
-The challenge is to identify further factors on which both the wet-day mean precipitation and the wet-day frequency depend.
-It's surprising to see such strong modes in the residual of the wet-day frequency still.

### 2016-11-28 EU-Circle/eu-circle_case-study-3.Rmd
updated case study 3

--- 2016-10-19 EU-Circle/eu-circle_case-study-3.Rmd
@@
----
-output: pdf_document
----
 **EU-Circle case study 3
@@
 ## Information about the system and session
+sessionInfo()
 ```
@@
 ---------------------------------
+```{r,eval=FALSE}
+## Get MIDAS preciptation data that has laready been retrieved 
+## from the MIDAS data base (UK Met Office)
+load('pr.eu-circle-torbay.rda')
+```
 
@@
 
-```{r function-probability}
+```{r}
 PrexpPr <- function(mu,fw,x0=10) {
@@
   attr(Pr,'unit') <- 'probability'
-  attr(Pr,'longname') <- paste('Probability of exceeding',x0,'mm')
+  attr(Pr,'longname') <- paste('Probability of exceeding',x0)
   class(Pr) <- class(mu)
@@
 
-```{r read-precip-data}
-## Get MIDAS preciptation data that has laready been retrieved 
-## from the MIDAS data base (UK Met Office)
+```{r}
 load('pr.eu-circle-torbay.rda') 
@@
 x <- subset(pr,is=(nok> 15000))
-
-## Quality check: remove outliers days with more than 300 mm - set to NA
-xc <- coredata(x); xc[xc > 300] <- NA; xc -> coredata(x); rm('xc')
 
@@
 
-```{r ecad-data}
+```{r}
  ss <- select.station(param='precip',src='ecad',cntr=c('United Kingdom','France'),lat=c(45,55),lon=c(-10,0),nmin=50)
@@
 
-```{r combine-stations}
+```{r}
 mu <- combine.stations(mu,mu.ecad)
@@
 
-```{r fill in missing data}
+```{r}
 ##Fill in the remaining gaps using a PCA-based interpolation scheme 
@@
 
-```{r probabilities}
-Pr.gt.30 <- PrexpPr(mu,fw,x0=30)
-plot(Pr.gt.30); grid()
-for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.30[,i]))
+```{r}
+Pr.gt.10 <- PrexpPr(mu,fw,x0=10)
+plot(Pr.gt.10); grid()
+for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.10[,i]))
 ```
@@
 
-another attempt to search for systematic influences on the precipitation statistics ($\mu$) is to examine the mean seasonal cycle. This type of approach may
+Another attempt to search for systematic influences on the precipitation statistics ($\mu$) is to examine the mean seasonal cycle. This type of approach may
 
-```{r aggregate}
+```{r}
 ## Estimate the mean seasonal cycle in the wet-day mean precipitaiton
@@
 
-```{r precipitation-intensity}
+```{r}
 ## Show the mean seasonal cycle in the wet-day mean precipitation
@@
 
-```{r wet-day frequency}
+```{r}
 ## Show the mean seasonal cycle in the wet-day frequency
@@
 
-```{r sesonal-precipitation}
+```{r}
 ## Show the mean seasonal cycle in the total precipitation
@@
 
-```{r winter-rain}
+```{r}
 ## Estimate the PCA for the annual cycle to identify the general features:
@@
 
-```{r downscale-ensemble}
+```{r}
 if (!file.exists('dse.mu.eu-circle-worstcase.rda')) {
   dse.mu <- DSensemble.mu.worstcase(x,biascorrect=TRUE,mask=TRUE,
-                                lon=c(-80,20),lat=c(10,60),
-                                rel.cord=FALSE,verbose=TRUE)
+                                    lon=c(-80,20),lat=c(10,60),rel.cord=FALSE,verbose=TRUE)
   save(file='dse.mu.eu-circle-worstcase.rda',dse.mu)
 } else load('dse.mu.eu-circle-worstcase.rda')
-#plot(dse.mu[[1]])
-plot.zoo(dse.mu[[1]],plot.type='single',col=rgb(1,0,0,0.05))
-grid()
+plot(dse.mu[[1]])
 ```
@@
 index(newlyn) <- year(newlyn)
-index(sonel) <- year(sonel)
+index(sonel) <- year(sonal)
 index(gloss) <- year(gloss)
@@
 ```{r slp-trend-map}
-slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg.nc',lon=c(-6,0),lat=c(48,52))
+slp.eobs <- retrieve('data.ECAD/pp_0.25deg_reg_v12.0.nc',lon=c(-6,0),lat=c(48,52))
 slp.eobs <- 10000*slp.eobs ## The data have some funny scaling factor
@@
 
-Pick three grid points from which we can estiamte the geostropic wind
+Pick three grid points from which we can estimate the geostropic wind.
 
 ```{r trend-map}
-## Fudge it - pick some windstations, but do not use their data. Reset the
-## coordinates to select a triangle of SLP from the EOBS data.
-ss <- select.station(param='slp',src='ecad',cntr=c('France','UNITED KINGDOM'))
-fx <- station(ss)
-triangle <- subset(fx,is=1:3)
+triangle <- subset(Z,is=c(2,3,7))
 ## The stations are not situated quite where there is valid slp data in the EOBS dataset
@@
 plot(annual(uv,nmin=300),main='zonal and meridional wind components',
-     sub='Estimated from triangulation of mean sea-level pressure from EOBS',zoom=7)
+     sub='Estimated from triangulation of mean sea-level pressure from EOBS')
 ```
@@
 ```{r predictand}
-pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(-6,-1),lat=c(50,52))
+pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(-6,-1),lat=c(50,52))
 mu <- annual(pr.eobs,FUN='wetmean')
@@
 ## The wet-day frequency and SLP
-ds.fw <- DS(subset(eof.fw,ip=1:4),eof.slp,eofs=1:5)
+ds.fw <- DS(subset(eof.fw,pattern=1:4),eof.slp,eofs=1:5)
 plot(ds.fw)
-ds.mu <- DS(subset(eof.mu,ip=1:4),eof.es,eofs=1:7)
+ds.mu <- DS(subset(eof.mu,pattern=1:4),eof.es,eofs=1:7)
 plot(ds.mu)

### 2016-11-28 CixPAG/CixPAG-imd-precip.Rmd

--- 2016-10-19 CixPAG/CixPAG-imd.Rmd
@@
 
-## Reading data predictand: maximum temperature
+## Reading data predictand: Precipitation
 
-```{r IMD maxmum temperature}
-tx <- retrieve('data/IMD/maxT_IMD.1951-2013.nc')
-map(tx)
-```
-
-Some initial exploration of the data - the mean summer maximum daily maximum temperature
-
-```{r}
-Tx.maxjja <- subset(as.4seasons(tx,FUN='max'),it='jja')
-eof.tx <- EOF(Tx.maxjja)
-plot(eof.tx,new=FALSE)
-```
-
-Trend in the men summer maximum daily maximum temperature
-
-```{r}
-map(Tx.maxjja,FUN='trend',new=FALSE)
-```
-
-The mean number of days with daily maximum temperature exceeding 40C.
-
-```{r}
-Tx.nhd40 <- annual(tx,FUN='count',threshold=40)
-map(Tx.nhd40,new=FALSE)
-```
-
-## Precipitation
-
-```{r IMD maxmum temperature}
+```{r IMD_precip, warning=FALSE}
 rr <- retrieve('data/IMD/tp_IMD.1951-2007.nc')
@@
 Monsoon rain statistics for the past - any interesting trends?
+Derive aggregated statistics for wet-day mean $\mu$ (`mu`), wet-day frequency $f_w$ (`fw`), and the annual total amount (`pt`).
+The threshold value for a wet day was set to 1mm/day.
 
-```{r}
+```{r aggregated_precip, warning=FALSE}
 mu <- aggregate(subset(rr,it=month.abb[6:9]),year,FUN='wetmean')
@@
 
-The precipitation intencity during the monsoon months
+The highest total precipitation amounts are found along the western coast of India and in the far north-eastern corner. The mean annual can be as high as 4500 mm/year.
+
+The mean wet-day mean precipitation (intensity) during the monsoon months:
+
+```{r map_mu, warning=FALSE}
+attr(mu,'variable') <- 'mu'
+map(mu,new=FALSE)
+```
+
+The mean intensity $\mu$ over the monsoon season (Jun--Sep) suggests that part of the total rainfall amounts are due to higher intensities in the same region.
 
 ```{r}
-map(mu,new=FALSE)
-map(mu,FUN='trend',new=FALSE)
+map(mu,FUN='trend',new=FALSE,
+    colbar=list(pal='t2m',breaks=seq(-2.5,2.5,by=0.1),rev=TRUE))
 ```
 
-The precipitation frequency during the monsoon months
+Trend estimates over 1951--2007 give a mixed picture with increases in some locations and decreases in others.
+
+The mean precipitation frequency during the monsoon months:
+
+```{r map_fw, warning=FALSE}
+attr(fw,'variable') <- 'f[w]'
+attr(fw,'unit') <- '%'
+fw <- fw*100
+map(fw,new=FALSE)
+```
+
+The regions with highest annual rainfall also are associated with highest wet-day frequencies, with rain up to 90\% of the days during the monsoon season.
+There are also some regions with fewer rainy days in the north west and east of the rainbelt (rain shadow) along the west coast.
 
 ```{r}
-map(fw,new=FALSE)
-map(fw,FUN='trend',new=FALSE)
+map(fw,FUN='trend',new=FALSE,
+    colbar=list(pal='t2m',breaks=seq(-2.5,2.5,by=0.1),rev=TRUE))
 ```
+
+Trend analysis for the monsoon wet-day frequency suggests a general reduction in the number of wet days over 1951--2007, but there have also been patches with increase in the number of rainy days in parts of southern India and along the northern border.
+
+## Predictands:
+
+Prepare the predictands - use EOFs for the gridded wet-day mean precipitation for the monsoon season:.
+
+```{r mu_eof, warning=FALSE,fig.width=8, fig.height=8}
+index(mu) <- year(mu)
+eof.mu <- EOF(mu,n=5)
+plot(eof.mu,new=FALSE)
+```
+
+The leading EOF accounts for 22\% of the variance and its spatial structure has a uniform character over most of the country, with the exception in the far eastern part which is anti-correlated to the rest of the country.
+The principal component (PC) does not reveal any particular and clear trend.
+Second mode (not shown) has strongest weights along the west coast, also with little trend.
+The PC for the third mode (9.6\% of variance) has a suspect (spurious?) spike towards the end of the record.
+
+EOFs for the gridded wet-day frequency for the monsoon season:
+
+```{r fw_eof, warning=FALSE,fig.width=8, fig.height=8}
+index(fw) <- year(fw)
+eof.fw <- EOF(fw,n=5)
+plot(eof.fw,new=FALSE)
+```
+
+The leading EOF for $f_w$ has greatest weights over the Mumbai region.
+The PC of the second mode has a suspicious peak towards the end of the record, associated with weights over southeastern India, which also were found to be a region with increasing rainy days in the trend analysis.
+The trend in this region is doubtful, however.
+The third EOF (9.2\% of the variance) has to a large extent geographically uniform weights, and is associated with a slight increasing trend.
+
+## Predictors:
+
+Use different predictors for $\mu$ and $f_w$ because we expect different dependencies to large-scale conditions due to physical reasons.
+Use the large-scale vapour saturation pressure $e_s$ (`es`) for $\mu$ and mean sea-level pressure (SLP) for $f_w$.
+The leading EOF for $e_s$ accounts for 78.7\% of the variance and dominates the variability.
+The greatest weights are found east off Srilanka and southern India, and it is also associated with an increasing trend.
+
+```{r predictor_for_mu, warning=FALSE,fig.width=8, fig.height=8}
+predictor <- retrieve('air.mon.mean.nc',lon=c(60,100),lat=c(0,25))
+predictor <- mask(predictor,land=TRUE)
+es <- C.C.eq(predictor)
+attr(es,'variable') <- 'e[s]'
+attr(es,'unit') <- 'Pa'
+attr(es,'longname') <- 'saturated vapour pressure'
+predictor <- annual(subset(es,it=month.abb[6:9]),nmin=3)
+eof.es <- EOF(predictor,n=10)
+plot(eof.es,new=FALSE)
+```
+
+The leading EOF for the Jun--Sep SLP accounts for 82\% of the variance and has a character of long-term variability - possibly a trend.
+The spatial structure resembles a dipole with greater weights over the northwestern parts of India and the typical monsoon SLP pattern.
+
+```{r predictor_for_fw, warning=FALSE,fig.width=8, fig.height=8}
+predictor <- retrieve('slp.mon.mean.nc',lon=c(60,100),lat=c(0,25))
+predictor <- annual(subset(predictor,it=month.abb[6:9]),nmin=3)
+eof.slp <- EOF(predictor,n=10)
+plot(eof.slp,new=FALSE)
+```
+
+The second EOF (7.6\%) has a southwest-northeast dipole structure, but little trend in the PC.
+
+## Downscaling
+
+The EOFs for $e_s$ from the NCEP/NCAR 1 reanalysis was used as a predictor to check the link between large and small scales for $\mu$, based on multiple regression (linear model, LM).
+The model was assessed through cross-validation analysis, which returned a correlation of 0.21.
+Changes in $\mu$ could to some extent be associated with increases in $e_s$ near Mumbai, and decreases elsewhere.
+
+```{r DS_test_mu, warning=FALSE,fig.width=8, fig.height=8}
+class(eof.mu) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+class(eof.es) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+index(eof.mu) <- as.Date(paste(year(eof.mu),'-07-01',sep=''))
+index(eof.es) <- as.Date(paste(year(eof.es),'-07-01',sep=''))
+ds.test <- DS(eof.mu,eof.es,ip=1:6)
+plot(ds.test,new=FALSE)
+```
+
+The link between large and small scales for $f_w$ was analysed in a similar fashion, and the cross-validation found a close association between monsoon-type SLP anomalies and the number of rainy days over large parts over India, with a correlation score of 0.76.
+
+```{r DS_test_fw, warning=FALSE,fig.width=8, fig.height=8}
+class(eof.fw) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+class(eof.slp) <- c("eof", "field", "season", "zoo") # Fix some incorrect settings
+index(eof.fw) <- as.Date(paste(year(eof.fw),'-07-01',sep=''))
+index(eof.slp) <- as.Date(paste(year(eof.slp),'-07-01',sep=''))
+ds.test <- DS(eof.fw,eof.slp,ip=1:6)
+plot(ds.test,new=FALSE)
+```
+
+The total precipitation is the product $p_t = n f_w \mu$, where $n$ is the total number of days.
+The skill associated with downscaling precipitation is to a large extent associated with wet-day frequency rather than the intensity.
+However, the probability of heavy rain events taking place is proportional to the wet-day frequency is the wet-day amounts approximately follow the exponential distribution: $Pr(X>x) = f_w e^{-x/\mu}$.

### 2016-11-28 CixPAG/CixPAG-imd.Rmd

--- 2016-10-20 CixPAG/CixPAG-imd.Rmd
@@
     dse.Tx <- DSensemble.eof(eof.Y,eof.X,ip=1:6,it='jja',
-                             predictor='air.mon.mean.nc',verbose=FALSE)
+                             predictor='air.mon.mean.nc',verbose=TRUE)
     save(dse.Tx,file='dse.Tx.jja.cixpag.rda')

### 2016-11-28 CixPAG/CixPAG-spell.Rmd

--- 2016-06-13 CixPAG/CixPAG-aphrodite.Rmd
@@
 ---
-title: "CixPAG-aphrodites"
+title: "CixPAG - spells"
 author: "Rasmus Benestad"
-date: "May 11, 2016"
-output: html_document
+date: "November 3, 2016"
+output: pdf_document
 ---
@@
 
-```{r}
-library(esd)
+```{r cars}
+summary(cars)
 ```
@@
 
-```{r}
-odurl <- 'http://iridl.ldeo.columbia.edu/SOURCES/.RIHN/.aphrodite/.V1101/.APHRO_MA_050deg/.precip/dods'
-con <- nc_open(url)
-print.ncdf4(con)
+You can also embed plots, for example:
+
+```{r pressure, echo=FALSE}
+plot(pressure)
 ```
+
+Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+
+Scatter plots: mean versus standard deviation of anomalies. 
+Combine several stations to link mean with mean spell length?
+
+Downscaling of heat wave spell length. Paper. Wheat.

### 2016-11-28 EU-Circle/eu-circle_case-study-3.Rmd

--- 2016-11-17 EU-Circle/eu-circle_case-study-3.Rmd
@@
 
-another attempt to search for systematic influences on the precipitation statistics ($\mu$) is to examine the mean seasonal cycle. This type of approach may
+Another attempt to search for systematic influences on the precipitation statistics ($\mu$) is to examine the mean seasonal cycle. This type of approach may
 
@@
 
-Pick three grid points from which we can estiamte the geostropic wind
+Pick three grid points from which we can estimate the geostropic wind.
 

### 2016-11-28 KSS/ds-pr4KSS.Rmd

@@
   ss <- select.station(param='precip',src='ecad',
-                       lon=c(5,35),lat=c(55,72),nmin=60)
+                       lon=c(5,30),lat=c(55,72),nmin=60)
   precip <- station(ss)

### 2016-11-29 EU-Circle/eu-circle_case-study-1.Rmd
EU-Circle new analysis

@@
 date: "May 2, 2016"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 bibliography: bibliography.bib
@@
 amncd <- annual(ncd)
-plot(amncd,col=c(rgb(0,0,1),rgb(1,0.3,0)),new=FALSE)
+plot(amncd,col=c(rgb(0,0,1),rgb(1,0.3,0)),new=FALSE,map.show=FALSE)
 grid()
@@
 ```{r,warning=FALSE}
-X <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(4,8),lat=c(43,46))
+X <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
 map(X,new=FALSE)

### 2016-11-29 EU-Circle/eu-circle_case-study-5.Rmd
Added a set with new Rmd scripts
EU-Circle new analysis

@@
 date: "May 2, 2016"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+
+## Analysis: precipitation
+
+Data from ECA&D: precipitation. Snow depth?
+
+```{r}
+library(esd)
+ss <- select.station(param='precip',cntr='Germany',alt=250,nmin=75)
+map(ss,new=FALSE)
+Y <- subset(station(ss),it=c(1950,2015))
+diagnose(Y)
+```
+
+Area total precipitation from gridded EOBS.
+
+```{r}
+rr <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(5,15),lat=c(45,55))
+map(rr,new=FALSE)
+Arr <- aggregate.area(rr)
+plot(rr,new=FALSE)
+```
+
+## Analysis: temperature
+
+```{r}
+library(esd)
+ss <- select.station(param='temperature',cntr='Germany',alt=500,nmin=75)
+map(ss)
+X <- subset(station(ss),it='mam')
+diagnose(X)
+```

### 2016-12-07 paper58/paper58.Rmd
new updates - Climatrans
updated paper58.Rmd

@@
-*Benestad et al - Upper-limit estimation of future precipitation return-values* 
+*Benestad et al - Simple and approximate estimation of future precipitation return-values* 
 -----------------------------------
@@
 library(esd)
-## Information about the system and session
-Sys.info()
+## Information about the session
 sessionInfo()
@@
 
-**Simple functions which estimate various statistics**
-
-  Functions to extract the numbers for mean,min,max,wettest month,driest month, estimate percentiles assuming an exponential distribution, skill-scores, and regression coefficients.
+**Simple functions to estimate various statistics**
+
+  Functions to extract the numbers for mean, min, max, wettest month, driest month, estimate percentiles assuming an exponential distribution, skill-scores, and regression coefficients.
 These are used to make the R-code simpler, shorter, and improve clarity.
@@
 
-**Generation of graphics/maps**
+**Generation of graphics**
 
@@
 
+##**Calculations and analysis**
+
 The definition of functions is followed by code that carry out the analysis based on these.  First get the data needed - if they are not stored locally, download from Figshare where they are stored.:
+
+**Access and prepare data**
 
@@
 
-Now the data is ready for the analysis. Calibrate the regression models and extract the regression coeffieients. Use 'apply' to speed up the process for multiple stations.
+**Statistical downscaling and analysis**
+
+Now the data is ready for the analysis. Calibrate the regression models and extract the regression coefficients. Use 'apply' to speed up the process for multiple stations.
 
@@
 
-Need to collect aggregated results based on the ensembles of GCM projections.
+Collect aggregated GCM projections (CMIP5) and apply the regression models to the ensembles.
 
@@
 ```
-Extract the statistics of the different emission scenarios and time slices
+Extract statistics of the wet-daymean associated with different emission scenarios and time slices and calculate the changes from the present (2010) to the far future (2100).
 
@@
 ```
-
-**Supporting material/analysis (SM)**
-
-Extract the essential information
+##** ----------- Supporting Material ------------------------**
+
+The supporting material includes diagnostics and plots that checks a number of assumptions made in this study.
+
+**Reorganise data and statistical information**
+Extract and reorganise essential information from previous analysis that will be used in the SM.
 
@@
 
-##** ----------- Supporting Material ------------------------**
-
-The supporting material includes a number of diagnostics and plots that chencks a number of assumptions made in this study.
-
-##**figure SM2**
-
-Presents the statistics of $R^2$ from the model calibration based on the different locations.
+##**Figure SM1**
+
+Figure SM1 shows a test of the assumption that that the wet-day daily precipitation is approximately exponentially distributed, by comparing the actual percentiles with quantiles estimated for different samples
+with different annual mean wet-day precipitation using the formula for exponentially distributed data: $q_p = -ln(1-p)\mu$.
+
+```{r}
+## Figure SM1.
+## test: See if the quantiles are consistent when the mean mu varies.
+qtest <- aggregate(y1,year,FUN='qqexp')
+qx <- c(coredata(qtest[,1:101]))
+qy <- c(coredata(qtest[,102:202]))
+
+par(bty='n')
+plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
+     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
+     main='Test: exponential distribution & changing mean',
+     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
+lines(c(0,40),c(0,40),col='grey')
+figlab('Figure SM1')
+```
+
+##**Figure SM2**
+
+Figure SM2 presents the $R^2$ statistics from the model calibration (i.e., the regression between the seasonal cycles of $\mu$ and $e_s$) for different locations.
 
@@
 ```
+
+##**Figure SM3**
+
+Figure SM3 compares the observed and predicted historical trends in the wet-day mean preciptation.
+
+```{r}
+par(bty='n',col.sub='grey',cex.sub=0.8)
+plot(trend.mux,trend.pre,pch=19,col=rgb(0.6,0.2,0,0.3),cex=1.5,
+     xlab=expression(paste('Observed trend in ',mu,' (mm/decade)')),
+     ylab=expression(paste('Predicted trend in ',mu,'(mm/decade)')),
+     xlim=xylim,ylim=xylim,
+     sub=paste('Mean correlation for local year-to-year variations over t=[',
+       start(MUx),',',end(MUx),
+       '] is ',round(mean(r.eval),2),' (',round(quantile(r.eval,0.05),2),', ',
+       round(quantile(r.eval,0.95),2),
+       ')',sep='') )
+grid()
+
+polygon(c(xylim[1],xylim[2],xylim[1],xylim[1]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
+        col=rgb(0.2,0.6,1,0.1),border=NA)
+polygon(c(xylim[1],xylim[2],xylim[2],xylim[2]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
+        col=rgb(1,0.2,0.2,0.1),border=NA)
+points(trend.mux,trend.pre,pch=1,col=rgb(0,0,0,0.1),cex=1.5)
+
+## Plot error bars 
+apply(rbind(trend.mux,trend.pre,trenderr.mux,trenderr.pre),2,
+      FUN=function(x) {lines(x[1]+c(-2,2)*x[3],x[2]+c(0,0),col=rgb(0.6,0.2,0,0.1))
+                       lines(x[1]+c(0,0),x[2]+c(-2,2)*x[4],col=rgb(0.6,0.2,0,0.1))
+                       lines(x[1]+c(-1,1)*0.01,x[2]+c(2,2)*x[4],col=rgb(0.6,0.2,0,0.05))
+                       lines(x[1]+c(-1,1)*0.01,x[2]+c(-2,-2)*x[4],col=rgb(0.6,0.2,0,0.05))
+                       lines(x[1]+c(-2,-2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
+                       lines(x[1]+c(2,2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
+                     })
+figlab('Figure SM3')
+```
+
+##**Figure SM4**
+
+Figure SM4 shows the historical trends in the wet-day mean preciptation.
+
+```{r}
+## Map showing trends in mu
+map(MU,FUN='trend',colbar=list(pal="burd",breaks=seq(-1,1,length=21)),cex=cex)
+figlab('Figure SM4',ypos=0.999)
+figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
+```
+
+## Figure SM5 is not used in either the main paper of supplementary material
+<##**Figure SM5**>
+
+#Figure SM5 is a histogram of the  shows the historical trends in the wet-day mean preciptation.
+
+#```{r}
+## 99-percentile wet-day mean and typical wet-day frequency
+## Prob(X>x) for annual maximum for 24hr precip
+## Pr(X > x) = 1/365.25 = fw*(1-p): p = 1 - 1/(365.25*fw)
+#par(bty='n')
+#W <- apply(fw,2,FUN='muclim')
+#hist(100*(1-1/(W[1,]*365.25)),col='steelblue',lwd=2,breaks=seq(90,100,by=0.1),
+#     main='Wet-day percentile for annual maximum 24-hr precipitation',
+#     xlab='p (%)')
+#grid()
+#figlab('Figure SM5')
+#```
+
+##**Figure SM6**
+
+Figure SM6 displays a histogram of historical trends in the wet-day frequency for the period... for the stations in Europe shown in Figure SM7.
+
+```{r}
+## Statistics of trend in wet-day frequency
+print('Wet-day frequency statistics')
+hist(fw.trend,breaks=seq(-50,50,by=1),col='grey',lwd=2,
+     main='Trend in wet-day frequency',
+     xlab=expression(paste(f[w],' (%/decade)')))
+grid()
+figlab('Figure SM6')
+```
+
+##**Figure SM7**
+
+Figure SM7 shows the spatial distribution of the historical trends in wet-day frequency on a map.
+
+```{r}
+## Map showing trends in fw
+#dev.new()
+map(FW,FUN='trend',colbar=list(breaks=seq(-0.05,0.05,length=21),rev=TRUE),cex=cex)
+figlab('Figure SM7',ypos=0.999)
+figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
+```
+
+##**Figure SM9**
+
+```{r}
+## Print the numbers:
+print('--- Changes in 20% returnvalue in terms of % from 2010:')
+print(lapply(mu.2100,summary))
+print(paste('Summary for',length(mu.2100[[1]]),'locations'))
+
+## Compare the regression coefficients derived from individual
+## seasonal cycle with that derived from mean climatology.
+
+data(mu.eq.f.tx)
+
+col <- rgb(0.1,0.1,0.7,0.25)
+mutx <- summary(mu.eq.f.tx)$coefficients[c(2,4)]
+b1 <- as.numeric(lapply(Beta,function(x) x[1]))
+e1 <- as.numeric(lapply(Beta,function(x) x[2]))
+
+par(bty='n')
+plot(b1,pch=19,col=col,cex=cexr2,
+     main=expression(paste('Scaling coefficient for ',mu,' and ',e[s])),
+     xlab='',ylab=expression(beta))
+grid()
+for (i in 1:length(b1)) {
+  lines(rep(i,2),b1[i]+e1[i]*c(-2,2),col=col)
+  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(-2,-2),col=col)
+  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(2,2),col=col)
+}
+polygon(c(1,rep(length(b1),2),rep(1,2)),
+        mutx[1]+mutx[2]*c(-2,-2,2,2,-2),
+        border=rgb(0.5,0.5,0.5,0.4),col=rgb(0.5,0.5,0.5,0.3))
+lines(c(1,length(b1)),rep(mutx[1],2),lwd=3,col=rgb(0.5,0.5,0.5,0.3))
+figlab('Figure SM9')
+```
+
+##**Figure SM10: Map of the future return values**
+
+```{r}
+print('Map of return values for 2100 with RCP4.5')
+rv <- MUx
+## x2100u gives the percentage of mu,
+coredata(rv) <- t((x2100u/100)*apply(coredata(MUx),2,'mean')*log(365.25*t(coredata(FWx))))
+cexr2 <- 1.5*c(as.numeric(r2)) + 0.2
+map(rv,FUN='mean',cex=cexr2,colbar=list(breaks=seq(30,100,by=1)))
+figlab('Figure SM10',ypos=0.999)
+figlab('Return values for 2100 assuming RCP4.5',xpos=0.3,ypos=0.999)
+```
+
 ##**Figure SM12**
@@
 
-##**Figure SM3**
-
-Historical trends in the wet-day mean preciptation
-
-```{r}
-par(bty='n',col.sub='grey',cex.sub=0.8)
-plot(trend.mux,trend.pre,pch=19,col=rgb(0.6,0.2,0,0.3),cex=1.5,
-     main=expression(paste('Trends in ',mu,': observed and predicted upper limit')),
-     xlab='Observed trend (mm/decade)',ylab='predicted trend (mm/decade)',
-     xlim=xylim,ylim=xylim,
-     sub=paste('Mean correlation for local year-to-year variations over t=[',
-       start(MUx),',',end(MUx),
-       '] is ',round(mean(r.eval),2),' (',round(quantile(r.eval,0.05),2),', ',
-       round(quantile(r.eval,0.95),2),
-       ')',sep='') )
-grid()
-
-polygon(c(xylim[1],xylim[2],xylim[1],xylim[1]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
-        col=rgb(0.2,0.6,1,0.1),border=NA)
-polygon(c(xylim[1],xylim[2],xylim[2],xylim[2]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
-        col=rgb(1,0.2,0.2,0.1),border=NA)
-points(trend.mux,trend.pre,pch=1,col=rgb(0,0,0,0.1),cex=1.5)
-
-## Plot error bars 
-apply(rbind(trend.mux,trend.pre,trenderr.mux,trenderr.pre),2,
-      FUN=function(x) {lines(x[1]+c(-2,2)*x[3],x[2]+c(0,0),col=rgb(0.6,0.2,0,0.1))
-                       lines(x[1]+c(0,0),x[2]+c(-2,2)*x[4],col=rgb(0.6,0.2,0,0.1))
-                       lines(x[1]+c(-1,1)*0.01,x[2]+c(2,2)*x[4],col=rgb(0.6,0.2,0,0.05))
-                       lines(x[1]+c(-1,1)*0.01,x[2]+c(-2,-2)*x[4],col=rgb(0.6,0.2,0,0.05))
-                       lines(x[1]+c(-2,-2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
-                       lines(x[1]+c(2,2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
-                     })
-figlab('Figure SM3')
-```
-
-##**Figure SM4**
-
-```{r}
-## Map showing trends in mu
-map(MU,FUN='trend',colbar=list(breaks=seq(-1,1,length=21),rev=TRUE),cex=cex)
-figlab('Figure SM4',ypos=0.999)
-figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
-```
-##**Figure SM5**
-
-```{r}
-## 99-percentile wet-day mean and typical wet-day frequency
-## The Tellus paper on specification approximately annual maximum.
-## Prob(X>x) for annual maximum for 24hr precip
-## Pr(X > x) = 1/365.25 = fw*(1-p): p = 1 - 1/(365.25*fw)
-par(bty='n')
-W <- apply(fw,2,FUN='muclim')
-hist(100*(1-1/(W[1,]*365.25)),col='steelblue',lwd=2,breaks=seq(90,100,by=0.1),
-     main='Wet-day percentile for annual maximum 24-precipitation',
-     xlab='p (%)')
-grid()
-figlab('Figure SM5')
-```
-
-##**Figure SM6**
-
-```{r}
-## Statistics of trend in wet-day frequency
-print('Wet-day frequency statistics')
-hist(fw.trend,breaks=seq(-50,50,by=1),col='grey',lwd=2,
-     main='Trend in wet-day frequency',
-     xlab=expression(paste(f[w],' (%/decade)')))
-grid()
-figlab('Figure SM6')
-```
-##**Figure SM7**
-
-```{r}
-## Map showing trends in fw
-#dev.new()
-map(FW,FUN='trend',colbar=list(breaks=seq(-0.05,0.05,length=21),rev=TRUE),cex=cex)
-figlab('Figure SM7',ypos=0.999)
-figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
-```
-
-##**Figure SM1**
-
-```{r}
-## Figure SM1.
-## test: See if the quantiles are consistent when the mean mu varies.
-qtest <- aggregate(y1,year,FUN='qqexp')
-qx <- c(coredata(qtest[,1:101]))
-qy <- c(coredata(qtest[,102:202]))
-
-par(bty='n')
-plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
-     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
-     main='Test: exponential distribution & changing mean',
-     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
-lines(c(0,40),c(0,40),col='grey')
-figlab('Figure SM1')
-```
-
 ##**Figure SM14: compare the mean seasonal variations in the different precipitation statistics**
@@
 
-##**Figure SM10: Map of the future return values**
-
-```{r}
-print('Map of return values for 2100 with RCP4.5')
-rv <- MUx
-## x2100u gives the percentage of mu,
-coredata(rv) <- t((x2100u/100)*apply(coredata(MUx),2,'mean')*log(365.25*t(coredata(FWx))))
-cexr2 <- 1.5*c(as.numeric(r2)) + 0.2
-map(rv,FUN='mean',cex=cexr2,colbar=list(breaks=seq(30,100,by=1)))
-figlab('Figure SM10',ypos=0.999)
-figlab('Return values for 2100 assuming RCP4.5',xpos=0.3,ypos=0.999)
-```
-
-##**Figure SM9**
-
-```{r}
-## Print the numbers:
-print('--- Changes in 20% returnvalue in terms of % from 2010:')
-print(lapply(mu.2100,summary))
-print(paste('Summary for',length(mu.2100[[1]]),'locations'))
-
-## Compare the regression coefficients derived from individual
-## seasonal cycle with that derived from mean climatology.
-
-data(mu.eq.f.tx)
-
-col <- rgb(0.1,0.1,0.7,0.25)
-mutx <- summary(mu.eq.f.tx)$coefficients[c(2,4)]
-b1 <- as.numeric(lapply(Beta,function(x) x[1]))
-e1 <- as.numeric(lapply(Beta,function(x) x[2]))
-
-par(bty='n')
-plot(b1,pch=19,col=col,cex=cexr2,
-     main=expression(paste('Scaling coefficient for ',mu,' and ',e[s])),
-     xlab='',ylab=expression(beta))
-grid()
-for (i in 1:length(b1)) {
-  lines(rep(i,2),b1[i]+e1[i]*c(-2,2),col=col)
-  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(-2,-2),col=col)
-  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(2,2),col=col)
-}
-polygon(c(1,rep(length(b1),2),rep(1,2)),
-        mutx[1]+mutx[2]*c(-2,-2,2,2,-2),
-        border=rgb(0.5,0.5,0.5,0.4),col=rgb(0.5,0.5,0.5,0.3))
-lines(c(1,length(b1)),rep(mutx[1],2),lwd=3,col=rgb(0.5,0.5,0.5,0.3))
-figlab('Figure SM9')
-```
-
 ##**Figure SM13**
@@
 
-Additional analysis testing the connection between local temperature and wet-day mean precipitation
-
-```{r}
-## Regression analysis between the wet-day mean mu and the mean temperature.
-
-corhalf <- function(x) {
-  n <- length(x)
-  x1 <- x[1:(n/2)]; x2 <- x[(n/2+1):n]
-  ok <- is.finite(x1) & is.finite(x2)
-  r <- cor(x1[ok],x2[ok])
-  return(r)
-}
-```
-
+Additional analysis testing the connection between local temperature and wet-day mean precipitation.
+
+**Regression analysis between the wet-day mean mu and the mean temperature**
+
+Load South American temperature and precipitation data from the climate database CLARIS and COST-VALUE(?). 
 ```{r}
@@
 ```
+
+Load and prepare temperature and precipitation data from North America from the GDCN database.
 
@@
 
+Combine and aggregate the observational data from North America and South America, and then fit a regression model based on comparing the annual mean $\mu$ and $e_s$ at the different stations.  
 ```{r}
@@
 
-#map(MU,FUN='mean',cex=0)
-#map(MU,FUN='trend',cex=0,colbar=list(rev=TRUE,breaks=seq(-1,1,length=21)))
-#map(FW,FUN='trend',cex=0,colbar=list(rev=TRUE,breaks=seq(-0.01,0.01,length=21)))
 calmu <- data.frame(x=as.numeric(apply(coredata(es),2,'mean',na.rm=TRUE)),
@@
 ## Estimate the year-by-year correlation in es and mu
+corhalf <- function(x) {
+  n <- length(x)
+  x1 <- x[1:(n/2)]; x2 <- x[(n/2+1):n]
+  ok <- is.finite(x1) & is.finite(x2)
+  r <- cor(x1[ok],x2[ok])
+  return(r)
+}
 

### 2016-12-07 EU-Circle/eu-circle_case-study-5.Rmd

--- 2016-05-02 EU-Circle/eu-circle_case-study-5.Rmd
@@
 This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+
+## Analysis: precipitation
+
+Data from ECA&D: precipitation. Snow depth?
+
+```{r}
+library(esd)
+ss <- select.station(param='precip',cntr='Germany',alt=250,nmin=75)
+map(ss,new=FALSE)
+Y <- station(ss)
+Y <- subset(Y,it=c(1950,2015))
+diagnose(Y)
+nv <- apply(coredata(Y),2,'nv')
+Y <- subset(Y,is=nv > 24000)
+```
+
+```{r}
+y <- subset(Y,is=1)
+coredata(y) <- rowMeans(coredata(Y))
+attr(y,'location') <- 'German higher elevation'
+attr(y,'longitude') <- mean(lon(Y))
+attr(y,'latitude') <- mean(lat(Y))
+attr(y,'altitude') <- min(alt(Y))
+plot(y)
+lines(trend(y),lty=2)
+grid()
+```
+
+Area total precipitation from gridded EOBS.
+
+```{r}
+rr <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=range(lon(Y)),lat=range(lat(Y)))
+map(rr,new=FALSE)
+Arr <- aggregate.area(rr)
+plot(rr,new=FALSE)
+```
+
+Wet day mean and frequency
+
+```{r}
+mu <- annual(rr,FUN='wetmean')
+fw <- annual(rr,FUN='wetfreq')
+MU <- aggregate.area(mu,FUN='mean')
+FW <- aggregate.area(fw,FUN='mean')
+plot(MU,new=FALSE)
+lines(trend(MU),lty=2)
+grid()
+plot(FW,new=FALSE)
+grid()
+lines(trend(FW),lty=2)
+```
+
+Plot the probability of more than 50mm/day
+
+```{r}
+Pr.50 <- FW*exp(-50/MU)
+plot(Pr.50)
+lines(trend(Pr.50),lty=2)
+grid()
+```
+
+## Analysis: temperature
+
+Check for melting.
+
+```{r}
+ss <- select.station(param='tn',cntr='Germany',alt=500,nmin=75)
+map(ss)
+X <- subset(station(ss),it=month.abb[1:4])
+diagnose(X)
+X <- subset(X,it=c(1940,2015))
+```
+
+```{r}
+tnx <- annual(X,FUN='max',nmin=100)
+plot(tnx,new=FALSE)
+grid()
+```

### 2016-12-07 EU-Circle/eu-circle_case-study-2.Rmd
EU-Circle and paper58

---
title: "Case2"
author: "Rasmus Benestad"
date: "December 7, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Storm and Sea Surge at a Baltic Sea Port , Gdynia Poland

### Scenario 1: Oil and Natural Gas Transport in port

The oil piping transportation system is operating at one of the Baltic Oil Terminals that is designated for the reception from ships, the storage and sending by carriages or cars the oil products. It is also designated for receiving from carriages or cars, the storage and loading the tankers with oil products such like petrol and oil. On the basis of the piping system operation and safety statistical data coming from its operators its safety will be modelled, identified and predicted. Further, under the assumption of the possibility of the stress of weather influence on the operation conditions in the form of maritime storm and/or other hard sea conditions existence, the piping system safety will be examined and the results will be compared with the previous results. The piping system safety and operation optimization will be performed and practical suggestions and procedures improving its  safety will be worked out.
Presentation: EU-CIRCLE Case Study 2 – Scenario 1

### Scenario 2: Chemical Spill due to extreme sea surges

The sea transport of dangerous chemicals is pretty safe at normal environmental conditions. However, the transported goods may be swept overboard as a result of bad weather and hard sea conditions. The released chemicals may create the threat for the crew and the ship as well as pollute the seawater and the coast. The Baltic Sea and nearby ecosystems are vulnerable to pollution and contamination as a result of sea accident during the dangerous goods transportation. Today, one major accident at the Baltic Sea happens every year approximately. There are more than 50,000 ships entering and leaving the Baltic Sea every year and about 2,000 vessels are at the Baltic Sea at any given moment as illustrated in Figure 1. This huge traffic across the Baltic Sea will be observed.

On the basis of the statistical data coming from reports of chemical accidents at sea, the risk of dangerous chemicals accidents at sea and their dangerous consequences will be modelled, identified and predicted. Further, under the assumption of the stress of weather influence on the operation conditions in the form of maritime storm and/or other hard sea  conditions existence, the risk of chemical spills at sea will be examined and the results will be compared with the previous results. The risk of chemical spills at sea the environment degradation optimization will be performed and practical suggestions and procedures decreasing the risk of the environment degradation will be worked out.

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r esd}
library(esd)
```

## analysis

Wind, stormtracks, wave heights, ans sea level.

```{r pressure}
data(imilast.M03)
```

### 2016-12-07 EU-Circle/eu-circle_torbay-precip.Rmd
EU-Circle and paper58

@@
 date: "April 18, 2016"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 library(esd)
-pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(-6,-1),lat=c(50,52))
+pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(-6,-1),lat=c(50,52))
 mu <- annual(pr.eobs,FUN='wetmean')
@@
 plot(mu,new=FALSE)
+grid()
+lines(trend(mu),lty=2)
 ## The area mean wet-day frequency
 plot(fw,new=FALSE)
+grid()
+lines(trend(mu),lty=2)
 ## The early part of the record looks suspect:
@@
 
-```{r EOFs-in-precip-mu}
+```{r EOFs-in-precip-mu, fig.width=8, fig.height=8}
 ## The area mean wet-day mean precipitation
@@
 
-```{r EOFs-in-precip-fw}
+```{r EOFs-in-precip-fw, fig.width=8, fig.height=8}
 ## Show maps of trends for the wet-day mean precipitation
@@
 
-```{r cca}
+```{r cca, fig.width=8, fig.height=8}
 ## The wet-day frequency and SLP
@@
 
-```{r DS}
+```{r DS, fig.width=8, fig.height=8}
 ## The wet-day frequency and SLP
-ds.fw <- DS(subset(eof.fw,pattern=1:4),eof.slp,eofs=1:5)
+ds.fw <- DS(subset(eof.fw,ip=1:4),eof.slp,ip=1:5)
 plot(ds.fw)
-ds.mu <- DS(subset(eof.mu,pattern=1:4),eof.es,eofs=1:7)
+ds.mu <- DS(subset(eof.mu,ip=1:4),eof.es,ip=1:7)
 plot(ds.mu)
@@
 
-```{r residual for mu}
+```{r residual for mu, fig.width=8, fig.height=8}
 ## The residual of the wet-day precipitation
@@
 For comparison the original data look like this:
-```{r EOF for mu}
+```{r EOF for mu, fig.width=8, fig.height=8}
 ## The residual of the wet-day precipitation
@@
 
-```{r residual for fw}
+```{r residual for fw, fig.width=8, fig.height=8}
 ## Repeat with the wet-day frequency for the benefit of comparison
@@
 For comparison the original data look like this:
-```{r EOF for fw}
+```{r EOF for fw, fig.width=8, fig.height=8}
 ## The residual of the wet-day precipitation

### 2016-12-07 EU-Circle/eu-circle_case-study-4.Rmd

---
title: "Case 4"
author: "Rasmus Benestad"
date: "December 7, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Bangladesh: tropical cyclones and stormsurges, heavi precipitation.

Bangladesh is highly vulnerable to climate induced hazrds and disasters and its coastal part are mostly threatened for the impacts of climate change in case of cyclone, storm surges, flood, salinity and tsunami disaster. Even last twenty years the coastal peoples have suffered due to super cyclone SIDR and AILA devastations. Last two deacdes a lots of disaster coping interventions have been provided to the community to achieve resiliency in this areas. So really it will be a field laboratory for the research of Disaster Resilience and Climate Change: Science and innovation for adaptation to climate  change: from assessing costs, risks and opportunities to demonstration of options and practices.

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r esd}
library(esd)
```

## Analysis

```{r pressure, echo=FALSE}
# Daily precipitation from BMD 
load('bmd.rda')
```

Scale the

### 2016-12-07 EU-Circle/eu-circle_case-study-SL.Rmd

--- 2016-11-28 CixPAG/CixPAG-spell.Rmd
@@
 ---
-title: "CixPAG - spells"
+title: "Sea level"
 author: "Rasmus Benestad"
-date: "November 3, 2016"
+date: "December 7, 2016"
 output: pdf_document
@@
 
+Evidence for coherent the local sea level rise based on PCA. Relate the local response to the global, expecting a physical connection
+
 ## R Markdown
@@
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
+```{r esd}
+library(esd)
+library(RgoogleMaps)
 ```
 
-## Including Plots
+## Analysis
 
@@
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+```{r sea-level-data}
+gsl <- GSL()
+gloss <- station.gcloss()
+sonel <- station.sonel()
+newlin <- station.newlin()
 ```
@@
 Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
-
-Scatter plots: mean versus standard deviation of anomalies. 
-Combine several stations to link mean with mean spell length?
-
-Downscaling of heat wave spell length. Paper. Wheat.

### 2016-12-07 paper58/paper58.R
paper58 - added files

@@ -1,1 +1,907 @@
-/home/rasmusb/Dropbox/Public/LaTeX/paper58/paper58.R
+## Rasmus Benestad, 2015-02-23.
+## R-script for the analysis presented in the manuscript  
+# Estimation of worst-case changes in precipitation
+
+rm(list=ls())
+#xlim <- c(-10,40); ylim <- c(35,70)
+xlim <- c(0,35); ylim <- c(45,70)
+readecad <- FALSE
+figshare=TRUE
+nmin=50
+
+## Check if you need to get the esd-package:
+install.esd <- ("esd" %in% rownames(installed.packages()) == FALSE)
+
+if (install.esd) {
+  print('Need to install the esd package')
+  ## Need online access.
+  ## Use the devtools-package for simple facilitation of installing.
+  if ("devtools" %in% rownames(installed.packages()) == FALSE)
+    install.packages('devtools')
+  library(devtools)
+  ## Install esd directly from github
+  install_github('metno/esd')
+  print('The latest version of esd has been installed')
+}
+
+## Start the esd-library:
+library(esd)
+
+## ----------------------------------------------------------------------
+## Function that produces figures as well as statistics for the
+## wet-day mean mu.
+
+
+
+## Figures/diagnostics: 
+## - Time series of annual means for displaying trends
+## - Seasonal variation / climatology
+
+
+## Return statistics for mean,min,max,wettest month,driest month
+muclim <- function(x) {
+  y <- coredata(x)
+  iX <- mean((1:12)[is.element(y,max(y))])
+  iN <- mean((1:12)[is.element(y,min(y))])
+  stats <- c(mean(y,na.rm=TRUE),min(y,na.rm=TRUE),max(y,na.rm=TRUE),iX,iN)
+  names(stats) <- c('mean','min','max','wettest month','driest month')
+  stats
+}
+
+## Produce a set of percentiles and their counterparts for the exponential distribution
+qqexp <- function(x,x0=1) {
+  x[x < x0] <- NA
+  if (sum(is.finite(x))>0) {
+    mu <- mean(x,na.rm=TRUE)
[…]

### 2016-12-07 paper58/paper58.Rmd
updated paper58.Rmd

@@
 
-The following functions are used for the presentation of the results
+The following functions are used for the presentation of the results.
 
@@
 ```{r,echo=TRUE}
-## Extract the mean, min, max, wettest month, and driest months in terms of mu
-## for all stations:
+## Extract the mean, min, max, wettest month, and driest months 
+##  in terms of mu for all stations:
 print('muclim')
@@
 
-Estimate trend statistics and extract only the results for locations with a good match.
-The models calibrated here only seem to be valid for regions dominated by convective precipitation and are poor over regions where orographically forced precipitation dominate.
+Estimate trend statistics if the historical wet-day mean and wet-day frequency.
 
@@
                       apply(MU,2,'mean',na.rm=TRUE)
-
+## Check data availability
 diagnose(MU)
-
+```
+
+Extract results (regression models, statistical information) only for locations where there is a good match between the seasonal cycles of $\mu$ and $e_s$ ($R^2 > 0.6$).
+The models calibrated here only seem to be valid for regions dominated by convective precipitation and are poor over regions where orographically forced precipitation dominate.
+
+```{r,echo=TRUE}
 ## Select only the sites which have an R2 greater than 0.6:
@@
 
-Collect aggregated GCM projections (CMIP5) and apply the regression models to the ensembles.
+Collect aggregated GCM projections (CMIP5) and apply the empirical models to the ensembles.
 
@@
 ```
-Extract statistics of the wet-daymean associated with different emission scenarios and time slices and calculate the changes from the present (2010) to the far future (2100).
+Extract statistics of the wet-day mean associated with different emission scenarios and time slices and calculate the changes from the present (2010) to the far future (2100).
 
@@
 mu2010.rcp2.6 <- lapply(Z.rcp2.6,'window',start=2010,end=2010)
-x2010 <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[3]]))
-x2010u <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[2]]))
+x2010 <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[3]])) ## Median value
+x2010u <- as.numeric(lapply(mu2010.rcp4.5,function(x) x[[2]])) ## 95th percentile
 y2010 <- as.numeric(lapply(mu2010.rcp8.5,function(x) x[[3]]))
@@
 
-## Data frams with changes in percentages:
-mu.2100 <- data.frame(mean.RCP4.5=x2100 - x2010,q95.RCP4.5=x2100u - x2010u,
-                      mean.RCP2.6=z2100 - z2010,q95.RCP4.5=z2100u - z2010u,
-                      mean.RCP8.5=y2100 - y2010,q95.RCP8.5=y2100u - y2010u)
+## Data frames with changes in percentages:
+mu.2100 <- data.frame(median.RCP4.5=x2100 - x2010,q95.RCP4.5=x2100u - x2010u,
+                      median.RCP2.6=z2100 - z2010,q95.RCP4.5=z2100u - z2010u,
+                      median.RCP8.5=y2100 - y2010,q95.RCP8.5=y2100u - y2010u)
 ```
@@
 ## Plot a map of projected values:
-## Map showing RCP4.5 ensemble mean and the upper 95% change in the
+## Map showing RCP4.5 ensemble median and the upper 95% change in the
 ## outer part of the symbol. Also an insert with box-plot diagram
@@
 print("Plot a map of projected values:")
-map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',gridlines=FALSE,
+map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',col='grey70',gridlines=FALSE,
     colbar=list(col=cols,n=12,type="p",h=0.6,v=1))
@@
 
-Load South American temperature and precipitation data from the climate database CLARIS and COST-VALUE(?). 
+Load South American temperature and precipitation data from the climate database CLARIS and European observations that were used in COST-VALUE experiment 1. 
 ```{r}
@@
 
-Combine and aggregate the observational data from North America and South America, and then fit a regression model based on comparing the annual mean $\mu$ and $e_s$ at the different stations.  
+Combine the observational data from North America, Europe, and South America. Then fit a linear regression model comparing the mean $\mu$ and $e_s$ values at different locations.  
 ```{r}
@@
 
-Assess the connection between temperature and the wet-day mean precipitation.
+Assess the connection between temperature and the wet-day mean precipitation in space, based on the regression models defined above.
 

### 2016-12-08 EU-Circle/eu-circle_case-study-4.Rmd
More material for paper58 + more scripts for EU-Circle

@@
 
-Scale the
+Scale the precipitation according to the indididual wet-day mean. Then take the mean of the station series to get an index of large-scale precipitation intensity.
+
+```{r}
+mu <- apply(coredata(pr.bmd),2,'wetmean')
+nv <- apply(coredata(pr.bmd),1,'nv')
+X <- zoo(rowMeans(t(t(pr.bmd)/mu),na.rm=TRUE),order.by=index(pr.bmd))
+X[nv < 11] <- NA
+X <- as.station(X,param='rain-index',unit='dimensionless',loc='Bangladesh',
+                lon=mean(lon(pr.bmd)),lat=mean(lat(pr.bmd)),alt=0)
+```
+
+Examine the index:
+
+```{r}
+plot(X,new=FALSE)
+```
+
+Check the seasonality
+
+```{r}
+wheel(X,new=FALSE)
+```
+
+The long-term perspective of the rainfall intensity
+
+```{r}
+plot(annual(X,'wetmean'),new=FALSE)
+grid()
+lines(trend(annual(X,'wetmean')),lty=2)
+```
+
+The long-term perspective of the wet-day frequency
+
+```{r}
+plot(annual(X,'wetfreq'),new=FALSE)
+grid()
+lines(trend(annual(X,'wetfreq')),lty=2)
+```
+
+Get data on storm tracks... IMILAST
+Correlation between storm tracks and stations.

### 2016-12-08 EU-Circle/eu-circle_case-study-SL.Rmd
More material for paper58 + more scripts for EU-Circle

@@
 
+```{r}
+dm <- function(x) {
+  if (!is.null(dim(x))) y <- zoo(t(t(coredata(x)) -                      colMeans(coredata(x),na.rm=TRUE)),order.by=index(x)) else
+                        y <- zoo(x - mean(x,na.rm=TRUE), order.by=index(x))
+  y <- attrcp(x,y)
+  class(y) <- class(x)
+  y
+}
+```
+
 ## Analysis
@@
 gsl <- GSL()
-gloss <- station.gcloss()
+gloss <- station.gloss()
 sonel <- station.sonel()
-newlin <- station.newlin()
+#newlyn <- station.newlyn()
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Examine the mean seasonal cycle in the tide gauge records:
+
+```{r}
+plot(aggregate(dm(gloss),month,'mean'),map.show=FALSE,new=FALSE)
+lines(aggregate(dm(gsl),month,'mean'))
+
+plot(aggregate(dm(sonel),month,'mean'),new=FALSE)
+```
+
+Examine globally distributed tide gauge data.
+
+```{r, fig.height=8}
+gloss <- subset(gloss,it=c(1950,1992))
+nv <- apply(coredata(gloss),2,'nv')
+gloss <- subset(gloss,is=nv > 400)
+gloss <- pcafill(gloss)
+pca <- PCA(annual(gloss))
+plot(pca,new=FALSE)
+```
+
+Tide gauge data from the coast of northern France. There are some missing data gaps, but ignore those.
+
+```{r, fig.height=8}
+sonel <- annual(sonel,nmin=275)
+sonel <- subset(sonel,it=c(1976,2015))
+nv <- apply(coredata(sonel),2,'nv')
+sonel <- subset(sonel,is=nv > 30)
+sonel <- pcafill(sonel)
+pca <- PCA(annual(sonel),n=3)
+plot(pca,new=FALSE)
+```
+
+Connect to the global mean sea level
+
+```{r}
+gsl <- annual(gsl); index(gsl) <- year(gsl)
+xy <- merge(zoo(annual(gsl)),zoo(pca),all=TRUE)
+cal.sl <- data.frame(y = xy[,2],x=xy[,1])
+slfit <- lm(y ~ x, data=cal.sl)
+print(summary(slfit))
+```
+
+Construct the effect of a global sea level rise on the local tide levels using the leading PC only:
+
+```{r}
+z <- gsl; dim(z) <- c(length(z),1)
+U <- attr(pca,'pattern')[,1]; dim(U) <- c(1,length(U))
+## Need to apply the matrix product to recover the original data. 
+y <- zoo((attr(pca,'eigenvalues')[1] * slfit$coefficient[2]) * (z %*% U),order.by=index(gsl))
+plot(y,plot.type='single')
+grid()
+```
+
+This regression model can now be applied to projected global mean sea levels from e.g.
+CMIP.
+Need to use the `slfit$coefficient[2]` together with `U`, `attr(pca,'eigenvalues')[1]` and the global sea level.

### 2016-12-08 paper58/paper58.Rmd
updated paper58.Rmd

@@
 ## Use subset - in the PCA there is something strange in the 1990s.
-fw.trend <- 100*apply(subset(FW,it=c(1950,1990)),2,'trend.coef')/
+## But we use the data from the 1990s and 2000s in the rest of the analysis!
+## It makes no sense to exclude it here, and the results are similar.
+fw.trend <- 100*apply(subset(FW,it=c(1961,2014)),2,'trend.coef')/
                       apply(FW,2,'mean',na.rm=TRUE)
-mu.trend <- 100*apply(subset(MU,it=c(1950,1990)),2,'trend.coef')/
+mu.trend <- 100*apply(subset(MU,it=c(1961,2014)),2,'trend.coef')/
                       apply(MU,2,'mean',na.rm=TRUE)
@@
 ```
-
-**Plot figure 3**
+Print the projected changes in return-values from 2010 to 2100.
+
+```{r}
+## Print the numbers:
+print('--- Changes in 20% returnvalue in terms of % from 2010:')
+print(lapply(mu.2100,summary))
+print(paste('Summary for',length(mu.2100[[1]]),'locations'))
+```
+
+**Plot Figure 3**
+
+Figure 3 shows the projected changes in the median (inner marker) and 95th precentile of the wet-day mean from the present (2010) to the far future (2100).
 
@@
 ## Evaluation:
-## Use the calibration strategy to predict the annual mu based on the
-## predictor (t2m -> e_s)
+## Use the calibration strategy to predict the annual mu based on the predictor (t2m -> e_s)
 print('evaluation: correlation')
@@
 Figure SM3 compares the observed and predicted historical trends in the wet-day mean preciptation.
+First extract information about trends in $\mu$.
+
+```{r,echo=TRUE}
+## Trend in projected wet-day mean
+trendbeta <- unlist(lapply(Z.rcp4.5,function(x) trend.coef(x[,3])))
+## Strange results:
+print((1:N)[trendbeta < 0])
+
+## test: Does the model predict observed trends?
+print('evaluation: trends')
+
+## Make sure to compare series with data for same times
+mask <- !is.finite(coredata(MUx))
+class(MUz) <- class(MUx)
+muz <- coredata(MUz)
+muz[mask] <- NA; dim(muz) <- dim(MUz)
+coredata(MUz) <- muz
+
+## Only look at stations with more than 50 years with data
+ok <- (apply(coredata(MUx),2,nv) > 50)
+MUz <- subset(MUz,is=ok)
+MUx <- subset(MUx,is=ok)
+FWx <- subset(FWx,is=ok)
+
+trend.mux <- apply(MUx,2,'trend.coef')
+trend.pre <- apply(MUz,2,'trend.coef')
+trenderr.mux <- apply(MUx,2,'trend.err')
+trenderr.pre <- apply(MUz,2,'trend.err')
+
+## Need to get a picture whether the predictions gives a plausible
+## upper limit.
+trend.sense <- data.frame(x=c(-trend.mux,trend.mux),
+                          y=c(-trend.pre,trend.pre))
+
+xylim <- max(abs(c(trend.mux,trend.pre)))*c(-1,1)
+```
+
+Then make a scatterplot comparing the observed and predicted trends.
 
@@
 
-## Figure SM5 is not used in either the main paper of supplementary material
-<##**Figure SM5**>
-
-#Figure SM5 is a histogram of the  shows the historical trends in the wet-day mean preciptation.
-
-#```{r}
-## 99-percentile wet-day mean and typical wet-day frequency
-## Prob(X>x) for annual maximum for 24hr precip
-## Pr(X > x) = 1/365.25 = fw*(1-p): p = 1 - 1/(365.25*fw)
-#par(bty='n')
-#W <- apply(fw,2,FUN='muclim')
-#hist(100*(1-1/(W[1,]*365.25)),col='steelblue',lwd=2,breaks=seq(90,100,by=0.1),
-#     main='Wet-day percentile for annual maximum 24-hr precipitation',
-#     xlab='p (%)')
-#grid()
-#figlab('Figure SM5')
-#```
-
-##**Figure SM6**
-
-Figure SM6 displays a histogram of historical trends in the wet-day frequency for the period... for the stations in Europe shown in Figure SM7.
+##**Figure SM5**
+
+Figure SM5 displays a histogram of historical trends in the wet-day frequency for the period... for the stations in Europe shown in Figure SM6.
 
@@
 grid()
-figlab('Figure SM6')
-```
-
-##**Figure SM7**
-
-Figure SM7 shows the spatial distribution of the historical trends in wet-day frequency on a map.
+figlab('Figure SM5')
+```
+
+##**Figure SM6**
+
+Figure SM6 shows the spatial distribution of the historical trends in wet-day frequency on a map.
 
@@
 #dev.new()
-map(FW,FUN='trend',colbar=list(breaks=seq(-0.05,0.05,length=21),rev=TRUE),cex=cex)
-figlab('Figure SM7',ypos=0.999)
+map(FW,FUN='trend',colbar=list(pal="t2m",breaks=seq(-0.05,0.05,length=21)),cex=cex)
+figlab('Figure SM6',ypos=0.999)
 figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
@@
 
+Figures SM7-8 require additional data and will follow after Figures SM9-11
+
 ##**Figure SM9**
 
-```{r}
-## Print the numbers:
-print('--- Changes in 20% returnvalue in terms of % from 2010:')
-print(lapply(mu.2100,summary))
-print(paste('Summary for',length(mu.2100[[1]]),'locations'))
-
-## Compare the regression coefficients derived from individual
-## seasonal cycle with that derived from mean climatology.
-
-data(mu.eq.f.tx)
-
-col <- rgb(0.1,0.1,0.7,0.25)
-mutx <- summary(mu.eq.f.tx)$coefficients[c(2,4)]
-b1 <- as.numeric(lapply(Beta,function(x) x[1]))
-e1 <- as.numeric(lapply(Beta,function(x) x[2]))
-
-par(bty='n')
-plot(b1,pch=19,col=col,cex=cexr2,
-     main=expression(paste('Scaling coefficient for ',mu,' and ',e[s])),
-     xlab='',ylab=expression(beta))
-grid()
-for (i in 1:length(b1)) {
-  lines(rep(i,2),b1[i]+e1[i]*c(-2,2),col=col)
-  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(-2,-2),col=col)
-  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(2,2),col=col)
-}
-polygon(c(1,rep(length(b1),2),rep(1,2)),
-        mutx[1]+mutx[2]*c(-2,-2,2,2,-2),
-        border=rgb(0.5,0.5,0.5,0.4),col=rgb(0.5,0.5,0.5,0.3))
-lines(c(1,length(b1)),rep(mutx[1],2),lwd=3,col=rgb(0.5,0.5,0.5,0.3))
-figlab('Figure SM9')
-```
-
-##**Figure SM10: Map of the future return values**
-
-```{r}
-print('Map of return values for 2100 with RCP4.5')
-rv <- MUx
-## x2100u gives the percentage of mu,
-coredata(rv) <- t((x2100u/100)*apply(coredata(MUx),2,'mean')*log(365.25*t(coredata(FWx))))
-cexr2 <- 1.5*c(as.numeric(r2)) + 0.2
-map(rv,FUN='mean',cex=cexr2,colbar=list(breaks=seq(30,100,by=1)))
-figlab('Figure SM10',ypos=0.999)
-figlab('Return values for 2100 assuming RCP4.5',xpos=0.3,ypos=0.999)
-```
-
-##**Figure SM12**
-
-A figure showing one example of projections for one location, showing all three emission scenarios.
+Figure SM9 shows an example of projections of the wet-day mean for one location, showing all three emission scenarios.
 
@@
 ## Example of estimates for 2050:
-print('Example plot - evolution')
+print('Example plot - projections of mu')
 par(bty='n')
@@
 grid()
-figlab('Figure SM12')
-```
-
-Further processing for Fig SM3: extract information about trends in $\mu$.
-
-```{r,echo=TRUE}
-## Trend in projected wet-day mean
-trendbeta <- unlist(lapply(Z.rcp4.5,function(x) trend.coef(x[,3])))
-## Strange results:
-print((1:N)[trendbeta < 0])
-
-## test: Does the model predict observed trends?
-print('evaluation: trends')
-
-## Make sure to compare series with data for same times
-mask <- !is.finite(coredata(MUx))
-class(MUz) <- class(MUx)
-muz <- coredata(MUz)
-muz[mask] <- NA; dim(muz) <- dim(MUz)
-coredata(MUz) <- muz
-
-## Only look at stations with more than 50 years with data
-ok <- (apply(coredata(MUx),2,nv) > 50)
-MUz <- subset(MUz,is=ok)
-MUx <- subset(MUx,is=ok)
-FWx <- subset(FWx,is=ok)
-
-trend.mux <- apply(MUx,2,'trend.coef')
-trend.pre <- apply(MUz,2,'trend.coef')
-trenderr.mux <- apply(MUx,2,'trend.err')
-trenderr.pre <- apply(MUz,2,'trend.err')
-
-## Need to get a picture whether the predictions gives a plausible
-## upper limit.
-trend.sense <- data.frame(x=c(-trend.mux,trend.mux),
-                          y=c(-trend.pre,trend.pre))
-
-xylim <- max(abs(c(trend.mux,trend.pre)))*c(-1,1)
-```
-
-##**Figure SM14: compare the mean seasonal variations in the different precipitation statistics**
-
-```{r}
+legend("topleft",legend=c("RCP4.5","RCP8.5","RCP2.6"),lty=1,lwd=2,
+       bty="o",box.lwd=0.2,
+       col=c(rgb(0.5,0.5,0.5,0.5),rgb(1,0.5,0.5,0.5),rgb(0.5,1,0.5,0.5)))
+figlab('Figure SM9')
+```
+
+##**Figure SM10**
+
+Figure SM10 presents a map of the NCEP reanalysis temperature for the North Atlantic region that is used as predictor for the downscaling (see e.g., Figure 1 and 3).
+
+```{r}
+## Show the predictor area:
+X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
+map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
+figlab('Figure SM10',xpos=0.8,ypos=0.999)
+```
+
+##**Figure SM11**
+
+Figure SM11 show the seasonal cycles of various precipitation statistics (mean, wet-day mean, wet-day frequency, mean consecutive wet days and dry days), calculated based on ECA&D observations for stations with long precipitation records.
+
+```{r}
+# y1 contains precipitation observations from ECA&D stations with long records
 pr.mean <- aggregate(y1,by=month,FUN='mean')
@@
                expression(bar(n[c*w*d])),expression(bar(n[c*d*d]))),bty='n',
-       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1))
-figlab('Figure SM9')
-```
-
-##**Figure SM13**
-
-```{r}
-## Show the predictor area:
-X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
-map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
-figlab('Figure SM13',xpos=0.8,ypos=0.999)
+       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1),
+       ncol=2)
+figlab('Figure SM11')
 ```
@@
 ## COST-VALUE
-load('VALUE_ECA_86_v2/stationsVALUE-exp1a.rda')
+load('stationsVALUE-exp1a.rda')
 Tx2 <- Tx
@@
 
-##** Figure SM8**
+```{r}
+## Estimate the correlation between the annual mean es and mu at the different sites
+corhalf <- function(x) {
+  n <- length(x)
+  x1 <- x[1:(n/2)]; x2 <- x[(n/2+1):n]
+  ok <- is.finite(x1) & is.finite(x2)
+  r <- cor(x1[ok],x2[ok])
+  return(r)
+}
+
+X <- matchdate(MU,es)
+Y <- matchdate(es,MU)
+w <- apply(coredata(fw),2,'mean')
+ok <- w > 0.25
+Z <- rbind(coredata(X),coredata(Y))
+r <- apply(Z,2,corhalf)
+hist(r)
+```
+
+##** Figure SM7**
 
@@
            cos(pi*calmu$lon/180)^2,
-           1-(1+sin(pi*calmu$lat/180))/2,0.15)
+           1-(1+sin(pi*calmu$lat/180))/2,0.45)
 cex <- 1.5*Fw/max(Fw,na.rm=TRUE)
@@
 points(calmu$lon,calmu$lat,pch=19,col=col,cex=0.3*cex)
-figlab('Figure SM8')
+figlab('Figure SM7')
 
@@
 attr(mu.eq.f.tx,'calibration_data') <- calmu
-save(file='esd/data/mu.eq.f.tx.rda',mu.eq.f.tx)
-```
-
-```{r}
-## Estimate the year-by-year correlation in es and mu
-corhalf <- function(x) {
-  n <- length(x)
-  x1 <- x[1:(n/2)]; x2 <- x[(n/2+1):n]
-  ok <- is.finite(x1) & is.finite(x2)
-  r <- cor(x1[ok],x2[ok])
-  return(r)
-}
-
-X <- matchdate(MU,es)
-Y <- matchdate(es,MU)
-w <- apply(coredata(fw),2,'mean')
-ok <- w > 0.25
-Z <- rbind(coredata(X),coredata(Y))
-r <- apply(Z,2,corhalf)
-hist(r)
-```
+save(file='mu.eq.f.tx.rda',mu.eq.f.tx)
+```
+
+##**Figure SM8**
+
+Figure SM9 compares the regression coefficients derived from the seasonal cycles of $\mu$ and the North Atlantic $\e_s$ (see e.g.
+Figures 1 and 3), to the regression coefficient from the comparison between the mean $\mu$ and $e_s$ from various sites (see Figure SM8).
+
+```{r}
+
+## Compare the regression coefficients derived from individual
+## seasonal cycles with that derived from mean climatology at different sites.
+
+data(mu.eq.f.tx)
+
+col <- rgb(0.1,0.1,0.7,0.25)
+mutx <- summary(mu.eq.f.tx)$coefficients[c(2,4)]
+b1 <- as.numeric(lapply(Beta,function(x) x[1]))
+e1 <- as.numeric(lapply(Beta,function(x) x[2]))
+
+par(bty='n')
+plot(b1,pch=19,col=col,cex=cexr2,
+     main=expression(paste('Scaling coefficient for ',mu,' and ',e[s])),
+     xlab='',ylab=expression(beta))
+grid()
+for (i in 1:length(b1)) {
+  lines(rep(i,2),b1[i]+e1[i]*c(-2,2),col=col)
+  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(-2,-2),col=col)
+  lines(i+c(-0.45,0.45),b1[i]+e1[i]*c(2,2),col=col)
+}
+polygon(c(1,rep(length(b1),2),rep(1,2)),
+        mutx[1]+mutx[2]*c(-2,-2,2,2,-2),
+        border=rgb(0.5,0.5,0.5,0.4),col=rgb(0.5,0.5,0.5,0.3))
+lines(c(1,length(b1)),rep(mutx[1],2),lwd=3,col=rgb(0.5,0.5,0.5,0.3))
+figlab('Figure SM8')
+```

### 2016-12-08 EU-Circle/eu-circle_torbay-precip.Rmd

@@
 grid()
-lines(trend(mu),lty=2)
+#lines(trend(mu),lty=2)
 ## The area mean wet-day frequency
@@
 grid()
-lines(trend(mu),lty=2)
+#lines(trend(fw),lty=2)
 ## The early part of the record looks suspect:

### 2016-12-09 paper58/paper58.Rmd
updated paper58

@@
 
-Below are definition of a number of functions used in the data processing and analysis. The first are simple functions used in ploting.
+##**Tools and functions**
+
+Below we define a number of functions used in the data processing and analysis. These functions are used to make the R-code more simple and improve clarity.
 
@@
 
-  Functions to extract the numbers for mean, min, max, wettest month, driest month, estimate percentiles assuming an exponential distribution, skill-scores, and regression coefficients.
-These are used to make the R-code simpler, shorter, and improve clarity.
+First we define the methods for calculating statistical measures of precipitation (mean, min, max, wettest month, and driest month), estimating percentiles assuming an exponential distribution, calculating skill-scores and regression coefficients.
 
@@
 
-Functions for reading the data and processing it so that it can readily be handelled in the analysis.
-
+Here we define functions for reading and processing data so that it can be readily handled in the analysis.
+
+```{r}
 ## Read the data from the CMIP5 GCMs
-```{r}
 readGCMs <- function(path='CMIP5.monthly/rcp45/',pattern='tas',
@@
 
-Main function that is used for calibrating the statistical model based on the mean seasonal cycle.
+This is the main function that is used for calibrating the statistical model based on the mean seasonal cycles of the predictand ($\mu$) and predictor ($e_s$, calculated from the temperature).
 
@@
 
-Functions for projection and prediction for the future/past:
+These functions apply the statistical model to GCM or reanalysis data in order to produce projections and predictions for the future/past:
 
@@
 
-The definition of functions is followed by code that carry out the analysis based on these.  First get the data needed - if they are not stored locally, download from Figshare where they are stored.:
+The definition of functions is followed by code that carry out the analysis based on these. First get the data needed - if they are not stored locally, download from Figshare where they are stored.
 
@@
 The convention here is that the wet-day mean precipitation (precipitation intensity) is represented by symbol $\mu$ and the variable name 'mu' in the R-scripts.
-The wet-day frequency $f_w$ is referret to as 'fw'.
+The wet-day frequency $f_w$ is referred to as 'fw'.
 Lower and upper case refer to the mean seasonal cycle and annually aggregated values respectively.
@@
 
-**Plot figure 1**
+**Figure 1**
 
@@
 ## Strange results:
+N <- length(Z.rcp4.5)
 print((1:N)[trendbeta < 0])
@@
            cos(pi*calmu$lon/180)^2,
-           1-(1+sin(pi*calmu$lat/180))/2,0.45)
+           1-(1+sin(pi*calmu$lat/180))/2,0.4)
 cex <- 1.5*Fw/max(Fw,na.rm=TRUE)
@@
 
-Figure SM9 compares the regression coefficients derived from the seasonal cycles of $\mu$ and the North Atlantic $\e_s$ (see e.g.
+Figure SM8 compares the regression coefficients derived from the seasonal cycles of $\mu$ and the North Atlantic $\e_s$ (see e.g.
 Figures 1 and 3), to the regression coefficient from the comparison between the mean $\mu$ and $e_s$ from various sites (see Figure SM8).
@@
 e1 <- as.numeric(lapply(Beta,function(x) x[2]))
+cex.r2 <- 1.5*unlist(r2) + 0.2
 
 par(bty='n')
-plot(b1,pch=19,col=col,cex=cexr2,
+plot(b1,pch=19,col=col,cex=cex.r2,xaxt="n",
      main=expression(paste('Scaling coefficient for ',mu,' and ',e[s])),
-     xlab='',ylab=expression(beta))
+     xlab='Observation site',ylab=expression(beta))
+axis(side=1, seq(0,1000,200), labels = FALSE)
 grid()
@@
 lines(c(1,length(b1)),rep(mutx[1],2),lwd=3,col=rgb(0.5,0.5,0.5,0.3))
+legend(200,-4e-3,legend=c("seasonal regression","spatial regression"),
+       col=c(col,rgb(0.5,0.5,0.5,0.4)),cex=0.8,lwd=2,box.lwd=0.5,
+       lty=c(1,1),pch=c(19,NA),fill=c(NA,rgb(0.5,0.5,0.5,0.3)),
+       border=c(NA,rgb(0.5,0.5,0.5,0.4)))
 figlab('Figure SM8')

### 2016-12-09 paper58/paper58.R

@@ -1,9 +1,5 @@
-## Rasmus Benestad, 2015-02-23.
-## R-script for the analysis presented in the manuscript  
-# Estimation of worst-case changes in precipitation
 
 rm(list=ls())
-#xlim <- c(-10,40); ylim <- c(35,70)
 xlim <- c(0,35); ylim <- c(45,70)
 readecad <- FALSE
 figshare=TRUE
@@ -26,17 +22,12 @@ if (install.esd) {
 
 ## Start the esd-library:
 library(esd)
+## Information about the session
+sessionInfo()
 
-## ----------------------------------------------------------------------
-## Function that produces figures as well as statistics for the
-## wet-day mean mu.
-
-
-
-## Figures/diagnostics: 
-## - Time series of annual means for displaying trends
-## - Seasonal variation / climatology
-
+##=============================================================
+## Define functions
+##=============================================================
 
 ## Return statistics for mean,min,max,wettest month,driest month
 muclim <- function(x) {
@@ -70,72 +61,6 @@ muskill <- function(x) {
   r2
 }
 
-## Calibrate a model for the wet-day mean mu using temperature as input
-mucal <- function(x,pre=NULL,lon=c(-100,30),lat=c(0,40),
-                  plot=FALSE,verbose=FALSE) {
-
-## If no pre, use the crude NCEP-reanalysis provided in esd 
-  if (is.null(pre)) {
-    if (verbose) print('default predictor')
-      t2m <- t2m.NCEP(lon=lon,lat=lat)
-      pre <- spatial.avg.field(C.C.eq(t2m))
-      if (plot) plot(EOF(t2m))
-  } else
-      if (is.character(pre))
-          pre <- spatial.avg.field(C.C.eq(retrieve(ncfile=pre,lon=lon,lat=lat))) else
-  if (inherits(pre,'field')) {
-      if (is.T(pre)) pre <- spatial.avg.field(C.C.eq(pre)) else
-      pre <- spatial.avg.field(pre)
-  } else if (inherits(pre,'station')) pre <- pre
-  z <- aggregate(pre,by=month,FUN='mean')
-    
-  cal <- data.frame(y=coredata(x),x=coredata(z))
[…]

### 2016-12-09 paper58/paper58.Rmd
updated paper58

@@
 ##**R set-up**
+
 First pre-amble that checks whether the esd-package is installed and installs it if needed.
@@
          ylim=ylim,xlim=xlim,
-         main='"Worst-case" fit based on seasonal variations',
+         main='Regression based on seasonal variations',
          sub=paste(loc(x),' (',round(lon(x),2),'E/',round(lat(x),2),'N; ',
@@
 ## Estimate trend statistics for both fw and mu:
-## Use subset - in the PCA there is something strange in the 1990s.
-## But we use the data from the 1990s and 2000s in the rest of the analysis!
-## It makes no sense to exclude it here, and the results are similar.
 fw.trend <- 100*apply(subset(FW,it=c(1961,2014)),2,'trend.coef')/
@@
 
-**Plot Figure 3**
+**Figure 3**
 
@@
 ```
-##** ----------- Supporting Material ------------------------**
+
+##**Supporting Material**
 
@@
 
-##**Figure SM1**
+Now prepare the supporting figures.
+
+```{r,echo=TRUE}
+## Supporting figures
+par(xaxt='n',yaxt='n',bty='n')
+plot(c(0,1),c(0,1),type='n',xlab='',ylab='')
+text(0.5,0.5,'Supporting figures',cex=2,font=2)
+```
+
+**Figure SM1**
 
@@
 
-##**Figure SM2**
+**Figure SM2**
 
@@
 ```{r,echo=TRUE}
-par(xaxt='n',yaxt='n',bty='n')
-plot(c(0,1),c(0,1),type='n',xlab='',ylab='')
-text(0.5,0.5,'Supporting figures',cex=2,font=2)
-
 ## Plot the statistics of R2:
@@
 
-##**Figure SM3**
+**Figure SM3**
 
@@
 
-##**Figure SM4**
+**Figure SM4**
 
@@
 
-##**Figure SM5**
+**Figure SM5**
 
@@
 
-##**Figure SM6**
+**Figure SM6**
 
@@
 
-Figures SM7-8 require additional data and will follow after Figures SM9-11
-
-##**Figure SM9**
+Figures SM7-8 require additional data and will follow after Figures SM9-11.
+
+**Figure SM9**
 
@@
 plot(Z.rcp4.5[[i]],plot.type='single',lty=c(2,2,1),lwd=c(1,1,2),
-     ylab='%',ylim=c(80,150),
+     xlab='Year',ylab='%',ylim=c(80,150),
      main=paste('Wet-day mean at',loc(subset(mux,is=i))))
@@
 
-##**Figure SM10**
+**Figure SM10**
 
@@
 
-##**Figure SM11**
+**Figure SM11**
 
@@
 
-##**Analysis from other continents and with local temperature**
+**Figures SM7-8: Analysis from other continents and with local temperature**
 
@@
 Combine the observational data from North America, Europe, and South America. Then fit a linear regression model comparing the mean $\mu$ and $e_s$ values at different locations.
+
 ```{r}
@@
 
-##** Figure SM7**
+**Figure SM7**
 
@@
 
-##**Figure SM8**
+**Figure SM8**
 

### 2016-12-09 paper58/paper58.R

@@ -136,7 +136,7 @@ mucal <- function(x,pre=NULL,lon=c(-100,30),lat=c(0,40),
                ylab=expression(paste(mu,' (mm/day)')),
                xlab=expression(paste(e[s],' (Pa)')),
                ylim=ylim,xlim=xlim,
-               main='"Worst-case" fit based on seasonal variations',
+               main='Regression based on seasonal variations',
                sub=paste(loc(x),' (',round(lon(x),2),'E/',round(lat(x),2),'N; ',
                          alt(x),'m.a.s.l.)',sep=''))
           segments(x0=cal$x,y0=cal$y,x1=cal$x,y1=cal$y+2*attr(x,'standard.error'),
@@ -387,9 +387,11 @@ pca <- svd(Mu)
 
 ## Plot maps with PCs:
 print('Maps with PCs')
+dev.new()
 mupcamap(mu,pca,1,xlim,ylim,r2)
 figlab('Figure 2a')
 dev.copy2pdf(file='figure2a.pdf')
+dev.new()
 mupcamap(mu,pca,2,xlim,ylim,r2)
 figlab('Figure 2b')
 dev.copy2pdf(file='figure2b.pdf')
@@ -410,6 +412,7 @@ fw.trend <- 100*apply(subset(FW,it=c(1961,2014)),2,'trend.coef')/
 mu.trend <- 100*apply(subset(MU,it=c(1961,2014)),2,'trend.coef')/
   apply(MU,2,'mean',na.rm=TRUE)
 ## Check data availability
+dev.new()
 diagnose(MU)
 
 ## Extract results (regression models, statistical information) for locations 
@@ -488,6 +491,7 @@ coly <- cols[cx2100u]
 Cex <- 1.25
 
 print("Plot a map of projected values:")
+dev.new()
 map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',col='grey70',gridlines=FALSE,
     colbar=list(col=cols,n=12,type="p",h=0.6,v=1))
 points(lon(mux),lat(mux),pch=19,col=colx,cex=Cex)
@@ -529,12 +533,20 @@ FWx <- subset(FWx,it=MUz)
 r.eval <- apply(rbind(coredata(MUz),coredata(MUx)),2,'cormu')
 print(summary(r.eval))
 
[…]

### 2016-12-13 paper58/paper58.Rmd
updated paper58

@@
 
-Figure SM1 shows a test of the assumption that that the wet-day daily precipitation is approximately exponentially distributed, by comparing the actual percentiles with quantiles estimated for different samples
+Figure SM1 show the seasonal cycles of various precipitation statistics (mean, wet-day mean, wet-day frequency, mean consecutive wet days and dry days), calculated based on ECA&D observations for stations with long precipitation records.
+
+```{r}
+# y1 contains precipitation observations from ECA&D stations with long records
+pr.mean <- aggregate(y1,by=month,FUN='mean')
+pr.mu <- aggregate(y1,by=month,FUN='wetmean')
+pr.fw <- aggregate(y1,by=month,FUN='wetfreq')
+y1.l <- spell(y1,threshold=1)
+pr.wet <- aggregate(subset(y1.l,is=1),by=month,FUN='mean')
+pr.dry <- aggregate(subset(y1.l,is=2),by=month,FUN='mean')
+
+par(bty='n',xaxt='n')
+plot(merge(pr.mean,pr.mu,10*pr.fw,pr.wet,pr.dry),plot.type='single',
+     col=c('steelblue','darkblue','grey','darkgreen','red'),
+     lwd=c(3,3,1,1,1),ylab="",xlab="Calendar month",main=loc(y1))
+grid()
+par(yaxt='s',xaxt='s')
+axis(1,at=1:12,labels=month.abb,cex.lab=0.7, col='grey')
+axis(4,at=10*pretty(pr.fw),pretty(pr.fw),col='grey')
+
+legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
+               expression(bar(n[c*w*d])),expression(bar(n[c*d*d]))),bty='n',
+       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1),
+       ncol=2)
+figlab('Figure SM1')
+```
+
+**Figure SM2**
+
+Figure SM2 shows a test of the assumption that that the wet-day daily precipitation is approximately exponentially distributed, by comparing the actual percentiles with quantiles estimated for different samples
 with different annual mean wet-day precipitation using the formula for exponentially distributed data: $q_p = -ln(1-p)\mu$.
@@
 ```{r}
-## Figure SM1.
+## Figure SM2.
 ## test: See if the quantiles are consistent when the mean mu varies.
@@
 lines(c(0,40),c(0,40),col='grey')
-figlab('Figure SM1')
-```
-
-**Figure SM2**
-
-Figure SM2 presents the $R^2$ statistics from the model calibration (i.e., the regression between the seasonal cycles of $\mu$ and $e_s$) for different locations.
-
-```{r,echo=TRUE}
-## Plot the statistics of R2:
-hist(100*as.numeric(r2),breaks=seq(0,100,by=5),lwd=2,col=rgb(0,0.3,0.5),
-     xlab=expression(paste(R^2,' (%)')),freq=TRUE,
-     main="Summary of regression scores")
-grid()
 figlab('Figure SM2')
@@
 
-Figure SM3 compares the observed and predicted historical trends in the wet-day mean preciptation.
-First extract information about trends in $\mu$.
-
-```{r,echo=TRUE}
-## Trend in projected wet-day mean
-trendbeta <- unlist(lapply(Z.rcp4.5,function(x) trend.coef(x[,3])))
-## Strange results:
-N <- length(Z.rcp4.5)
-print((1:N)[trendbeta < 0])
-
-## test: Does the model predict observed trends?
-print('evaluation: trends')
-
-## Make sure to compare series with data for same times
-mask <- !is.finite(coredata(MUx))
-class(MUz) <- class(MUx)
-muz <- coredata(MUz)
-muz[mask] <- NA; dim(muz) <- dim(MUz)
-coredata(MUz) <- muz
-
-## Only look at stations with more than 50 years with data
-ok <- (apply(coredata(MUx),2,nv) > 50)
-MUz <- subset(MUz,is=ok)
-MUx <- subset(MUx,is=ok)
-FWx <- subset(FWx,is=ok)
-
-trend.mux <- apply(MUx,2,'trend.coef')
-trend.pre <- apply(MUz,2,'trend.coef')
-trenderr.mux <- apply(MUx,2,'trend.err')
-trenderr.pre <- apply(MUz,2,'trend.err')
-
-## Need to get a picture whether the predictions gives a plausible
-## upper limit.
-trend.sense <- data.frame(x=c(-trend.mux,trend.mux),
-                          y=c(-trend.pre,trend.pre))
-
-xylim <- max(abs(c(trend.mux,trend.pre)))*c(-1,1)
-```
-
-Then make a scatterplot comparing the observed and predicted trends.
-
-```{r}
-par(bty='n',col.sub='grey',cex.sub=0.8)
-plot(trend.mux,trend.pre,pch=19,col=rgb(0.6,0.2,0,0.3),cex=1.5,
-     xlab=expression(paste('Observed trend in ',mu,' (mm/decade)')),
-     ylab=expression(paste('Predicted trend in ',mu,'(mm/decade)')),
-     xlim=xylim,ylim=xylim,
-     sub=paste('Mean correlation for local year-to-year variations over t=[',
-       start(MUx),',',end(MUx),
-       '] is ',round(mean(r.eval),2),' (',round(quantile(r.eval,0.05),2),', ',
-       round(quantile(r.eval,0.95),2),
-       ')',sep='') )
-grid()
-
-polygon(c(xylim[1],xylim[2],xylim[1],xylim[1]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
-        col=rgb(0.2,0.6,1,0.1),border=NA)
-polygon(c(xylim[1],xylim[2],xylim[2],xylim[2]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
-        col=rgb(1,0.2,0.2,0.1),border=NA)
-points(trend.mux,trend.pre,pch=1,col=rgb(0,0,0,0.1),cex=1.5)
-
-## Plot error bars 
-apply(rbind(trend.mux,trend.pre,trenderr.mux,trenderr.pre),2,
-      FUN=function(x) {lines(x[1]+c(-2,2)*x[3],x[2]+c(0,0),col=rgb(0.6,0.2,0,0.1))
-                       lines(x[1]+c(0,0),x[2]+c(-2,2)*x[4],col=rgb(0.6,0.2,0,0.1))
-                       lines(x[1]+c(-1,1)*0.01,x[2]+c(2,2)*x[4],col=rgb(0.6,0.2,0,0.05))
-                       lines(x[1]+c(-1,1)*0.01,x[2]+c(-2,-2)*x[4],col=rgb(0.6,0.2,0,0.05))
-                       lines(x[1]+c(-2,-2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
-                       lines(x[1]+c(2,2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
-                     })
-figlab('Figure SM3')
+Figure SM3 presents a map of the NCEP reanalysis temperature for the North Atlantic region that is used as predictor for the downscaling (see e.g., Figure 1 and 3).
+
+```{r}
+## Show the predictor area:
+X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
+map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
+figlab('Figure SM3',xpos=0.8,ypos=0.999)
 ```
@@
 
-Figure SM4 shows the historical trends in the wet-day mean preciptation.
-
-```{r}
-## Map showing trends in mu
-map(MU,FUN='trend',colbar=list(pal="burd",breaks=seq(-1,1,length=21)),cex=cex)
-figlab('Figure SM4',ypos=0.999)
-figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
-```
-
-**Figure SM5**
-
-Figure SM5 displays a histogram of historical trends in the wet-day frequency for the period... for the stations in Europe shown in Figure SM6.
-
-```{r}
-## Statistics of trend in wet-day frequency
-print('Wet-day frequency statistics')
-hist(fw.trend,breaks=seq(-50,50,by=1),col='grey',lwd=2,
-     main='Trend in wet-day frequency',
-     xlab=expression(paste(f[w],' (%/decade)')))
-grid()
-figlab('Figure SM5')
-```
-
-**Figure SM6**
-
-Figure SM6 shows the spatial distribution of the historical trends in wet-day frequency on a map.
-
-```{r}
-## Map showing trends in fw
-#dev.new()
-map(FW,FUN='trend',colbar=list(pal="t2m",breaks=seq(-0.05,0.05,length=21)),cex=cex)
-figlab('Figure SM6',ypos=0.999)
-figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
-```
-
-Figures SM7-8 require additional data and will follow after Figures SM9-11.
-
-**Figure SM9**
-
-Figure SM9 shows an example of projections of the wet-day mean for one location, showing all three emission scenarios.
+Figure SM4 shows an example of projections of the wet-day mean for one location, showing all three emission scenarios.
 
@@
        col=c(rgb(0.5,0.5,0.5,0.5),rgb(1,0.5,0.5,0.5),rgb(0.5,1,0.5,0.5)))
-figlab('Figure SM9')
-```
-
-**Figure SM10**
-
-Figure SM10 presents a map of the NCEP reanalysis temperature for the North Atlantic region that is used as predictor for the downscaling (see e.g., Figure 1 and 3).
-
-```{r}
-## Show the predictor area:
-X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
-map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
-figlab('Figure SM10',xpos=0.8,ypos=0.999)
-```
-
-**Figure SM11**
-
-Figure SM11 show the seasonal cycles of various precipitation statistics (mean, wet-day mean, wet-day frequency, mean consecutive wet days and dry days), calculated based on ECA&D observations for stations with long precipitation records.
-
-```{r}
-# y1 contains precipitation observations from ECA&D stations with long records
-pr.mean <- aggregate(y1,by=month,FUN='mean')
-pr.mu <- aggregate(y1,by=month,FUN='wetmean')
-pr.fw <- aggregate(y1,by=month,FUN='wetfreq')
-y1.l <- spell(y1,threshold=1)
-pr.wet <- aggregate(subset(y1.l,is=1),by=month,FUN='mean')
-pr.dry <- aggregate(subset(y1.l,is=2),by=month,FUN='mean')
-
-par(bty='n',xaxt='n')
-plot(merge(pr.mean,pr.mu,10*pr.fw,pr.wet,pr.dry),plot.type='single',
-     col=c('steelblue','darkblue','grey','darkgreen','red'),
-     lwd=c(3,3,1,1,1),ylab="",xlab="Calendar month",main=loc(y1))
+figlab('Figure SM4')
+```
+
+**Figure SM5**
+
+Figure SM5 presents the $R^2$ statistics from the model calibration (i.e., the regression between the seasonal cycles of $\mu$ and $e_s$) for different locations.
+
+```{r,echo=TRUE}
+## Plot the statistics of R2:
+hist(100*as.numeric(r2),breaks=seq(0,100,by=5),lwd=2,col=rgb(0,0.3,0.5),
+     xlab=expression(paste(R^2,' (%)')),freq=TRUE,
+     main="Summary of regression scores")
 grid()
-par(yaxt='s',xaxt='s')
-axis(1,at=1:12,labels=month.abb,cex.lab=0.7, col='grey')
-axis(4,at=10*pretty(pr.fw),pretty(pr.fw),col='grey')
-
-legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
-               expression(bar(n[c*w*d])),expression(bar(n[c*d*d]))),bty='n',
-       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1),
-       ncol=2)
-figlab('Figure SM11')
-```
-
-**Figures SM7-8: Analysis from other continents and with local temperature**
+figlab('Figure SM5')
+```
+
+**Figure SM6**
+
+Figure SM6 compares the observed and predicted historical trends in the wet-day mean preciptation.
+First extract information about trends in $\mu$.
+
+```{r,echo=TRUE}
+## Trend in projected wet-day mean
+trendbeta <- unlist(lapply(Z.rcp4.5,function(x) trend.coef(x[,3])))
+## Strange results:
+N <- length(Z.rcp4.5)
+print((1:N)[trendbeta < 0])
+
+## test: Does the model predict observed trends?
+print('evaluation: trends')
+
+## Make sure to compare series with data for same times
+mask <- !is.finite(coredata(MUx))
+class(MUz) <- class(MUx)
+muz <- coredata(MUz)
+muz[mask] <- NA; dim(muz) <- dim(MUz)
+coredata(MUz) <- muz
+
+## Only look at stations with more than 50 years with data
+ok <- (apply(coredata(MUx),2,nv) > 50)
+MUz <- subset(MUz,is=ok)
+MUx <- subset(MUx,is=ok)
+FWx <- subset(FWx,is=ok)
+
+trend.mux <- apply(MUx,2,'trend.coef')
+trend.pre <- apply(MUz,2,'trend.coef')
+trenderr.mux <- apply(MUx,2,'trend.err')
+trenderr.pre <- apply(MUz,2,'trend.err')
+
+## Need to get a picture whether the predictions gives a plausible
+## upper limit.
+trend.sense <- data.frame(x=c(-trend.mux,trend.mux),
+                          y=c(-trend.pre,trend.pre))
+
+xylim <- max(abs(c(trend.mux,trend.pre)))*c(-1,1)
+```
+
+Then make a scatterplot comparing the observed and predicted trends.
+
+```{r}
+par(bty='n',col.sub='grey',cex.sub=0.8)
+plot(trend.mux,trend.pre,pch=19,col=rgb(0.6,0.2,0,0.3),cex=1.5,
+     xlab=expression(paste('Observed trend in ',mu,' (mm/decade)')),
+     ylab=expression(paste('Predicted trend in ',mu,'(mm/decade)')),
+     xlim=xylim,ylim=xylim,
+     sub=paste('Mean correlation for local year-to-year variations over t=[',
+       start(MUx),',',end(MUx),
+       '] is ',round(mean(r.eval),2),' (',round(quantile(r.eval,0.05),2),', ',
+       round(quantile(r.eval,0.95),2),
+       ')',sep='') )
+grid()
+
+polygon(c(xylim[1],xylim[2],xylim[1],xylim[1]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
+        col=rgb(0.2,0.6,1,0.1),border=NA)
+polygon(c(xylim[1],xylim[2],xylim[2],xylim[2]),c(xylim[1],xylim[2],xylim[2],xylim[1]),
+        col=rgb(1,0.2,0.2,0.1),border=NA)
+points(trend.mux,trend.pre,pch=1,col=rgb(0,0,0,0.1),cex=1.5)
+
+## Plot error bars 
+apply(rbind(trend.mux,trend.pre,trenderr.mux,trenderr.pre),2,
+      FUN=function(x) {lines(x[1]+c(-2,2)*x[3],x[2]+c(0,0),col=rgb(0.6,0.2,0,0.1))
+                       lines(x[1]+c(0,0),x[2]+c(-2,2)*x[4],col=rgb(0.6,0.2,0,0.1))
+                       lines(x[1]+c(-1,1)*0.01,x[2]+c(2,2)*x[4],col=rgb(0.6,0.2,0,0.05))
+                       lines(x[1]+c(-1,1)*0.01,x[2]+c(-2,-2)*x[4],col=rgb(0.6,0.2,0,0.05))
+                       lines(x[1]+c(-2,-2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
+                       lines(x[1]+c(2,2)*x[3],x[2]+c(-1,1)*0.01,col=rgb(0.6,0.2,0,0.05))
+                     })
+figlab('Figure SM6')
+```
+
+**Figure SM7**
+
+Figure SM7 shows the historical trends in the wet-day mean preciptation.
+
+```{r}
+## Map showing trends in mu
+map(MU,FUN='trend',colbar=list(pal="burd",breaks=seq(-1,1,length=21)),cex=cex)
+figlab('Figure SM7',ypos=0.999)
+figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
+```
+
+**Figure SM8**
+
+Figure SM8 displays a histogram of historical trends in the wet-day frequency for the period... for the stations in Europe shown in Figure SM8.
+
+```{r}
+## Statistics of trend in wet-day frequency
+print('Wet-day frequency statistics')
+hist(fw.trend,breaks=seq(-50,50,by=1),col='grey',lwd=2,
+     main='Trend in wet-day frequency',
+     xlab=expression(paste(f[w],' (%/decade)')))
+grid()
+figlab('Figure SM8')
+```
+
+**Figure SM9**
+
+Figure SM9 shows the spatial distribution of the historical trends in wet-day frequency on a map.
+
+```{r}
+## Map showing trends in fw
+#dev.new()
+map(FW,FUN='trend',colbar=list(pal="t2m",breaks=seq(-0.05,0.05,length=21)),cex=cex)
+figlab('Figure SM9',ypos=0.999)
+figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
+```
+
+**Analysis from other continents and with local temperature**
 
@@
 
-**Figure SM7**
+**Figure SM10**
 
@@
 points(calmu$lon,calmu$lat,pch=19,col=col,cex=0.3*cex)
-figlab('Figure SM7')
+figlab('Figure SM10')
 
@@
 
-**Figure SM8**
-
-Figure SM8 compares the regression coefficients derived from the seasonal cycles of $\mu$ and the North Atlantic $\e_s$ (see e.g.
-Figures 1 and 3), to the regression coefficient from the comparison between the mean $\mu$ and $e_s$ from various sites (see Figure SM8).
+**Figure SM11**
+
+Figure SM11 compares the regression coefficients derived from the seasonal cycles of $\mu$ and the North Atlantic $\e_s$ (see e.g.
+Figures 1 and 3), to the regression coefficient from the comparison between the mean $\mu$ and $e_s$ from various sites (see Figure SM10).
 
@@
        border=c(NA,rgb(0.5,0.5,0.5,0.4)))
-figlab('Figure SM8')
-```
+figlab('Figure SM11')
+```

### 2016-12-13 paper58/paper58.R

@@ -540,7 +540,32 @@ plot(c(0,1),c(0,1),type='n',xlab='',ylab='')
 text(0.5,0.5,'Supporting figures',cex=2,font=2)
 dev.copy2pdf(file='figureSM0.pdf')
 
-## Figure SM1.
+## Figure SM1
+# y1 contains precipitation observations from ECA&D stations with long records
+pr.mean <- aggregate(y1,by=month,FUN='mean')
+pr.mu <- aggregate(y1,by=month,FUN='wetmean')
+pr.fw <- aggregate(y1,by=month,FUN='wetfreq')
+y1.l <- spell(y1,threshold=1)
+pr.wet <- aggregate(subset(y1.l,is=1),by=month,FUN='mean')
+pr.dry <- aggregate(subset(y1.l,is=2),by=month,FUN='mean')
+
+dev.new()
+par(bty='n',xaxt='n')
+plot(merge(pr.mean,pr.mu,10*pr.fw,pr.wet,pr.dry),plot.type='single',
+     col=c('steelblue','darkblue','grey','darkgreen','red'),new=FALSE,
+     lwd=c(3,3,1,1,1),ylab="",xlab="Calendar month",main=loc(y1))
+grid()
+par(yaxt='s',xaxt='s')
+axis(1,at=1:12,labels=month.abb,cex.lab=0.7, col='grey')
+axis(4,at=10*pretty(pr.fw),pretty(pr.fw),col='grey')
+legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
+               expression(bar(n)[c*w*d]),expression(bar(n)[c*d*d])),bty='n',
+       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1),
+       ncol=2)
+figlab('Figure SM1')
+dev.copy2pdf(file='figureSM1.pdf')
+
+## Figure SM2.
 ## test: See if the quantiles are consistent when the mean mu varies.
 qtest <- aggregate(y1,year,FUN='qqexp')
 qx <- c(coredata(qtest[,1:101]))
@@ -553,20 +578,48 @@ plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
      main='Test: exponential distribution & changing mean',
[…]

### 2016-12-16 paper58/paper58.Rmd
updated paper58

@@
   r2 <- as.numeric(r2)
-  colpc <- rev(colscal("t2m",n=100))
+  colpc <- rev(colscal("budrd",n=100))
+  print(colpc)
   cz <- round(100*abs(pca$v[,ipca])/quantile(abs(pca$v[,ipca]),0.95))
@@
   pch[!mo] <- 4
-  cex <- 1.5*c(r2) + 0.2
-
-  map(mu, xlim=xlim,ylim=ylim,bg='grey80',new=FALSE)
+  cex <- 1.2*c(r2) + 0.5
+  map(mu, xlim=xlim,ylim=ylim,bg='grey90',col='grey90',cex=0.5,new=FALSE)
   points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
@@
 X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
-map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
+map(X,projection='sphere',colbar=list(pal="budrd",breaks=seq(8,28,by=0.5)))
 figlab('Figure SM3',xpos=0.8,ypos=0.999)

### 2016-12-16 paper58/paper58.R

@@ -217,9 +217,11 @@ mupredict <- function(x,pre,verbose=FALSE,prct=TRUE) {
 
 ## Generate a map for the PC weights for the different modes of the
 ## wet-day mean mu annual cycle.
+
+
 mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
   r2 <- as.numeric(r2)
-  colpc <- rev(colscal("t2m",n=100))
+  colpc <- rev(colscal("budrd",n=100))
   cz <- round(100*abs(pca$v[,ipca])/quantile(abs(pca$v[,ipca]),0.95))
   cz[cz > 100] <- 100; cz[cz < 1] <- 1
   col <- colpc[cz]
@@ -228,9 +230,9 @@ mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
   hi <- c(r2) > 0.6
   pch[!hi] <- 1
   pch[!mo] <- 4
-  cex <- 1.5*c(r2) + 0.2
+  cex <- 1.2*c(r2) + 0.5
   
-  map(mu, xlim=xlim,ylim=ylim,bg='grey80',new=FALSE)
+  map(mu, xlim=xlim,ylim=ylim,bg='grey90',col='grey90',cex=0.5,new=FALSE)
   points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
   par(xpd=TRUE)
   text(10,73,paste('Annual cycle in PC',ipca,
@@ -585,7 +587,7 @@ dev.copy2pdf(file='figureSM2.pdf')
 ## Show the predictor area:
 X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
 dev.new()
-map(X,projection='sphere',colbar=list(breaks=seq(8,28,by=0.5)))
+map(X,projection='sphere',colbar=list(pal="budrd",breaks=seq(8,28,by=0.5)))
 figlab('Figure SM3',xpos=0.8,ypos=0.999)
 dev.copy2pdf(file='figureSM3.pdf')
 

### 2016-12-16 paper58/paper58.Rmd
updated minor bug in paper58.rmd

@@
   colpc <- rev(colscal("budrd",n=100))
-  print(colpc)
   cz <- round(100*abs(pca$v[,ipca])/quantile(abs(pca$v[,ipca]),0.95))

### 2016-12-16 paper58/paper58.Rmd
updated paper58

@@
+---
+output: html_document
+---
 *Benestad et al - Simple and approximate estimation of future precipitation return-values* 
@@
 
-  colbar(pretty(pca$v[,ipca],n=100),colpc,fig = c(0.05, 0.1, 0.05, 0.2))
+  colbar(pretty(pca$v[,ipca],n=100),colpc,fig = c(0.08, 0.12, 0.05, 0.2))
 
@@
 ## Map showing trends in mu
-map(MU,FUN='trend',colbar=list(pal="burd",breaks=seq(-1,1,length=21)),cex=cex)
+map(MU,FUN='trend',cex=1,colbar=list(pal="budrd",breaks=pretty(c(-0.5,0.5),n=21)))
 figlab('Figure SM7',ypos=0.999)
@@
 #dev.new()
-map(FW,FUN='trend',colbar=list(pal="t2m",breaks=seq(-0.05,0.05,length=21)),cex=cex)
+map(FW,FUN='trend',cex=1,colbar=list(pal="budrd",breaks=pretty(c(-0.03,0.03),n=21)))
 figlab('Figure SM9',ypos=0.999)
@@
 
-Figure SM11 compares the regression coefficients derived from the seasonal cycles of $\mu$ and the North Atlantic $\e_s$ (see e.g.
+Figure SM11 compares the regression coefficients derived from the seasonal cycles of $\mu$ and the North Atlantic $e_s$ (see e.g.
 Figures 1 and 3), to the regression coefficient from the comparison between the mean $\mu$ and $e_s$ from various sites (see Figure SM10).

### 2016-12-16 paper58/paper58.R

@@ -238,7 +238,7 @@ mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
   text(10,73,paste('Annual cycle in PC',ipca,
                    'with variance of',round(100*pca$d[ipca]^2/sum(pca$d^2)),'%'),pos=4)
   
-  colbar(pretty(pca$v[,ipca],n=100),colpc,fig = c(0.05, 0.1, 0.05, 0.2))
+  colbar(pretty(pca$v[,ipca],n=100),colpc,fig = c(0.08, 0.12, 0.05, 0.2))
   
   par(xaxt = "n", yaxt = "s", fig = c(0.05,0.25,0.80,0.95),
       mar = c(0, 1, 0, 0), new = TRUE, las = 1, cex.axis = 0.5,bty='n')
@@ -690,7 +690,8 @@ dev.copy2pdf(file='figureSM6.pdf')
 ## Figure SM7
 ## Map showing trends in mu
 dev.new()
-map(MU,FUN='trend',colbar=list(pal="burd",breaks=seq(-1,1,length=21)),cex=cex)
+map(MU,FUN='trend',cex=1,
+    colbar=list(pal="budrd",breaks=pretty(c(-0.5,0.5),n=21)))
 figlab('Figure SM7',xpos=0.8,ypos=0.999)
 figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),ypos=0.999)
 figlab(paste(range(year(MU)),collapse="-"),ypos=0.96)
@@ -710,7 +711,8 @@ dev.copy2pdf(file='figureSM8.pdf')
 ## Figure SM9
 ## Map showing trends in fw
 dev.new()
-map(FW,FUN='trend',colbar=list(pal="t2m",breaks=seq(-0.05,0.05,length=21)),cex=cex)
+map(FW,FUN='trend',cex=1,
+    colbar=list(pal="budrd",breaks=pretty(c(-0.03,0.03),n=12)))
 figlab('Figure SM9',xpos=0.8,ypos=0.999)
 figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),ypos=0.999)
 figlab(paste(range(year(MU)),collapse="-"),ypos=0.96)

### 2016-12-22 EU-Circle/eu-circle_case-study-2.Rmd
More material for paper58 + more scripts for EU-Circle

@@
 ---
-title: "Case2"
-author: "Rasmus Benestad"
-date: "December 7, 2016"
-output: pdf_document
+title: "eu-circle_case-study-2"
+author: "Kajsa Parding"
+date: "August 1, 2016"
+output: html_document
+bibliography: bibliography.bib
 ---
@@
 
-## Storm and Sea Surge at a Baltic Sea Port , Gdynia Poland
-
-### Scenario 1: Oil and Natural Gas Transport in port
-
-The oil piping transportation system is operating at one of the Baltic Oil Terminals that is designated for the reception from ships, the storage and sending by carriages or cars the oil products.
-It is also designated for receiving from carriages or cars, the storage and loading the tankers with oil products such like petrol and oil.
-On the basis of the piping system operation and safety statistical data coming from its operators its safety will be modelled, identified and predicted.
-Further, under the assumption of the possibility of the stress of weather influence on the operation conditions in the form of maritime storm and/or other hard sea conditions existence, the piping system safety will be examined and the results will be compared with the previous results.
-The piping system safety and operation optimization will be performed and practical suggestions and procedures improving its  safety will be worked out.
-Presentation: EU-CIRCLE Case Study 2 – Scenario 1
-
-### Scenario 2: Chemical Spill due to extreme sea surges
-
-The sea transport of dangerous chemicals is pretty safe at normal environmental conditions.
-However, the transported goods may be swept overboard as a result of bad weather and hard sea conditions.
-The released chemicals may create the threat for the crew and the ship as well as pollute the seawater and the coast.
-The Baltic Sea and nearby ecosystems are vulnerable to pollution and contamination as a result of sea accident during the dangerous goods transportation.
-Today, one major accident at the Baltic Sea happens every year approximately.
-There are more than 50,000 ships entering and leaving the Baltic Sea every year and about 2,000 vessels are at the Baltic Sea at any given moment as illustrated in Figure 1.
-This huge traffic across the Baltic Sea will be observed.
-
-On the basis of the statistical data coming from reports of chemical accidents at sea, the risk of dangerous chemicals accidents at sea and their dangerous consequences will be modelled, identified and predicted.
-Further, under the assumption of the stress of weather influence on the operation conditions in the form of maritime storm and/or other hard sea  conditions existence, the risk of chemical spills at sea will be examined and the results will be compared with the previous results.
-The risk of chemical spills at sea the environment degradation optimization will be performed and practical suggestions and procedures decreasing the risk of the environment degradation will be worked out.
-
-## R Markdown
-
 This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
 
-```{r esd}
+## Storm and Sea Surge at a Baltic Sea Port
+
+An oil piping transportation system is operating at Gdynia Port, Poland, one of the Baltic Oil Terminals.
+The system is designated to receive, store and send carriages and cars containing oil products and loading oil tankers.
+The purpose of the case study is to identify and predict safety risks associated with maritime transportation of petrol and oil.
+The transport of dangerous chemicals on sea is reasonably safe at normal environmental conditions.
+However, the transported goods may be swept overboard as a result of rough weather and hard sea conditions.
+The Baltic Sea and nearby ecosystems are vulnerable to pollution and contamination as a result of maritime accidents during the transporation of dangerous goods.
+Approximately one major accident happen at the Baltic Sea every year.
+There are more than 50,000 ships entering and leaving the Baltic Sea every year and about 2,000 vessels are at the Baltic Sea at any given moment.
+
+Here, we consider the potential stress of weather on the operation conditions, focusing specifically on maritime storms and hard sea conditions.
+Environmental conditions such as the sea level, wave height, wind speed and direction, and the occurence of storms affect the operation of the oil piping transporation system.
+Meteorological observations of these parameters are used to investigate the weather conditions at Gdynia Port and the Baltic Sea region.
+
+**Wind**
+
+Observations of the wind are sparse in the Baltic region, but approximate values of wind speed and direction can be calculated from the sea level pressure (slp).
+A gridded slp product from E-OBS dataset is used which is made available by the EU-FP6 project ENSEMBLES (http://ensembles-eu.metoffice.com) and the data providers in the ECA&D project (http://www.ecad.eu).
+
+***Import E-OBS data***
+
+Load the R-library ```esd``` which is used for data analysis.
+
+```{r,message=FALSE}
 library(esd)
@@
 
-## analysis
+Import the E-OBS slp data using the retrieve function of the esd package. The data is represented with lower-case $x$.
 
-Wind, stormtracks, wave heights, ans sea level.
+```{r,warning=FALSE}
+if (!file.exists('eu-circle_case-study-2_slp.rda')) {
+  x <- retrieve('pp_0.25deg_reg_v13.1.nc',path='~/Data/data.EOBS',
+                lon=c(10,30),lat=c(53,66))
+  save(file='eu-circle_case-study-2_slp.rda',x)
+} else load('eu-circle_case-study-2_slp.rda')
+## Only keep the ECAD data
+map(x,FUN='q95',new=TRUE)
+diagnose(x)
+```
 
-```{r pressure}
-data(imilast.M03)
+***Geostrophic wind***
+
+Calculate the geostrophic wind based on the slp data.
+
+```{r,warning=FALSE}
+if (!file.exists('eu-circle_case-study-2_slp.rda')) {
+  x <- retrieve('pp_0.25deg_reg_v13.1.nc',path='~/Data/data.EOBS',
+                lon=c(10,30),lat=c(53,66))
+  save(file='eu-circle_case-study-2_slp.rda',x)
+} else load('eu-circle_case-study-2_slp.rda')
+## Only keep the ECAD data
+map(x,FUN='q95',new=TRUE)
+diagnose(x)
 ```
+
+**Ship accidents in the Baltic Sea**
+baltic.events <- c(19931010,20000328,20010328,20040313,20060330,20081017,20081202,20091002,20100128)
+
+**Precipitation**
+
+Examine the trend in the wet-day mean precipitation $\mu$
+
+```{r,warning=FALSE}
+mu <- annual(x,'wetmean',nmin=250)
+wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
+q95 <- annual(x,'wq95')
+plot(mu,new=FALSE)
+for (i in 1:dim(mu)[2]) lines(trend(subset(mu,is=i)),lty=2)
+grid()
+
+```
+
+There appears to be a slight increasing trend in $\mu$ for most stations, according to a linear regression analysis against time.
+There are also pronounced year-to-year variations in $\mu$ with high "spikes" in some years.
+
+Need to check to see if $\mu$ is correlated with number of rainy days, e.g. can be high due to smaller samples and higher sampling fluctuations:
+
+```{r}
+x1 <- zoo(subset(annual(x,'count',threshold=1),is=1))
+y1 <- zoo(subset(mu,is=1))
+plot(x1,y1,xlab=expression(n[X>1*mm]),ylab=expression(mu),xlim=c(30,150),ylim=c(5,30),
+     main='Test: wet-day mean precipitation dependency on sample size')
+for (i in 2:dim(mu)[2]) {
+  x1 <- zoo(subset(annual(x,'count',threshold=1),is=i))
+  y1 <- zoo(subset(mu,is=i))
+  points(x1,y1)
+}
+grid()
+```
+
+There is no indication suggesting that high value of annual $\mu$ is connected with a low number of rain events.
+
+For many purposes, it may be useful to assume that the daily precipitation amount is approximately exponentially distributed for the days when it rains and that the probability of the precipitation exceedin a threshold value $x_0$ can be estimated according to $Pr(X>x_0) = f_w e^{-x_0/\mu}$ where $f_w$ is the wet-day frequency.
+Below is a test for whether the amount is close to exponential for a wet day.
+
+```{r,warning=FALSE}
+## check the relationship between wet-day 95-percentile and wet-day mean
+plot(-log(0.05)*zoo(mu[,2]),zoo(q95[,1]),pch=19,xlim=c(0,120),ylim=c(0,120),
+     main='Wet-day mean v.s. 95th wet percentile',
+     xlab=expression(paste(-ln(1-0.95)*mu,' (mm/day)')), ylab=expression(paste(q[95],' (mm/day)')))
+points(-log(0.05)*zoo(mu[,1]),zoo(q95[,2]),pch=19,col=rgb(1,0,0,0.2))
+points(-log(0.05)*zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
+points(-log(0.05)*zoo(mu[,4]),zoo(q95[,4]),pch=19,col=rgb(0,1,0,0.2))
+points(-log(0.05)*zoo(mu[,5]),zoo(q95[,5]),pch=19,col=rgb(0,0.7,0.7,0.2))
+points(-log(0.05)*zoo(mu[,6]),zoo(q95[,6]),pch=19,col=rgb(0.7,0,0.7,0.2))
+points(-log(0.05)*zoo(mu[,7]),zoo(q95[,7]),pch=19,col=rgb(0.5,0.5,0.5,0.2))
+grid()
+legend(0,120,loc(mu),pch=19,col=c('black',rgb(1,0,0,0.2),rgb(0,0,1,0.2),rgb(0,1,0,0.2),rgb(0,0.7,0.7,0.2),rgb(0.7,0,0.7,0.2),rgb(0.5,0.5,0.5,0.2)),bty='n',cex=0.5)
+```
+
+The scatter plot suggests that the points are mainly clustered around the diagonal and hence the 24-hr precipitation amoun is approximately exponentially distributed.
+The Marseille data seem to be less similar to the exponential distribution than the other sites.
+
+We also need to examine the trend in the wet-day frequency:
+
+```{r,warning=FALSE}
+fw <- annual(x,'wetfreq',nmin=250)
+plot(fw,new=FALSE)
+for (i in 1:dim(fw)[2]) lines(trend(subset(fw,is=i)),lty=2)
+grid()
+```
+
+The rain gauge data suggests decreasing trends in $f_w$ for most locations, except for the highest station (360 m a.s.l.) further inland (Faverges-de-la-Tour) and with a short record.
+
+Estimate the evolution in the probability for heavy precipitation: $Pr(X > x) = f_w e^{-x_0/\mu}$.
+
+```{r,warning=FALSE}
+x0 <- 100 #mm
+Pr <- zoo(1-pbinom(0,size=365,prob=coredata(fw)*exp(-x0/coredata(mu))),order.by=year(fw))
+class(Pr) <- class(mu)
+Pr <- attrcp(mu,Pr)
+attr(Pr,'variable') <- 'Pr'
+attr(Pr,'unit') <- 'probability'
+attr(Pr,'longname') <- paste('probability of at least one day with more than',x0,'mm') 
+plot(Pr,ylab=expression(Pr(n>0)),xlab='',new=FALSE,errorbar=FALSE,
+          main=paste('Probability of at least one day with more than',x0,'mm'))
+for (i in 1:dim(Pr)[2]) lines(trend(subset(Pr,is=i)),lty=2)
+grid()
+```
+
+The results indicate large annual variations in the probability of extreme precipitation ($X > 100mm$), suggesting a high sensitivity to both $\mu$ and $f_w$.
+This sensitivity is also seen in the different estimates for different locations, even if the difference between the mean $\mu$ is not as dramatic.
+
+One relevant parameter is the duration of dry and wet spells (number of dry/wet consecutive days).
+The grapghic below shows the length of dry (red) and wet (blue) intervals (dry are plotted with negative sign for clarity).
+
+```{r,warning=FALSE}
+z <- subset(x,it=c(1900,2016),is=1)
+ncd <- spell(z,threshold=1)
+plot(ncd)
+grid()
+```
+
+The dry runs usually last longer than the wet runs.
+some of the longest dry spells have lasted for ~100 days, but it has rarely rained more than 5 days in a row.
+One question is whether the dry/wet spell statistics can be considered to be a result of random/stochastic processes, or if they contain more structure.
+
+If there is only a fixed probability for wet/dry day that detemines the outcome, then we can use the geometric distribution to represent the statistics of the wet/dry spells:  <https://en.wikipedia.org/wiki/Geometric_distribution>.
+The question then is how closely the duration of wet/dry spells follow the geometric distribution.
+The histogram below compares the data to the pdfs for the geometric distribution and suggest a good match.
+
+```{r,warning=FALSE}
+hist(ncd)
+```
+
+The mean value if the parameter that describes the shape of the geometric distribution. The comparison suggests that it does not give a bad representation of the observed charactereistics.
+
+The second question is then: Are there any trends in the annual mean durations?
+
+```{r,warning=FALSE}
+amncd <- annual(ncd)
+plot(amncd,col=c(rgb(0,0,1),rgb(1,0.3,0)),new=FALSE)
+grid()
+lines(trend(subset(amncd,is=1)),lty=2)
+mcdd <- trend(subset(amncd,is=2))
+lines(mcdd,lty=1)
+print(attr(mcdd,'coefficients'))
+```
+
+There is a slight increasing trend in the annual mean duration of dry spells, which also has a consequence for the probability of long-lasting meteorological droughts.
+
+```{r,warning=FALSE}
+plot(1-pgeom(1:100,1/mean(mcdd)),type='l',xlab='consecutive dry days',ylab=expression(Pr(X > x)))
+plot(index(mcdd),1-pgeom(30, 1/mcdd),type='l',xlab='',ylab=expression(Pr(X > 30*days)),
+     main='Estimated probability for 30-day dry spell',ylim=c(0,0.05))
+for (i in 1:dim(Pr)[2]) lines(trend(subset(Pr,is=i)),lty=2)
+grid()
+```
+
+Linear trend analysis applied to both the duration of dry spells (solid line for one site in figure above) and the probability of more than 100 mm rain per day (dashed line showing several sites) both exhibit long-term increases over the last ~100 years.
+This tendency can be interpreted as the risk of long dry periods increases while the amount that falls as rain gets more extreme for the wet days.
+In other words, there has been a growing risk for both droughts and flash floods.
+
+***Gridded precipitation***
+
+The geographical pattern of trend in dry spell duration can be analysed using gridded data from EOBs (based on the ECA&D station records) Here gridded precipitation is represented with upper-case $X$.
+Gridded data are not ideal when it comes to extremes, as the gridding process makes the data spatially inhomogeneous.
+There are geographical variations in the climatological precipitation, with more precipitation over the higher elevations and the southern Alps.
+
+```{r,warning=FALSE}
+X <- retrieve('data.ECAD/rr_0.25deg_reg_v12.0.nc',lon=c(4,8),lat=c(43,46))
+map(X,new=FALSE)
+```
+
+A similar analysis of the mean dry spell duration indictes that the dry periods typically last longer near the coast.
+
+```{r}
+mean.dry <- function(x,na.rm=TRUE) mean(subset(spell(x,threshold=1),is=2),na.rm=na.rm)
+map(X,FUN='mean.dry',new=FALSE,colbar=list(pal='warm'))
+```
+
+It is also along the coast where there has been a strongest trend in the mean duration of the dry spells, which has an implication for the probability of long dry periods according to the geometric distribution.
+
+```{r}
+trend.dry <- function(x,na.rm=TRUE) trend.coef(subset(spell(x,threshold=1),is=2))
+map(X,FUN='trend.dry',new=FALSE,colbar=list(pal='warm'))
+```
+
+## Temperature
+
+For heat waves, use a similar approach as described in the supporting material of where a GLM was used to estimate the number of events with a basis of the 'Poisson' distribution.

### 2017-01-25 EU-Circle/eu-circle_case-study-5.Rmd
updated EU-Circle case study5

@@
 
+Test to see if the long-wave outgoing radiation (OLR) can be used as a predictor for the winter-time wet-day mean precipitation over Germany.
+It contains information about cloud tops, and the hypothesis is that heavy precipitation comes from clouds with high (and cold) cloud tops.
+
+```{r, fig.width=8, fig.height=8}
+olr <- retrieve('~/Downloads/ERAINT-OLR-mon.nc', lon=c(-50,40),lat=c(40,70))
+olr <- subset(as.4seasons(olr),it='djf')
+eof.olr<- EOF(olr,n=5)
+class(eof.olr)[3] <- "season"
+mu.djf <- subset(as.4seasons(Y,FUN='wetmean'),it='djf')
+pca.mu <- PCA(pcafill(mu.djf),n=5)
+ds.mu.orl <- DS(pca.mu,eof.olr)
+plot(ds.mu.orl,new=FALSE)
+```
+
 ## Analysis: temperature

### 2017-01-25 SPECS/specs-decade-mos.Rmd
updated EU-Circle case study5

@@
 library(esd)
-library(ncdf)
+#library(ncdf)
 setwd("~/R")
@@
 
-source("~/Dropbox/Public/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
+source("/home/rasmusb/disk1/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
 
@@
 aggregatedailyfield <- function(fname,cline='cdo monmean',
-                                output='mu_0.25deg_reg_v12.0.nc') {
+                                output='mu_0.25deg_reg.nc') {
   print('This function only works if you have CDO installed')
@@
   print('calculate mask')
-  system(paste('cdo gtc,1',ifile,'data/mask.nc'))
+  system(paste('cdo gtc,1',ifile,'~/data/mask.nc'))
   ## her blir ofile en mask med 1 i alle tidsskritt med data over konstanten (satt til 0),
@@
   X <- aggregatedailyfield(fname=fname,cline=cline)
-  Y <- aggregatedailyfield(fname=,'data/mask.nc',cline=cline,output='fw_0.25deg_reg_v12.0.nc')
-  file.remove(fname); file.remove('data/mask.nc')
+  Y <- aggregatedailyfield(fname=,'~/data/mask.nc',cline=cline,output='fw_0.25deg_reg.nc')
+  file.remove(fname); file.remove('~/data/mask.nc')
 }
@@
 print('Get the predictands: mu & fw from EOBS')
-if (!file.exists('mu_0.25deg_reg_v12.0.nc'))
-  mu.eobs <-  wetdaymean('data.ECAD/rr_0.25deg_reg_v12.0.nc') else
-  mu.eobs <- retrieve('mu_0.25deg_reg_v12.0.nc')
-fw.eobs <- retrieve('fw_0.25deg_reg_v12.0.nc')
+if (!file.exists('~/data/mu_0.25deg_reg.nc'))
+  mu.eobs <-  wetdaymean('~/data/data.ECAD/rr_0.25deg_reg.nc') else
+  mu.eobs <- retrieve('~/data/mu_0.25deg_reg.nc')
+fw.eobs <- retrieve('~/data/fw_0.25deg_reg.nc')
 attr(mu.eobs,'variable') <- 'mu'

### 2017-01-25 SPECS/specs_decadal-exp-d52.3_esd-loocv.R

@@ -1,14 +1,15 @@
 ## Script that applies MOS to SPECS decadal forecasts
 
 library(esd)
-library(ncdf)
+#library(ncdf)
 
 downloadfromunican <- FALSE   # Preparations - get the data
 downscaledreanalysis <- FALSE # SM: Assessment of large-small scale link
 looexperiments <- FALSE        # Cross-validation MOS on forecasts
 diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts
 
-source("~/Dropbox/Public/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
+#source("~/Dropbox/Public/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
+source('~/disk1/SPECS/R/udg2esd_met.R')
 
 ## This variable is needed because CDO did not fix the times properly
 ## for the predictands (EOBS)...
@@ -123,7 +124,7 @@ Xpredictor <- function(x,m=1:3,it=NULL,is=NULL,mask=FALSE,
 }
 
 aggregatedailyfield <- function(fname,cline='cdo monmean',
-                                output='mu_0.25deg_reg_v12.0.nc') {
+                                output='mu_0.25deg.nc') {
   print('This function only works if you have CDO installed')
   print(paste(cline,fname,output))
   system(paste(cline,fname,output))
@@ -144,7 +145,7 @@ wetdaymean <- function(ifile,cline='cdo yearmean',fname='data/wetdaymean.nc') {
   ##  print(paste('cdo ifthen -gtc,0',ifile,ifile,fname))
   ##  system(paste('cdo ifthen -gtc,0',ifile,ifile,fname))
   X <- aggregatedailyfield(fname=fname,cline=cline)
-  Y <- aggregatedailyfield(fname=,'data/mask.nc',cline=cline,output='fw_0.25deg_reg_v12.0.nc')
+  Y <- aggregatedailyfield(fname=,'data/mask.nc',cline=cline,output='fw_0.25deg.nc')
   file.remove(fname); file.remove('data/mask.nc')
 }
 
@@ -383,9 +384,9 @@ if (downloadfromunican) {
 
 print('Get the predictands: mu & fw from EOBS')
 if (!file.exists('mu_0.25deg_reg_v12.0.nc'))
[…]

### 2017-01-25 EU-Circle/eu-circle_case-study-5.Rmd
updated EU-Circle case study5

@@
 
+The calibration is fairly successful in identifying a connection between OLR and precipitation intencity.
+To see if other factors such as moisture has an effect, we estimate the residual from this regression analysis and repeat the analysis using saturation water vapour as predictor.
+
+```{r, fig.width=8, fig.height=8}
+## Estimate the residuals
+pca.res <- subset(pca.mu,it=ds.mu.orl)
+coredata(pca.res) <- coredata(pca.res) - coredata(ds.mu.orl)
+plot(pca.res)
+```
+
+```{r, fig.width=8, fig.height=8}
+mat <- retrieve('~/Downloads/ERAINT-t2m-mon.nc', lon=c(-90,40),lat=c(30,70))
+es <- C.C.eq(mat)
+es <- subset(as.4seasons(es),it='djf')
+es <- mask(es,land=TRUE)
+eof.es<- EOF(es,n=10)
+class(eof.es)[3] <- "season"
+ds.mu.es <- DS(pca.res,eof.es)
+plot(ds.mu.es,new=FALSE)
+```
+
 ## Analysis: temperature

### 2017-04-06 CixPAG/spell-statistics.Rmd
Cleansed version

---
title: "spell-statistics"
author: "Rasmus Benestad"
date: "February 7, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Assess the duration of warm spells in Indian tmax data.

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r tmax}
library(esd)
if (!file.exists("tmax.india.rda")) {
  ss <- select.station(param='tmax',src='GHCND',cntr='India',nmin=50)
  tmax <- station(ss)
  map(tmax,FUN='nv')
} else load("tmax.india.rda")
```

## Data quality

You can also embed plots, for example:

```{r data-quality-control}
diagnose(tmax)
## Remove periods with mostly missing data and stations with few valid data
Y <- subset(tmax,it=c(1970,2015))
nv <- apply(coredata(Y),2,FUN='nv')
Y <- subset(Y,is=nv > 10290)
```

## Spell sttatistics

Wheat: sowing: Upper temperature limit: 15.7C; T>31C for 5 days induces pollen sterility before the flowering phase; 
```{r, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
d <- dim(Y)
for (i in 1:d[2]) {
  sds <- spell(subset(Y,is=i),threshold=31)
  if (i==1) lws <- subset(sds,is=1) else
            lws <- combine.stations(lws,subset(sds,is=1))
  hist(sds,new=FALSE)
}
```

```{r}
tmax.jja <- subset(as.4seasons(Y),it='jja')
lws.jja <- subset(as.4seasons(lws),it='jja')
```

```{r}
## Figure 4.
xy <- merge(zoo(lws.jja),zoo(tmax.jja),all=FALSE)
cal.tmax.lws <- data.frame(x=c(coredata(xy)[,36:70]),y=c(coredata(xy[,1:35])))
cal.tmax.lws.log <- data.frame(x=log(c(coredata(xy)[,36:70])),y=log(c(coredata(xy[,1:35]))))
summary(lm(y ~x, data=cal.tmax.lws.log))
fit <- glm(y ~ x, data=cal.tmax.lws,family='poisson')
plot(cal.tmax.lws,xlab=expression(bar(T[max])),ylab=expression(bar(L[hot])))
pre <- data.frame(x=seq(28,42,by=0.1))
lines(pre$x,exp(predict(fit,newdata=pre)),col=rgb(1,0,0,0.4),lwd=3)
grid()
```

## Downscaling

### Prepare rthe preictand

```{r, fig.width=8, fig.height=8}
pca.lws <- PCA(pcafill(lws.jja),n=5)
pca.tmax <- PCA(pcafill(tmax.jja),n=5)
plot(pca.lws,new=FALSE)
plot(pca.tmax,new=FALSE)
```

```{r}
if (!file.exists("dse.tmax.india.rda")) {
  T2m <- subset(as.4seasons(retrieve("~/Downloads//air.mon.mean.nc",lon=c(65,95),lat=c(5,27))),it='jja')
  dse.tmax.india <- DSensemble.pca(pca.tmax,predictor=T2m,biascorrect = TRUE,ip=1:4)
  save(dse.tmax.india,file="dse.tmax.india.rda")
} else load("dse.tmax.india.rda")

```

## Supporting analysis

```{r}
#3 Function: autocorrelation

AR <- function(n,mean=1,sd=1,a1=0.8) {
  rn <- rnorm(n,mean=mean,sd=sd)
  for (i in 2:n) rn[i] <- (a1*rn[i-1] + (1-a1)*rn[i])
  invisible(rn)
}

```

Compare the mean temperature and spell length statistics for data in Europe.

```{r, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
## Script that reads European temperature data and explores the connection between 
## the seasonal mean temperature and the mean length of the warm/cold spells 
it <- 'jja'
if (it =='djf') {
  cold <- TRUE
  threshold <- 0
  is <- 2
} else {
  cold <- FALSE
  threshold <- 20
  is <- 1
}

ss <- select.station(src='ecad',param='tg',nmin=75)
d <- dim(ss)
x <- rep(NA,d[1]); y <- x; std <- y
q.spell <- rep(NA,d[1]*10); dim(q.spell) <- c(d[1],10); q.geom <- q.spell

if (!file.exists(paste('ecad.tg.',it,'.rda',sep=''))) {
  for (i in 1:d[1]) {
   z <- station(ss[i,])
   print(loc(z))
   ## Make sure that there are values above and below the given threshold - otherwise
   ## spell will not work.
   if ( (sum(z > threshold,na.rm=TRUE)>1000) & (sum(z < threshold,na.rm=TRUE)> 1000) ) {
    s <- spell(z,threshold=threshold)
    y[i] <- mean(subset(subset(s,is=is),it=it),na.rm=TRUE)

    ## Compare the spell-distribution with a geometric distribution
    q.spell[i,] <- quantile(subset(subset(s,is=is),it=it),probs=seq(0.05,0.95,by=0.1),na.rm=TRUE)
    q.geom[i,] <- qgeom(p=seq(0.05,0.95,by=0.1),prob=1/y[i])
    std[i] <- sd(subset(z,it=it),na.rm=TRUE)
    x[i] <- mean(subset(z,it=it),na.rm=TRUE)
    }
  }
  save(x,y,s,std,q.spell,q.geom,file=paste('ecad.tg.',it,'.rda',sep=''))
} else load(paste('ecad.tg.',it,'.rda',sep=''))
## Plot results

plot(x,y,main=paste('Mean',it,'temperature & mean length of intervals',
                    c('below','above')[c(cold,!cold)],threshold,'C'),
     sub='source: ECA&D',pch=19,col=rgb(0.5,0,0,0.3),
     xlab=expression(bar(T)),ylab=expression(bar(tau)))
grid()

## Monte-Carlo simulations to compare spell length with 
mstd <- 2*max(std,na.rm=TRUE)
if (!cold) mx <- mean(x,na.rm=TRUE) else mx <- 0
nmc <- 300
ymc <- rep(NA,nmc); xmc <- ymc
for (m in seq(mx-mstd,mx+mstd,length=nmc)) {
  coredata(z) <- AR(length(z),mean=m,sd=mstd)
  s <- spell(z,threshold=threshold)
  ymc[i] <- mean(subset(subset(s,is=is),it=it),na.rm=TRUE)
  xmc[i] <- m
  points(xmc,ymc,pch=19,col='grey75')
}

points(x,y,pch=19,col=rgb(0.5,0,0,0.3))

ix <- order(x); x <- x[ix]; y <- y[ix]
ok <- is.finite(x) & is.finite(y)
x <- x[ok]; y <- y[ok]; std <- std[ok]
calfit <- data.frame(x=x,y=y)
attr(calfit,'max(x)') <- max(x,na.rm=TRUE)
#fit <- lm(y ~ I(x) + I(x^2),data=calfit)
fit <- glm(y ~ x,data=calfit,family='poisson')
lines(calfit$x,exp(predict(fit)),col='red')

attr(x,'description') <- paste(it,'mean temperature (degC)')
if (cold) attr(y,'description') <- 'mean cold spell length (days)' else
          attr(y,'description') <- 'mean warm spell length (days)'
attr(y,'threshold') <- threshold
attr(x,'label') <- expression(bar(T))
attr(x,'Monte-Carlo') <- xmc
attr(y,'label') <- expression(bar(tau[T < T0]))
attr(y,'Monte-Carlo)') <- ymc
meanspell <- data.frame(meanT=x,meanL=y,std=std)
attr(meanspell,'fit') <- fit
attr(meanspell,'geometric.fit') <- data.frame(q.spell=q.spell,q.geom = q.geom)
save(meanspell,file=paste('meanspell',it,c('below','above')[c(cold,!cold)],threshold,'.rda',sep=''))

## Test if the spell length statistics is close to geometric
plot(c(q.spell),c(q.geom),main='Spell length statistics & the geometric distribution',
     xlim=range(q.spell,q.geom,na.rm=TRUE),
     ylim=range(q.spell,q.geom,na.rm=TRUE),pch=19,col=rgb(0,0,0,0.2),
     xlab=expression(q[p]),ylab='qgeom(p,1/mean)')
grid()
lines(range(q.spell,q.geom,na.rm=TRUE),range(q.spell,q.geom,na.rm=TRUE),col='red')

```

### 2017-04-06 EU-Circle/eu-circle_case-study-2.Rmd
Cleansed version

@@
 This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+
+Load the R-library ```esd``` which is used for data analysis.
+
+```{r,message=FALSE}
+library(esd)
+```
 
@@
 
-Observations of the wind are sparse in the Baltic region, but approximate values of wind speed and direction can be calculated from the sea level pressure (slp).
-A gridded slp product from E-OBS dataset is used which is made available by the EU-FP6 project ENSEMBLES (http://ensembles-eu.metoffice.com) and the data providers in the ECA&D project (http://www.ecad.eu).
+Synoptic storms are associated with winds, storm surges and waves.
+
+```{r}
+load('cyclones.baltic.rda')
+map(cyclones.baltic)
+```
+
+Plot the number of trajectories over time
+
+```{r}
+plot(cyclones.baltic)
+```
 
@@
 
-Load the R-library ```esd``` which is used for data analysis.
-
-```{r,message=FALSE}
-library(esd)
-```
+Observations of the wind are sparse in the Baltic region, but approximate values of wind speed and direction can be calculated from the sea level pressure (slp).
+A gridded slp product from E-OBS dataset is used which is made available by the EU-FP6 project ENSEMBLES (http://ensembles-eu.metoffice.com) and the data providers in the ECA&D project (http://www.ecad.eu).
 
@@
 if (!file.exists('eu-circle_case-study-2_slp.rda')) {
-  x <- retrieve('pp_0.25deg_reg_v13.1.nc',path='~/Data/data.EOBS',
+  x <- retrieve('pp_0.25deg_reg.nc',path='~/storeB/data/data.ECAD',
                 lon=c(10,30),lat=c(53,66))
@@
 
-```{r,warning=FALSE}
-if (!file.exists('eu-circle_case-study-2_slp.rda')) {
-  x <- retrieve('pp_0.25deg_reg_v13.1.nc',path='~/Data/data.EOBS',
-                lon=c(10,30),lat=c(53,66))
-  save(file='eu-circle_case-study-2_slp.rda',x)
-} else load('eu-circle_case-study-2_slp.rda')
-## Only keep the ECAD data
-map(x,FUN='q95',new=TRUE)
-diagnose(x)
+**Ship accidents in the Baltic Sea**
+
+```{r}
+baltic.events <- c(19931010,20000328,20010328,20040313,20060330,20081017,20081202,20091002,20100128)
+uv <- geostrophicwind(x)
 ```
-
-**Ship accidents in the Baltic Sea**
-baltic.events <- c(19931010,20000328,20010328,20040313,20060330,20081017,20081202,20091002,20100128)
 

### 2017-04-06 EU-Circle/eu-circle_case-study-3.Rmd
Cleansed version

@@
 ## Information about the system and session
+```
+
+```{r, eval=FALSE}
+## Get files on the outgoing long-wave radiation (OLR).
+cmip5.urls <- function(experiment='rcp45',varid='rlut',
+                       url="http://climexp.knmi.nl/CMIP5/monthly/", 
+                       path=NULL,off=FALSE,force=FALSE,verbose=FALSE) {
+  urlfiles <- "NA"
+  if(verbose) print("cmip5.urls")
+  if(is.null(path)) path <- getwd()
+  for (iexp in experiment) {
+    if(verbose) print(iexp)
+    for (ivar in varid) {
+      if(verbose) print(ivar)
+      ## Loop on the number of experiments
+      for (irun in 0:110) { ## 
+        if(verbose) print(paste(irun))
+        ## Update experiment number
+        if (irun < 10) run.id = paste("00",as.character(irun),sep="")
+        else if (irun < 100) run.id = paste("0",as.character(irun),sep="")
+        else run.id <- as.character(irun)
+
+        ## Create output directory for the climate experiment
+        path.exp <- file.path(path,experiment[grep(iexp,experiment)],
+                              fsep = .Platform$file.sep)
+        if (!file.exists(path.exp)) dir.create(path.exp)
+        if (verbose) print(path.exp[grep(iexp,experiment)])
+        ## Define the output file
+        destfile <- paste(path.exp,varid[grep(ivar,varid)],sep="/") 
+        destfile <- paste(destfile,"_Amon_ens_",sep="")
+        destfile <- paste(destfile,iexp,sep="")
+        destfile <- paste(destfile,run.id,sep="_")
+        destfile <- paste(destfile,".nc",sep="")
+
+        if (!file.exists(destfile) | force) {
+          ## Update output filename with attributes:
+          urlfile  <- paste(url,ivar,sep="")             # add var directory
+          urlfile  <- paste(urlfile,ivar,sep="/")        # add v.name
+          urlfile  <- paste(urlfile,"_Amon_ens_",sep="") # add text
+          urlfile  <- paste(urlfile,iexp,sep="")         # add exp.name
+          urlfile  <- paste(urlfile,run.id,sep="_")      # add exp ID number
+          urlfile  <- paste(urlfile,".nc",sep="")        # add file ext
+        }
+        urlfiles <- c(urlfiles,urlfile)
+        if (verbose) print(urlfile)
+      }
+    } # End for   
+  }
+  return(urlfiles[-1])
+}
+
+## download the CMIP5-files from the KNMI ClimateExplorer
+urls <- cmip5.urls()
+for (i in 1:length(urls)) if (!file.exists(substr(urls[i],43,70))) download.file(urls[i],substr(urls[i],43,70))
+
+```
+
+**Wind**
+
+Synoptic storms are associated with winds, storm surges and waves.
+
+```{r}
+load('cyclones.uk.rda')
+map(cyclones.uk,new=FALSE)
+```
+
+Plot the number of trajectories over time
+
+```{r}
+plot(cyclones.uk,new=FALSE)
 ```
@@
 ## The mean sea-level pressure
-slp <- retrieve('slp.mon.mean.nc',lon=c(-80,0),lat=c(20,70))
+slp <- retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-80,0),lat=c(20,70))
 ## The surface temperature
-t2m <- retrieve('air.mon.mean.nc',lon=c(-80,0),lat=c(10,70))
+t2m <- retrieve('~/Downloads/air.mon.mean.nc',lon=c(-80,0),lat=c(10,70))
 ## Saturation water pressure estimated from maritime temperatures
@@
 ```{r newlyn, results="hide", warning=FALSE, message=FALSE}
+if (!file.exists('newlyn.rda')) {
   newlyn <- station.newlyn()
   newlyn <- aggregate(newlyn,year,FUN='max',na.rm=TRUE)
+  save(newlyn,file='newlyn.rda')
+} else load('newlyn.rda')
 ```
@@
 Z <- subset(Z,is=(nv > 50))
-Z <- subset(Z,it=c(1950,2015))
+Z <- subset(Z,it=c(1966,2015))
+coredata(Z)[!is.finite(Z)] <- NA
+coredata(Z) <- t(t(coredata(Z)) - colMeans(coredata(Z),na.rm=TRUE))
 plot(Z); grid()
@@
 
-```{r sl-trend-map, warning=FALSE, message=FALSE}
+```{r sl-trend-map, warning=FALSE, message=FALSE, fig.height=8}
 zc <- coredata(Z)
@@
 zc -> coredata(Z)
-map(Z,FUN='trend',cex=-2)
+map(Z,pch=19,FUN='trend',cex=-2,add.text=TRUE)
+```
+
+Get the annual mean global mean sea-level and analyse the correspondense between that and the local sea level - we can expect a link from heuristic physics.
+Also include an account of the mean sea-level pressure.
+
+```{r, warning=FALSE, message=FALSE, fig.height=8}
+gsl <- annual(GSL())
+plot(gsl)
+slp <- annual(retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-20,10),lat=c(45,60)))
+eof.slp <- EOF(slp,n=5)
+## Manipulate the data: replace the last SLP EOF with the global sea-level and use DS to explore connections
+i1 <- is.element(year(eof.slp),year(gsl))
+i2 <- is.element(year(gsl),year(eof.slp))
+coredata(eof.slp)[i1,5] <- coredata(gsl)[i2]
+attr(eof.slp,'pattern')[,,5] <- 1
+attr(eof.slp,'eigenvalue') <- 1
+plot(eof.slp,ip=5,new=FALSE)
+```
+
+Prepare the local tidal data: PCA
+
+```{r, warning=FALSE, message=FALSE, fig.height=8}
+nv <- apply(coredata(Z),2,'nv')
+Z <- subset(Z,is=nv > 45)
+Z <- pcafill(Z)
+pca.Z <- PCA(Z)
+plot(pca.Z, new=FALSE)
 ```
@@
 ## Trend in upper wind direction
-SLP <- retrieve('slp.mon.mean.nc',lon=c(-50,30),lat=c(40,60))
+SLP <- retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-50,30),lat=c(40,60))
 eof.slp<- EOF(annual(SLP))
@@
 ## Estimate grostrophich wind
-UV <- geostrophicwind(SLP,progressbar=FALSE)
+#UV <- geostrophicwind(SLP,progressbar=FALSE)
 ```
@@
 ## wind rose
-windrose(uv)
+#windrose(UV)
 ```

### 2017-04-06 EU-Circle/eu-circle_case-study-4.Rmd
Bangladesh tide measurements
Cleansed version

@@
 
-Bangladesh: tropical cyclones and stormsurges, heavi precipitation.
+Bangladesh: tropical cyclones and stormsurges, heavy precipitation.
 
@@
 load('bmd.rda')
+map(pr.bmd,new=FALSE)
 ```
@@
 Correlation between storm tracks and stations.
+
+## Tidal measurements
+
+Time series and map of the tidal data from
+
+```{r}
+load('Bangladesh-tidal-stations.rda')
+plot(Z,zoom=8,new=FALSE)
+```

### 2017-04-06 EU-Circle/eu-circle_case-study-5.Rmd
Cleansed version

@@
 ## Analysis: precipitation
+
+```{r, eval=FALSE}
+## Get files on the outgoing long-wave radiation (OLR).
+cmip5.urls <- function(experiment='rcp45',varid='rlut',
+                       url="http://climexp.knmi.nl/CMIP5/monthly/", 
+                       path=NULL,off=FALSE,force=FALSE,verbose=FALSE) {
+  urlfiles <- "NA"
+  if(verbose) print("cmip5.urls")
+  if(is.null(path)) path <- getwd()
+  for (iexp in experiment) {
+    if(verbose) print(iexp)
+    for (ivar in varid) {
+      if(verbose) print(ivar)
+      ## Loop on the number of experiments
+      for (irun in 0:110) { ## 
+        if(verbose) print(paste(irun))
+        ## Update experiment number
+        if (irun < 10) run.id = paste("00",as.character(irun),sep="")
+        else if (irun < 100) run.id = paste("0",as.character(irun),sep="")
+        else run.id <- as.character(irun)
+
+        ## Create output directory for the climate experiment
+        path.exp <- file.path(path,experiment[grep(iexp,experiment)],
+                              fsep = .Platform$file.sep)
+        if (!file.exists(path.exp)) dir.create(path.exp)
+        if (verbose) print(path.exp[grep(iexp,experiment)])
+        ## Define the output file
+        destfile <- paste(path.exp,varid[grep(ivar,varid)],sep="/") 
+        destfile <- paste(destfile,"_Amon_ens_",sep="")
+        destfile <- paste(destfile,iexp,sep="")
+        destfile <- paste(destfile,run.id,sep="_")
+        destfile <- paste(destfile,".nc",sep="")
+
+        if (!file.exists(destfile) | force) {
+          ## Update output filename with attributes:
+          urlfile  <- paste(url,ivar,sep="")             # add var directory
+          urlfile  <- paste(urlfile,ivar,sep="/")        # add v.name
+          urlfile  <- paste(urlfile,"_Amon_ens_",sep="") # add text
+          urlfile  <- paste(urlfile,iexp,sep="")         # add exp.name
+          urlfile  <- paste(urlfile,run.id,sep="_")      # add exp ID number
+          urlfile  <- paste(urlfile,".nc",sep="")        # add file ext
+        }
+        urlfiles <- c(urlfiles,urlfile)
+        if (verbose) print(urlfile)
+      }
+    } # End for   
+  }
+  return(urlfiles[-1])
+}
+
+## download the CMIP5-files from the KNMI ClimateExplorer
+urls <- cmip5.urls()
+for (i in 1:length(urls)) if (!file.exists(substr(urls[i],43,70))) download.file(urls[i],substr(urls[i],43,70))
+
+```
 
@@
 
+A projection of future wet-day mean precipitation based on DSensemble will be carried out in two steps.
+First use the PCA from the original data as predictands in a set of runs with OLR as predictors.
+Then use the residual as estimated above as predictand for a set of DSensembles runs with the water vapour pressure as predictors.
+The final result is the sum of these two steps.The wet-day mean precipitation is projected through s single stemp where DSensemble uses SLP as predictor.
+The seasonal means are then estimated as the product between the wet-day frequency and the wet-day mean precipitation.
+
 ## Analysis: temperature

### 2017-04-06 EU-Circle/eu-circle_torbay-precip.Rmd
Cleansed version

@@
 
+```{r, eval=FALSE}
+## Get files on the outgoing long-wave radiation (OLR).
+cmip5.urls <- function(experiment='rcp45',varid='rlut',
+                       url="http://climexp.knmi.nl/CMIP5/monthly/", 
+                       path=NULL,off=FALSE,force=FALSE,verbose=FALSE) {
+  urlfiles <- "NA"
+  if(verbose) print("cmip5.urls")
+  if(is.null(path)) path <- getwd()
+  for (iexp in experiment) {
+    if(verbose) print(iexp)
+    for (ivar in varid) {
+      if(verbose) print(ivar)
+      ## Loop on the number of experiments
+      for (irun in 0:110) { ## 
+        if(verbose) print(paste(irun))
+        ## Update experiment number
+        if (irun < 10) run.id = paste("00",as.character(irun),sep="")
+        else if (irun < 100) run.id = paste("0",as.character(irun),sep="")
+        else run.id <- as.character(irun)
+
+        ## Create output directory for the climate experiment
+        path.exp <- file.path(path,experiment[grep(iexp,experiment)],
+                              fsep = .Platform$file.sep)
+        if (!file.exists(path.exp)) dir.create(path.exp)
+        if (verbose) print(path.exp[grep(iexp,experiment)])
+        ## Define the output file
+        destfile <- paste(path.exp,varid[grep(ivar,varid)],sep="/") 
+        destfile <- paste(destfile,"_Amon_ens_",sep="")
+        destfile <- paste(destfile,iexp,sep="")
+        destfile <- paste(destfile,run.id,sep="_")
+        destfile <- paste(destfile,".nc",sep="")
+
+        if (!file.exists(destfile) | force) {
+          ## Update output filename with attributes:
+          urlfile  <- paste(url,ivar,sep="")             # add var directory
+          urlfile  <- paste(urlfile,ivar,sep="/")        # add v.name
+          urlfile  <- paste(urlfile,"_Amon_ens_",sep="") # add text
+          urlfile  <- paste(urlfile,iexp,sep="")         # add exp.name
+          urlfile  <- paste(urlfile,run.id,sep="_")      # add exp ID number
+          urlfile  <- paste(urlfile,".nc",sep="")        # add file ext
+        }
+        urlfiles <- c(urlfiles,urlfile)
+        if (verbose) print(urlfile)
+      }
+    } # End for   
+  }
+  return(urlfiles[-1])
+}
+
+## download the CMIP5-files from the KNMI ClimateExplorer
+urls <- cmip5.urls()
+for (i in 1:length(urls)) if (!file.exists(substr(urls[i],43,70))) download.file(urls[i],substr(urls[i],43,70))
+
+```
+
 ```{r predictor}

### 2017-04-06 paper58/paper58.Rmd
Cleansed version

@@
 ---
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
   cex <- 1.2*c(r2) + 0.5
-  map(mu, xlim=xlim,ylim=ylim,bg='grey90',col='grey90',cex=0.5,new=FALSE)
+  map(mu, xlim=xlim,ylim=ylim,bg='grey90',col='grey90',cex=0.5,gridlines=FALSE,new=FALSE)
   points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
@@
 par(xpd=TRUE)
-text(20,73,'Wet-day mean: 2100')
+text(20,72,'Wet-day mean: 2100')
 legend(20,32,c(expression(bar(x)),expression(q[95])),
@@
 
-**Figure SM1**
+**Figure SM2**
+
+Figure SM2 shows a test of the assumption that that the wet-day daily precipitation is approximately exponentially distributed, by comparing the actual percentiles with quantiles estimated for different samples
+with different annual mean wet-day precipitation using the formula for exponentially distributed data: $q_p = -ln(1-p)\mu$.
+
+```{r}
+## Figure SM1.
+## test: See if the quantiles are consistent when the mean mu varies.
+qtest <- aggregate(y1,year,FUN='qqexp')
+qx <- c(coredata(qtest[,1:101]))
+qy <- c(coredata(qtest[,102:202]))
+
+par(bty='n')
+plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
+     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
+     main='Test: exponential distribution & changing mean',
+     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
+lines(c(0,40),c(0,40),col='grey')
+figlab('Figure SM1')
+```
+
+**Figure SM2**
+
+Figure SM3 presents a map of the NCEP reanalysis temperature for the North Atlantic region that is used as predictor for the downscaling (see e.g., Figure 1 and 3).
+
+```{r}
+## Show the predictor area:
+X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
+map(X,projection='sphere',colbar=list(pal="budrd",breaks=seq(8,28,by=0.5)))
+figlab('Figure SM2',xpos=0.8,ypos=0.999)
+```
+
+**Figure SM3**
 
@@
        ncol=2)
-figlab('Figure SM1')
-```
-
-**Figure SM2**
-
-Figure SM2 shows a test of the assumption that that the wet-day daily precipitation is approximately exponentially distributed, by comparing the actual percentiles with quantiles estimated for different samples
-with different annual mean wet-day precipitation using the formula for exponentially distributed data: $q_p = -ln(1-p)\mu$.
-
-```{r}
-## Figure SM2.
-## test: See if the quantiles are consistent when the mean mu varies.
-qtest <- aggregate(y1,year,FUN='qqexp')
-qx <- c(coredata(qtest[,1:101]))
-qy <- c(coredata(qtest[,102:202]))
-
-par(bty='n')
-plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
-     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
-     main='Test: exponential distribution & changing mean',
-     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
-lines(c(0,40),c(0,40),col='grey')
-figlab('Figure SM2')
-```
-
-**Figure SM3**
-
-Figure SM3 presents a map of the NCEP reanalysis temperature for the North Atlantic region that is used as predictor for the downscaling (see e.g., Figure 1 and 3).
-
-```{r}
-## Show the predictor area:
-X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
-map(X,projection='sphere',colbar=list(pal="budrd",breaks=seq(8,28,by=0.5)))
-figlab('Figure SM3',xpos=0.8,ypos=0.999)
+figlab('Figure SM3')
 ```
@@
 Figure SM7 shows the historical trends in the wet-day mean preciptation.
-
-```{r}
+Locations with trends that are statistically significant at the 5\%-level are shown with a ring around the symbol.
+The significance test was based on a regression analysis and the p-value associated with the fitted slope of a linear fit.
+
+```{r, fig.height=8}
 ## Map showing trends in mu
 map(MU,FUN='trend',cex=1,colbar=list(pal="budrd",breaks=pretty(c(-0.5,0.5),n=21)))
+pval <- apply(coredata(MU),2,'trend.pval') <= 0.05
+points(lon(subset(MU,is=pval)),lat(subset(MU,is=pval)),cex=1,col=rgb(0,0,0,0.2))
 figlab('Figure SM7',ypos=0.999)
@@
 
-Figure SM9 shows the spatial distribution of the historical trends in wet-day frequency on a map.
-
-```{r}
+Figure SM9 shows the spatial distribution of the historical trends in wet-day frequency on a map, similar to Fig SM7.
+
+```{r, fig.height=8}
 ## Map showing trends in fw
-#dev.new()
 map(FW,FUN='trend',cex=1,colbar=list(pal="budrd",breaks=pretty(c(-0.03,0.03),n=21)))
+pval <- apply(coredata(FW),2,'trend.pval') <= 0.05
+points(lon(subset(FW,is=pval)),lat(subset(FW,is=pval)),cex=1.2,col=rgb(0,0,0,0.75))
+#points(lon(FW),lat(FW),cex=1.2,col=rgb(0,0,0,0.3))
 figlab('Figure SM9',ypos=0.999)

### 2017-04-06 paper58/paper58.R

@@ -232,10 +232,10 @@ mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
   pch[!mo] <- 4
   cex <- 1.2*c(r2) + 0.5
   
-  map(mu, xlim=xlim,ylim=ylim,bg='grey90',col='grey90',cex=0.5,new=FALSE)
+  map(mu, xlim=xlim,ylim=ylim,bg='grey90',col='grey90',cex=0.5,gridlines=FALSE,new=FALSE)
   points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
   par(xpd=TRUE)
-  text(10,73,paste('Annual cycle in PC',ipca,
+  text(10,72,paste('Annual cycle in PC',ipca,
                    'with variance of',round(100*pca$d[ipca]^2/sum(pca$d^2)),'%'),pos=4)
   
   colbar(pretty(pca$v[,ipca],n=100),colpc,fig = c(0.08, 0.12, 0.05, 0.2))
@@ -499,7 +499,7 @@ map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',col='grey70',gridlines=FALSE,
 points(lon(mux),lat(mux),pch=19,col=colx,cex=Cex)
 points(lon(mux),lat(mux),pch=21,col=coly,cex=Cex,lwd=2)
 par(xpd=TRUE)
-text(20,73,'Wet-day mean: 2100')
+text(20,72,'Wet-day mean: 2100')
 legend(20,32,c(expression(bar(x)),expression(q[95])),
        pch=c(21,19),bty='n',col='grey',text.col='grey',horiz=TRUE)
 
@@ -564,8 +564,8 @@ legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
                expression(bar(n)[c*w*d]),expression(bar(n)[c*d*d])),bty='n',
        col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1),
        ncol=2)
-figlab('Figure SM1')
-dev.copy2pdf(file='figureSM1.pdf')
+figlab('Figure SM3')
+dev.copy2pdf(file='figureSM3.pdf')
 
 ## Figure SM2.
 ## test: See if the quantiles are consistent when the mean mu varies.
@@ -580,16 +580,16 @@ plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
[…]

### 2017-04-14 TLT/tropo-temp.Rmd
TLT

---
title: "tropospheric-temperature"
author: "Rasmus E Benestad"
date: "April 14, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Evaluation of tropospheric temperatures

This script compares tropospheric temperatures (TLT) simulated from gloal climate models (CMIP5) and data from satellites (MSU).

### R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

```{r}
library(esd)
```

## CMIP5 simulations

### functions extracting the TLT

## Vertical weigthing functions

Retrieve the weighting loads for the different heights corresponding to the TLT temperature. More on the vertical weighting function from <http://www.remss.com/measurements/upper-air-temperature> and <http://tropic.ssec.wisc.edu/real-time/amsu/explanation.html>.

```{r}
if (!file.exists('tlt.W.rda')) {
  colnames <- c('level','h(m)','T(K)','P(pa)','PV(pa)','Weight')
  ## Weighting over land:
  ## Surface Weight                                  0.15104
  tlt.wl <- read.table('http://data.remss.com/msu/weighting_functions/std_atmosphere_wt_function_chan_tlt_land.txt',skip=7,
                 col.names = colnames)
  attr(tlt.wl,'Surface Weight') <- 0.15104
  ## Weighting over ocean
  ## Surface Weight                                  0.11863
  tlt.wo <- read.table('http://data.remss.com/msu/weighting_functions/std_atmosphere_wt_function_chan_tlt_ocean.txt',skip=7,
                 col.names=colnames)
  attr(tlt.wo,'Surface Weight') <- 0.11863
  tlt.W <- list(tlt.wl=tlt.wl,tlt.wo=tlt.wo)
  save(tlt.W,file='tlt.W.rda')
} else load('tlt.W.rda')
```

```{r}
plot(tlt.W$tlt.wo$Weight,tlt.W$tlt.wo$h.m.,type='l',lwd=2,col='grey')
lines(tlt.W$tlt.wl$Weight,tlt.W$tlt.wl$h.m.,col='red',lty=2)
grid()
```

The air temperature needs to be weighted according to surface area (latitude) and height before it is aggregated to a product that is comparable to that of the satellite TLT.

```{r}
## Estimate area weighted mean temperature 
areamean <- function(x,W.lat,d) {
  y <- x*c(W.lat)
  dim(y) <- c(d[1],d[2])
  z <- colSums(y,na.rm=TRUE)/sum(W.lat[,1],na.rm=TRUE)
  return(z)
}

ta2tlt <- function(fname,tlt.W,xlat=c(-90,90),varid='ta') {
  require(ncdf4)
  ncid <- nc_open(fname)
  X <- ncvar_get(ncid,varid)
  lat <- ncvar_get(ncid,'lat')
  plev <- ncvar_get(ncid,'plev')
  tim <- ncvar_get(ncid,'time')
  tunit <- ncatt_get(ncid,'time','units')
  model.id <- ncatt_get(ncid,0,'model_id')
  rcp <- ncatt_get(ncid,0,'experiment')
  ## Different GCMs use different calendars: 365-day, Gregorian, ets.
  calendar <- ncatt_get(ncid,'time','calendar')
  nc_close(ncid)
  ## Extract selected latitude band
  iy <- (lat >= min(xlat)) & (lat <= max(xlat))
  lat <- lat[iy]
  X <- X[iy,,]
  d <- dim(X)

  ## Area weights - latitude
  W.lat <- matrix(rep(cos(pi*lat/180),d[2]),d[1],d[2])
  ## Prepare for fast matrix operations
  dim(X) <- c(d[1]*d[2],d[3])
  ## Z contains the global mean temperature at different vertical levels
  Z <- apply(X,2,'areamean',W.lat,d)

  ## Estimate the TLT-temperature based on the TLT weights
  data(etopo5)
  etopo5 <- subset(etopo5,is=list(lat=xlat))
  ## Weighted average of land and ocean
  nl <- sum(etopo5 >=0)/length(etopo5); no <- sum(etopo5 <0)/length(etopo5)
  W0 <- no*tlt.W$tlt.wo$Weight + nl*tlt.W$tlt.wl$Weight
  W <- approx(tlt.W$tlt.wo$P.pa.,W0,plev,rule=2)$y

  tlt <- apply(Z,2,function(x,W) sum(x*W,na.rm=TRUE)/sum(W,na.rm=TRUE),W)

  ## Use the first time stamp and the time orgigin to set the first date, and then
  ## Assume every month since that (there are monthly mean data after all)
  t1 <- as.Date(tim[1],origin = sub('days since ','',tunit$value))
  TLT <- zoo(tlt, order.by=seq(t1,by='month',length.out=length(tim)))
  attr(TLT,'model_id') <- model.id
  attr(TLT,'rcp') <- rcp
  if (sum(is.finite(tlt))==0) browser()
  return(TLT)
}
```

Download CMIP5 data from the KNMI Climate Explorer and estimate the TLT temperature. The data is stored as zonal mean temperature with the dimensions latitude and pressure-level.

```{r pressure, echo=FALSE}
N <- 102  ## Number of model runs with taz available at ClimateExplorer
if (!file.exists('tlt.cmip5.rda')) {
  model.id <- rep('',N); rcp <- model.id
  for (i in 1:N) {
   if (i <= 10)  fname <- paste('taz_Amon_one_rcp45to85_00',i-1,'.nc',sep='') else
   if (i <= 100) fname <- paste('taz_Amon_one_rcp45to85_0',i-1,'.nc',sep='') else
                 fname <- paste('taz_Amon_one_rcp45to85_',i-1,'.nc',sep='')
   print(fname)
   if (!file.exists(fname)) 
     download.file(paste('https://climexp.knmi.nl/CMIP5/monthly/taz/',fname,sep=''),destfile = fname)
   tlt.glob <- ta2tlt(fname,tlt.W,xlat=c(-70,82.5),varid='ta')  ## These match RSS
   tlt.trop <- ta2tlt(fname,tlt.W,xlat=c(-25,25),varid='ta')    ## These match RSS
   model.id[i] <- attr(tlt.glob,'model_id')
   rcp[i] <- attr(tlt.glob,'rcp')
   if (i==1) {
     TLT.glob <- tlt.glob; TLT.trop <- tlt.trop
   } else {
     TLT.glob <- merge(TLT.glob,tlt.glob)
     TLT.trop <- merge(TLT.glob,tlt.trop)
   }
  }

  tlt.cmip5 <- list(TLT.glob=TLT.glob,TLT.trop=TLT.trop)
  attr(tlt.cmip5,'model_id') <- model.id
  attr(tlt.cmip5,'rcp') <- rcp
  save(tlt.cmip5,file='tlt.cmip5.rda')
} else load('tlt.cmip5.rda')
plot(annual(anomaly(tlt.cmip5$TLT.glob)),plot.type='single',col=rgb(0,0,0,0.2))
```

## Data from satellites

### Get the lower tropospheric data from the RSS:

```{r}
if (!file.exists('RSS.glob.rda')) {
  rss.glob <- read.table('http://data.remss.com/msu/graphics/TLT/time_series/RSS_TS_channel_TLT_Global_Land_And_Sea_v03_3.txt',skip=5)
  rss.trop <- read.table('http://data.remss.com/msu/graphics/TLT/time_series/RSS_TS_channel_TLT_Tropics_Land_And_Sea_v03_3.txt',skip=5)
  rss.glob$V3[rss.glob$V3 <= -99] <- NA
  RSS.glob <- zoo(rss.glob$V3,order.by=as.Date(paste(rss.glob$V1,rss.glob$V2,'01',sep='-')))
  rss.trop$V3[rss.trop$V3 <= -99] <- NA
  RSS.trop <- zoo(rss.trop$V3,order.by=as.Date(paste(rss.trop$V1,rss.trop$V2,'01',sep='-')))
  save(RSS.glob,file='RSS.glob.rda')
  save(RSS.trop,file='RSS.trop.rda')
} else {load('RSS.glob.rda'); load('RSS.trop.rda')}
```

### University of Alabama Huntsville

```{r}
if (!file.exists('UAH.glob.rda')) {
  UAH <- readLines('http://www.nsstc.uah.edu/data/msu/t2lt/tltglhmam_5.6.txt')
  writeLines(UAH[-length(UAH)],con='UAHTLT.dat')
  uah <- read.table('UAHTLT.dat',skip=4,header=TRUE)
  UAH.glob <- zoo(uah$GLOBAL,order.by=as.Date(paste(uah$YEAR,uah$MON,'01',sep='-')))
  UAH.trop <- zoo(uah$TRPC,order.by=as.Date(paste(uah$YEAR,uah$MON,'01',sep='-')))
  save(UAH.glob,file='UAH.glob.rda')
  save(UAH.trop,file='UAH.trop.rda')
} else {load('UAH.glob.rda'); load('UAH.trop.rda')}
```

## Comapre the data

### The global mean TLT temperature

```{r}
X <- merge(annual(anomaly(RSS.glob,ref=1979:1990)),
           annual(anomaly(UAH.glob,ref=1979:1990)),
           annual(anomaly(window(tlt.cmip5$TLT.glob,start=start(RSS.glob),end=end(RSS.glob)),ref=1979:1990)))
par(bty='n')
plot(X[,-c(1,2)],plot.type='single',col=rep(rgb(0,0,0,0.15),N),main='Global lower Tropospheric Temperature (TLT)',ylab='Temperature')
grid()
lines(X[,1],lwd=3,col='red')
lines(X[,2],lwd=3,col='blue')
```

### The global mean TLT temperature

```{r}
X <- merge(annual(anomaly(RSS.trop,ref=1979:1990)),
           annual(anomaly(UAH.trop,ref=1979:1990)),
           annual(anomaly(window(tlt.cmip5$TLT.trop,start=start(RSS.trop),end=end(RSS.trop)),ref=1979:1990)))
par(bty='n')
plot(X[,-c(1,2)],plot.type='single',col=rep(rgb(0,0,0,0.15),N),main='Tropical lower Tropospheric Temperature (TLT)',ylab='Temperature')
grid()
lines(X[,1],lwd=3,col='red')
lines(X[,2],lwd=3,col='blue')
```

### 2017-04-14 TLT/tropo-temp.Rmd
TLT - update

@@
 This script compares tropospheric temperatures (TLT) simulated from gloal climate models (CMIP5) and data from satellites (MSU).
+When comparing model results to data derived from satellites, there are a number of issues to keep in mind.
+The satellite data provide a measure of the air teperature over a range of altitudes, and are also affected by the surface temperatures.
+Different vertical weights are used for data over ocean and land, respectively.
+To process the model results to mimic what really was observed by the sattelites, we need to apply these weights correctly.
+
+The model results used here were only the zonal means, which makes it impossible to derive the exact results.
+However, it is possible to estiamte the fraction with land and fraction with ocean, and use a combined vertical weighting function with approporaite weight on land and ocean.
+
+The global climate models (GCMs) provide results in the form of grid boxes, with latitude along one axis and pressure-level (vertical) along the other.
+The area of a latitude band near tthe equator is larger than a latitude band of the same width at higher latitude, due to earth's curvature (spherical shape).
+Hence, the global and tropical mean TLT must take into account the latitude-variable area and weight the model results accordingly.
+
+Another issue is that different GCMs operate with different calendars, e.g.
+the Gregorian (which is used in the real world), 365-day calendar (no leap-year), and 360-day calendar.
+If this is not taken into account, the months in the future will end up in the wrong season.
 
@@
 lines(X[,2],lwd=3,col='blue')
+legend(1979,1.5,c('CMIP5','RSS','UAH'),col=c(rgb(0,0,0,0.15),'red','blue'),bty='n',lwd=c(1,3,3))
 ```
@@
 lines(X[,2],lwd=3,col='blue')
+legend(1979,1.5,c('CMIP5','RSS','UAH'),col=c(rgb(0,0,0,0.15),'red','blue'),bty='n',lwd=c(1,3,3))
 ```
+
+### Global trend analysis
+
+```{r}
+z <- apply(coredata(annual(anomaly(window(tlt.cmip5$TLT.glob,start=start(RSS.glob),end=end(RSS.glob)),ref=1979:1990))),2,trend.coef)
+h <- hist(z,col='grey',freq=FALSE,main='',ylab='',
+          xlab='Trend (K/decade)')
+x <- seq(-max(abs(z)),max(abs(z)),length=100)
+lines(x,dnorm(x,mean=mean(z),sd=sd(z)),lwd=3,col=rgb(1,0,0,0.4))
+points(trend.coef(annual(anomaly(RSS.glob,ref=1979:1990))),0,pch=19,cex=2,col='red')
+points(trend.coef(annual(anomaly(UAH.glob,ref=1979:1990))),0,pch=19,cex=2,col='blue')
+grid()
+par(new=TRUE,fig=c(0.75,0.98,0.75,0.98),mar=rep(0,4),cex.axis=0.7)
+qqnorm(z,main=''); qqline(z,col='red')
+```
+
+### Tropics trend analysis
+
+```{r}
+z <- apply(coredata(annual(anomaly(window(tlt.cmip5$TLT.trop,start=start(RSS.trop),end=end(RSS.trop)),ref=1979:1990))),2,trend.coef)
+h <- hist(z,col='grey',freq=FALSE,main='',ylab='',
+          xlab='Trend (K/decade)')
+x <- seq(-max(abs(z)),max(abs(z)),length=100)
+lines(x,dnorm(x,mean=mean(z),sd=sd(z)),lwd=3,col=rgb(1,0,0,0.4))
+points(trend.coef(annual(anomaly(RSS.trop,ref=1979:1990))),0,pch=19,cex=2,col='red')
+points(trend.coef(annual(anomaly(UAH.trop,ref=1979:1990))),0,pch=19,cex=2,col='blue')
+grid()
+par(new=TRUE,fig=c(0.75,0.98,0.75,0.98),mar=rep(0,4),cex.axis=0.7)
+qqnorm(z,main=''); qqline(z,col='red')
+```

### 2017-04-14 TLT/tropo-temp.Rmd
TLT - update

@@
 
-### functions extracting the TLT
+## Vertical weigthing functions
 
-## Vertical weigthing functions
+The GCM results are stored in a 4D grid with the dimensions longitude, latitude, vertical levels, and time, whereas the satellite microwave sounding unit (MSU) data measure temperatures over different altitudes.
+Any comparison between these two aspects requires some data processing before we get comparable measures.
+THe MSU data are equivalent to a vertical-weighted average temperature.
 
@@
 
+A small test to see whether the vertical weights look right and to see the difference in the weights over land and ocean.
+
 ```{r}
@@
 
-The air temperature needs to be weighted according to surface area (latitude) and height before it is aggregated to a product that is comparable to that of the satellite TLT.
+The two weights differ most near the surface.
+
+### Functions extracting the TLT
+
+The following R-code describe a set of functions used for processing the GCM results to make them more comparable with the MSU data.
+
+#### Global and tropical means.
+
+The air temperature also needs to be weighted according to surface area (latitude) and height before it is aggregated to a product that is comparable to that of the satellite TLT.
 
@@
 
+## Retrieveing and processing the GCM data
+
 Download CMIP5 data from the KNMI Climate Explorer and estimate the TLT temperature. The data is stored as zonal mean temperature with the dimensions latitude and pressure-level.
@@
      TLT.glob <- merge(TLT.glob,tlt.glob)
-     TLT.trop <- merge(TLT.glob,tlt.trop)
+     TLT.trop <- merge(TLT.trop,tlt.trop)
    }
@@
 
-### University of Alabama Huntsville
+### Get the lower tropospheric data from University of Alabama Huntsville
 
@@
 
-## Comapre the data
+The data are svaed locally so that they can be accessed even with no Internet.
+
+## Compare the data
 
 ### The global mean TLT temperature
+
+The comparison between time evolution of the GCM results and the MSU data is based on anomalies with respect to the 1979--1990 period.
 
@@
 ```
+
+A comparison between the MSU data and the GCM results (CMIP5) suggests that the GCMs somewhat over-estimate the recent warming, but the satellite data are nevertheless within the spread of the model ensemble.
+These results suggests that there is a fair agreement between the two.
 
@@
 
+for the tropical region, the GCM results are still consistent with the satellite data, but the latter is further out on the fridge of the model spread.
+
 ### Global trend analysis
@@
 qqnorm(z,main=''); qqline(z,col='red')
+
+## RSS - how well does it compare with GCM results?
+print(100*pnorm(trend.coef(annual(anomaly(RSS.glob,ref=1979:1990))),mean=mean(z),sd=sd(z)))
+
+## UAH - how well does it compare with GCM results?
+print(100*pnorm(trend.coef(annual(anomaly(UAH.glob,ref=1979:1990))),mean=mean(z),sd=sd(z)))
 ```
+
+A comparison of the global linear trend estiamtes (based on a least-squares fit to a linear function in time) suggests that the satellite data are well within the model results and the fitted normal distribution.
+The sample of model results are roughly normally distributed except for the lower part of the spread (the "lower tail"" of the distribution).
 
@@
 qqnorm(z,main=''); qqline(z,col='red')
+
+## RSS - how well does it compare with GCM results?
+print(100*pnorm(trend.coef(annual(anomaly(RSS.trop,ref=1979:1990))),mean=mean(z),sd=sd(z)))
+
+## UAH - how well does it compare with GCM results?
+print(100*pnorm(trend.coef(annual(anomaly(UAH.trop,ref=1979:1990))),mean=mean(z),sd=sd(z)))
 ```
+
+A comparison of the tropical trends also suggests that the satellite data are within the model results and the fitted normal distribution, but further out on the fringe compared to the global trends.
+The comparison between the trends in the MSU data and CMIP5 suggests a p-value of 0.9\% and 0.3\%, which indicates that the trends in the GCM results and the satellites are statistically different  at the 1\% level for the tropical region.
+This region is also characterised by deep moist convection (the upward branch of the Hadley cell), and one of the hot spots for the hydrological cycle.
+Moreover, this limited region is not expected to have a simple association with the global warming (Benestad, 2016 <http://link.springer.com/article/10.1007%2Fs00704-016-1732-y>).
+The distribution of trend estimates is close to being normally distributefd for the tropics.
+
+## Summary
+
+GCMs are designed to represent the main features of earth's climate, and the most important feature that they are expected to reproduce is the increased global mean temperatures.
+These models are not perfect, and there will be some details - limited regions or variables - which are not as well reproduced.
+Here, the results suggests that the global TLT produced by the GCMs are consistent with those estimated from the datallites (MSU).
+The two are more different for more regional selections, but this is not unexpected.
+The estimate of the temperature trends may be tricky for both models and satellites  the tropical regions dominated by deep moist convection and clouds.

### 2017-04-20 SPECS/specs-decade-mos.Rmd
R-code

@@
 library(esd)
+#library(knitr)
+#purl('git/esd_Rmarkdown/SPECS/specs-decade-mos.Rmd', output='git/esd_Rmarkdown/SPECS/specs-decade-mos.R')
 #library(ncdf)
@@
 ```{r wetdaymean}
-wetdaymean <- function(ifile,cline='cdo yearmean',fname='data/wetdaymean.nc') {
+wetdaymean <- function(ifile,cline='cdo yearmean',fname='wetdaymean.nc') {
   print('calculate mask')
-  system(paste('cdo gtc,1',ifile,'~/data/mask.nc'))
+  system(paste('cdo gtc,1',ifile,'mask.nc'))
   ## her blir ofile en mask med 1 i alle tidsskritt med data over konstanten (satt til 0),
@@
   print('apply mask')
-  system(paste('cdo ifthen data/mask.nc',ifile, fname))
+  system(paste('cdo ifthen mask.nc',ifile, fname))
   ## her appliseres masken på ifile og lagres i ofile. (Øvrige punkter får verdien NA.)
@@
   X <- aggregatedailyfield(fname=fname,cline=cline)
-  Y <- aggregatedailyfield(fname=,'~/data/mask.nc',cline=cline,output='fw_0.25deg_reg.nc')
-  file.remove(fname); file.remove('~/data/mask.nc')
+  Y <- aggregatedailyfield(fname=,'mask.nc',cline=cline,output='fw_0.25deg_reg.nc')
+  file.remove(fname); file.remove('mask.nc')
 }
@@
 ```{r compare}
-comparemannual <- function(tas_DECA,reanalysis='air.mon.mean.nc',mask=TRUE,l= 3,eofs=NULL) {
+comparemannual <- function(tas_DECA,reanalysis='~/Downloads/air.mon.mean.nc',mask=TRUE,l= 3,eofs=NULL) {
 
@@
 ## retrieve did not get the right time stamp:
-  es <- annual(C.C.eq(retrieve('air.mon.mean.nc',lon=c(-60,40),lat=c(20,70))))
+  es <- annual(C.C.eq(retrieve('~/Downloads/air.mon.mean.nc',lon=c(-60,40),lat=c(20,70))))
   es.eof <- EOF(mask(es,land=TRUE))
@@
 
-print('Get the decadal predictors saved locally in "data/gfdl_DECA.rda"')
-load('data/gfdl_DECA.rda')
+print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
+load('~/data/gfdl_DECA.rda')
 
@@
 ## Scaling dependency: time - space
-t2m <- retrieve('~/data/ncep/air.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
+t2m <- retrieve('~/Downloads/air.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
 t2m <- anomaly(t2m)

### 2017-04-20 SPECS/specs-decade-mos.R

@@ -1,0 +1,653 @@
+## ----setup, include=FALSE------------------------------------------------
+knitr::opts_chunk$set(echo = TRUE)
+
+## ----initialise----------------------------------------------------------
+## Script that applies MOS to SPECS decadal forecasts
+
+library(esd)
+#library(ncdf)
+setwd("~/R")
+
+downloadfromunican <- FALSE    # Preparations - get the data
+downscaledreanalysis <- FALSE  # SM: Assessment of large-small scale link
+looexperiments <- FALSE        # Cross-validation MOS on forecasts
+diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts
+
+source("/home/rasmusb/disk1/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
+
+## This variable is needed because CDO did not fix the times properly
+## for the predictands (EOBS)...
+time = c(364, 729, 1095, 1460, 1825, 2190, 2556, 2921, 3286, 3651, 4017, 4382, 
+    4747, 5112, 5478, 5843, 6208, 6573, 6939, 7304, 7669, 8034, 8400, 8765, 
+    9130, 9495, 9861, 10226, 10591, 10956, 11322, 11687, 12052, 12417, 12783, 
+    13148, 13513, 13878, 14244, 14609, 14974, 15339, 15705, 16070, 16435, 
+    16800, 17166, 17531, 17896, 18261, 18627, 18992, 19357, 19722, 20088, 
+    20453, 20818, 21183, 21549, 21914, 22279, 22644, 23010, 23375, 23740, 
+    23921) ;
+
+
+
+## ----functions-----------------------------------------------------------
+#--- functions ---
+
+fcst2field <- function(x,it0=1,im=1,verbose=FALSE) {
+
+  if (verbose) print(paste('fcst2field',it0,im))
+  lon <- x$xyCoords$x; nx <- length(lon)
+  lat <- x$xyCoords$y; ny <- length(lat)
+  index <- x$Dates$start; nt <- length(index)
+
+## The data is scrambled with unconvntional x-y-order - generate unscramble index:
+  usi <- t(matrix(1:(nx*ny),ny,nx))
+  param <- x$Variable$varName
+  unit <- attr(x$Variable,'units')
+  if (verbose) print(paste(param,unit))
+
+## one field
+  df <- matrix(x$Data[it0,im,,,],nt,nx*ny)
+  y <- zoo(df[,c(usi)],order.by=as.Date(index))
[…]

### 2017-04-28 ClimaTrans/climatrans.Rmd
ClimaTrans

@@
 
+To extract just the R-code:
+
+```{r, eval=FALSE}
+library(knitr)
+purl('~/git/esd_Rmarkdown/ClimaTrans/climatrans.Rmd', output='~/git/esd_Rmarkdown/ClimaTrans/climatrans.R')
+```
+
 Load the esd and ncdf packages.
@@
 library(esd)
-library(ncdf)
+library(ncdf4)
 setwd('~/R')
@@
 ## time origin: minutes since 1901-01-01 00:00
-library(ncdf)
-ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/delhi.nc')
-delhi <- get.var.ncdf(ncid,varid='p')
-time <- get.var.ncdf(ncid,varid='time')
-close.ncdf(ncid)
+library(ncdf4)
+ncid <- nc_open('~/Dropbox/Public/ClimaTrans/delhi.nc')
+delhi <- ncvar_get(ncid,varid='p')
+time <- ncvar_get(ncid,varid='time')
+nc_close(ncid)
 delhi <- zoo(x=delhi,order.by=as.Date(time/(24*60),origin='1901-01-01'))
@@
 
-ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bombay.nc')
-bombay <- get.var.ncdf(ncid,varid='p')
-time <- get.var.ncdf(ncid,varid='time')
-close.ncdf(ncid)
+ncid <- nc_open('~/Dropbox/Public/ClimaTrans/bombay.nc')
+bombay <- ncvar_get(ncid,varid='p')
+time <- ncvar_get(ncid,varid='time')
+nc_close(ncid)
 bombay <- zoo(x=bombay,order.by=as.Date(time/(24*60),origin='1901-01-01'))
@@
 
-ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bangalore.nc')
-bangalore <- get.var.ncdf(ncid,varid='p')
-time <- get.var.ncdf(ncid,varid='time')
-close.ncdf(ncid)
+ncid <- nc_open('~/Dropbox/Public/ClimaTrans/bangalore.nc')
+bangalore <- ncvar_get(ncid,varid='p')
+time <- ncvar_get(ncid,varid='time')
+nc_close(ncid)
 
@@
 ## time origin: minutes since 1969-01-01 00:00
-ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/delhi_tmax.nc')
-delhi <- get.var.ncdf(ncid,varid='temp')
-time <- get.var.ncdf(ncid,varid='time')
-close.ncdf(ncid)
+ncid <- nc_open('~/Dropbox/Public/ClimaTrans/delhi_tmax.nc')
+delhi <- ncvar_get(ncid,varid='temp')
+time <- ncvar_get(ncid,varid='time')
+nc_close(ncid)
 delhi <- zoo(x=delhi,order.by=as.Date(time/(24*60),origin='1969-01-01'))
@@
 
-ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bombay_tmax.nc')
-bombay <- get.var.ncdf(ncid,varid='temp')
-time <- get.var.ncdf(ncid,varid='time')
-close.ncdf(ncid)
+ncid <- nc_open('~/Dropbox/Public/ClimaTrans/bombay_tmax.nc')
+bombay <- ncvar_get(ncid,varid='temp')
+time <- ncvar_get(ncid,varid='time')
+nc_close(ncid)
 bombay <- zoo(x=bombay,order.by=as.Date(time/(24*60),origin='1969-01-01'))
@@
 
-ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bangalore_tmax.nc')
-bangalore <- get.var.ncdf(ncid,varid='temp')
-time <- get.var.ncdf(ncid,varid='time')
-close.ncdf(ncid)
+ncid <- nc_open('~/Dropbox/Public/ClimaTrans/bangalore_tmax.nc')
+bangalore <- ncvar_get(ncid,varid='temp')
+time <- ncvar_get(ncid,varid='time')
+nc_close(ncid)
 bangalore <- zoo(x=bangalore,order.by=as.Date(time/(24*60),origin='1969-01-01'))
@@
 grid()
-print(trend(fw,result='pval')) 
+print(as.numeric(trend(fw,result='coef'))) 
+print(as.numeric(trend(fw,result='pval'))) 
 ```
@@
 grid()
-print(trend(mu,result='pval'))
+print(as.numeric(trend(mu,result='coef'))) 
+print(as.numeric(trend(mu,result='pval')))
 ```
@@
 grid()
-print(trend(Pr,result='pval'))
+print(as.numeric(trend(Pr,result='coef')))
+print(as.numeric(trend(Pr,result='pval')))
 ```
@@
 grid()
-print(trend(pt,result='pval'))
+print(as.numeric(trend(pt,result='pval')))
+print(as.numeric(trend(pt,result='pval')))
 ```
@@
 z <- subset(fw,is=2)
+attr(z,'unit') <- 'correlation'
 corfield(z,slp,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)),new=FALSE)
@@
 
-```{r}
+```{r, fig.height=8,fig.width=8}
 ds.fw <- DS(pca.fw,eof.slp,eofs=1:20)
@@
 
-```{r pca-of-residual-fw}
+```{r pca-of-residual-fw, fig.height=8,fig.width=8}
 ## Examine the PCA of residuals for the cold monthsto see if there are any structures left
@@
 
-```{r}
+```{r, fig.height=8,fig.width=8}
   ds.tx <- DS(pca.tx,eof.t2m)
@@
 
-```{r}
+```{r, fig.height=8,fig.width=8}
 ## Check the residuals
@@
 diag.slp <- biasdiag(slp,it=month.abb[season],verbose=FALSE)
+mean.diff <- unlist(lapply(diag.slp,function(x) x$mean.diff[,1]))
+sd.ratio <- 1 - unlist(lapply(diag.slp,function(x) x$sd.ratio[,1]))
+ar.ratio <- 1 - unlist(lapply(diag.slp,function(x) x$autocorr.ratio[,1]))
+summary(sd.ratio)
+summary(ar.ratio)
 ```

### 2017-04-28 EU-Circle/eu-circle_case-study-1.Rmd
ClimaTrans

@@
 This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+
+```{r, eval=FALSE}
+library(knitr)
+purl('~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.Rmd', output='~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.R')
+```
 
@@
 ```{r,warning=FALSE}
-X <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
+X <- retrieve('~/data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
 map(X,new=FALSE)
@@
 
+Calculate EOFs which can serve as predictors for the downscaling.
+
+```{r, fig.height=8}
+dryseason <- function(x,na.rm=TRUE) if (sum(is.finite(x))>1) mean(subset(spell(x,threshold=1),is=2),na.rm=na.rm) else NA
+if (!file.exists('case1-dryseason.rda')) {
+  z <- aggregate(subset(X,it='mjjas'),year,FUN='dryseason')
+  save(z,file='case1-dryseason.rda')
+} else load('case1-dryseason.rda')
+eof.dry <- EOF(z)
+plot(eof.dry)
+```
+
+Predictors for calibration: check for any dependency
+
+```{r}
+T2M <- aggregate(subset(retrieve('~/Downloads/air.mon.mean.nc',lon=c(-12,35),lat=c(27,50)),it='mjjas'),year,'mean')
+SLP <- aggregate(subset(retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-12,35),lat=c(27,50)),it='mjjas'),year,'mean')
+eof.t2m <- EOF(T2M)
+eof.slp <- EOF(SLP)
+```
+
+Any connection between temperature and dry spells?
+
+```{r, fig.height=8}
+ds.t2m <- DS(eof.dry,eof.t2m)
+plot(ds.t2m)
+```
+
+Any connection between sea-level pressure anomalies and dry spells?
+
+```{r, fig.height=8}
+ds.slp <- DS(eof.dry,eof.slp)
+plot(ds.slp)
+```
+
 ## Temperature

### 2017-04-28 SPECS/specs-decade-mos.Rmd
ClimaTrans

@@
     print(paste('Only keep the large-scale structures; EOFs:',paste(eofs,collapse=' ,')))
-    Xm.eofs <- subset(EOF(Xm),pattern=eofs)
+    Xm.eofs <- subset(EOF(Xm),ip=eofs)
     Xm <- eof2field(Xm.eofs)

### 2017-04-28 ClimaTrans/climatrans.R

@@ -1,0 +1,888 @@
+## ----eval=FALSE----------------------------------------------------------
+## library(rmarkdown); render('climatrans.Rmd',pdf_document())
+
+## ---- eval=FALSE---------------------------------------------------------
+## library(knitr)
+## purl('~/git/esd_Rmarkdown/ClimaTrans/climatrans.Rmd', output='~/git/esd_Rmarkdown/ClimaTrans/climatrans.R')
+
+## ----echo=TRUE-----------------------------------------------------------
+library(esd)
+library(ncdf4)
+setwd('~/R')
+#rm(list=ls())
+## Information about the system and session
+
+## ------------------------------------------------------------------------
+tas2es <- function(x,mask=TRUE,land=TRUE,season=5:10,FUN='min') {
+  if (mask) x <- mask(x,land=land)
+  es <- subset(C.C.eq(x),it=month.abb[season])
+  es <- aggregate(es,year,FUN=FUN)
+  index(es) <- year(es)
+  invisible(es)
+}
+
+## ------------------------------------------------------------------------
+PrexpPr <- function(mu,fw,x0=10) {
+  Pr <- zoo(coredata(fw)*exp(-x0/coredata(mu)),order.by=year(fw))
+
+  Pr.mu <- zoo(mean(coredata(fw))*exp(-x0/coredata(mu)),order.by=year(fw))
+  Pr.fw <- zoo(coredata(fw)*exp(-x0/mean(coredata(mu))),order.by=year(fw))
+  attr(Pr,'prob.f(mu)') <- Pr.mu
+  attr(Pr,'prob.f(fw)') <- Pr.fw
+  Pr <- attrcp(mu,Pr)
+  attr(Pr,'variable') <- 'Pr'
+  attr(Pr,'unit') <- 'probability'
+  attr(Pr,'longname') <- paste('Probability of exceeding',x0)
+  class(Pr) <- class(mu)
+  invisible(Pr)
+}
+
+## ------------------------------------------------------------------------
+RMSE <- function(x,y) sqrt(sum((x-y)^2))/length(x)
+
+precipcmip <- function(predictor='data/ERAINT/era-int-precip-mon.nc',
+                    path="CMIP5.monthly/",rcp='rcp45',pattern="pr_Amon_ens_",
[…]

### 2017-04-28 EU-Circle/eu-circle_case-study-1.R

@@ -1,0 +1,122 @@
+## ----setup, include=FALSE------------------------------------------------
+knitr::opts_chunk$set(echo = TRUE)
+
+## ---- eval=FALSE---------------------------------------------------------
+## library(knitr)
+## purl('~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.Rmd', output='~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.R')
+
+## ----message=FALSE-------------------------------------------------------
+library(esd)
+
+## ----warning=FALSE-------------------------------------------------------
+if (!file.exists('eu-circle_case-study-1_precio.rda')) {
+  ss <- select.station(param='precip',lon=c(5,10),lat=c(42,45),nmin=50,src=c('ecad','ghcnd'))
+  x <- station(ss)
+  x <- subset(x,is=list(alt=-400))
+  save(file='eu-circle_case-study-1_precio.rda',x)
+} else load('eu-circle_case-study-1_precio.rda')
+## Only keep the ECAD data
+x <- subset(x,is=is.element(src(x),'ECAD'))
+map(x,FUN='q95',new=TRUE)
+diagnose(x)
+
+## ----warning=FALSE-------------------------------------------------------
+plot(x,new=FALSE)
+grid()
+
+## ----warning=FALSE-------------------------------------------------------
+mu <- annual(x,'wetmean',nmin=250)
+wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
+q95 <- annual(x,'wq95')
+plot(mu,new=FALSE)
+for (i in 1:dim(mu)[2]) lines(trend(subset(mu,is=i)),lty=2)
+grid()
+
+
+## ------------------------------------------------------------------------
+x1 <- zoo(subset(annual(x,'count',threshold=1),is=1))
+y1 <- zoo(subset(mu,is=1))
+plot(x1,y1,xlab=expression(n[X>1*mm]),ylab=expression(mu),xlim=c(30,150),ylim=c(5,30),
+     main='Test: wet-day mean precipitation dependency on sample size')
+for (i in 2:dim(mu)[2]) {
[…]

### 2017-04-28 SPECS/nc2rda.R

@@ -1,0 +1,46 @@
+## Generate rda-files for the decadal forecasts - organise the data in the same way as the GFDL data
+##> summary(psl_DECA)
+##                    Length   Class  Mode     
+##Variable                   2 -none- list     
+##Data                79704000 -none- numeric  
+##xyCoords                   2 -none- list     
+##Dates                      2 -none- list     
+##InitializationDates        9 -none- list     
+##Members                   10 -none- character
+##
+##> dim(psl_DECA$Data)
+## [1]   9  10 480  41  45
+##      ini mem tim lat lon
+#> range(tas_DECA$xyCoords$x)
+#[1] -88.75  21.25
+#> range(tas_DECA$xyCoords$y)
+#[1] -1.011236 79.887640
+#> 
+  
+## Each netCDF file: float tas(time, ensemble, latitude, longitude) - global
+#dimensions:
+#  latitude = 73 ;
+#  longitude = 96 ;
+#  time = UNLIMITED ; // (120 currently)
+#  ensemble = 10 ;
+  
+library(ncdf4)
+
+getvar <- function(fname,varid,lons=c(-88.75, 21.25),lats=c(-1.0, 80.0)) {
+  ## Extract the region
+  ncid <- nc_open(fname)
+  
+  lon <- ncvar_get(ncid,'longitude')
+  lat <- ncvar_get(ncid,'latitude')
+  ens <- ncvar_get(ncid,'ensemble')
+  tim <- ncvar_get(ncid,'time')
+  ix <- (1:length(lon))[(lon >= lons[1]) & (lon <= lons[2])]
+  iy <- (1:length(lat))[(lat >= lats[1]) & (lat <= lats[2])]
+  lon <- lon[ix]; lat <- lat[iy]
+  x <- ncvar_get(ncid,varid,start=c(1,1,min(iy),min(ix)),count(length(tim),ength(ens),length(iy),length(ix)))
+  
+}
+
+varid <- 'tas'
+files <- list.files(path='~/Downloads/SPECS-decadal-fcst_hadcm3_i2p1',pattern=varid,full.names=TRUE)
+#files <- list.files(path='~/Downloads/SPECS-decadal-fcst_MPI',pattern=varid,full.names=TRUE)

### 2017-04-28 SPECS/prep-dec-fcst.R

@@ -1,0 +1,18 @@
+## Organise decadal forecasts for SPECS
+## Rasmus.Benestad@met.no
+## Did not manage to install the R-package, so I source the R-code instead
+rscripts <- list.files(path='~/Downloads/loadR/',pattern='.R',full.names = TRUE)
+rscripts <- rscripts[-grep('.Rd',rscripts,fixed=TRUE)]
+rscripts <- rscripts[-grep('.dic',rscripts,fixed=TRUE)]
+rscripts <- rscripts[-grep('DESCRIPTION',rscripts,fixed=TRUE)]
+for (i in 1:length(rscripts)) source(rscripts[i])
+
+tasfiles <- list.files(path="~/Downloads/SPECS-decadal-fcst_hadcm3_i2p1/",pattern="tas",full.names = TRUE)
+
+tasDECA <- loadDecadalForecast(dataset = tasfiles[1], # or  "mydata.nc"
+                               latLim = c(0,80),
+                               lonLim = c(-90,20),
+                               var = "tas", 
+                               dictionary = F,
+                               years = 1971:2010,
+                               season = 1:12)

### 2017-05-05 EU-Circle/eu-circle_case-study-1.Rmd
EU-Circle

@@
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+This is an R Markdown document.
+Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
+For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+To strip off all the text and make a pure R-script, then use `purl`as shown below:
 
 ```{r, eval=FALSE}
+## Extract just the R-code
 library(knitr)
@@
 
-***Gridded precipitation***
+### Gridded precipitation ###
 
@@
 ```{r,warning=FALSE}
-X <- retrieve('~/data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
+X <- retrieve('~/data/data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
 map(X,new=FALSE)
@@
 
-Predictors for calibration: check for any dependency
+Only the 4 leading EOFs for the seasonal mean dry spell length are relevant.
+The higher orders are noise.
+This is according to expectations, i.e.
+that droughts and heat waves are large-scale phenomena.
+
+```{r}
+eof.dry <- subset(eof.dry,ip=1:4)
+```
+
+### Predictors for calibration: check for any dependency
 
@@
 
-Any connection between temperature and dry spells?
+Examine the connection between temperature and dry spells:
 
-```{r, fig.height=8}
+```{r, fig.height=8, warning=FALSE}
 ds.t2m <- DS(eof.dry,eof.t2m)
@@
 
-Any connection between sea-level pressure anomalies and dry spells?
+Cross-validation scores of 0.33 for the leading EOF of the mean duration of dry spells and a temperature anomaly over Iberia and souther France suggest a real link.
 
-```{r, fig.height=8}
+There is also a connection between sea-level pressure anomalies and dry spells.
+
+```{r, fig.height=8, warning=FALSE}
 ds.slp <- DS(eof.dry,eof.slp)
 plot(ds.slp)
+```
+
+Cross-validation score of 0.43 and high SLP anomalies over France. The summertime temperature and SLP are expected to be related:
+
+```{r, fig.height=8, warning=FALSE}
+cca.slp.t2m <- CCA(eof.slp,eof.t2m)
+plot(cca.slp.t2m)
+```
+
+The CCA suggests that there is covariance between SLP and temperature, and we need to avoid double counting of these effects.
+
+### Residual:
+
+```{r, fig.width=8, fig.height=8}
+## Estimate the residuals
+eof.res <- subset(eof.dry,it=ds.slp)
+coredata(eof.res) <- coredata(eof.res) - coredata(ds.slp)
+```
+
+Check to see if the residual contains covariance with the temperature
+
+```{r, fig.height=8, warning=FALSE}
+ds.res <- DS(eof.res,eof.t2m)
+plot(ds.res)
+```
+
+### Dry conditions and wild fire statistics
+
+Use the EOFs describing the dry conditions as predictors and wild fire statistics as predictand to assess the connection
+
+```{r}
+# z <- DS(wfire,eof.dry)
+```
+
+### Projections for the future
+
+Need to read the predictors again to get all months
+
+```{r}
+T2M <- retrieve('~/Downloads/air.mon.mean.nc',lon=c(-12,35),lat=c(27,50))
+SLP <- retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-12,35),lat=c(27,50))
+## Also - fix the index of eof - set to year.
+index(eof.dry) <- year(eof.dry)
+```
+
+Two seperate analyses based on SLP and temperature respectively, assuming the RCP 4.5 emission scenario. Start with the SLP and the original data
+
+```{r, warning=FALSE}
+if (!file.exists('dse.mld.slp.rda')) {
+  dse.mld.slp <- DSensemble.eof(eof.dry,predictor=SLP,it='mjjas',path="~/data/CMIP5.monthly",pattern = "psl_Amon_ens_")
+  save(dse.mld.slp,file = 'dse.mld.slp.rda')
+} else load('dse.mld.slp.rda')
+plot(dse.mld.slp)
+```
+
+Follow up with the residuals and the large-scale temperture
+
+```{r, warning=FALSE}
+if (!file.exists('dse.mld.t2m.rda')) {
+  dse.mld.t2m <- DSensemble.eof(eof.res,predictor=T2M,it='mjjas',path="~/data/CMIP5.monthly",pattern = "tas_Amon_ens_")
+  save(dse.mld.t2m,file = 'dse.mld.t2m.rda')
+} else load('dse.mld.t2m.rda')
+plot(dse.mld.t2m)
+```
+
+Add the the two components to get an estimate of the total change. The `dse`objects are lists which contain `zoo` time series:
+
+```{r}
+dse.mld <- dse.mld.slp
+gcmnm <- names(dse.mld)
+gcmnm <- gcmnm[-grep('info',gcmnm)]; gcmnm <- gcmnm[-grep('eof',gcmnm)]
+for (nm in gcmnm) {
+  coredata(dse.mld[[nm]]) <- coredata(dse.mld.slp[[nm]]) + coredata(dse.mld.t2m[[nm]])
+}
 ```

### 2017-05-05 EU-Circle/eu-circle_case-study-2.Rmd
EU-Circle

--- 2016-12-07 EU-Circle/eu-circle_case-study-2.Rmd
@@
 
-## Storm and Sea Surge at a Baltic Sea Port , Gdynia Poland
-
-### Scenario 1: Oil and Natural Gas Transport in port
-
-The oil piping transportation system is operating at one of the Baltic Oil Terminals that is designated for the reception from ships, the storage and sending by carriages or cars the oil products.
-It is also designated for receiving from carriages or cars, the storage and loading the tankers with oil products such like petrol and oil.
-On the basis of the piping system operation and safety statistical data coming from its operators its safety will be modelled, identified and predicted.
-Further, under the assumption of the possibility of the stress of weather influence on the operation conditions in the form of maritime storm and/or other hard sea conditions existence, the piping system safety will be examined and the results will be compared with the previous results.
-The piping system safety and operation optimization will be performed and practical suggestions and procedures improving its  safety will be worked out.
-Presentation: EU-CIRCLE Case Study 2 – Scenario 1
-
-### Scenario 2: Chemical Spill due to extreme sea surges
-
-The sea transport of dangerous chemicals is pretty safe at normal environmental conditions.
-However, the transported goods may be swept overboard as a result of bad weather and hard sea conditions.
-The released chemicals may create the threat for the crew and the ship as well as pollute the seawater and the coast.
-The Baltic Sea and nearby ecosystems are vulnerable to pollution and contamination as a result of sea accident during the dangerous goods transportation.
-Today, one major accident at the Baltic Sea happens every year approximately.
-There are more than 50,000 ships entering and leaving the Baltic Sea every year and about 2,000 vessels are at the Baltic Sea at any given moment as illustrated in Figure 1.
-This huge traffic across the Baltic Sea will be observed.
-
-On the basis of the statistical data coming from reports of chemical accidents at sea, the risk of dangerous chemicals accidents at sea and their dangerous consequences will be modelled, identified and predicted.
-Further, under the assumption of the stress of weather influence on the operation conditions in the form of maritime storm and/or other hard sea  conditions existence, the risk of chemical spills at sea will be examined and the results will be compared with the previous results.
-The risk of chemical spills at sea the environment degradation optimization will be performed and practical suggestions and procedures decreasing the risk of the environment degradation will be worked out.
-
-## R Markdown
-
 This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
 
-```{r esd}
+Load the R-library ```esd``` which is used for data analysis.
+
+```{r,message=FALSE}
 library(esd)
@@
 
-## analysis
+## Storm and Sea Surge at a Baltic Sea Port
 
-Wind, stormtracks, wave heights, ans sea level.
+An oil piping transportation system is operating at Gdynia Port, Poland, one of the Baltic Oil Terminals.
+The system is designated to receive, store and send carriages and cars containing oil products and loading oil tankers.
+The purpose of the case study is to identify and predict safety risks associated with maritime transportation of petrol and oil.
+The transport of dangerous chemicals on sea is reasonably safe at normal environmental conditions.
+However, the transported goods may be swept overboard as a result of rough weather and hard sea conditions.
+The Baltic Sea and nearby ecosystems are vulnerable to pollution and contamination as a result of maritime accidents during the transporation of dangerous goods.
+Approximately one major accident happen at the Baltic Sea every year.
+There are more than 50,000 ships entering and leaving the Baltic Sea every year and about 2,000 vessels are at the Baltic Sea at any given moment.
 
-```{r pressure}
-data(imilast.M03)
+Here, we consider the potential stress of weather on the operation conditions, focusing specifically on maritime storms and hard sea conditions.
+Environmental conditions such as the sea level, wave height, wind speed and direction, and the occurence of storms affect the operation of the oil piping transporation system.
+Meteorological observations of these parameters are used to investigate the weather conditions at Gdynia Port and the Baltic Sea region.
+
+**Wind**
+
+Synoptic storms are associated with winds, storm surges and waves.
+
+```{r}
+load('cyclones.baltic.rda')
+map(cyclones.baltic)
 ```
+
+Plot the number of trajectories over time
+
+```{r}
+plot(cyclones.baltic)
+```
+
+***Import E-OBS data***
+
+Observations of the wind are sparse in the Baltic region, but approximate values of wind speed and direction can be calculated from the sea level pressure (slp).
+A gridded slp product from E-OBS dataset is used which is made available by the EU-FP6 project ENSEMBLES (http://ensembles-eu.metoffice.com) and the data providers in the ECA&D project (http://www.ecad.eu).
+
+Import the E-OBS slp data using the retrieve function of the esd package. The data is represented with lower-case $x$.
+
+```{r,warning=FALSE}
+if (!file.exists('eu-circle_case-study-2_slp.rda')) {
+  x <- retrieve('pp_0.25deg_reg.nc',path='~/storeB/data/data.ECAD',
+                lon=c(10,30),lat=c(53,66))
+  save(file='eu-circle_case-study-2_slp.rda',x)
+} else load('eu-circle_case-study-2_slp.rda')
+## Only keep the ECAD data
+map(x,FUN='q95',new=TRUE)
+diagnose(x)
+```
+
+***Geostrophic wind***
+
+Calculate the geostrophic wind based on the slp data.
+
+**Ship accidents in the Baltic Sea**
+
+```{r}
+baltic.events <- c(19931010,20000328,20010328,20040313,20060330,20081017,20081202,20091002,20100128)
+uv <- geostrophicwind(x)
+```

### 2017-05-05 EU-Circle/eu-circle_case-study-5.Rmd
EU-Circle

@@
 To see if other factors such as moisture has an effect, we estimate the residual from this regression analysis and repeat the analysis using saturation water vapour as predictor.
+This way, we avoid correlation/overlap between the two factors.
 
@@
 
+## Projections
+
+Need to read the predictors again to get all months
+
+```{r}
+T2M <- retrieve('~/Downloads/ERAINT-t2m-mon.nc', lon=c(-90,40),lat=c(30,70))
+OLR <- retrieve('~/Downloads/ERAINT-OLR-mon.nc', lon=c(-50,40),lat=c(40,70))
+## Also - fix the index of eof - set to year.
+index(eof.dry) <- year(eof.dry)
+```
+
+Two seperate analyses based on OLR and saturation vapour pressure respectively, assuming the RCP 4.5 emission scenario.
+The first iteration uses the original predictand whereas the latter uses the residuals.
+
+```{r, warning=FALSE}
+if (!file.exists('dse.mu.olr.rda')) {
+  dse.mu.olr <- DSensemble.eof(pca.mu,predictor=OLR,it='djf',path="~/data/CMIP5.monthly",pattern = "rlut_Amon_ens_")
+  save(dse.mu.olr,file = 'dse.mu.olr.rda')
+} else load('dse.mu.olr.rda')
+plot(dse.mu.olr)
+```
+
+```{r, warning=FALSE}
+if (!file.exists('dse.mu.es.rda')) {
+  dse.mu.es <- DSensemble.eof(pca.res,predictor=SLP,it='djf',path="~/data/CMIP5.monthly",pattern = "tas_Amon_ens_")
+  save(dse.mu.es,file = 'dse.mu.es.rda')
+} else load('dse.mu.es.rda')
+plot(dse.mld.slp)
+```
+
+Need to add the two together.
+
 ## Analysis: temperature

### 2017-05-05 SPECS/specs-decade-mos.Rmd
EU-Circle

@@
 load('~/data/gfdl_DECA.rda')
-
+psl_DECA.gfdl <- psl_DECA; rm('psl_DECA'); tas_DECA.gfdl <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
+#> dim(psl_DECA.gfdl$Data)
+#[1]   9  10 480  41  45
+print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
+load('~/data/hadcm3_DECA.rda')
+psl_DECA.hadcm3 <- psl_DECA; rm('psl_DECA'); tas_DECA.hadcm3 <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
+#> dim(psl_DECA.hadcm3$Data)
+#[1]   6  10 480  33  30
+load('~/data/miroc5_DECA.rda')
+psl_DECA.miroc5 <- psl_DECA; rm('psl_DECA'); tas_DECA.miroc5 <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
+#> dim(psl_DECA.miroc5$Data)
+#[1]   6   6 480  59  79
+rm("psl_DECA_annual","tas_DECA_annual"); gc(reset=TRUE)
+## Need to regrid (regrid.matrix) and combine the different forecasts into one object:
 ```

### 2017-05-05 EU-Circle/eu-circle_case-study-1.R

@@ -2,6 +2,7 @@
 knitr::opts_chunk$set(echo = TRUE)
 
 ## ---- eval=FALSE---------------------------------------------------------
+## ## Extract just the R-code
 ## library(knitr)
 ## purl('~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.Rmd', output='~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.R')
 
@@ -104,7 +105,7 @@ for (i in 1:dim(Pr)[2]) lines(trend(subset(Pr,is=i)),lty=2)
 grid()
 
 ## ----warning=FALSE-------------------------------------------------------
-X <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
+X <- retrieve('~/data/data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
 map(X,new=FALSE)
 
 ## ------------------------------------------------------------------------
@@ -116,7 +117,72 @@ trend.dry <- function(x,na.rm=TRUE) trend.coef(subset(spell(x,threshold=1),is=2)
 map(X,FUN='trend.dry',new=FALSE,colbar=list(pal='warm'))
 
 ## ---- fig.height=8-------------------------------------------------------
-dryseason <- function(x,na.rm=TRUE) mean(subset(spell(x,threshold=1),is=2),na.rm=na.rm)
-eof.dry <- EOF(annual(X,FUN='dryseason'))
+dryseason <- function(x,na.rm=TRUE) if (sum(is.finite(x))>1) mean(subset(spell(x,threshold=1),is=2),na.rm=na.rm) else NA
+if (!file.exists('case1-dryseason.rda')) {
+  z <- aggregate(subset(X,it='mjjas'),year,FUN='dryseason')
+  save(z,file='case1-dryseason.rda')
+} else load('case1-dryseason.rda')
+eof.dry <- EOF(z)
 plot(eof.dry)
+
+## ------------------------------------------------------------------------
+eof.dry <- subset(eof.dry,ip=1:4)
+
+## ------------------------------------------------------------------------
+T2M <- aggregate(subset(retrieve('~/Downloads/air.mon.mean.nc',lon=c(-12,35),lat=c(27,50)),it='mjjas'),year,'mean')
[…]

### 2017-05-05 EU-Circle/eu-circle_case-study-1.Rmd
EU-Circle R-markdowns

@@
 
-```{r, warning=FALSE}
+```{r, warning=FALSE, eval=FALSE,echo=FALSE}
 if (!file.exists('dse.mld.t2m.rda')) {
@@
 
-```{r}
+```{r, eval=FALSE,echo=FALSE}
 dse.mld <- dse.mld.slp

### 2017-05-05 EU-Circle/eu-circle_case-study-2.Rmd
EU-Circle R-markdowns

@@
 load('cyclones.baltic.rda')
-map(cyclones.baltic)
+map(cyclones.baltic,new=FALSE)
 ```
@@
 ```{r}
-plot(cyclones.baltic)
+plot(cyclones.baltic,new=FALSE)
 ```
@@
 ## Only keep the ECAD data
-map(x,FUN='q95',new=TRUE)
-diagnose(x)
+map(x,FUN='q95',new=FALSE)
 ```
@@
 
-Calculate the geostrophic wind based on the slp data.
+Calculate the geostrophic wind based on the slp data from EOBS.
 
@@
 
-```{r}
+NOT FINISHED
+
+```{r, eval=FALSE,echo=FALSE}
 baltic.events <- c(19931010,20000328,20010328,20040313,20060330,20081017,20081202,20091002,20100128)
+if (!file.exists('uv.rda')) {
   uv <- geostrophicwind(x)
+  save(uv,file='uv.rda')
+} else load('uv.rda')
 ```

### 2017-05-11 EU-Circle/eu-circle_case-study-1.Rmd
EU-Circle cast 4

@@
 For heat waves, use a similar approach as described in the supporting material of where a GLM was used to estimate the number of events with a basis of the 'Poisson' distribution.
+
+## Wildfire risk
+
+The fire season: May-October (allowing for longer season in the future).
+
+Hazards: Smoke (Area/density) and burned area.
+Use historic data or model results (RCM)?
+Damage functions - has not been used and need to come up with some sensible choices (e.g.
+mean/minimum density of assets?).
+The assets are distributed unevenly (roads and power lines).
+Model risk in a conditioned Monte-Carlo type of fashion?
+
+Data: number of events $n_f$, burned area $A_f$, area of smoke $A_s$, location of event $(x,y)$, the day of the event, number of multiple events $n_i\ i =[2,3,..]$.
+
+### Fire indices
+
+Gridded data which can be downscaled.
+
+### Events
+
+Model number of evens $n_f$ and their area $A_f$ seperately, a little like precipitation ($\mu$ and $f_w$).
+The area has an impact on the risk of affecting the number of assets.
+Base this on historic data and relate to large-scale conditions (predictors).
+Model as a partly stochastic process where the likelihood depends on an external variable.
+Poisson model (`GLM`) of $n_f$ and the nature of the $A_f$ distribution dictates the model strategy for the area.
+
+Also model number of simultaneous events, which strain the fire fighters, as they need to split/divert resources.
+Based on historic data and the assumption of a Poisson process dependent on external condition.

### 2017-05-11 EU-Circle/eu-circle_case-study-2.Rmd
EU-Circle cast 4

@@
 
-```{r,warning=FALSE}
+```{r,warning=FALSE,eval=FALSE}
 if (!file.exists('eu-circle_case-study-2_slp.rda')) {

### 2017-05-11 EU-Circle/eu-circle_case-study-4.Rmd
EU-Circle cast 4

@@
 Bangladesh is highly vulnerable to climate induced hazrds and disasters and its coastal part are mostly threatened for the impacts of climate change in case of cyclone, storm surges, flood, salinity and tsunami disaster.
-Even last twenty years the coastal peoples have suffered due to super cyclone SIDR and AILA devastations.
+Even last twenty years the coastal peoples have suffered due to super cyclone SIDR (2007-11-15) and AILA (YYYY-MM-DD) devastations.
 Last two deacdes a lots of disaster coping interventions have been provided to the community to achieve resiliency in this areas.
@@
 
+## HURDAT2
+
+```{r}
+url <- 'http://www.aoml.noaa.gov/hrd/hurdat/hurdat2-1851-2016-apr2017.txt'
+
+```
+
 ## Tidal measurements

### 2017-05-11 EU-Circle/bangladesh-imilast.R

@@ -1,0 +1,12 @@
+## Rasmus Benestad
+## Read IMILAST storms - select Bangladesh
+## The IMILAST project excluded tropical cyclones.
+
+library(esd)
+# 03 000001 001 1989010100 1989    1    1   0   -170   61.5    961.24
+imilast.storms <- '~/disk1/IMILAST/imilast.cyclones/ERAinterim_1.5_NH_M03_19890101_20090331_ST.txt'
+Z <- events2trajectory(read.imilast(imilast.storms))
+##Z <- subset(Z,is=list(lon=c(80,120),lat=c(10,30)))
+map(Z)
+
+## Use the hurdat2 data.

### 2017-05-19 EU-Circle/eu-circle_case-study-1.Rmd
Progress on EU-circle case-study 1

@@
 ```{r,warning=FALSE}
+
+cols <- c(rgb(1,0,0,0.2),rgb(0,0,1,0.2),rgb(0,1,0,0.2),rgb(0,0.7,0.7,0.2),rgb(0.7,0,0.7,0.2),rgb(0.5,0.5,0.5,0.2))
 ## check the relationship between wet-day 95-percentile and wet-day mean
-plot(-log(0.05)*zoo(mu[,2]),zoo(q95[,1]),pch=19,xlim=c(0,120),ylim=c(0,120),
+plot(-log(0.05)*zoo(mu[,1]),zoo(q95[,1]),pch=19,xlim=c(0,120),ylim=c(0,120),
      main='Wet-day mean v.s. 95th wet percentile',
      xlab=expression(paste(-ln(1-0.95)*mu,' (mm/day)')), ylab=expression(paste(q[95],' (mm/day)')))
-points(-log(0.05)*zoo(mu[,1]),zoo(q95[,2]),pch=19,col=rgb(1,0,0,0.2))
-points(-log(0.05)*zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
-points(-log(0.05)*zoo(mu[,4]),zoo(q95[,4]),pch=19,col=rgb(0,1,0,0.2))
-points(-log(0.05)*zoo(mu[,5]),zoo(q95[,5]),pch=19,col=rgb(0,0.7,0.7,0.2))
-points(-log(0.05)*zoo(mu[,6]),zoo(q95[,6]),pch=19,col=rgb(0.7,0,0.7,0.2))
-points(-log(0.05)*zoo(mu[,7]),zoo(q95[,7]),pch=19,col=rgb(0.5,0.5,0.5,0.2))
+for (i in 1:dim(mu[2])) points(-log(0.05)*zoo(mu[,i]),zoo(q95[,i]),pch=19,col=cols[i])
 grid()
-legend(0,120,loc(mu),pch=19,col=c('black',rgb(1,0,0,0.2),rgb(0,0,1,0.2),rgb(0,1,0,0.2),rgb(0,0.7,0.7,0.2),rgb(0.7,0,0.7,0.2),rgb(0.5,0.5,0.5,0.2)),bty='n',cex=0.5)
+legend(0,120,loc(mu),pch=19,col=c('black',cols[1:dim(mu)[2]]),bty='n',cex=0.5)
 ```
@@
 } else load('case1-dryseason.rda')
+attr(z,'variable') <- 'duration'
+attr(z,'longname') <- 'mean dry spell duration'
 eof.dry <- EOF(z)
-plot(eof.dry)
+plot(eof.dry,new=FALSE)
 ```
@@
 ds.t2m <- DS(eof.dry,eof.t2m)
-plot(ds.t2m)
+plot(ds.t2m,new=FALSE)
 ```
@@
 ds.slp <- DS(eof.dry,eof.slp)
-plot(ds.slp)
+plot(ds.slp,new=FALSE)
 ```
@@
 cca.slp.t2m <- CCA(eof.slp,eof.t2m)
-plot(cca.slp.t2m)
+plot(cca.slp.t2m,new=FALSE)
 ```
@@
 ds.res <- DS(eof.res,eof.t2m)
-plot(ds.res)
+plot(ds.res,new=FALSE)
 ```
 
-### Dry conditions and wild fire statistics
-
-Use the EOFs describing the dry conditions as predictors and wild fire statistics as predictand to assess the connection
-
-```{r}
-# z <- DS(wfire,eof.dry)
-```
+The downscaling exercise applied to the residual suggests that there is little to gain from using both SLP and temperature for downscaling the statistics of duration of dry days.
+Hence, a projection using SLP from GCMs as predictor may capture most of the future change.
+The temperature will nevertheless still be important for fires, as hotter temperatures are more favouralbe for fires.
 
@@
 if (!file.exists('dse.mld.slp.rda')) {
-  dse.mld.slp <- DSensemble.eof(eof.dry,predictor=SLP,it='mjjas',path="~/data/CMIP5.monthly",pattern = "psl_Amon_ens_")
+  dse.mld.slp <- DSensemble.eof(eof.dry,predictor=SLP,it='mjjas',ip=1:10,path="~/data/CMIP5.monthly",pattern = "psl_Amon_ens_")
   save(dse.mld.slp,file = 'dse.mld.slp.rda')
 } else load('dse.mld.slp.rda')
-plot(dse.mld.slp)
+plot(dse.mld.slp,new=FALSE)
+grid()
 ```
 
-Follow up with the residuals and the large-scale temperture
+Plot maps of changes in the mean dry spell duration
 
-```{r, warning=FALSE, eval=FALSE,echo=FALSE}
-if (!file.exists('dse.mld.t2m.rda')) {
-  dse.mld.t2m <- DSensemble.eof(eof.res,predictor=T2M,it='mjjas',path="~/data/CMIP5.monthly",pattern = "tas_Amon_ens_")
-  save(dse.mld.t2m,file = 'dse.mld.t2m.rda')
-} else load('dse.mld.t2m.rda')
-plot(dse.mld.t2m)
-```
-
-Add the the two components to get an estimate of the total change. The `dse`objects are lists which contain `zoo` time series:
-
-```{r, eval=FALSE,echo=FALSE}
-dse.mld <- dse.mld.slp
-gcmnm <- names(dse.mld)
-gcmnm <- gcmnm[-grep('info',gcmnm)]; gcmnm <- gcmnm[-grep('eof',gcmnm)]
-for (nm in gcmnm) {
-  coredata(dse.mld[[nm]]) <- coredata(dse.mld.slp[[nm]]) + coredata(dse.mld.t2m[[nm]])
-}
+```{r}
+ zmap1 <- map(dse.mld.slp,FUN="mean",FUNX='mean',it=c(2000,2009),plot=FALSE)
+ zmap2 <- map(dse.mld.slp,FUN="mean",FUNX='mean',it=c(2090,2099),plot=FALSE)
+ zmap <- zoo(coredata(zmap2) - coredata(zmap1),order.by=index(zmap2))
+ zmap <- attrcp(zmap1,zmap)
+ class(zmap) <- class(zmap1)
+ attr(zmap,'variable') <- 'mean dry duration'
+ attr(zmap,'unit') <- 'days'
+ rm('zmap1','zmap2')
+ map(zmap,colbar=list(pal='warm'),main='Change in mean duration of dry spells from 2000-2009',new=FALSE)
 ```
@@
 
-### Events
+#### Events
 
 Model number of evens $n_f$ and their area $A_f$ seperately, a little like precipitation ($\mu$ and $f_w$).
+When it comes to number of events, we can look at the total number or the number of events associated with a threshold area - the number of sevre fires.
+
 The area has an impact on the risk of affecting the number of assets.
@@
 Based on historic data and the assumption of a Poisson process dependent on external condition.
+
+```{r}
+## Read wild fire data
+colnames <- c('year','number','type','department','code-INSEE','municipality','something','DFCI.code','Alert','origin','area')
+fires <- read.table('~/Downloads/liste_incendies_ du_09_05_2017.csv',sep=';',col.names=colnames,header=FALSE)
+h <- hist(fires$year,col='grey',breaks=1973:2016)
+```
+
+Examine the fire event statistics: is it Poisson distributed?
+
+```{r}
+breaks <- seq(0,3000,by=250)
+lambda <- mean(h$counts)
+hist(h$counts,breaks=breaks,freq = FALSE,ylim=c(0,0.02))
+lines(seq(0,3000,by=10),dpois(seq(0,3000,by=10),lambda=lambda),lwd=3,col='red')
+```
+
+The annual number is not Poisson distributed. Examine the statistics of the area.
+
+```{r}
+summary(fires$area)
+plot(fires$area,log='y')
+```
+
+Compare the fire statistics with the mean duration of dry intervals from ECAD (eof.dry), the temperature and SLP.
+
+```{r}
+## Number of annual fires
+naf <- zoo(h$counts,order.by=trunc(h$mids))
+naf[naf==0] <- NA
+index(naf) <- year(naf)
+xy <- merge(naf,zoo(eof.dry),all=FALSE)
+cal.naf <- data.frame(y = coredata(xy$naf), x1=coredata(xy$x.1), x2=coredata(xy$x.2),
+                      x3= coredata(xy$x.3), x4=coredata(xy$x.4))
+naf.model <- lm(y ~ x1 + x2 + x3 + x4, data=cal.naf)
+z.naf <- zoo(predict(naf.model),order.by=trunc(h$mids))
+plot(naf,z.naf)
+```
+
+#### Burned area
+
+```{r}
+firearea <- data.frame(i=rep(1,length(fires$year)),area=as.numeric(as.character(fires$area)),year=fires$year)
+## Only include fires where more than 1 km^2 has burned down
+Af <- aggregate(firearea,by=list(firearea$year,firearea$area > 100000),FUN=sum)
+
+## Number of events with Af > 1km^2 per year
+plot(Af$Group.1[Af$Group.2],Af$i[Af$Group.2])
+
+## Number of events with Af > 1km^2 per year
+plot(Af$Group.1[Af$Group.2],Af$area[Af$Group.2])
+
+## Relationship between the number of events and total area.
+col <- rgb((2015 - Af$Group.1[Af$Group.2])/42,0,1- (2015 - Af$Group.1[Af$Group.2])/42,0.4)
+plot(Af$i[Af$Group.2],Af$area[Af$Group.2],col=col,pch=19)
+## annual total area of fires
+ataf <- as.station(zoo(Af$area[Af$Group.2],order.by=Af$Group.1[Af$Group.2]),loc='Southern France',param='area',unit='m^2',
+                  longname='burned area')
+plot(ataf,new=FALSE)
+```
+
+The number of wild fire events appears to have decreased over time and there burned area is characterised by a weak decline in years with moderate fire activity.
+There are also some exceptional years with large burned area, but the sample of these extreme cases is too small for a proper trend analysis.
+The peak years in fire activity may suggest that the fire hazard is sensitive to certain factors or that increases abruptly when some factors exceed a critical threshold state.
+
+The high sensitivity in the wild fire activity can be compared with the predictors if we take ln(Af) rather than using a linear relationship:
+
+```{r}
+index(ataf) <- year(ataf)
+index(eof.dry) <- year(eof.dry)
+eof.dry <- subset(eof.dry,ip=1:4)
+ds.ataf.dry <- DS(log(ataf),eof.dry)
+ds.ataf.t2m <- DS(log(ataf),eof.t2m)
+ds.ataf.slp <- DS(ataf,eof.slp)
+```
+
+Examine the connection between the mean dry spell duration and the area of burned forest
+
+```{r, height=8}
+plot(ds.ataf.dry,new=FALSE)
+```
+
+The regression results suggest a connection between the mean duration of the dry intervals and ln(Af) as the cross-validation correlation is 0.49.
+
+Also examine the connection between high summer temperature and the burned area:
+
+```{r, height=8}
+plot(ds.ataf.t2m,new=FALSE)
+```
+
+There is also a connection between ln(Af) and the summer temperature, and the cross-validation correlation is 0.4.
+
+The same analysis repeated for the SLP fails to find a connection for ln(Af), but the linear relationship between SLP and Af captures a dependency
+
+```{r, height=8}
+plot(ds.ataf.slp,new=FALSE)
+```
+
+The connection between SLP and burned area is weaker than for the the mean dry interval duration and summer temperature if we use ln(Af), with a cross-validation correlation of 0.35.
+It is particularly the peak years which are associated with high-pressure systems (often associated with dry conditions and heatwaves).
+
+The time series suggest an increasing bias over time, which possibly can be explained by more efficient fire service since 1990.
+This can be checked to some extent by examining the budget of the fire services to see if there have been more efforts and resources devoted to firefighting over time.
+It is also possible that the vegetation has changed or that the landscape has been altered (e.g.
+more fields or more paved area) so that the potential for wild fires has been reduced.
+Also, once there has been a fire, the are is burned down and it takes time for it to recover so that new fires may occur.
+
+### Projections of future wild fire activity
+
+We can combine the projections of the mean dry interval duration with the downscaling result using the same quantity as predictor.
+The dry duration statistics was downscaled using the SLP as predictor, so the SLP is in the equation.
+
+```{r}
+## Convert the projected data into an EOF for the multi-model ensemble
+eof.dry.z <- as.eof(dse.mld.slp)
+## Use this EOF as input with the downscaling model calibrated with mean dry interval duration to make projections of ln(fire area)
+ds.Af <- exp(predict(ds.ataf.dry,newdata=eof.dry.z))
+plot(ds.Af,new=FALSE)
+```
+
+#### Simultaneos fires
+
+```{r}
+fevents <- data.frame(i=rep(1,length(fires$year)),area=as.numeric(as.character(fires$area)),year=fires$year,date=as.Date(fires$Alert))
+## Only include fires where more than 1 km^2 has burned down
+se <- aggregate(fevents$i,by=list(fevents$date,fevents$area > 100000),FUN=sum)
+
+## Examine the statistical distribution of simultaneous events
+hist(se$x, col='grey', main='Number of simultaneous wild fires')
+
+plot(year(se$Group.1),se$x,pch=19,col=rgb(1,0,0,0.2),main='Number of simultaneous wild fires by year',ylab='Simultaneous events',xlab='')
+points(year(se$Group.1),se$x,pch=21,col=rgb(0.5,0.5,0.5))
+grid()
+```

### 2017-05-19 EU-Circle/eu-circle_case-study-4.Rmd
Progress on EU-circle case-study 1

@@
 
-Get data on storm tracks... IMILAST
-Correlation between storm tracks and stations.
+Get data on storm tracks.
+IMILAST data do not cover the region as the analysis is only for storms north of 20N and south of 20S.
+Tropical cyclones are reported in the HURDAT2 dataset \url{http://www.aoml.noaa.gov/hrd/hurdat/hurdat2-1851-2016-apr2017.txt}, but only for the North-Atlantic.
+There are some data in the best-track data base \url{http://www.usno.navy.mil/NOOC/nmfc-ph/RSS/jtwc/best_tracks/}.
 
-## HURDAT2
+## Best Track
 
 ```{r}
-url <- 'http://www.aoml.noaa.gov/hrd/hurdat/hurdat2-1851-2016-apr2017.txt'
+url <- 'http://www.usno.navy.mil/NOOC/nmfc-ph/RSS/jtwc/best_tracks/'
+```
 
-```
+Correlation between storm tracks and stations.
 

### 2017-05-19 EU-Circle/eu-circle_case-study-1.R

@@ -25,6 +25,8 @@ diagnose(x)
 plot(x,new=FALSE)
 grid()
 
+print('wet-day mean')
+
 ## ----warning=FALSE-------------------------------------------------------
 mu <- annual(x,'wetmean',nmin=250)
 wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
@@ -46,6 +48,8 @@ for (i in 2:dim(mu)[2]) {
 }
 grid()
 
+print('quantiles')
+
 ## ----warning=FALSE-------------------------------------------------------
 ## check the relationship between wet-day 95-percentile and wet-day mean
 plot(-log(0.05)*zoo(mu[,2]),zoo(q95[,1]),pch=19,xlim=c(0,120),ylim=c(0,120),
@@ -55,16 +59,20 @@ points(-log(0.05)*zoo(mu[,1]),zoo(q95[,2]),pch=19,col=rgb(1,0,0,0.2))
 points(-log(0.05)*zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
 points(-log(0.05)*zoo(mu[,4]),zoo(q95[,4]),pch=19,col=rgb(0,1,0,0.2))
 points(-log(0.05)*zoo(mu[,5]),zoo(q95[,5]),pch=19,col=rgb(0,0.7,0.7,0.2))
-points(-log(0.05)*zoo(mu[,6]),zoo(q95[,6]),pch=19,col=rgb(0.7,0,0.7,0.2))
-points(-log(0.05)*zoo(mu[,7]),zoo(q95[,7]),pch=19,col=rgb(0.5,0.5,0.5,0.2))
+#points(-log(0.05)*zoo(mu[,6]),zoo(q95[,6]),pch=19,col=rgb(0.7,0,0.7,0.2))
+#points(-log(0.05)*zoo(mu[,7]),zoo(q95[,7]),pch=19,col=rgb(0.5,0.5,0.5,0.2))
 grid()
 legend(0,120,loc(mu),pch=19,col=c('black',rgb(1,0,0,0.2),rgb(0,0,1,0.2),rgb(0,1,0,0.2),rgb(0,0.7,0.7,0.2),rgb(0.7,0,0.7,0.2),rgb(0.5,0.5,0.5,0.2)),bty='n',cex=0.5)
+
+print('Wet-day frequency')
 
[…]

### 2017-07-07 SPECS/specs-decade-mos-all.Rmd
SPECS decadal multi-model analysis

--- 2017-04-28 SPECS/specs-decade-mos.Rmd
@@
 ```
+
+Use `purl()` to generate a plain R-script.
 
@@
 downloadfromunican <- FALSE    # Preparations - get the data
-downscaledreanalysis <- FALSE  # SM: Assessment of large-small scale link
-looexperiments <- FALSE        # Cross-validation MOS on forecasts
-diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts
+downscaledreanalysis <- TRUE  # SM: Assessment of large-small scale link
+looexperiments <- TRUE        # Cross-validation MOS on forecasts
+diagnoseFCST <- TRUE          # Show correlation & common EOFs of reanalysis+fcsts
 
-source("/home/rasmusb/disk1/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
+#source("/home/rasmusb/disk1/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
 
@@
     ## member number
+    if (verbose) print(year(Y1))
     if (d[2]<=12) index(Y1) <- as.Date(paste(year(Y1),im,'01',sep='-'),
                                        origin = as.Date("1950-01-01")) else
-                  index(Y1) <- as.Date(paste(year(Y1),'01-01',sep='-')+im,
+                  index(Y1) <- as.Date(as.Date(paste(year(Y1),'01-01',sep='-'))+im,
                                        origin = as.Date("1950-01-01"))
@@
                           origin = as.Date("1950-01-01")) else
-      index(y) <- as.Date(paste(year(y),'01-01',sep='-')+i,
+      index(y) <- as.Date(as.Date(paste(year(y),'01-01',sep='-'))+i,
                           origin = as.Date("1950-01-01"))
@@
     }
+  } else {
+    jday <- julian(index(x))-julian(as.Date(paste(year(x),'01-01',sep='-')))
+    m <- as.integer(rownames(table(jday)))
+    if (verbose) print(paste('separate ensemble members by day 1 --',m))
+    j <- 1
+    for (i in m) {
+      itday <- is.element(jday,i)
+      if (verbose) print(c(i,j,sum(itday)))
+      y <- pentad(subset(X,it=itday),l=l)
+      index(y) <- as.Date(paste(year(y),i,'01',sep='-'))
+      if (j==1) Y <- zoo(y) else
+                Y <- c(Y,zoo(y))
+      j <- j+1
+    }
   }
@@
 ```{r compare}
-comparemannual <- function(tas_DECA,reanalysis='~/Downloads/air.mon.mean.nc',mask=TRUE,l= 3,eofs=NULL) {
+comparemannual <- function(tas_DECA,reanalysis='~/R/air.mon.mean.nc',mask=TRUE,l= 3,eofs=NULL) {
 
@@
   load(results)  
-  eval(parse(text=paste('zmos <-',substr(results,14,nchar(results)-4))))
+  zl <- regexpr('.z.',results)
+  print(paste('zmos <-',substr(results,zl+1,nchar(results)-4)))
+  eval(parse(text=paste('zmos <-',substr(results,zl+1,nchar(results)-4))))
   attr(zmos,'variable') <- varid(y)
@@
 print('Get the predictands: mu & fw from EOBS')
-if (!file.exists('~/data/mu_0.25deg_reg.nc'))
+if (!file.exists('~/storeA/mu_0.25deg_reg.nc'))
   mu.eobs <-  wetdaymean('~/data/data.ECAD/rr_0.25deg_reg.nc') else
-  mu.eobs <- retrieve('~/data/mu_0.25deg_reg.nc')
-fw.eobs <- retrieve('~/data/fw_0.25deg_reg.nc')
+  mu.eobs <- retrieve('~/storeA/mu_0.25deg_reg.nc')
+fw.eobs <- retrieve('~/storeA/fw_0.25deg_reg.nc')
 attr(mu.eobs,'variable') <- 'mu'
@@
 attr(fw.eobs,'unit') <- 'fraction'
-index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
-index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
+#index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
+#index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
 
@@
 ## retrieve did not get the right time stamp:
-  es <- annual(C.C.eq(retrieve('~/Downloads/air.mon.mean.nc',lon=c(-60,40),lat=c(20,70))))
+  es <- annual(C.C.eq(retrieve('~/R/air.mon.mean.nc',lon=c(-60,40),lat=c(20,70))))
   es.eof <- EOF(mask(es,land=TRUE))
@@
 ## Also get the mean sea-level pressure for the wet-day frequency 
-  slp <- annual(retrieve('slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
+  slp <- annual(retrieve('~/R/slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
   slp.eof <- EOF(slp)
@@
   class(mu.eof) <- class(es.eof)
+  index(mu.eof) <- year(mu.eof)
   mu.ds <- DS(mu.eof,es.eof)
   class(fw.eof) <- class(slp.eof)
+  index(fw.eof) <- year(fw.eof)
   fw.ds <- DS(fw.eof,slp.eof)
@@
   plot(mu.ds,colbar1=list(breaks=seq(-0.025,0.025,by=0.005),rev=TRUE),
-       colbar2=list(breaks=seq(-350,350,by=25)),type='fill',new=FALSE,)
+       colbar2=list(breaks=seq(-350,350,by=25)),type='fill',new=FALSE)
   figlab('ESD(NCEP/NCAR)',ypos=0.999)
@@
 print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
-load('~/data/gfdl_DECA.rda')
-
+load('~/R/mme_DECA.rda')
 ```
@@
   print('Check results for LOO experiment - if absent do it')
-  if (!file.exists('specsdecadal.z.fw.y1.rda')) {
+  if (!file.exists('~/R/specsdecadal.z.fw.y1.rda')) {
     print('do specsdecadal.z.fw.y1.rda')
     z.fw.y1 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1,it=NULL,is=NULL,mask=FALSE)
-    save(file='specsdecadal.z.fw.y1.rda',z.fw.y1)
+    save(file='~/R/specsdecadal.z.fw.y1.rda',z.fw.y1)
     rm("z.fw.y1"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.mu.y1.rda')) {
+  if (!file.exists('~/R/specsdecadal.z.mu.y1.rda')) {
     print('do specsdecadal.z.mu.y1.rda')
     z.mu.y1 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1,it=NULL,is=NULL,mask=TRUE)
-    save(file='specsdecadal.z.mu.y1.rda',z.mu.y1)
+    save(file='~/R/specsdecadal.z.mu.y1.rda',z.mu.y1)
     rm("z.mu.y1"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.fw.y1.3.rda')) {
+  if (!file.exists('~/R/specsdecadal.z.fw.y1.3.rda')) {
     print('do specsdecadal.z.fw.y1.3.rda')
     z.fw.y1.3 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:3,it=NULL,is=NULL,mask=FALSE)
-    save(file='specsdecadal.z.fw.y1.3.rda',z.fw.y1.3)
+    save(file='~/R/specsdecadal.z.fw.y1.3.rda',z.fw.y1.3)
     rm("z.fw.y1.3"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.mu.y1.3.rda')) {
+  if (!file.exists('~/R/specsdecadal.z.mu.y1.3.rda')) {
     print('do specsdecadal.z.mu.y1.3.rda')
     z.mu.y1.3 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:3,it=NULL,is=NULL,mask=TRUE)
-    save(file='specsdecadal.z.mu.y1.3.rda',z.mu.y1.3)
+    save(file='~/R/specsdecadal.z.mu.y1.3.rda',z.mu.y1.3)
     rm("z.mu.y1.3"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.fw.y1.5.rda')) {
+  if (!file.exists('~/R/specsdecadal.z.fw.y1.5.rda')) {
     print('do specsdecadal.z.fw.y1.5.rda')
     z.fw.y1.5 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:5,it=NULL,is=NULL,mask=FALSE)
-    save(file='specsdecadal.z.fw.y1.5.rda',z.fw.y1.5)
+    save(file='~/R/specsdecadal.z.fw.y1.5.rda',z.fw.y1.5)
     rm("z.fw.y1.5"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.mu.y1.5.rda')) {
+  if (!file.exists('~/R/specsdecadal.z.mu.y1.5.rda')) {
     print('do specsdecadal.z.mu.y1.5.rda')
     z.mu.y1.5 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:5,it=NULL,is=NULL,mask=TRUE)
-    save(file='specsdecadal.z.mu.y1.5.rda',z.mu.y1.5)
+    save(file='~/R/specsdecadal.z.mu.y1.5.rda',z.mu.y1.5)
     rm("z.mu.y1.5"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.fw.y1.9.rda')) {
+  if (!file.exists('~/R/specsdecadal.z.fw.y1.9.rda')) {
     print('do specsdecadal.z.fw.y1.9.rda')
      z.fw.y1.9 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:9,it=NULL,is=NULL,mask=FALSE)
-     save(file='specsdecadal.z.fw.y1.9.rda',z.fw.y1.9)
+     save(file='~/R/specsdecadal.z.fw.y1.9.rda',z.fw.y1.9)
      rm("z.fw.y1.9"); gc(reset=TRUE)
    }
-  if (!file.exists('specsdecadal.z.mu.y1.9.rda')) {
+  if (!file.exists('~/R/specsdecadal.z.mu.y1.9.rda')) {
     print('do specsdecadal.z.mu.y1.9.rda')
     z.mu.y1.9 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:9,it=NULL,is=NULL,mask=TRUE)
-    save(file='specsdecadal.z.mu.y1.9.rda',z.mu.y1.9)
+    save(file='~/R/specsdecadal.z.mu.y1.9.rda',z.mu.y1.9)
     rm("z.mu.y1.9"); gc(reset=TRUE)
@@
   print('diagnose the MOS results')
-  load('specsdecadal.z.mu.y1.3.rda')
+  load('~/R/specsdecadal.z.mu.y1.3.rda')
   it0 <- range(year(z.mu.y1.3))
   print(paste('time period',paste(it0,collapse='-')))
-  score.mu.y1 <- diagnoseMOS('specsdecadal.z.mu.y1.rda',mu.eobs)
-  score.fw.y1 <- diagnoseMOS('specsdecadal.z.fw.y1.rda',fw.eobs)
-  score.mu.y1.3 <- diagnoseMOS('specsdecadal.z.mu.y1.3.rda',mu.eobs,l=3)
-  score.fw.y1.3 <- diagnoseMOS('specsdecadal.z.fw.y1.3.rda',fw.eobs,l=3)
-  score.mu.y1.5 <- diagnoseMOS('specsdecadal.z.mu.y1.5.rda',mu.eobs,l=5)
-  score.fw.y1.5 <- diagnoseMOS('specsdecadal.z.fw.y1.5.rda',fw.eobs,l=5)
-  score.mu.y1.9 <- diagnoseMOS('specsdecadal.z.mu.y1.9.rda',mu.eobs,l=9)
-  score.fw.y1.9 <- diagnoseMOS('specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)
+  score.mu.y1 <- diagnoseMOS('~/R/specsdecadal.z.mu.y1.rda',mu.eobs)
+  score.fw.y1 <- diagnoseMOS('~/R/specsdecadal.z.fw.y1.rda',fw.eobs)
+  score.mu.y1.3 <- diagnoseMOS('~/R/specsdecadal.z.mu.y1.3.rda',mu.eobs,l=3)
+  score.fw.y1.3 <- diagnoseMOS('~/R/specsdecadal.z.fw.y1.3.rda',fw.eobs,l=3)
+  score.mu.y1.5 <- diagnoseMOS('~/R/specsdecadal.z.mu.y1.5.rda',mu.eobs,l=5)
+  score.fw.y1.5 <- diagnoseMOS('~/R/specsdecadal.z.fw.y1.5.rda',fw.eobs,l=5)
+  score.mu.y1.9 <- diagnoseMOS('~/R/specsdecadal.z.mu.y1.9.rda',mu.eobs,l=9)
+  score.fw.y1.9 <- diagnoseMOS('~/R/specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)
   while (dev.cur()>1) dev.off()
@@
   comparemannual(tas_DECA,mask=TRUE,l= 8)
-  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 1)
-  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 2)
-  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 3)
-  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 4)
-  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 5)
-  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 6)
-  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 7)
-  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 8)
+  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 1)
+  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 2)
+  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 3)
+  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 4)
+  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 5)
+  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 6)
+  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 7)
+  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 8)
   rmap.1.10.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:10)
@@
 ## Scaling dependency: time - space
-t2m <- retrieve('~/Downloads/air.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
+t2m <- retrieve('~/R/air.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
 t2m <- anomaly(t2m)
@@
 ## Repeat for SLP
-slp <- retrieve('~/data/ncep/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
+slp <- retrieve('~/R/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
 slp <- anomaly(slp)

### 2017-07-07 KSS/dse4KSS.R

@@ -5,23 +5,28 @@ args=commandArgs(trailingOnly=TRUE)
 print('--- R-job for downscaling Nordic temperatures for KSS ---')
 print(args)
 
+#library(devtools)
+#install.packages('esd',repos=NULL)
 library(esd)
 
 ## Function definitions:
 
 downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
                       period=c(1950,2015),plot=FALSE,rcp='rcp45',verbose=FALSE,
-                      lon=c(-20,40),lat=c(50,80),ip=1:2,n=3,
+                      lon=c(-20,40),lat=c(50,80),ip=1:7,n=5,
                       pattern='tas_Amon_ens',
                       rel.cord=FALSE,select=NULL) {
   
-  print('downscale')
+  if (verbose) print(paste('downscale',param,it,FUN,FUNX))
   
   ## Use a time and space window:  
+  if (verbose) print(paste('subset',paste(period,collapse='-')))
   Y <- subset(Y,it=period)
   
   ## Estimate seasonal means & weed out stations with little data
+  if (verbose) print('season')
   Y4 <- subset(as.4seasons(Y,FUN=FUN),it=it)
+  if (verbose) print('weed out missing data')
   ok <- apply(coredata(Y4),1,nv)
   Y4 <- subset(Y4,it=ok>0)
   nok <- apply(coredata(Y4),2,nv)
@@ -35,17 +40,20 @@ downscale <- function(Y,predictor,it='djf',param='t2m',FUN='mean',FUNX='mean',
   #print(paste(nmiss,'% missing',sep=''))
   
   ## Fill missing data using PCA-based regression
+  if (verbose) print('pcafill')
   Z <- pcafill(Y4)
+  ## Negative precipitation is impossible - clip to zero
+  if (is.precip(Z)) coredata(Z)[Z<0]<- 0
   
+  if (verbose) print('pca')
   pca <- PCA(Z,n=n)
   if (plot) plot(pca)
   
   ## Downscale results
-  print('DSensemble')
+  print(paste('DSensemble',varid(predictor)))
[…]

### 2017-07-07 KSS/dse4KSS.sh

@@ -2,14 +2,14 @@
 
 echo "Submit downscaling jobs for KSS as parallel tasks"
 for PARAM in t2m mu fw
-#for PARAM in mu fw
-#for PARAM in t2m #TEST
+#for PARAM in mu 
+#for PARAM in t2m # TEST
 do  
   for SEASON in djf mam jja son ondjfm amjjas
   #for SEASON in djf # TEST
   do
     for RCP in rcp45 rcp85 rcp26
-    #for RCP in rcp45
+    #for RCP in rcp45 # TEST
     do
       echo dse4KSS.$PARAM.$SEASON.$RCP.job
       sed 's|param|'$PARAM'|;

### 2017-07-07 SPECS/specs-decade-mos.R

@@ -1,5 +1,5 @@
 ## ----setup, include=FALSE------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE)
+#knitr::opts_chunk$set(echo = TRUE)
 
 ## ----initialise----------------------------------------------------------
 ## Script that applies MOS to SPECS decadal forecasts
@@ -287,7 +287,7 @@ comparemannual <- function(tas_DECA,reanalysis='~/Downloads/air.mon.mean.nc',mas
   ## Apply an EOF-filter to emphasise only the large scale anomalies:
   if (!is.null(eofs)) {
     print(paste('Only keep the large-scale structures; EOFs:',paste(eofs,collapse=' ,')))
-    Xm.eofs <- subset(EOF(Xm),pattern=eofs)
+    Xm.eofs <- subset(EOF(Xm),ip=eofs)
     Xm <- eof2field(Xm.eofs)
     land <- paste(land,'eof_',paste(eofs,collapse='+'))
   }

### 2017-08-01 SPECS/specs-decade-mos.Rmd
EU-Circle case 1

--- 2017-04-28 SPECS/specs-decade-mos.Rmd
@@
 setwd("~/R")
+experiments <- c('~/data/gfdl_DECA.rda','~/data/hadcm3_DECA.rda','~/data/miroc5_DECA.rda','~/data/mpi_DECA.rda')
 
 downloadfromunican <- FALSE    # Preparations - get the data
-downscaledreanalysis <- FALSE  # SM: Assessment of large-small scale link
-looexperiments <- FALSE        # Cross-validation MOS on forecasts
-diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts
+downscaledreanalysis <- TRUE  # SM: Assessment of large-small scale link
+looexperiments <- TRUE        # Cross-validation MOS on forecasts
+diagnoseFCST <- TRUE          # Show correlation & common EOFs of reanalysis+fcsts
 
@@
   d <- dim(x$Data)
+  nmembers <- length(x$members)
   if (verbose) {print(summary(x)); print(d)}
@@
       ## Read the entire field
-      y1 <- fcst2field(x,it0=10-ii,im=im,verbose=verbose)
+      y1 <- fcst2field(x,it0=nmembers-ii,im=im,verbose=verbose)
       ## Extract a sub-region
@@
   X <- aggregatedailyfield(fname=fname,cline=cline)
-  Y <- aggregatedailyfield(fname=,'mask.nc',cline=cline,output='fw_0.25deg_reg.nc')
+  Y <- aggregatedailyfield(fname='mask.nc',cline=cline,output='fw_0.25deg_reg.nc')
   file.remove(fname); file.remove('mask.nc')
@@
   Z <- subset(Z,it=range(year(X)))
-  for (i in 1:10) {
+  nmembers <- length(table(month(X)))
+  for (i in 1:nmembers) {
     z <- subset(X,it=month.abb[i])
@@
   load(results)  
-  eval(parse(text=paste('zmos <-',substr(results,14,nchar(results)-4))))
+  zl <- regexpr('.z.',results)
+  eval(parse(text=paste('zmos <-',substr(results,zl+1,nchar(results)-4))))
   attr(zmos,'variable') <- varid(y)
@@
 print('Get the predictands: mu & fw from EOBS')
-if (!file.exists('~/data/mu_0.25deg_reg.nc'))
+if (!file.exists('~/mu_0.25deg_reg.nc'))
   mu.eobs <-  wetdaymean('~/data/data.ECAD/rr_0.25deg_reg.nc') else
-  mu.eobs <- retrieve('~/data/mu_0.25deg_reg.nc')
-fw.eobs <- retrieve('~/data/fw_0.25deg_reg.nc')
+  mu.eobs <- retrieve('~/mu_0.25deg_reg.nc')
+fw.eobs <- retrieve('~/fw_0.25deg_reg.nc')
 attr(mu.eobs,'variable') <- 'mu'
@@
 attr(fw.eobs,'unit') <- 'fraction'
-index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
-index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
-
+#index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
+#index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
 ```
 
-```{r ds}
+```{r ds,fig.height=8}
 
@@
 ## Also get the mean sea-level pressure for the wet-day frequency 
-  slp <- annual(retrieve('slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
+  slp <- annual(retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
   slp.eof <- EOF(slp)
@@
 
-print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
-load('~/data/gfdl_DECA.rda')
+for (iexp in 1:1) {
+  print(paste('Get the decadal predictors saved locally in',experiments[iexp]))
+  load(experiments[iexp])
 
-```
-
-Leave-one-out cross-validation experiments
-
-```{r loo}
+## Leave-one-out cross-validation experiments
+## Names of file names with results stored  
+fw1name <- paste('specsdecadal.z.fw.y1.gcm',iexp,'rda',sep='.')
+mu1name <- paste('specsdecadal.z.mu.y1.gcm',iexp,'rda',sep='.')
+fw3name <- paste('specsdecadal.z.fw.y1.3.gcm',iexp,'rda',sep='.')
+mu3name <- paste('specsdecadal.z.mu.y1.3.gcm',iexp,'rda',sep='.')
+fw5name <- paste('specsdecadal.z.fw.y1.5.gcm',iexp,'rda',sep='.')
+mu5name <- paste('specsdecadal.z.mu.y1.5.gcm',iexp,'rda',sep='.')
+fw9name <- paste('specsdecadal.z.fw.y1.9.gcm',iexp,'rda',sep='.')
+mu9name <- paste('specsdecadal.z.mu.y1.9.gcm',iexp,'rda',sep='.')
 
@@
   print('Check results for LOO experiment - if absent do it')
-  if (!file.exists('specsdecadal.z.fw.y1.rda')) {
-    print('do specsdecadal.z.fw.y1.rda')
+  if (!file.exists(fw1name)) {
+    print(fw1name)
     z.fw.y1 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1,it=NULL,is=NULL,mask=FALSE)
-    save(file='specsdecadal.z.fw.y1.rda',z.fw.y1)
+    save(file=fw1name,z.fw.y1)
     rm("z.fw.y1"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.mu.y1.rda')) {
-    print('do specsdecadal.z.mu.y1.rda')
+
+  if (!file.exists(mu1name)) {
+    print(mu1name)
     z.mu.y1 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1,it=NULL,is=NULL,mask=TRUE)
-    save(file='specsdecadal.z.mu.y1.rda',z.mu.y1)
+    save(file=mu1name,z.mu.y1)
     rm("z.mu.y1"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.fw.y1.3.rda')) {
-    print('do specsdecadal.z.fw.y1.3.rda')
+  if (!file.exists(fw3name)) {
+    print(fw3name)
     z.fw.y1.3 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:3,it=NULL,is=NULL,mask=FALSE)
-    save(file='specsdecadal.z.fw.y1.3.rda',z.fw.y1.3)
+    save(file=fw3name,z.fw.y1.3)
     rm("z.fw.y1.3"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.mu.y1.3.rda')) {
-    print('do specsdecadal.z.mu.y1.3.rda')
+  if (!file.exists(mu3name)) {
+    print(mu3name)
     z.mu.y1.3 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:3,it=NULL,is=NULL,mask=TRUE)
-    save(file='specsdecadal.z.mu.y1.3.rda',z.mu.y1.3)
+    save(file=mu3name,z.mu.y1.3)
     rm("z.mu.y1.3"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.fw.y1.5.rda')) {
-    print('do specsdecadal.z.fw.y1.5.rda')
+  if (!file.exists(fw5name)) {
+    print(fw5name)
     z.fw.y1.5 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:5,it=NULL,is=NULL,mask=FALSE)
-    save(file='specsdecadal.z.fw.y1.5.rda',z.fw.y1.5)
+    save(file=fw5name,z.fw.y1.5)
     rm("z.fw.y1.5"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.mu.y1.5.rda')) {
-    print('do specsdecadal.z.mu.y1.5.rda')
+  if (!file.exists(mu5name)) {
+    print(mu5name)
     z.mu.y1.5 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:5,it=NULL,is=NULL,mask=TRUE)
-    save(file='specsdecadal.z.mu.y1.5.rda',z.mu.y1.5)
+    save(file=mu5name,z.mu.y1.5)
     rm("z.mu.y1.5"); gc(reset=TRUE)
   }
-  if (!file.exists('specsdecadal.z.fw.y1.9.rda')) {
-    print('do specsdecadal.z.fw.y1.9.rda')
+  if (!file.exists(fw9name)) {
+    print(fw9name)
      z.fw.y1.9 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:9,it=NULL,is=NULL,mask=FALSE)
-     save(file='specsdecadal.z.fw.y1.9.rda',z.fw.y1.9)
+     save(file=fw9name,z.fw.y1.9)
      rm("z.fw.y1.9"); gc(reset=TRUE)
    }
-  if (!file.exists('specsdecadal.z.mu.y1.9.rda')) {
-    print('do specsdecadal.z.mu.y1.9.rda')
+  if (!file.exists(mu9name)) {
+    print(mu9name)
     z.mu.y1.9 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:9,it=NULL,is=NULL,mask=TRUE)
-    save(file='specsdecadal.z.mu.y1.9.rda',z.mu.y1.9)
+    save(file=mu9name,z.mu.y1.9)
     rm("z.mu.y1.9"); gc(reset=TRUE)
@@
   score.fw.y1.9 <- diagnoseMOS('specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)
-  while (dev.cur()>1) dev.off()
+
 ## Repeat and make a table of the results...
-
 }
 
-```
+## Diagnose the forecasts
 
-Diagnose the forecasts
-
-```{r diagnosefcst}
 ## Common EOFs to diagnose the results?
@@
 }
-```
 
-```{r rmaps}
   ## Test the forecasts of the larg-scale anomalies/gravest EOFs
@@
   rmap.1.9.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:9)
-```
 
-```{r rdiff}
   rdiff <- rmap.1.3.nomask.3 - rmap.nomask.3
@@
   figlab('TAS forecast (3 years): correlations EOFs 1-9 vs full field',ypos=0.99)
+}
 ```
@@
 ```{r slp}
-
 ## Repeat for SLP
-slp <- retrieve('~/data/ncep/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
+slp <- retrieve('~/data/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
 slp <- anomaly(slp)

### 2017-08-01 EU-Circle/eu-circle_case-study-1.R

@@ -25,8 +25,6 @@ diagnose(x)
 plot(x,new=FALSE)
 grid()
 
-print('wet-day mean')
-
 ## ----warning=FALSE-------------------------------------------------------
 mu <- annual(x,'wetmean',nmin=250)
 wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
@@ -48,31 +46,22 @@ for (i in 2:dim(mu)[2]) {
 }
 grid()
 
-print('quantiles')
-
 ## ----warning=FALSE-------------------------------------------------------
+
+cols <- c(rgb(1,0,0,0.2),rgb(0,0,1,0.2),rgb(0,1,0,0.2),rgb(0,0.7,0.7,0.2),rgb(0.7,0,0.7,0.2),rgb(0.5,0.5,0.5,0.2))
 ## check the relationship between wet-day 95-percentile and wet-day mean
-plot(-log(0.05)*zoo(mu[,2]),zoo(q95[,1]),pch=19,xlim=c(0,120),ylim=c(0,120),
+plot(-log(0.05)*zoo(mu[,1]),zoo(q95[,1]),pch=19,xlim=c(0,120),ylim=c(0,120),
      main='Wet-day mean v.s. 95th wet percentile',
      xlab=expression(paste(-ln(1-0.95)*mu,' (mm/day)')), ylab=expression(paste(q[95],' (mm/day)')))
-points(-log(0.05)*zoo(mu[,1]),zoo(q95[,2]),pch=19,col=rgb(1,0,0,0.2))
-points(-log(0.05)*zoo(mu[,3]),zoo(q95[,3]),pch=19,col=rgb(0,0,1,0.2))
-points(-log(0.05)*zoo(mu[,4]),zoo(q95[,4]),pch=19,col=rgb(0,1,0,0.2))
-points(-log(0.05)*zoo(mu[,5]),zoo(q95[,5]),pch=19,col=rgb(0,0.7,0.7,0.2))
-#points(-log(0.05)*zoo(mu[,6]),zoo(q95[,6]),pch=19,col=rgb(0.7,0,0.7,0.2))
-#points(-log(0.05)*zoo(mu[,7]),zoo(q95[,7]),pch=19,col=rgb(0.5,0.5,0.5,0.2))
[…]

### 2017-08-01 paper58/paper58.R

@@ -564,7 +564,7 @@ legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
                expression(bar(n)[c*w*d]),expression(bar(n)[c*d*d])),bty='n',
        col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1),
        ncol=2)
-figlab('Figure SM3')
+figlab('Figure S3')
 dev.copy2pdf(file='figureSM3.pdf')
 
 ## Figure SM2.
@@ -580,7 +580,7 @@ plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
      main='Test: exponential distribution & changing mean',
      xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
 lines(c(0,40),c(0,40),col='grey')
-figlab('Figure SM1')
+figlab('Figure S1')
 dev.copy2pdf(file='figureSM1.pdf')
 
 ## Figure SM3
@@ -588,7 +588,7 @@ dev.copy2pdf(file='figureSM1.pdf')
 X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
 dev.new()
 map(X,projection='sphere',colbar=list(pal="budrd",breaks=seq(8,28,by=0.5)))
-figlab('Figure SM2',xpos=0.8,ypos=0.999)
+figlab('Figure S2',xpos=0.8,ypos=0.999)
 dev.copy2pdf(file='figureSM2.pdf')
 
 ## Figure SM4
@@ -608,7 +608,7 @@ grid()
 legend("topleft",legend=c("RCP4.5","RCP8.5","RCP2.6"),lty=1,lwd=2,
        bty="o",box.lwd=0.2,
        col=c(rgb(0.5,0.5,0.5,0.5),rgb(1,0.5,0.5,0.5),rgb(0.5,1,0.5,0.5)))
-figlab('Figure SM4')
+figlab('Figure S4')
 dev.copy2pdf(file='figureSM4.pdf')
 
 ## Figure SM5
@@ -618,7 +618,7 @@ hist(100*as.numeric(r2),breaks=seq(0,100,by=5),lwd=2,col=rgb(0,0.3,0.5),
      xlab=expression(paste(R^2,' (%)')),freq=TRUE,
      main="Summary of regression scores")
 grid()
[…]

### 2017-08-08 EU-Circle/bangladesh/bangladesh-imilast.R
updated script for Bangladesh cyclones

@@ -1,0 +1,115 @@
+## Rasmus Benestad
+## Read IMILAST storms - select Bangladesh
+## The IMILAST project excluded tropical cyclones.
+
+library(esd)
+
+find.file <- function(filename) {
+  command <- paste("find -L $HOME -name",filename,sep=" ")
+  fullpath <- system(command,intern=TRUE)
+  if(length(fullpath)==0) return(FALSE)
+  return(fullpath)
+}
+
+path <- sub("/bangladesh-imilast.R","",find.file("bangladesh-imilast.R")[1])
+
+## HURDAT2 CANNOT BE USED FOR BANGLADESH.
+## IT CONTAINS ONLY ATLANTIC (AND SOME PACIFIC?) STORMS. 
+#hurdat2.storms <- find.file('hurdat2-1851-2016-apr2017.txt')[1]
+#Y <- read.hurdat2(hurdat2.storms)
+
+## IMILAST data for Bangladesh and the surrounding region
+is <- list(lon=c(60,120),lat=c(0,40))
+fname.imilast <- file.path(path,'imilast_bangladesh_ERAinterim_1.5.rda')
+if(is.character(file.exists(fname.imilast))) {
+  load(fname.imilast)
+} else {
+  fname <- find.file('ERAinterim_1.5_NH_*.txt')
+  fname <- fname[!duplicated(gsub(".*/","",fname))]
+  M <- gsub("_.*","",gsub(".*M","",fname))
+  imilast <- list()
+  for (i in seq(length(fname))) {
+    print(paste("imilast M",M[i],sep=""))
+    x <- subset(read.imilast(fname[i]),is=is)
+    imilast[[paste("M",M[i],sep="")]] <- x
+  }
+  save(file=fname.imilast,imilast)
+}
+
+## Map cyclone trajectories
+save.fig <- TRUE
+is <- list(lon=c(75,105),lat=c(0,30))
+## Cyclone 1BOB01, April 29, 1991
+t1 <- seq(strptime("19910428 12",format="%Y%m%d %H"), 
+          strptime("19910501 00",format="%Y%m%d %H"), by = "12 hours")
+## May 1997 Bangladesh cyclone
+t2 <- seq(strptime("19970515 12",format="%Y%m%d %H"), 
+          strptime("19970518 00",format="%Y%m%d %H"), by = "12 hours")
+## Cyclone Sidr, November 15, 2007
+t3 <- seq(strptime("20071114 12",format="%Y%m%d %H"), 
[…]

### 2017-08-10 SPECS/specs-decade-mos.R
SPECS MOS R-script

@@ -548,7 +548,6 @@ if (looexperiments) {
   score.fw.y1.9 <- diagnoseMOS('specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)
   while (dev.cur()>1) dev.off()
 ## Repeat and make a table of the results...
-  
 }
   
 

### 2017-08-10 SPECS/specs-decade-mos.Rmd
SPECS MOS R-script

@@
 #library(knitr)
-#purl('git/esd_Rmarkdown/SPECS/specs-decade-mos.Rmd', output='git/esd_Rmarkdown/SPECS/specs-decade-mos.R')
+#purl('~/git/esd_Rmarkdown/SPECS/specs-decade-mos.Rmd', output='~/git/esd_Rmarkdown/SPECS/specs-decade-mos.R')
 #library(ncdf)
@@
   es <- annual(C.C.eq(retrieve('~/Downloads/air.mon.mean.nc',lon=c(-60,40),lat=c(20,70))))
+  index(es) <- year(es)
   es.eof <- EOF(mask(es,land=TRUE))
@@
   slp <- annual(retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
+  index(slp) <- year(slp)
   slp.eof <- EOF(slp)
@@
   load(experiments[iexp])
+  dir <- substr(experiments[iexp],1,nchar(experiments[iexp])-4)
+  if (!file.exists(dir)) dir.create(dir)
 
@@
 ## Names of file names with results stored  
-fw1name <- paste('specsdecadal.z.fw.y1.gcm',iexp,'rda',sep='.')
-mu1name <- paste('specsdecadal.z.mu.y1.gcm',iexp,'rda',sep='.')
-fw3name <- paste('specsdecadal.z.fw.y1.3.gcm',iexp,'rda',sep='.')
-mu3name <- paste('specsdecadal.z.mu.y1.3.gcm',iexp,'rda',sep='.')
-fw5name <- paste('specsdecadal.z.fw.y1.5.gcm',iexp,'rda',sep='.')
-mu5name <- paste('specsdecadal.z.mu.y1.5.gcm',iexp,'rda',sep='.')
-fw9name <- paste('specsdecadal.z.fw.y1.9.gcm',iexp,'rda',sep='.')
-mu9name <- paste('specsdecadal.z.mu.y1.9.gcm',iexp,'rda',sep='.')
+fw1name <- paste(dir,'/specsdecadal.z.fw.y1.gcm.',iexp,'rda',sep='')
+mu1name <- paste(dir,'/specsdecadal.z.mu.y1.gcm.',iexp,'rda',sep='')
+fw3name <- paste(dir,'/specsdecadal.z.fw.y1.3.gcm.',iexp,'rda',sep='')
+mu3name <- paste(dir,'/specsdecadal.z.mu.y1.3.gcm.',iexp,'rda',sep='')
+fw5name <- paste(dir,'/specsdecadal.z.fw.y1.5.gcm.',iexp,'rda',sep='')
+mu5name <- paste(dir,'/specsdecadal.z.mu.y1.5.gcm.',iexp,'rda',sep='')
+fw9name <- paste(dir,'/specsdecadal.z.fw.y1.9.gcm.',iexp,'rda',sep='')
+mu9name <- paste(dir,'/specsdecadal.z.mu.y1.9.gcm.',iexp,'rda',sep='')
 

### 2017-08-10 SPECS/specs-decade-mos.R

@@ -1,17 +1,20 @@
 ## ----setup, include=FALSE------------------------------------------------
-#knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE)
 
 ## ----initialise----------------------------------------------------------
 ## Script that applies MOS to SPECS decadal forecasts
 
 library(esd)
+#library(knitr)
+#purl('~/git/esd_Rmarkdown/SPECS/specs-decade-mos.Rmd', output='~/git/esd_Rmarkdown/SPECS/specs-decade-mos.R')
 #library(ncdf)
 setwd("~/R")
+experiments <- c('~/data/gfdl_DECA.rda','~/data/hadcm3_DECA.rda','~/data/miroc5_DECA.rda','~/data/mpi_DECA.rda')
 
 downloadfromunican <- FALSE    # Preparations - get the data
-downscaledreanalysis <- FALSE  # SM: Assessment of large-small scale link
-looexperiments <- FALSE        # Cross-validation MOS on forecasts
-diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts
+downscaledreanalysis <- TRUE  # SM: Assessment of large-small scale link
+looexperiments <- TRUE        # Cross-validation MOS on forecasts
+diagnoseFCST <- TRUE          # Show correlation & common EOFs of reanalysis+fcsts
 
 source("/home/rasmusb/disk1/SPECS/R/udg2esd_met.R") ## Upload udg2esd function
 
@@ -64,6 +67,7 @@ Xpredictor <- function(x,m=1:3,it=NULL,is=NULL,mask=FALSE,
   ## specifies e.g. months
   ## Get the dimensions of the data structure
   d <- dim(x$Data)
+  nmembers <- length(x$members)
   if (verbose) {print(summary(x)); print(d)}
   ## initialisations members times nx ny
 
@@ -73,7 +77,7 @@ Xpredictor <- function(x,m=1:3,it=NULL,is=NULL,mask=FALSE,
     for (ii in m) {
       if (verbose) print(paste('member',im,'initialisation',ii))
       ## Read the entire field
[…]

### 2017-08-10 SPECS/specs-decade-mos.Rmd
SPECS MOS R-script

@@
   mu.eof <- EOF(mu.eobs)
+  index(mu.eof) <- year(mu.eof)
   plot(mu.eof,colbar=list(pal='t2m',breaks=seq(-0.02,0.02,by=0.001),type='fill',rev=TRUE),new=FALSE)
@@
   fw.eof <- EOF(fw.eobs)
+  index(fw.eof) <- year(fw.eof)
   plot(fw.eof,colbar=list(pal='t2m',breaks=seq(-0.01,0.015,by=0.001),type='fill',rev=TRUE),new=FALSE)
@@
 
-for (iexp in 1:1) {
+for (iexp in 1:4) {
   print(paste('Get the decadal predictors saved locally in',experiments[iexp]))

### 2017-08-10 SPECS/specs-decade-mos.R

@@ -1,5 +1,5 @@
 ## ----setup, include=FALSE------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE)
+##knitr::opts_chunk$set(echo = TRUE)
 
 ## ----initialise----------------------------------------------------------
 ## Script that applies MOS to SPECS decadal forecasts
@@ -421,12 +421,14 @@ attr(fw.eobs,'unit') <- 'fraction'
 ## Display the downscaled reanalysis data for assessing link between large and small scales  - in the shape of EOFs
 if (downscaledreanalysis) {
   mu.eof <- EOF(mu.eobs)
+  index(mu.eof) <- year(mu.eof)
   plot(mu.eof,colbar=list(pal='t2m',breaks=seq(-0.02,0.02,by=0.001),type='fill',rev=TRUE),new=FALSE)
   figlab('EOBS',ypos=0.999)
 
 ## The wet-day frequency:
   
   fw.eof <- EOF(fw.eobs)
+  index(fw.eof) <- year(fw.eof)
   plot(fw.eof,colbar=list(pal='t2m',breaks=seq(-0.01,0.015,by=0.001),type='fill',rev=TRUE),new=FALSE)
   figlab('EOBS',ypos=0.999)
 
@@ -489,16 +491,16 @@ for (iexp in 1:4) {
 
 ## Leave-one-out cross-validation experiments
 ## Names of file names with results stored  
-  fw1name <- paste(dir,'/specsdecadal.z.fw.y1.gcm.',iexp,'rda',sep='')
-  mu1name <- paste(dir,'/specsdecadal.z.mu.y1.gcm.',iexp,'rda',sep='')
-  fw3name <- paste(dir,'/specsdecadal.z.fw.y1.3.gcm.',iexp,'rda',sep='')
-  mu3name <- paste(dir,'/specsdecadal.z.mu.y1.3.gcm.',iexp,'rda',sep='')
-  fw5name <- paste(dir,'/specsdecadal.z.fw.y1.5.gcm.',iexp,'rda',sep='')
-  mu5name <- paste(dir,'/specsdecadal.z.mu.y1.5.gcm.',iexp,'rda',sep='')
-  fw9name <- paste(dir,'/specsdecadal.z.fw.y1.9.gcm.',iexp,'rda',sep='')
[…]

### 2017-08-15 EU-Circle/eu-circle_case-study-3.Rmd
EU-Circle case study 3 - updated analysis

@@
 
+**Summary**
+
+Synoptic storms (low-pressure systems) represent an important factor that may affect flooding and innuation in the Torbay region, as they are both associated with precipitation, wind and storm surges.
+Their influence depends on their intensity as well as the amount of rain that they bring.
+The storm statistics exhibits a seasonal dependency, with more frequent and intense storms during autumn and winter.
+The number of storms and intensity are sensitive to sea surface temperatures and large-scale atmospheric conditions such as the North Atlantic Oscillation and the location of the upper air jet.
+The severity of storm surges that they cause is increased with a sea-level rise.
+
+Other phenomena that may cause local flooding can be connected to convective events and cloud bursts.
+Observations so far have suggested that the amount of precipitation from such local phenomena has been secondary to synoptic storms for places like Torbay.
+
 **Analysis** 
@@
 ```
+
+# Definition of functions
 
@@
 
-**Wind**
-
-Synoptic storms are associated with winds, storm surges and waves.
-
-```{r}
-load('cyclones.uk.rda')
-map(cyclones.uk,new=FALSE)
-```
-
-Plot the number of trajectories over time
-
-```{r}
-plot(cyclones.uk,new=FALSE)
-```
-
----------------------------------
-
 Function to estimate probabilty for daily amount of exceeding a threshold x0 given that the distribution follows an exponential distribution and the records of the wet-day mean precipitation and frequency.
@@
 
-Load the data and weed out stations with lots of missing values
+# The data analysis
+
+## Precipitation
+
+First, we examine historical observed precipitation to get a picture of what has happened until now.
+Global warming has taken place for a few decades (especially since the 1980s), and we ought to see some trends already as a concequence of an increased greenhouse effect.
+Past changes will give some indication of how sensitive the local consequences are to the large-scale changes.
+The following lines load the data and weed out stations with lots of missing values
 
@@
 
-Fill in missing values, but since there are too many missing values, we need to get some more data from the region.
+Need to fill in missing values to proceed with the analysis, but since there are too many missing values, we need to get some more data from the region.
 This is ok since we are working with annual aggregated values, which may imply larer spatial scales.
@@
 
-supplement these to the MIDAS data:
+supplement ECA\&D to the MIDAS data:
 
@@
 
+The five stations with the highest mean wet-day precipitation suggest a decline in the intensity between 1960-2010, but most of the stations with smaller means suggest an increase over the same period.
+
 ```{r}
-plot(fw); grid()
-for (i in 1:dim(mu)[2]) lines(trend(mu[,i]))
+plot(fw,new=FALSE); grid()
+for (i in 1:dim(mu)[2]) lines(trend(fw[,i]))
 ```
+
+The number of rainy days appears to be fairly stable with a frequency of around 0.35, or 120-130 days per year (it may rain for only a fraction of the day).
 
@@
 Pr.gt.30 <- PrexpPr(mu,fw,x0=30)
-plot(Pr.gt.30); grid()
+plot(Pr.gt.30,new=FALSE); grid()
 for (i in 1:dim(mu)[2]) lines(trend(Pr.gt.30[,i]))
@@
 
-Another attempt to search for systematic influences on the precipitation statistics ($\mu$) is to examine the mean seasonal cycle. This type of approach may
+Another attempt to search for systematic influences on the precipitation statistics ($\mu$) is to examine the mean seasonal cycle.
+This type of approach may give som indication of how sensitive precipitation is to changing conditions.
 
@@
 ## Show the mean seasonal cycle in the wet-day mean precipitation
-plot(MU)
+plot(MU,new=FALSE)
 grid()
@@
 There is some indication of a mean seasonal cycle in $\mu$, with minimum values in late winter spring and higher values in late summer and autumn.
+This suggest that the intensity is driven more by precipitation associated with autumn storms (low-pressure systems) than summer-time convective events.
 
@@
 ## Show the mean seasonal cycle in the wet-day frequency
-plot(FW)
+plot(FW,new=FALSE)
 grid()
@@
 
-There is a clear seasonal cycle in the wet-day occurrence - more rainy days in the uatumn.
+There is a clear seasonal cycle in the wet-day occurrence - more rainy days in the autumn.
+This too may be connected to storms.
+The minimum tends to be in June-August when convective events are expected to be more dominant.
 
@@
 attr(MP,'unit') <- 'mm/month'
-plot(MP)
+plot(MP,new=FALSE)
 grid()
@@
 pca.MU <- PCA(MU,n=10)
-plot(pca.MU)
+plot(pca.MU,new=FALSE)
 ```
@@
 Most stations have the same features when it comes to a mean seasonal cycle in $\mu$, with the leading PCA accounting for 80% of the variance.
+There is some variations in the amplitude from place to place, and the mean precipitation is influenced by geography and the local environment.
 
@@
 
-## EOBS gridded data
+Empirical-statistical downscaling of the wet-day mean precipitation (intensity) suggests a slight increase for the future, although the analysis is very simple and far from perfect.
+The method is described in Benestatet al.
+(2017) <http://www.nat-hazards-earth-syst-sci.net/17/993/2017/>.
+This method was appropriate for convective events, but not suitable for precipitation associated with synoptic storms/low pressure systems (see below for storms).
+
+## EOBS gridded daily precipitation
 
@@
 ```{r predictand, warning=FALSE, message=FALSE}
-pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(-6,-1),lat=c(50,52))
+pr.eobs <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(-6,-1),lat=c(50,51.5))
 mu <- annual(pr.eobs,FUN='wetmean')
@@
 ## The area mean wet-day mean precipitation
-plot(mu,new=FALSE)
+plot(aggregate.area(mu,FUN='mean'),new=FALSE)
 ## The area mean wet-day frequency
-plot(fw,new=FALSE)
+plot(aggregate.area(fw,FUN='mean'),new=FALSE)
 ## The early part of the record looks suspect:
@@
 map(mu,FUN='trend',
-    colbar=list(breaks=seq(-0.12,0.12,by=0.02),pal='t2m',rev=TRUE))
+    colbar=list(breaks=seq(-0.12,0.12,by=0.02),pal='t2m',rev=TRUE),new=FALSE)
 ## Map of trends in wet-day frequency:
 map(fw,FUN='trend',
-    colbar=list(breaks=seq(-0.01,0.01,by=0.001),pal='t2m',rev=TRUE))
+    colbar=list(breaks=seq(-0.01,0.01,by=0.001),pal='t2m',rev=TRUE),new=FALSE)
 ```
@@
 ## The area mean wet-day frequency
-plot(eof.mu)
+plot(eof.mu,new=FALSE)
 
@@
 ## Map of trends in wet-day frequency:
-plot(eof.fw)
+plot(eof.fw,new=FALSE)
 ```
@@
 ## Predictor
+
+The predictor describes larg-scale conditions on which the local precipitation is expected to depend.
+The occurrence (frequency) of precipitation depends om the atmospheric circulation pattern (captured by mean sea-level pressure - SLP - patterns).
+The intensity is assumed to be influenced by the atmospheric humidity, which is influenced by the rate of evaporation and hence maritime temperatures.
 
@@
 
-```{r DS, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
+```{r DS-fw, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
 ## The wet-day frequency and SLP
@@
 plot(ds.fw,new=FALSE)
+```
+
+```{r DS-mu, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
 ds.mu <- DS(subset(eof.mu,ip=1:4),eof.es,eofs=1:7)
@@
 eof.res.mu <- EOF(res.mu)
-plot(eof.res.mu)
+plot(eof.res.mu,new=FALSE)
 ```
@@
 For comparison the original data look like this:
+
 ```{r EOF-for-mu, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
 ## The original of the wet-day precipitation
-plot(EOF(mu))
+plot(EOF(mu),new=FALSE)
 ```
@@
 eof.res.fw <- EOF(res.fw)
-plot(eof.res.fw)
+plot(eof.res.fw,new=FALSE)
 ```
@@
 For comparison the original data look like this:
+
 ```{r EOF-for-fw, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
 ## The original of the wet-day precipitation
-plot(EOF(fw))
+plot(EOF(fw),new=FALSE)
 ```
@@
 
-Analyse the sea-level trends and connection to global temperatures to assess future outcome. First the stations from the GLOSS dataset (global monthly data):
+Coastal innundation is expected to worsen with rising sea-levels, given similar types of mid-latitude storms as at present.
+To explore this type of effect, we can analyse the sea-level trends and connection to global temperatures to assess future outcome.
+First the stations from the GLOSS dataset (global monthly data):
 
@@
 ```{r}
-plot(gloss)
+plot(gloss,new=FALSE)
 grid()
@@
 ```{r}
-plot(newlyn)
+plot(newlyn,new=FALSE)
 grid()
 ```
+
+The curve suggests a longterm increase since the early 20th century, but the data stops before 2000.
 
@@
 sonel <- aggregate(station.sonel(),year,FUN='max',na.rm=TRUE)
-plot(sonel)
+plot(sonel,new=FALSE)
 grid()
@@
 coredata(Z) <- t(t(coredata(Z)) - colMeans(coredata(Z),na.rm=TRUE))
-plot(Z); grid()
+plot(Z,new=FALSE); grid()
 ```
+
+There is a general increasing trend for the majority of the tidal stations, but the picture is clutterend by missing-data gaps, dupious outliers, and intermittent data availability.
 
@@
 zc -> coredata(Z)
-map(Z,pch=19,FUN='trend',cex=-2,add.text=TRUE)
+map(Z,pch=19,FUN='trend',cex=-2,add.text=TRUE,new=FALSE)
 ```
+
+A trend analysis yields inconsistent results because missing-data gaps, of the inclusion of filled-in data by interpolation, and variable data quality.
+The map indicates different trends in the British Channel and the near-by parts of the Bay of Biscay.
 
@@
 
+The PCA suggests stronger amplitude in the interannual variations in the sea-level near the Bay of Biscay, and less variation in the British Channel.
+
 ## Wind
+
+Wind affects wave heights and can cause flooding and innundation through storm surges.
+Wind measurements are often unavailable, but it is possible to reconstruct wind informatnion from barometric pressure - sea-level pressure (SLP).
 
@@
 attr(slp.eobs,'unit') <- 'Pa'
-map(slp.eobs,FUN='trend')
+map(slp.eobs,FUN='trend',new=FALSE)
 ```
+
+The trend analysis indicates non-zero changes near the British Channel, although these may be due to natural variability.
 
@@
 plot(annual(uv,nmin=300),main='zonal and meridional wind components',
-     sub='Estimated from triangulation of mean sea-level pressure from EOBS',zoom=7)
+     sub='Estimated from triangulation of mean sea-level pressure from EOBS',zoom=7,new=FALSE)
 grid()
@@
 
-The estimated annual mean geostrophic wind reveals that something changed between 2006 and 2010. The change seems to have taken place over 4 years.
-
-```{r trend-event-2006, warning=FALSE, message=FALSE}
-plot(annual(uv,nmin=300),main='zonal and meridional wind components',
-     xlim=c(2000,2010),map.show=FALSE)
-grid()
-```
+THe level of the wind components seam fairly stable between 1950 and 2050, albeit with a number of peaks.
 
 This event also is seen in the upper tail of the wind speed distribution.
-There are four features concerning these wind speed estimates: a downward trend from moderate high wind speeds in the 1950s, two years with a large drop in values in the 1970s, a number of years with high values around year 2010, and a slight increasing trend in the upper wind speeds over the record.
+There are four features concerning these wind speed estimates: a downward trend from moderate high wind speeds in the 1950s, two years with a large drop in values in the 1970s, a number of years with low values around year 2010, and a slight decreasing trend in the upper wind speeds over the record.
 
@@
 plot(annual(ff,"quantile",probs=0.98),lwd=2,
-     main='Wind speed: annual 98-percentile (m/s)')
+     main='Wind speed: annual 98-percentile (m/s)',new=FALSE)
 lines(trend(annual(ff,"quantile",probs=0.98)),lty=2)
@@
 plot(annual(dd,FUN='mean'),lwd=2,ylim=c(-180,180),
-     main='Annual mean wind direction (degrees)')
+     main='Annual mean wind direction (degrees)',new=FALSE)
 axis(4,at=c(-180,-90,0,90,180),labels=c('S','E','N','W','S'))
@@
 
-The change in 2006-2010 is also visible in the wind direction.
-This could be an event caused by a change in circulation, but it cluld also be due to changes in instrumentation, data used in the gridding or observational practices.
 It is possible to examine whether there were changes in the criculation by using ESD to compare the windspeeds with SLP from reanalyses.
@@
 ds.u1 <- DS(uam,eof.slp,eofs=1:20)
-plot(ds.u1)
+plot(ds.u1,new=FALSE)
 ```
-
-The results suggest a mismatch between the reanalysis and the EOBS data, and the change after 2005 is not corroberated by other sources. We can therefore say that this feature is suspect.
-
-```{r zonal-wind-check2, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
-## Trend in upper wind direction
-ds.u2 <- DS(subset(uam,it=c(1950,2005)),eof.slp,eofs=1:20)
-plot(ds.u2)
-```
-
-When the test is repeated for the shorter period excluding data after 2005, there is a reasonably good match between the EOBS data and the reanalysis.
-The trend analysis can be repeated with this event excluded
-
-```{r wind-speed-trend-2, warning=FALSE, message=FALSE}
-## Trend in upper wind speed
-plot(subset(annual(ff,"quantile",probs=0.98),it=c(1950,2005)),lwd=2,
-     main='Wind speed: annual 98-percentile (m/s)')
-lines(trend(subset(annual(ff,"quantile",probs=0.98),it=c(1950,2005))),lty=2)
-grid()
-```
-
-The exclusion of the suspicious event also affects the long-term trend estiamte - now it's zero with a slight reduction.
 
@@
 
+The windspeed statistics has a distribution that is close to being Weibul, although with a slight bias.
+
 ```{r reanalysis-wind-speed, warning=FALSE, message=FALSE}
@@
 ```
+
+## Storms
+
+Synoptic storms are associated with winds, storm surges, precipitation, and waves.
+The storms (low pressure systems) tend to appear over the North Atlantic ("cyclogenesis"") between the coast of Newfoundland and the British Isles, and travel eastward and die off once they hit land (Scandinavia).
+These storms are a result of wind shear and temperature differences, due to barolinic instability, and tend to cluset around the upper air jet stream.
+The storm statistics is also influenced by the North Atlantic Oscillation (NAO).
+Since the storms travel a fair distance over the North Atlantic, they are expected to be subject to air-sea heat and moisture exchange before hitting the UK.
+The intensity of the storms may potentially be affected by affected by sea surface temperatures through air-sea interactions - higher temperatures favour more evaporation and more "fuel" for the storms.
+
+```{r}
+load('cyclones.uk.rda')
+map(cyclones.uk,new=FALSE)
+```
+
+A plot of the number of trajectories over time suggests a weak declining trend over 1980-2015 in the number of storms sweeping over the Torbay region.
+
+```{r}
+plot(cyclones.uk,new=FALSE)
+grid()
+```
+
+This curve does not indicate how the mean intensity of the storms has varied over the same period.
+
+```{r}
+cdat <- as.character(cyclones.uk[,2])
+press <- annual(zoo(apply(cyclones.uk[,is.element(colnames(cyclones.uk),'pcent')],1,min),
+             order.by=as.Date(paste(substr(cdat,1,4),'-',substr(cdat,5,6),'-',
+                                    substr(cdat,7,8),':',substr(cdat,9,10),':00',sep=''))),'min')
+plot(press,main='Annual minimum central storm sea-level pressure',ylab='pressure (hPa)')
+grid()
+```
+
+The central pressure is more closely connected to the wind speeds in the storm systems, but the connection between amount of precipitation caused by such storms is less clear.
+The precipitation intensity is expected to be influenced by the air moisture and the rate of evaporation over the ocean where the storms cross.

### 2017-08-15 SPECS/specs-decade-mos.R

@@ -486,7 +486,7 @@ if (downscaledreanalysis) {
 for (iexp in 1:4) {
   print(paste('Get the decadal predictors saved locally in',experiments[iexp]))
   load(experiments[iexp])
-  dir <- substr(experiments[iexp],1,nchar(experiments[iexp])-4)
+  dir <- substr(experiments[iexp],3,nchar(experiments[iexp])-4)
   if (!file.exists(dir)) dir.create(dir)
 
 ## Leave-one-out cross-validation experiments

### 2017-09-12 EU-Circle/eu-circle_case-study-4.Rmd
EU-Circle case study 4

@@
 
-Bangladesh is highly vulnerable to climate induced hazrds and disasters and its coastal part are mostly threatened for the impacts of climate change in case of cyclone, storm surges, flood, salinity and tsunami disaster.
+Bangladesh is highly vulnerable to climate induced hazards and disasters and its coastal part are mostly threatened for the impacts of climate change in case of cyclone, storm surges, flood, salinity and tsunami disaster.
 Even last twenty years the coastal peoples have suffered due to super cyclone SIDR (2007-11-15) and AILA (YYYY-MM-DD) devastations.
@@
 map(pr.bmd,new=FALSE)
+```
+
+```{r}
+plot(pr.bmd,new=FALSE)
 ```
@@
 ```
+
+Check the annual cycle in the tidal time series. Is it caused by seasonally changing wind patterns? That would be reflected in the SLP.
+
+```{r}
+SLP <- retrieve('~/Downloads/slp.mon.mean.nc',lon=c(75,105),lat=c(5,30))
+amslp <- mask(aggregate(SLP,month,FUN='mean'),land=TRUE)
+eof.amslp <- EOF(amslp,n=3)
+plot(eof.amslp)
+```
+
+Examine the seasonal cycle in the tidal data: extract the mean seasonal cycle using `aggregate`
+
+```{r}
+amsl <- aggregate(Z,month,'mean')
+pca.amsl <- PCA(amsl,n=3)
+plot(pca.amsl)
+```
+
+Use `DS` to apply a multiple regression analysis to the PCAs and EOFs describing the seasonal cycle.
+This is ok for short series when we know the connection is due to known physics - shifting wind patterns:
+
+```{r, fig.height=8}
+ds.asl <- DS(pca.amsl,eof.amslp,n=3)
+plot(ds.asl)
+```
+
+The seasonal variations are indeed caused by seasonally shifting winth patterns.
+This gives an idea of the sensitivity of the local sea level to changing winds in general, as the physical forces will be similar.
+We cal downscale the SLP from GCMs in multi-model ensembles to try to get a robust picture.
+There may be some problems in the SLP simulation over the Indian sub-continent, which means that we may need to mask out the land.

### 2017-09-12 SPECS/specs-decade-mos.R

@@ -413,6 +413,8 @@ fw.eobs <- retrieve('~/fw_0.25deg_reg.nc')
 attr(mu.eobs,'variable') <- 'mu'
 attr(fw.eobs,'variable') <- 'fw'
 attr(fw.eobs,'unit') <- 'fraction'
+attr(mu.eobs,'source') <- 'EOBS'
+attr(fw.eobs,'source') <- 'EOBS'
 #index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
 #index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
 

### 2017-09-15 EU-Circle/eu-circle_case-study-4.Rmd
EU-Circle case study 4
mimoza + Bangladesh

@@
 
-Correlation between storm tracks and stations.
+Look at correlation between storm tracks and stations.
+Also see if there is a connection between wind speed, cycvlone size, etc and the temperature (Indian Ocean, global mean).
+Possible to estimate a scaling function that can be applied to historical storms?
+Based on a large sample of data, such as the North Atlantic?
 
@@
 There may be some problems in the SLP simulation over the Indian sub-continent, which means that we may need to mask out the land.
+
+We can use similar approach as in Benestad et al.
+(2017) <http://www.nat-hazards-earth-syst-sci.net/17/993/2017/> to downscale the local sea level.
+This, however, does not account for the global increase, which needs to be analysed seperately and added to this.

### 2017-09-15 MIMOZA/mimoza.Rmd
mimoza + Bangladesh

---
title: "MIMOZA"
author: "Rasmus Benestad"
date: "September 14, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r esd}
library(esd)
```

## Tidal data and sea level

You can also embed plots, for example: <http://www.psmsl.org/data/obtaining/>,

```{r sealevel, echo=FALSE}
sl.maputo <- read.table('http://www.psmsl.org/data/obtaining/met.monthly.data/986.metdata',sep=';')
sl.island <- read.table('http://www.psmsl.org/data/obtaining/met.monthly.data/1054.metdata',sep=';')
```

## IMILAST storm tracks

```{r}
if (!file.exists('mimoza.imilast.rda')) {
  imilast.storms <- read.imilast('~/data/IMILAST/ERAinterim_0.75_SH_M07_19790101_20091231.txt',lon=c(30,90),lat=c(-50,-5))
  cyclones <- events2trajectory(imilast.storms)
  save(cyclones,file='mimoza.imilast.rda')
} load('mimoza.imilast.rda')
```

## GHCND temperature and precipitation

```{r}
## get the GHCND data
if (!file.exists('mimoza.ghcnd.rda')) {
  ## Daily maximum temperature
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='tmax')
  tmax <- station(ss,verbose=TRUE)
  ## Daily minimum temperature
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='tmin')
  tmin <- station(ss,verbose=TRUE)
  ## Daily precipitation
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='precip')
  precip <- station(ss,verbose=TRUE)
  ## Set bad (missing) data to NA
  z <- coredata(tmax); z[z < -20] <- NA; z[z > 70] <- NA; z -> coredata(tmax)
  z <- coredata(tmin); z[z < -20] <- NA; z[z > 70] <- NA; z -> coredata(tmin)
  z <- coredata(precip); z[z < 0] <- NA; z[z > 700] <- NA; z -> coredata(precip)
  ## Save in R-binary for convenience.
  save(tmax,tmin,precip,file='mimoza.ghcnd.rda')
} else load ('mimoza.ghcnd.rda')
```

### 2017-09-15 EU-Circle/bangladesh-imilast.R

@@ -4,7 +4,7 @@
 
 library(esd)
 # 03 000001 001 1989010100 1989    1    1   0   -170   61.5    961.24
-imilast.storms <- '~/disk1/IMILAST/imilast.cyclones/ERAinterim_1.5_NH_M03_19890101_20090331_ST.txt'
+imilast.storms <- '~/Disk1/IMILAST/imilast.cyclones/ERAinterim_1.5_NH_M03_19890101_20090331_ST.txt'
 Z <- events2trajectory(read.imilast(imilast.storms))
 ##Z <- subset(Z,is=list(lon=c(80,120),lat=c(10,30)))
 map(Z)

### 2017-09-20 R3/R3-WP4/R3_ESD_NVE.Rmd
R3 folder added into esd_Rmarkdown

Worshop on ESD 4 NVE 
-----------------------------------
title: "esd4nve_examples"
author: "Abdelkader Mezghani"
date: "March 11, 2016"
output: html_document
---

This is a generated R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

---------------------------------

## **Data & Structure**

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/DATA/data_esd_2016/fwddataforr3workshopenom2uker")
```

## Require the following libraries
```{r required R packages,echo=FALSE}
require(esd)
require(leaflet)
require(sp)
require(rgdal)
```

## R3 ESD & NVE
### Function for reading discharge data into the esd format

```{r readNVE,warning=FALSE}
## Create a function that converts NVE format to esd format
readNVE <- function(file='/data_esd_2016/atnsjo.csv',loc='Atnsjo',lat='61',lon='10',
                    alt=1204,stid = '2.32',plot=TRUE,...) {

  data <- read.csv(file=file,header=FALSE, comment.char="#", 
                   dec=".", sep=" ", stringsAsFactors=FALSE)

  dates <- as.Date(data$V1,format="%Y%m%d")
  xz <- zoo(data$V2,order.by=dates)

  y <- as.station(x=xz,loc=loc,lon=lon,lat=lat,alt=alt,stid=stid,
                  param="Q",
                  unit="m/s^{-2}",
                  longname="Discharge data",...)
  y[y==-9999] <- NA ## Replace missing values with NA	
  invisible(y)
}
```

#### Read meta data from the shape file

```{r}
library(rgdal)
library(leaflet)

rsf <- readOGR(dsn = 'Shape_filer_R3_WP4/R3_ESD_nbf_gruppe1.shp',layer = 'R3_ESD_nbf_gruppe1', GDAL1_integer64_policy = TRUE)
rsf.lonlat <- spTransform(rsf,CRSobj = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

str(rsf)

df <- data.frame(rsf@data, stringsAsFactors = FALSE)

xy.coords <- as.data.frame(t(apply(as.matrix(1:18), 1, function(i) unlist(rsf@polygons[[i]]@labpt))))
colnames(xy.coords) <- c('x','y')
coordinates(xy.coords) <- ~ x + y
#CRS(xy.coords) <- rsf@proj4string
proj4string(xy.coords) <- rsf@proj4string

lonlat.coords <- spTransform(xy.coords,CRSobj = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

lonlat <- attr(unclass(lonlat.coords),'coords')

meta <- data.frame(station_id = as.character(df$STASJON_NR), location = as.character(df$ST_NAVN), longitude = lonlat[1,] ,latitude = lonlat[,2])

icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = 'orange')

leaflet(rsf.lonlat) %>% 
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  addPolygons(data = rsf.lonlat,stroke = TRUE, weight = 1) %>%
  addAwesomeMarkers(lng = lonlat[,1] , lat = lonlat[,2],icon = icons)

```

#### Reading data into esd format
```{r,warning=FALSE} 
library(esd)

if (!file.exists('R3_WP4_Qdata_sep2017/q.nve.rda')) {
  lf <- list.files('R3_WP4_Qdata_sep2017',pattern = '.csv',full.names = TRUE) ## remember to update the file path name
  readName <- function(lf) {
    namefromfile <- function(i) unlist(strsplit(basename(lf),split = '_')[[i]][2])
    fname <- apply(as.matrix(1:length(lf)),1,namefromfile)
    invisible(substr(fname,1,nchar(fname) - 4))
  }

  locs <- tolower(rsf$ST_NAVN)
  staID <- substr(rsf$STASJON_NR,1,nchar(as.character(rsf$STASJON_NR))-2)
  ## Display list of discharge stations 
  listStationName <- readName(lf)
  for (i in 1 : length(lf)) {
    id <- grep(as.character(staID[i]), tolower(basename(lf)))
    ifile <- lf[id]
    cat(i, tolower(locs[i]),ifile,sep = '\n')
    #stopifnot(length(id) == 1)
    x <- readNVE(file= ifile,loc=locs[i],lon=lonlat[i,1],lat=lonlat[i,2], stid = staID[i], alt=NA , cntr='norway')
    if (i ==1) X <- x  else X <- combine.stations(X,x)
  }

  # Visual ckeck of the results
  plot(X,plot.type = 'mult',ylab = toupper(substr(locs,1,5)),cex.lab = 0.8,new = FALSE)

  # save to file 
  q.nve <- X 
  save(q.nve,file = 'R3_WP4_Qdata_sep2017/q.nve.rda')
} else 
  load('R3_WP4_Qdata_sep2017/q.nve.rda')
```

#### Q Data Analysis
```{r data Analysis, warning=FALSE,eval=TRUE, eval=TRUE, warning=FALSE}
setwd('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/')
load('q.nve.rda')
library(leaflet)

content <- paste(sep = "<br/>", paste("<b><a>",loc(q.nve),"</a></b>"))

# plot(climatology(as.monthly(q.nve)),new=FALSE) ## Plot the climatology/hydrograph

# compute annual max
Qall.ann.max <- annual(q.nve,FUN="max")
plot(Qall.ann.max,plot.type = 'mult')
grid()

# compute annual sum
Qall.ann.sum <- annual(q.nve,FUN="sum")
plot(Qall.ann.sum,plot.type = 'mult')
plot(Qall.ann.sum,plot.type = 'mult')

# perform a quick trend analysis
Qall.tr.ann.max <- apply(Qall.ann.max,2,FUN='trend.coef',na.rm=TRUE)
Qall.mean.ann.max <- apply(Qall.ann.max,2,FUN='mean',na.rm=TRUE)
tr.rel <- round(Qall.tr.ann.max / Qall.mean.ann.max *100,digits = 1) ## in % per decade

# test significancy
Qall.pval.ann.max <- round(apply(Qall.ann.max,2,FUN='trend.pval',na.rm=TRUE),digits = 2)
q.nve.sig <- subset(q.nve, is = which(Qall.pval.ann.max <= 0.05))

# Construct the map
content <- paste(sep = "<br/>", paste("<b><a>",toupper(loc(q.nve)),"</a></b>"), paste(as.character(tr.rel)," % "))
pal <- colorBin(rev(colscal(20,'t2m')),c(-10,10),bins = 8, pretty = TRUE)
df <- data.frame(lon = lon(q.nve),lat = lat(q.nve), trend = as.numeric(tr.rel))

leaflet() %>% 
  addProviderTiles(providers$Esri.WorldStreetMap,
                   options = providerTileOptions(noWrap = TRUE)) %>%  
  addCircleMarkers(lng  = df$lon, lat = df$lat, radius = 8, fillOpacity = 0.8,
                   stroke = TRUE, color = 'black', weight = 1, fill = TRUE, fillColor = pal(df$trend), popup = content) %>%
  addCircleMarkers(lng  = lon(q.nve.sig), lat = lat(q.nve.sig), radius = 2, fillOpacity = 0.8, stroke = TRUE, color = 'black', weight = 1, fill = TRUE, fillColor = 'black') %>%
  addLegend(position = 'bottomleft',pal = pal, values = pal(df$trend), 
            title = 'Trends in Q [%]')

```

## Downscaling monthly totals single discharge data @ Atnasjø

```{r,warning=FALSE,eval= FALSE}
library(esd)
load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
selQ <- subset(q.nve,is = 1)
```

## Search the nearest meteorological station 
```{r}
# select nearest preciptiation stations
path.data <- '~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017'

ssP <- select.station(src='metnod',param=c('precip'))

if (!file.exists(file.path(path.data,'selNearPre.rda'))) {
  mdis <- distAB(lon(selQ),lat(selQ),ssP$longitude,ssP$latitude)
  # 20 closest stations
  selP <- subset(as.data.frame(ssP), subset = is.element(order(mdis),c(1:20)))
  PRE <- station.metnod(stid = selP$station_id,param='precip',user = 'metno')
  save(PRE,file = file.path(path.data,'selNearPre.rda'))
} else {
  load(file.path(path.data,'selNearPre.rda'))
}

# Pgood <- clean.station(P,miss = 0.1,verbose = FALSE)
# plot(Pgood)
# title(paste(loc(Pgood)))
```

### Select nearest temperature station
```{r selNearT}
path.data <- '~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017'

ssT <- select.station(src='metnod',param=c('t2m'))

if (!file.exists(file.path(path.data,'selNearT2m.rda'))) {
  # select nearest temperature station
  mdis <- distAB(lon(selQ),lat(selQ),ssT$longitude,ssT$latitude)
  selT <- subset(as.data.frame(ssT), subset = is.element(order(mdis),c(1:20))) # top 20 stations
  T2m <- station.metnod(stid = selT$station_id, param = 't2m',user = 'metno')
  plot(T2m,plot.type = 'mult')
  Tgood <- subset(T2m, is = 16)
  #Tgood <- clean.station(T2m,miss = 0.1,verbose = FALSE)
  plot(Tgood)
  # title(paste(loc(Tgood)))
  save(T2m, file = file.path(path,'selNearT2m.rda'))
} else {
  load(file.path(path.data,'selNearT2m.rda'))
}
```

### Produce map with all layers
```{r selNearP}
labq <- paste(sep = "<br/>", paste("<b><a>",loc(q.nve),"</a></b>"))
labP <- paste(sep = "<br/>", paste("<b><a>", ssP$location,"</a></b>"))
labT <- paste(sep = "<br/>", paste("<b><a>", ssT$location,"</a></b>"))

leaflet() %>% 
  addProviderTiles(providers$Esri.WorldStreetMap,group = 'streetmap') %>%
  addPolygons(data = rsf.lonlat,stroke = TRUE, weight = 1 , group = 'NVE catchments') %>%
  setView(lng = 15, lat = 65, zoom = 4) %>%
  addMarkers(lng = lon(q.nve),lat = lat(q.nve), icon = icons, group = 'Discharge (red)',
             popup = labq) %>%
  #addAwesomeMarkers(lng = lon(q.nve),lat = lat(q.nve), icon = icons, group = 'Nearest') %>%
  addCircleMarkers(lng = ssT$longitude,lat = ssT$latitude, stroke = TRUE, weight = 1, 
                   radius= 3, color = 'orange', popup = labT, group = 'Thermometers (orange circles)') %>%
  addCircleMarkers(lng = ssP$longitude,lat = ssP$latitude, stroke = TRUE, weight = 1, 
                   radius= 3, color = 'blue', popup = labP, group = 'Raingauges (blue circles)') %>%
  addLayersControl(overlayGroups = c("Discharge (red)","Raingauges (blue circles)","Thermometers (orange circles)","NVE catchments"), options = layersControlOptions(collapsed = FALSE))
```

```{r}
load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
q1 <- subset(q.nve,is = 1)
Pmm <- subset(as.4seasons(precip.ERAINT(lon = c(0,30), lat = c(50,85))),it = 'djf') 
# use already defined predictors or download from https://climexp.knmi.nl/selectfield_rea.cgi?id=someone@somewhere
uam <- subset(as.4seasons(q1),it = 'djf') # predictand
ds.uam <- DS(uam,EOF(Pmm))
plot(ds.uam)
```

```{r}
load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
q1 <- subset(q.nve,is = 1)
Pmm <- subset(as.annual(subset(precip.ERAINT(lon = c(0,30), lat = c(50,85)),it = 'djf'),nmin=0),it = c(1979,2013)) 
# use already defined predictors or download from https://climexp.knmi.nl/selectfield_rea.cgi?id=someone@somewhere
uam <- subset(as.annual(subset(q1,it = 'djf'),nmin=0),it=c(1979,2013)) # predictand
ds.uam <- DS(uam,EOF(Pmm))
plot(ds.uam)
```

```{r}
load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
q1 <- subset(q.nve,is = 1)
Pmm <- subset(as.annual(subset(t2m.ERAINT(lon = c(0,30), lat = c(50,85)),it = 'djf'),nmin=0),it = c(1979,2013)) 
# use already defined predictors or download from https://climexp.knmi.nl/selectfield_rea.cgi?id=someone@somewhere
uam <- subset(as.annual(subset(q1,it = 'djf'),nmin=0),it=c(1979,2013)) # predictand
ds.uam <- DS(uam,EOF(Pmm))
plot(ds.uam)
```

```{r monthly flow prediction, eval = TRUE}
# find and extract common dates

# set the common time Period
commonPeriod <- c("1979-01-01","2002-03-31")
t2m <- subset(Tgood, it = commonPeriod)
precip <- subset(Pgood, it = commonPeriod)
q <- subset(selQ, it = commonPeriod)

# Perform a correlation map with large scale features.
T2M <- t2m.ERAINT(lon = c(0,30), lat = c(50,85))
map(T2M)
T2M <- subset(T2M,it = commonPeriod)
corT <- corfield(as.monthly(q),T2M)

PRE <- precip.ERAINT(lon = c(0,30), lat = c(50,85))
PRe <- subset(PRE, it = commonPeriod)
PRe.jan <- subset(PRe, it = 'jan')
q.jan <- subset(as.monthly(q), it = 'jan')
corP <- corfield(q.jan, PRe.jan,verbose = TRUE)

# plot the time series
Y <- combine.stations(precip,t2m,q)
plot(Y, plot.type = 'mult')

P <- as.zoo(precip)
t2m <- as.zoo(t2m)
u <- as.zoo(q)

par(bty="n")
plot(merge(P,u,as.anomaly(t2m)),plot.type='single',col=c("steelblue","black","red"),lwd=c(3,1,2))
dev2bitmap('hydroex-daily.png',res=150)

Pmm <- aggregate(P,as.yearmon,FUN='mean')
Tmm <- aggregate(t2m,as.yearmon,FUN='mean')
umm <- aggregate(subset(u, it = commonPeriod),as.yearmon,FUN='mean')

Pmm <- nearest.field(PRe, is = umm)
plot(Pmm)
Tmm <- nearest.field(T2m, is = umm)
plot(Tmm)

#plot(coredata(P),coredata(u))

cal <- data.frame(u=coredata(umm),t2m=coredata(Tmm),Pmm=coredata(Pmm))
fit <- lm(u ~ t2m + Pmm,data=cal)
print(summary(fit))
plot(coredata(umm),predict(fit,newdata = cal))
cor(coredata(umm),predict(fit,newdata = cal), use = 'complete.obs')
par(bty="n")
plot(coredata(umm),predict(fit,newdata = cal),
     main='Monthly river flow predicted from monthly T and P',
     ylim=range(coredata(Pmm)),
     xlab=expression(u[hydro]),ylab=expression(u==f(T[2*m],P)))
lines(range(coredata(Pmm)),range(coredata(Pmm)),col="red",lty=2)
grid()
dev2bitmap(file='hydroex.png',res=150)

temp <- as.station(t2m,lon=lon,lat=lat,alt=alt,unit='deg C',param='t2m',loc=loc,stid=stid)

if (FALSE) {
  trace(z.u <- try(DSensemble.t2m(umm, path = '~', predictor='data/ERA40/ERA40_precip_mon.nc', biascorrect=TRUE,plot=TRUE)))
  dev2bitmap('hydroex-dse-z.t2m.png',res=150)
  save(file='hydroex-dse-z.t2m.rda',z.t2m)
}

temp <- Tmm
if (FALSE) {
  trace(z.t2m <- try(DSensemble.t2m(temp, path = '~', predictor='~/ERA40_t2m_mon.nc', biascorrect=TRUE,plot=TRUE)))
  dev2bitmap('hydroex-dse-z.t2m.png',res=150)
  save(file='hydroex-dse-z.t2m.rda',z.t2m)
}

precip <- as.station(P,lon=lon,lat=lat,alt=alt,unit='mm/day',param='precip',loc=loc,stid=stid)
if (FALSE) {
  z.pre <- try(DSensemble.precip(precip,biascorrect=TRUE,plot=TRUE))
  dev2bitmap('hydroex-dse-z.pre.png',res=150)
  save(file='hydroex-dse-z.pre.rda',z.pre)
}

flow <- as.station(u,lon=lon,lat=lat,alt=alt,unit='m^3/s',param='u',loc=loc,stid=stid)

muam <- annual(precip,FUN='exceedance',threshold=1)
fwam <- annual(precip,FUN='wetfreq',threshold=1)
uamx <- annual(flow,FUN='count',threshold=30)
attr(uamx,'variable') <- 'n(X > m^3/s)'
attr(uamx,'unit') <- ' '

am <- data.frame(u=coredata(uamx),mu=coredata(muam),fw=coredata(fwam))
emu <- glm(u ~ mu + fw, data=am, family='poisson')

plot(uamx,lwd=3,ylim=c(0,70))
lines(zoo(exp(predict(emu)),order.by=index(uamx)))

text(1980,70,pos=4,'Number of annual events: u > 30 m³/s')
legend(1980,65,c("hydrologic model",expression(f(mu,f[w]))),
       lty=1,col=c("red","black"),lwd=c(3,1),bty="n")
dev2bitmap('hydroex-u.gt.30.png',res=150)

```

```{r}
PRE <- precip.ERAINT(lon = c(0,30),lat =c(50,80))
Q <- subset(q.nve,is = 1)

ds.Q <- DS(subset(annual(Q),it = c(1950,2015)), EOF(subset(annual(PRE),it=c(1950,2015))))
plot(ds.Q)
```
```{r}
cor(subset(annual(Q),it=c(1960,2010)), subset(annual(subset(PRE,is=1)),it = c(1960,2010)),use = 'complete.obs')
```

### 2017-09-20 CixPAG/spell-statistics.Rmd

@@
 
-## Spell sttatistics
+## Spell statistics
 
@@
 ```
+
+The mean summer daily maximum temperature and the length of summer-time warm spells.
 
@@
 
-### Prepare rthe preictand
+### Prepare the predictand
 
@@
 pca.lws <- PCA(pcafill(lws.jja),n=5)
+plot(pca.lws,new=FALSE)
+```
+
+```{r, fig.width=8, fig.height=8}
 pca.tmax <- PCA(pcafill(tmax.jja),n=5)
-plot(pca.lws,new=FALSE)
 plot(pca.tmax,new=FALSE)
@@
 } else load("dse.tmax.india.rda")
-
 ```
@@
 }
+```
+
+```{r}
 
@@
 ## Plot results
+```
 
+```{r}
 plot(x,y,main=paste('Mean',it,'temperature & mean length of intervals',
@@
 lines(range(q.spell,q.geom,na.rm=TRUE),range(q.spell,q.geom,na.rm=TRUE),col='red')
-
 ```

### 2017-09-20 EU-Circle/eu-circle_case-study-1.Rmd

@@
 eof.dry <- subset(eof.dry,ip=1:4)
+index(eof.dry) <- year(eof.dry)
 ```
@@
 
-```{r, height=8}
+```{r, fig.height=8}
 plot(ds.ataf.dry,new=FALSE)
@@
 
-```{r, height=8}
+```{r, fig.height=8}
 plot(ds.ataf.t2m,new=FALSE)
@@
 
-```{r, height=8}
+```{r, fig.height=8}
 plot(ds.ataf.slp,new=FALSE)
@@
 
-```{r}
+```{r, fig.height=8}
 ## Convert the projected data into an EOF for the multi-model ensemble

### 2017-09-20 EU-Circle/eu-circle_case-study-5.Rmd

@@
 
-```{r}
+```{r rain-gauges}
 library(esd)
@@
 ```{r}
-rr <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=range(lon(Y)),lat=range(lat(Y)))
+rr <- retrieve('~/storeB/data/data.ECAD/rr_0.25deg_reg.nc',lon=range(lon(Y)),lat=range(lat(Y)))
 map(rr,new=FALSE)

### 2017-09-20 SPECS/specs-decade-mos.Rmd

@@
 attr(fw.eobs,'variable') <- 'fw'
+attr(mu.eobs,'source') <- 'EOBS'
+attr(fw.eobs,'source') <- 'EOBS'
 attr(fw.eobs,'unit') <- 'fraction'

### 2017-09-20 SPECS/specs-decade-mos2.Rmd

--- 2017-04-28 SPECS/specs-decade-mos.Rmd
@@
     ## member number
+    if (verbose) print(yearY1)
     if (d[2]<=12) index(Y1) <- as.Date(paste(year(Y1),im,'01',sep='-'),
@@
   mu.eobs <-  wetdaymean('~/data/data.ECAD/rr_0.25deg_reg.nc') else
-  mu.eobs <- retrieve('~/data/mu_0.25deg_reg.nc')
-fw.eobs <- retrieve('~/data/fw_0.25deg_reg.nc')
+  mu.eobs <- retrieve('~/storeA/mu_0.25deg_reg.nc')
+fw.eobs <- retrieve('~/storeA/fw_0.25deg_reg.nc')
 attr(mu.eobs,'variable') <- 'mu'
@@
 ## Repeat for SLP
-slp <- retrieve('~/data/ncep/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
+slp <- retrieve('~/data/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
 slp <- anomaly(slp)

### 2017-09-20 R3/R3-WP4/R3_ESD_NVE.Rmd
Updated Rmd script

@@
-Worshop on ESD 4 NVE 
+R3 Workpackage 4 on Extension to hydrological simulations
 -----------------------------------
-title: "esd4nve_examples"
+title: "R3_WP4_Hydrological"
 author: "Abdelkader Mezghani"
@@
 
-## **Data & Structure**
+## Data Reading & Structure
 
-```{r "setup", include=FALSE}
+```{r "setup"}
 require("knitr")
-opts_knit$set(root.dir = "~/DATA/data_esd_2016/fwddataforr3workshopenom2uker")
+opts_knit$set(root.dir = "~/git/esd_Rmarkdown/R3/R3-WP4/")
+opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE) 
 ```
 
-## Require the following libraries
+### Require the following libraries
 ```{r required R packages,echo=FALSE}
@@
 
-## R3 ESD & NVE
 ### Function for reading discharge data into the esd format
@@
 
-#### Read meta data from the shape file
+### Read meta data from the shape file
 
@@
 
-#### Q Data Analysis
+## Discharge Data Analysis
 ```{r data Analysis, warning=FALSE,eval=TRUE, eval=TRUE, warning=FALSE}
-setwd('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/')
+setwd('R3_WP4_Qdata_sep2017')
 load('q.nve.rda')
@@
 
-## Downscaling monthly totals single discharge data @ Atnasjø
+### Downscaling monthly totals @ a single discharge station
 
@@
 library(esd)
-load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
+load('R3_WP4_Qdata_sep2017/q.nve.rda')
 selQ <- subset(q.nve,is = 1)
@@
 ## Search the nearest meteorological station 
-```{r}
+```{r,eval = FALSE}
 # select nearest preciptiation stations
-path.data <- '~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017'
+path.data <- 'R3_WP4_Qdata_sep2017'
 
@@
 ### Select nearest temperature station
-```{r selNearT}
-path.data <- '~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017'
+```{r selNearT,eval = FALSE}
+path.data <- 'R3_WP4_Qdata_sep2017'
 
@@
 ### Produce map with all layers
-```{r selNearP}
+```{r selNearP, eval = TRUE}
 labq <- paste(sep = "<br/>", paste("<b><a>",loc(q.nve),"</a></b>"))
@@
 
+### Downscaling Examples
+#### Example of Downscaling Seasonal values 
 ```{r}
-load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
+load('R3_WP4_Qdata_sep2017/q.nve.rda')
 q1 <- subset(q.nve,is = 1)
@@
 
+#### Example of downscaling annual means using ERAINT precip as predictor
 ```{r}
-load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
+load('R3_WP4_Qdata_sep2017/q.nve.rda')
 q1 <- subset(q.nve,is = 1)
@@
 
-```{r}
-load('~/DATA/data_esd_2016/fwddataforr3workshopenom2uker/R3_WP4_Qdata_sep2017/q.nve.rda')
+#### Example of downscaling annual means using ERAINT T2m as predictor
+```{r,eval = FALSE}
+load('R3_WP4_Qdata_sep2017/q.nve.rda')
 q1 <- subset(q.nve,is = 1)
@@
 
-```{r monthly flow prediction, eval = TRUE}
-# find and extract common dates
+### Downscaling of monthly discharge values # OLD example not run
+```{r monthly flow prediction, eval = FALSE}
+# This is an old example, not working for now !!!
 
@@
 ```
-
-```{r}
-PRE <- precip.ERAINT(lon = c(0,30),lat =c(50,80))
-Q <- subset(q.nve,is = 1)
-
-ds.Q <- DS(subset(annual(Q),it = c(1950,2015)), EOF(subset(annual(PRE),it=c(1950,2015))))
-plot(ds.Q)
-```
-```{r}
-cor(subset(annual(Q),it=c(1960,2010)), subset(annual(subset(PRE,is=1)),it = c(1960,2010)),use = 'complete.obs')
-```

### 2017-09-21 MIMOZA/mimoza.Rmd
added hydroex

@@
 
-## R Markdown
+## Introduction.
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+This is an R Markdown document that explores some climate data with relevance to Mozambique.
+Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
+For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
 
@@
 if (!file.exists('mimoza.imilast.rda')) {
-  imilast.storms <- read.imilast('~/data/IMILAST/ERAinterim_0.75_SH_M07_19790101_20091231.txt',lon=c(30,90),lat=c(-50,-5))
+  imilast.storms <- subset(read.imilast('~/data/IMILAST/ERAinterim_0.75_SH_M07_19790101_20091231.txt'),
+                           is=list(lon=c(20,90),lat=c(-40,-5)))
+# imilast.storms <- subset(read.imilast('~/data/IMILAST/ERAinterim_1.5_SH_M03_19890101_20090331_ST.txt'),
+#                           is=list(lon=c(30,90),lat=c(-40,-5)))
   cyclones <- events2trajectory(imilast.storms)
   save(cyclones,file='mimoza.imilast.rda')
-} load('mimoza.imilast.rda')
+} else load('mimoza.imilast.rda')
+ii <- is.element(colnames(cyclones),'amp_z850')
+mi <- apply(cyclones[,ii],1,min)
+storms <- cyclones[mi < 10,]
+storms <- attrcp(cyclones,storms)
+class(storms) <- class(cyclones)
+col <- rgb(0,0,1,1-(mi-min(mi))/diff(range(mi)))
+map(storms,verbose=TRUE,lonR=30,latR=0,col=col,lwd=3)
+```
+
+```{r}
+pca.storms <- PCA(storms,verbose=TRUE)
+plot(pca.storms[,1],type='l')
+plot(attr(pca.storms,'pattern')[,1],type='l')
+plot(attr(pca.storms,'eigenvalues'),type='b')
 ```
@@
   ## Daily precipitation
-  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='precip')
+  ss <- select.station(lon=c(31,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='precip')
   precip <- station(ss,verbose=TRUE)
@@
 ```
+
+Maximum temperature - mean seasonal cycle
+
+```{r}
+plot(aggregate(tmax,month),new=FALSE)
+plot(aggregate(tmax,month,FUN='sd'),new=FALSE)
+```
+
+Minimum temperature - mean seasonal cycle
+
+```{r}
+plot(aggregate(tmin,month),new=FALSE)
+plot(aggregate(tmin,month,FUN='sd'),new=FALSE)
+```
+
+Precipitation - mean seasonal cycle
+
+```{r}
+plot(aggregate(precip,month,FUN='sum'),new=FALSE)
+plot(aggregate(precip,month,FUN='wetfreq'),new=FALSE)
+plot(aggregate(precip,month,FUN='wetmean'),new=FALSE)
+```
+
+Precipitation - long-term variations
+
+```{r}
+plot(annual(precip,FUN='sum'),new=FALSE); grid()
+plot(annual(precip,FUN='wetfreq'),new=FALSE); grid()
+plot(annual(precip,FUN='wetmean'),new=FALSE); grid()
+```
+
+Spatial patterns of variability in annual precipitation totals
+
+```{r, fig.height=8}
+Y <- subset(annual(precip,FUN='sum'),it=c(1950,1990))
+nv <- apply(coredata(Y),2,'nv')
+Y <- subset(Y,is=nv > 36)
+Y <- pcafill(Y)
+tp.pca<- PCA(Y)
+plot(tp.pca,new=FALSE)
+```
+
+Spatial patterns of variability in annual wet-day frequency
+
+```{r, fig.height=8}
+Y <- subset(annual(precip,FUN='wetfreq'),it=c(1950,1990))
+nv <- apply(coredata(Y),2,'nv')
+Y <- subset(Y,is=nv > 36)
+Y <- pcafill(Y)
+fw.pca<- PCA(Y)
+plot(fw.pca,new=FALSE)
+```
+
+Spatial patterns of variability in annual wet-day mean precipitation
+
+```{r, fig.height=8}
+Y <- subset(annual(precip,FUN='wetmean'),it=c(1950,1990))
+nv <- apply(coredata(Y),2,'nv')
+Y <- subset(Y,is=nv > 36)
+Y <- pcafill(Y)
+mu.pca<- PCA(Y)
+plot(mu.pca,new=FALSE)
+```
+
+## Copernicus land-cover information - 100m resolution over Africa
+
+<http://land.copernicus.eu/global/products/lc>

### 2017-09-21 dipper/dipper.R

@@ -5,6 +5,7 @@ library(ncdf4)
 
 examine <- FALSE
 
+print('read dipper data')
 dipper <- read.table('~/Dropbox/data/dipper.csv',header=TRUE)
 
 dipper <- zoo(x=dipper[[2]],order.by=dipper[[1]])
@@ -18,6 +19,7 @@ predictor <- retrieve("air.mon.mean.nc",lon=c(-1,17),lat=c(57,62))
 predictor <- aggregate(subset(predictor,it='djf'),year,FUN='mean')
 
 if (examine) {
+  print('dipper statistical distribution')
   ## Check the distribution
   n <- seq(0,150,by=10); mu <- mean(dipper)
   hist(coredata(dipper),breaks=n,freq=FALSE,col='grey')
@@ -32,7 +34,7 @@ if (examine) {
 ## Downscale the dipper population directly based on the large-scale
 ## annual mean temperature
 
-dev.new()
+print('Downscaled results')
 if (!file.exists('dipper.Z.rcp45.rda')) {
   Z.rcp45 <- DSensemble.annual(y,biascorrect=TRUE,
                                predictor=predictor,
@@ -57,6 +59,7 @@ if (!file.exists('dipper.Z.rcp26.rda')) {
   save(file='dipper.Z.rcp26.rda',Z.rcp26)
 } else load('dipper.Z.rcp26.rda')
 
+print('ensemble statistics')
 year <- year(Z.rcp45)
 ci90.rcp45 <- apply(coredata(Z.rcp45),1,quantile,
                     probs=c(0.05,0.95),na.rm=TRUE)
@@ -65,6 +68,8 @@ ci90.rcp26 <- apply(coredata(Z.rcp26),1,quantile,
 ci90.rcp85 <- apply(coredata(Z.rcp85),1,quantile,
                     probs=c(0.05,0.95),na.rm=TRUE)
 
+print('plotting')
+dev.new()
 par(bty='n')
 plot(range(year),range(ci90.rcp45,ci90.rcp26,ci90.rcp85,y),
      type='n',xlab='',ylab='Population',main='Dipper',

### 2017-09-21 hydroex/hydroex.R

@@ -1,0 +1,78 @@
+## Rasmus.Benestad@met.no
+## A demonstration showing how river discharge can be downscaled using empirical-statistical downscaling
+
+library(esd)
+# VIKEDAL (met station no. 46850, LAT 59.5558N, LON 5.9955E
+hydroIO <- read.table("~/Dropbox/data/CERAD_to RASMUS METNO.csv",sep=";",header=TRUE)
+loc <- 'Vikedal'
+lon <- 5.9955
+lat <- 59.5558
+alt <- 159
+stid <- 46850
+
+P <- zoo(x=hydroIO$Precipitation..mm,order.by=as.Date(hydroIO$Date))
+t2m <- zoo(x=hydroIO$Temperature.C,order.by=as.Date(hydroIO$Date))
+u <- zoo(x=hydroIO$Flow..m3.s,order.by=as.Date(hydroIO$Date))
+
+par(bty="n")
+plot(merge(P,u,as.anomaly(t2m)),plot.type='single',col=c("steelblue","black","red"),lwd=c(3,1,2))
+dev2bitmap('hydroex-daily.png',res=150)
+
+Pmm <- aggregate(P,as.yearmon,FUN='mean')
+Tmm <- aggregate(t2m,as.yearmon,FUN='mean')
+umm <- aggregate(u,as.yearmon,FUN='mean')
+
+#plot(coredata(P),coredata(u))
+
+cal <- data.frame(u=coredata(umm),t2m=coredata(Tmm),Pmm=coredata(Pmm))
+fit <- lm(u ~ t2m + Pmm,data=cal)
+print(summary(fit))
+
+par(bty="n")
+plot(coredata(umm),predict(fit),
+     main='Monthly river flow predicted from monthly T and P',
+     ylim=range(coredata(Pmm)),
+     xlab=expression(u[hydro]),ylab=expression(u==f(T[2*m],P)))
+lines(range(coredata(Pmm)),range(coredata(Pmm)),col="red",lty=2)
+grid()
+dev2bitmap(file='hydroex.png',res=150)
+
+temp <- as.station(t2m,lon=lon,lat=lat,alt=alt,unit='deg C',param='t2m',loc=loc,stid=stid)
+
+if (FALSE) {
+z.t2m <- try(DSensemble.t2m(temp,biascorrect=TRUE,plot=TRUE))
+dev2bitmap('hydroex-dse-z.t2m.png',res=150)
[…]

### 2017-11-20 R3/R3-WP4/RCubed_wrk_hbe.Rmd
Helene B. E. R3 R-file for gmc and rcm

---
title: "RCubedwrk"
author: "Helene B. Erlandsen"
date: "November 16, 2017"
output: html_document
---

##GCM Files
downloaded from Norstore using python script .... and a modified version of ....py (provided by MP uib) to be compatible with Python version NN (a change in char byte-char handeling needed specification of utf8 format)

##RCM Files
9 years and one month of dynamically downscaled bias corrected NorESM-data have been downloaded and processed using cdo and nco (getRCMfiles.py) to provide files compatible with ESD retrieve.rcm. Note that the calender is noleap or 365day. The files are from the 3-hourly wrfxtrm-output, and has a res of about 4 km covering South Norway.

The processing includes extracting variables t2m,pr,u10,v10,q2,tskin,precip and aggregating to daily and montly means. For all variables but precip the monthly standard deviation is also stored.

Some cdo nco hacking is done to give the files 2-D latitude and longitude variables (cdo does not like wrf).

The daily precipitation is converted to mm per day, to allow aggregation to monthly wet-day mean and wet day frequence with a wet-day threshold of 1 mm/day. Remember weight by noleap month if aggregating from month to season.

The (nc4) RCM files:
wrf_sub3hr_NorESM_BC_2000_2010.nc 16GB compressed,
wrf_day_NorESM_BC_2000_2010.nc 2GB maybe comressed, do not remember
wrf_mon_NorESM_BC_2000_2010.nc 228M uncompressed

They are at lustre under A under my username

## Data Reading & Structure cp from am

```{r "setup"}
#require("knitr")
opts_knit$set(root.dir = "~/git/esd_Rmarkdown/R3/R3-WP4/")
opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE) 
```

### Require the following libraries
```{r required R packages,echo=FALSE}
require(esd)
#require(leaflet)
#require(sp)
#require(rgdal)
```

```{r setpath}
#install.packages("~/git/esd",repos=NULL,type="source")
library(esd)
source('~/git/esd/R/retrieve.R')
rcmdatapath <- '~/storah/RCubed/data/RCM'
gcmdatapath <- '~/storah/RCubed/data/GCMraw'
gcmbcdatapath <- '~/storah/RCubed/data/GCMBC'
eraidatapath<-'~/storah/RCubed/data/ERAI'
RCMmonp <- 'wrf_mon_NorESM_BC_2000_2010.nc' 
GCMBCmon <- 'NorESM_BC4R3_esd_2000_2010.nc'
GCMmon <- 'NorESM_nonBC4R3_esd_2000_2010.nc'
ERApl <- 'EraI_pl_4_esd_mdfa_197901to201706.nc' 
ERAsf <- 'EraI_sfc_4_esd_mdfa_197901to201706.nc'

```

```{r}
gridn <- par(mfrow=c(4, 5))
RCMt2m <- retrieve.rcm(ncfile=RCMmonp, path=rcmdatapath,param='T2MEAN',verbose = TRUE)
# just testing
map(RCMt2m,FUN ="mean")
par(gridn)
```

```{r}
RCMfw <- retrieve.rcm(ncfile=RCMmonp,path=rcmdatapath,param='fw',verbose = TRUE)
gridn <- par(mfrow=c(4, 5))
# just testing
map(RCMfw,FUN ="mean")
par(gridn)
```

```{r loadGMCBC}
GCMBCt2m <- retrieve(ncfile=GCMBCmon, path=gcmbcdatapath,param='tas',verbose = TRUE)
GCMBCt2m<-subset(GCMBCt2m,is=list(c(-20,15),c(45,72)),it=c("2000-12-01","2010-12-01"))
#GCMBCt2 <- attr(psl.nobc,'units') <- 'Pa'
map(GCMBCt2m,FUN ="mean")
```

### 2017-12-13 R3/R3-WP4/RCubed_wrk_hbe.Rmd
updated Rmd-file testing things and added a file with command lists to hack on wrf-files to makethem nice for esd

@@
 date: "November 16, 2017"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 downloaded from Norstore using python script ....
-and a modified version of ....py (provided by MP uib) to be compatible with Python version NN (a change in char byte-char handeling needed specification of utf8 format)
+and a modified version of *wps.py* (provided by MP, UiB) to be compatible with Python 3.6.3 (a change in char byte-char handeling needed specification of utf8-format)
 
@@
 
-Some cdo nco hacking is done to give the files 2-D latitude and longitude variables (cdo does not like wrf).
+For the dynamically downscalled bias corrected NorESM WRF run some cdo nco hacking is done to give the files 2-D latitude and longitude variables and dimensions x,y, and CF-time (cdo does not like wrf).
+The same is done to the WRF file with the geograhical features (geo_em).
+The cdo and nco commands are stored in scripts saved for later.
 
@@
 
-## Data Reading & Structure cp from am
+## Data Reading & Structure (copied parts from Abdelkader)
 
-```{r "setup"}
-#require("knitr")
+```{r setup, include=FALSE}
+require("knitr")
+knitr::opts_chunk$set(echo = TRUE)
 opts_knit$set(root.dir = "~/git/esd_Rmarkdown/R3/R3-WP4/")
@@
 ### Require the following libraries
+Also added a boolean for update esd (provided that a new version is pulled)
 ```{r required R packages,echo=FALSE}
-require(esd)
+needed<-c("ncdf4", "zoo")
+update_esd <- FALSE
+for (package in needed) { 
+  if (!require(package,character.only = TRUE, quietly = TRUE)) { 
+      install.packages(package);library(package,character.only = TRUE)
+    }
+}
+
+if ((!require('esd',character.only = TRUE, quietly = TRUE))| (update_esd) ){ 
+      install.packages("~/git/esd",repos=NULL,type="source");library('esd',character.only = TRUE)
+    }
+#require(esd)  
 #require(leaflet)
@@
 ```{r setpath}
-#install.packages("~/git/esd",repos=NULL,type="source")
+#
 library(esd)
-source('~/git/esd/R/retrieve.R')
+#source('~/git/esd/R/retrieve.R')
 rcmdatapath <- '~/storah/RCubed/data/RCM'
+rcmgeopath <-'~/storah/RCubed/data/RCM/WRF_filetypes'
 gcmdatapath <- '~/storah/RCubed/data/GCMraw'
@@
 eraidatapath<-'~/storah/RCubed/data/ERAI'
+RCMgeo<-'R3geo_emd03run1.nc'
 RCMmonp <- 'wrf_mon_NorESM_BC_2000_2010.nc' 
@@
 
-```{r}
+```{r, warning=FALSE}
 gridn <- par(mfrow=c(4, 5))
-RCMt2m <- retrieve.rcm(ncfile=RCMmonp, path=rcmdatapath,param='T2MEAN',verbose = TRUE)
+RCMt2m <- retrieve.rcm(ncfile=RCMmonp, path=rcmdatapath,param='T2MEAN',verbose = FALSE)
 # just testing
-map(RCMt2m,FUN ="mean")
+attr(RCMt2m,'greenwich')=TRUE
+RCM_oro <- retrieve.rcm(ncfile='geo_em_d03ED.nc',path=rcmgeopath,param='HGT_M',verbose = TRUE)
+
+#map.station(RCM_oro,FUN ="mean",cex=0.3,new=F, is=list(lon=c(3.75,15),lat=c(58,64)))
+
+attributes(RCM_oro) <- NULL
+attr(RCMt2m,'altitude') <- RCM_oro
+
+RCMt2m <- subset(RCMt2m, is=list(lon=c(3.75,15),lat=c(58,64)))
+map.station(RCMt2m,FUN ="mean",cex=0.3,new=F)
+par(gridn)
+
+#check later the stripes
+
+```
+
+```{r,warning=FALSE}
+RCMfw <- retrieve.rcm(ncfile=RCMmonp,path=rcmdatapath,param='fw',verbose = FALSE)
+RCMfw <- subset(RCMfw, is=list(lon=c(3.75,15),lat=c(58,64)))
+gridn <- par(mfrow=c(4, 5))
+# just testing
+map.station(RCMfw,FUN ="mean")
 par(gridn)
@@
 
-```{r}
-RCMfw <- retrieve.rcm(ncfile=RCMmonp,path=rcmdatapath,param='fw',verbose = TRUE)
+```{r loadGMCBC,warning=FALSE,fig.height=32}
+
+GCMBCt2m <- retrieve(ncfile=GCMBCmon, path=gcmbcdatapath,param='tas',verbose = FALSE)
+attr(GCMBCt2m,'Calendar') <- 'noleap'
+GCMBCt2m<-subset(GCMBCt2m,is=list(c(-10,15),c(45,72)),it=c("2000-12-01","2010-12-01"))
+
+map.field(GCMBCt2m,FUN ="mean", new=F)
+
+```
+
+```{r loadGMCnonBC,warning=FALSE}
+#loading non BC data
+GCMt2m <- retrieve(ncfile=GCMmon, path=gcmdatapath,param='tas',verbose = FALSE)
+attr(GCMt2m,'Calendar') <- 'noleap'
+GCMt2m<-subset(GCMt2m,is=list(c(-10,15),c(45,75)),it=c("2000-12-01","2010-12-01"))
+
+#map.field(GCMt2m,FUN ="mean", new=F)
+
+```
+
+```{r loadERAI,warning=FALSE}
+
+ERAIt2m <- retrieve(ncfile=ERAsf, path=eraidatapath,param='t2m',verbose =FALSE)
+ERAIt2m<-subset(ERAIt2m,is=list(c(-10,15),c(45,72)),it=c("2000-12-01","2010-12-01"))
+#GCMBCt2 <- attr(psl.nobc,'units') <- 'Pa'
 gridn <- par(mfrow=c(4, 5))
-# just testing
-map(RCMfw,FUN ="mean")
+map.field(ERAIt2m,FUN ="mean")
 par(gridn)
@@
 
-```{r loadGMCBC}
-GCMBCt2m <- retrieve(ncfile=GCMBCmon, path=gcmbcdatapath,param='tas',verbose = TRUE)
-GCMBCt2m<-subset(GCMBCt2m,is=list(c(-20,15),c(45,72)),it=c("2000-12-01","2010-12-01"))
-#GCMBCt2 <- attr(psl.nobc,'units') <- 'Pa'
-map(GCMBCt2m,FUN ="mean")
+```{r corrfield test, warning=FALSE,eval=FALSE}
+# the station object : pca
+# the field object : eof
+# later check out https://github.com/metno/esd_Rmarkdown/blob/master/KSS/dse4KSS.R
+# 4 similar examples, there Y is station values
+# LTR try DS on EOF RCM predictand on EOF GCM predictor
+# bc EOFs contain pattern of variance biases in different GCMs not very important, and can try same model for all benestad or try leave on out as in Walton
+# latr try common eof append [Re-analysis or GCM-BC | GCM raw] -->time  to find common eofs
+#need to consider that we have a short time-series. maybe need monthly data 12*9, monthly anomalies, or seasonal 4*9
+
+#Regridding and making seasonal instead of monthly to look at EOF
+RCMt2m.regrid <- regrid.station(RCMt2m,is=list(lon=seq(4,15,by=0.1),lat=seq(58,64,by=0.25)),verbose=T)
+# OBS if loaded demos The following object is masked from ‘package:esd’: unit, that is why unit in regrid.station does not work
+RCMt2m.seasons.regrid <- as.4seasons(RCMt2m.regrid)
+# Make EOFs of the RCM 
+eof.test <- EOF(RCMt2m.seasons.regrid)
+plot(eof.test)
+# Make EOF of the seasonal anaomaly (monthly?)
+eof.test <- EOF(as.anomaly(RCMt2m.seasons.regrid))
+plot(eof.test)
+plot(EOF(GCMBCt2m))
+
+eof.gcm <- EOF(as.anomaly(GCMBCt2m))
+plot(eof.gcm)
+plot(eof.gcm,ip=2)
+ds <- DS(eof.test,eof.gcm)
+plot(ds) #r=0.79
+plot(ds,ip=2)
+plot(eof.test,ip=2)
+
+#Slightly better result from BC model (which is boundary for dyn-DS)
+#eof.gcmnoBC <- EOF(as.anomaly(GCMt2m))
+#dsnoBC <- DS(eof.test,eof.gcmnoBC)
+#plot(dsnoBC) #r=0.76
+
+#Also trying ERA-I  really bad bc not source of dyn RCM
+eof.eraI <- EOF(as.anomaly(ERAIt2m))
+#dsEra <- DS(eof.test,eof.eraI)
+#plot(dsEra) #r=0.3
+#dev.new(); plot.zoo(subset(pca.y,ip=1:4))
+#dev.new(); plot.zoo(subset(pca.y,ip=1:4))
+
+RCMt2m.seasons.eof <- EOF(subset(RCMt2m.seasons.regrid,it='djf'),n=3)
+plot(RCMt2m.seasons.eof,ip=1)
+
 ```
+
+```{r pca_on_rcm_seasonal_anomalies}
+#Making seasonal and looking at PCA
+
+# Walton method was PCA on each of 60 warming patterns (not warming pattern here) with regional means removed
+RCMt2m.ses <-as.4seasons(as.monthly(RCMt2m))
+RCMt2m.ses.ano <-as.anomaly(RCMt2m.ses)
+RCMt2m.ses.anomaly.pca <- PCA(RCMt2m.ses.ano,n=3)
+plot.pca(RCMt2m.ses.anomaly.pca, new=F)
+plot.pca(RCMt2m.ses.anomaly.pca, new=F, ip=2)
+plot.pca(RCMt2m.ses.anomaly.pca, new=F, ip=3)
+attr(RCMt2m.ses.anomaly.pca,'greenwich')=TRUE
+#RCMtm.clim <- as.climatology(as.monthly(RCMt2m))
+```
+
+To do the converison of PCA to EOF we need to load some extra packages not loaded by defualt into *esd*.
+We call the demo-scripts for *esd* *gridmap* and *pca2eof* to load the packge *LatticeKrig*, *spam*, *grid*, *maps*, *fields* [*grid::unit* will replace the *esd*-function *unit*]
+
+```{r loadingaux, include=FALSE}
+
+#This this is based on a package from NCAR using kringing taking altitude into account, so need wrf-oro
+# since oro is lacking now making dummy oro
+
+#pca2eof uses etop5 min is around 8-10 km, lower than original grid of 4 km about does it matter?
+reload_auxpckg=F
+if (reload_auxpckg==TRUE){
+  demo(package='esd');demo(gridmap,ask=FALSE);demo(pca2eof,ask=FALSE)
+}
+
+```
+
+```{r rcmpca2eof_cca_gcm_eof}
+RCMt2m.seasons.eof.from.pca <- pca2eof(RCMt2m.ses.anomaly.pca) 
+#Compare not same bca regridding and dummy oro
+#dev.new()
+plot(RCMt2m.seasons.eof.from.pca)
+#dev.new()
+plot(RCMt2m.ses.anomaly.pca)
+
+#vis(RCMt2m.ses.anomaly.pca)
+#Compare
+#dev.new()
+#plot(RCMt2m.seasons.eof,ip=2)
+#dev.new()
+#plot(RCMt2m.seasons.pca,ip=2)
+
+eof.gcm <- EOF(as.anomaly(as.4seasons(GCMBCt2m)))
+cca <- CCA(RCMt2m.ses.anomaly.pca,eof.gcm)
+plot(cca,new=T) #RCMt2m.pca <- PCA(RCMt2m)
+
+```
+
+```{r test combine EI noresm}
+#know from before that ERAI has poor poor temporal covariance as expected with DS norESM
+comb <- combine(as.anomaly(as.4seasons(ERAIt2m)),as.anomaly(as.4seasons(GCMBCt2m)))
+ceof <- EOF(comb, n=5)
+plot(ceof,new=T)
+ds <- DS(RCMt2m.seasons.eof.from.pca,ceof)
+plot(ds, new=T)
+
+```

### 2017-12-13 R3/R3-WP4/Fix_RCM.py

@@ -1,0 +1,127 @@
+###
+# requires ncl and cdo
+# and the first part python module load python at lus
+# helenebe@met.no command list to average 3-hourly 4 km climate data
+# dynamically downscaled using wrf over 9 years + to small files to use for esd
+# will delete 3 hourly files, and keep variables Precip, t2, v10,u10,q2, tsk
+# to daily means -- kg m2 to mm/day for precip
+
+# makes monthly std of t2, v10,u10,q2, tsk
+# makes monthly means of Precip, t2, v10,u10,q2, tsk
+# makes monthly wet day mean and frequency
+
+# finally hacks on dimensions to get lat lon from other wrf file and make file
+# more cf-compatible, at least compatible for retrive.rcm in the
+# R packege 'esd' 
+
+import glob, os
+
+folder=open('foldernameNORESM.txt', 'r').read().replace('\n',"") # filepath given in other file without cuotation marks
+
+year=range(2000,2011,1)#  range(2000,2011,1) ##
+header=open('http_server.txt', 'r').read().replace('\n',"")
+for y in year:
+  dirf=header+folder+str(y) +'/'  #+'/wrfxtrm*'
+  os.system('wget -r --no-parent -A "wrfxtrm*" -nd -nc ' + dirf+ ' .')
+
+# The remaining commands may be put in bash-script or included in
+# python with os.system() or adding cdo and ncl
+# but for transperancy I cp+paste and check files using ncdump and ncview
+# as I go along now and then
+# the commands, filenames and such may not be an order as old script is lost
+# check carefully along the way!
+
+'''
+### merge all files with cat (works cause sorted alphanumerically) 
+cdo -f nc4 cat wrfxtrm_d03* wrf_DS_NorESM_BC_2000_2010.nc
+
[…]
